### 2019-06-04 numerics_paper/numerical_chj.tex
Add outline for numerical paper.

\documentclass{article}

\title{Numerical simulation of the shallow water circular hydraulic jump}

\author{
    David I. Ketcheson \and
    Manuel Quezada de Luna
}
\begin{document}

\maketitle

\begin{abstract}
We study numerical discretrizations of the shallow water equations
for the simulation of the circular hydraulic jump formed
when a vertical fluid jet impacts a horizontal plate.  
We focus on three numerical methods: finite volume Godunov-type
methods using either a Roe or HLLE approximate Riemann solver, and
a continuous Galerkin finite element method based on flux-corrected transport (FCT).
The finite volume methods are employed ona uniform cartesian grid
while the finite element method uses an unstructured triangular grid.
The three methods give completely different results, even for
highly-refined meshes.
\end{abstract}

\section{Introduction}

\section{The shallow water circular hydraulic jump}

\subsection{Semi-analytical solution under rotational symmetry}

\section{Numerical methods}

\subsection{Finite volume methods}

\subsubsection{Roe solver}

\subsubsection{HLLE solver}

\subsection{FCT finite element method}

\section{Numerical results}

\section{Conclusions}

\end{document}

### 2020-06-11 numerics_paper/numerical_chj.tex
Update outline.
Introductory material on the CHJ.

@@
 
+\subsection{The circular hydraulic jump}
+
+Perhaps the first reference to the observation of the circular hydraulic jump
+comes from Lord Rayleigh \cite{rayleigh1914theory}, who wrote that it
+``may usually be seen whenever a stream of water from a tap strikes a horizontal
+surface".  This phenomenon that is familiar in the everyday kitchen sink, is in
+fact highly nonlinear and unintuitive.  Near the jet, the flow is very shallow and
+fast, while further away it is deeper and slower.  The transition between these regions
+occurs in a very narrow region and takes the form of \emph{jump} or \emph{bore}
+that is roughly circular if the surface is flat; we refer to it herein as a
+circular hydraulic jump (CHJ).
+
+Early experimental work the CHJ began some time later
+\cite{kurihara1946hydraulic,tani1949water,watson1964radial}.
+Watson \cite{watson1964radial} worked out the jump radius
+implied by Rayleigh's approach and then worked out the vertical
+velocity profile in the inner region, assuming a no-slip
+boundary condition at the bottom.  He also studied the turbulent
+flow case and performed experiments.
+More detailed experiments revealed different qualitative classification
+of jumps  \cite{ishigai1977heat,craik1981circular}.
+Although later work incorporated more physical details (such as surface tension) into the models
+\cite{bush2003influence}, Bohr et. al. showed that important properties of the jump
+(particularly its radius) could be reasonably predicted using a simple shallow water
+model \cite{bohr1993shallow}.
+
+While the jump is roughly circular, it may deviate from this shape and
+deforming rapidly in time.
+Instability of the jump was observed from fairly early on \cite{craik1981circular}.
+Under special circumstances with more viscous fluids, the jump instability may lead to
+the formation of curious shapes such as polygons \cite{ellegaard1998creating}, but
+for a low-viscosity fluid like water the behavior is generally chaotic.
+The instability increases with the jet velocity and with the depth at the outside of the jump,
+and can be stable for sufficiently small velocities and depths.
+
+\subsection{Numerical shock instabilities}
+
 \section{The shallow water circular hydraulic jump}
@@
 
-\subsection{FCT finite element method}
+\subsubsection{HLLM solver}
+
+\subsection{Continuous Galerkin method}
+
+\subsection{Discontinuous Galerkin method}
 
@@
 
+\subsection{Stable jump}
+
+\subsection{Unstable jump}
+
 \section{Conclusions}
 
+\bibliographystyle{apalike}
+\bibliography{refs}
+
 \end{document}

### 2020-06-11 numerics_paper/numerical_chj.tex
Add introductory material.

@@
 
+In \cite{peery1988blunt} a particular instability was observed to
+appear in numerical simulation of a bow shock, near the symmetry plane.
+This phenomenon, known as a ``carbuncle", has been observed by many researchers
+in similar numerical experiments for the Euler equations, and many remedies
+have been proposed, mainly in the form of additional numerical dissipation
+\cite{quirk1997contribution,pandolfi2001numerical,dumbser2004matrix,chauvat2005shock,ismail2009proposed,shen2014stability}.
+Most notably, the HLLE Riemann solver suppresses the carbuncle instability \cite{quirk1997contribution}.
+For a recent review of numerical shock instability and work to alleviate it,
+we refer to \cite[Section 2.5]{simonnumerical}.
+
+Given the similarity of structure between the Euler equations and the shallow
+water equations, one might expect carbuncles to appear in numerical
+solutions of the latter as well.  Indeed, this has been observed \cite{kemm2014note}.
+The shallow water carbuncle behaves in a fashion very similar to the
+Euler carbuncle: it appears when the Roe solver is used, but not when
+the HLL solver is used, and at least some of the other methods that suppress
+the latter also suppress the former \cite{kemm2014note,bader2014carbuncle}.
+
+Meanwhile, some research has suggested that the carbuncle is not purely a
+numerical instability but instead is the manifestation of a true physical
+instability \cite{moschetta2001carbuncle,elling2009carbuncle}.
+
 \section{The shallow water circular hydraulic jump}

### 2020-06-15 numerics_paper/numerical_chj.tex
Add 3-wave solver.

@@
 \documentclass{article}
+
+\usepackage{amsmath}
+
+\newcommand{\W}{{\mathcal W}}
 
@@
 \section{The shallow water circular hydraulic jump}
+The shallow water equations in two (horizontal) dimensions are
+\begin{align}
+    h_t + (hu)_x + (hv)_y & = 0 \\
+    (hu)_t + \left(hu^2 + \frac{1}{2}gh^2\right)_x + (huv)_y & = 0 \\
+    (hv)_t + (huv)_x + \left(hv^2 + \frac{1}{2}gh^2\right)_y & = 0.
+\end{align}
+These can be written in vector form as
+\begin{align}
+    q_t + f(q)_x + g(q)_y & = 0.
+\end{align}
 
@@
 
-\subsubsection{HLLE solver}
+\subsubsection{HLLE (2-wave) solver}
+The HLLE solver approximates the Riemann solution using two waves with
+speeds $s_1, s_2$.  The middle state is
+\begin{align} \label{qm-HLL}
+    q_m^\textup{HLLE} = \frac{f(q_l) - f(q_r) + s_2 q_r - s_1 q_l}{s_2 - s_1}.
+\end{align}
 
-\subsubsection{HLLM solver}
+\subsubsection{A 3-wave solver}
+Refinements of the HLLE solver were subsequently introduced in order to better
+resolve contact and shear waves, which are strongly dissipated by the HLLE solver.
+In the shallow water equations, we have only a shear wave.  We can introduce a third
+wave in the approximate Riemann solver that carries the jump in the transverse velocity.
+For simplicity, we describe this solver for an x-interface; i.e. where $u$ is the normal
+velocity component and $v$ is the tangential velocity component.  Then the Riemann
+solution is assumed to consist of 3 jump-discontinuities $\W^1, \W^*, \W^2$ where $\W^{1,2}$
+carry all of the jump in $h$ and $hu$ while $\W^*$ carries all of the jump in $hv$.
+
+The values $h_m$ and $u_m$, along with the speeds for the waves $\W^1,\W^2$ 
+are thus determined by the same formula \eqref{qm-HLL} as in the HLL solver.
+The speed of the middle wave is determined by conservation:
+$$
+    s^* = \frac{h_r v_r - h_l v_l}{h_r u_r v_r - h_l u_l v_l}.
+$$
+The middle state transverse velocity $v_*$ is given by $v_l$ if $s^*>0$ or $v_r$ if $s^*<0$.
 

### 2020-06-18 numerics_paper/numerical_chj.tex
HLLEMCC solver description.

@@
 \usepackage{amsmath}
+\usepackage{graphicx}
 
@@
 
+\subsubsection{HLLEMCC solver}
+The HLLEMCC approximate Riemann solver for the shallow water equations was proposed
+in \cite{kemm2014note} specifically to deal with carbuncles.  The idea is to adjust
+the amount of dissipation applied to the shear wave, adding dissipation in regions
+where unphysical carbuncles would appear.  The approximate solution in this
+case consists of four waves; the faster waves use the HLLE wave speeds and carry
+jumps only in the depth and normal momentum, while the slower waves carry jumps
+only in the transverse momentum.  By adjusting the value of $\phi$, this solver
+can behave like the HLLE solver (if $\phi=1$) or like the Roe solver (if $\phi=0$).
+
+The middle states $q^1_m, q^2_m$ are the same as those used for the HLLE solver.
+Requiring conservation of the transverse momentum gives
+$$
+    q^3_m = \frac{f^3_l - f^3_r - (\hat{v}-\phi\hat{c})q^3_l + (\hat{v}+\phi\hat{c})q^3_r}{2\phi\hat{c}}.
+$$
+Here $\hat{v}, \hat{c}$ are the Roe-average normal velocity and sound speed, respectively.
+
+\begin{figure}
+    \center
+    \includegraphics[width=0.6\textwidth]{figures/hllemcc.pdf}
+    \caption{Structure of the HLLEMCC approximate Riemann solution, with four waves.}
+\end{figure}
+
 \subsection{Continuous Galerkin method}

### 2020-06-19 hydraulic_jump_2D.py
Update hydraulic_jump_2D.py
Add call to HLLC solver

@@ -162,12 +162,18 @@ def setup(h0, u0, r0=0.1, h_inf=0.15, g=1., num_cells=100, tfinal=1, solver_type
     
     from clawpack import riemann
     from clawpack import pyclaw
+    import shallow_hllc_2D
 
     riemann_solver = riemann_solver.lower()
 
     if solver_type == 'classic':
         if riemann_solver == 'hlle':
             solver = pyclaw.ClawSolver2D(riemann.shallow_hlle_2D)
+            solver.fwave = False
+        elif riemann_solver == 'hllc':
+            solver = pyclaw.ClawSolver2D(shallow_hllc_2D)
+            solver.num_eqn = 3
+            solver.num_waves = 3
             solver.fwave = False
         elif riemann_solver == 'hllem':
             # Not yet implemented

### 2020-06-19 rpn2_shallow_hllc.f90
Add files via upload
Fortran code/Makefile for the HLLC solver and a simple jupyter notebook to test it.

@@ -1,0 +1,164 @@
+! =====================================================
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =====================================================
+
+! HLLEM solver for the 2D shallow water equations.
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x-momentum
+!       3 y-momentum
+
+    implicit none
+
+    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
+    
+    double precision, dimension(meqn, mwaves, 1-mbc:maxm+mbc), intent(out) :: wave
+    double precision, dimension(mwaves, 1-mbc:maxm+mbc), intent(out) :: s
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(out) :: amdq, apdq
+    
+    double precision, dimension(2,1) :: q_hat
+    double precision :: u_l, u_r, c_l, c_r, u_hat, c_hat, v_l, v_r, v_star, v_hat, h_hat
+    double precision :: h_l, h_r, hsqrt_l, hsqrt_r, hsq2
+    double precision :: grav, a, b, c, aprime, bprime, root0, root
+    double precision :: h_ml, hu_ml, hv_ml, h_mr, hu_mr, hv_mr, s1, s2, s_star, u_ml, u_mr
+    integer :: depth, mu, mv
+    integer :: i, m, mw, j, maxiter
+    
+
+    common /cparam/  grav
+
+!     # set mu to point to  the component of the system that corresponds
+!     # to momentum in the direction of this slice, mv to the orthogonal
+!     # momentum:
+!
+    maxiter = 20
+    depth = 1
+    if (ixy.eq.1) then
+        mu = 2
+        mv = 3
+    else
+        mu = 3
+        mv = 2
+    endif
+
+    do i=2-mbc,mx+mbc
+        h_l = qr(depth,i-1)
+        h_r = ql(depth,i)
+        ! Velocity
[…]

### 2020-06-21 rpn2_shallow_hllc.f90
Update rpn2_shallow_hllc.f90
Just corrected some comments.

@@ -2,7 +2,7 @@
 subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
 ! =====================================================
 
-! HLLEM solver for the 2D shallow water equations.
+! HLLC solver for the 2D shallow water equations.
 
 ! waves: 3
 ! equations: 3
@@ -36,8 +36,9 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
 !     # set mu to point to  the component of the system that corresponds
 !     # to momentum in the direction of this slice, mv to the orthogonal
 !     # momentum:
-!
-    maxiter = 20
+
+    ! max num of iteration for the root-finding algorithm
+    maxiter = 20   
     depth = 1
     if (ixy.eq.1) then
         mu = 2
@@ -71,7 +72,7 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
         ! Speeds of non-shear waves
         s1 = min(u_l - c_l, u_hat - c_hat)
         s2 = max(u_r + c_r, u_hat + c_hat)
-        ! Speeds of shear wave
+        ! Speed of shear wave
         s_star = (h_r*v_r*u_r-h_l*v_l*u_l)/(h_r*v_r-h_l*v_l)
         
         if (s_star>0) then

### 2020-06-22 numerics_paper/numerical_chj.tex
Update numerical_chj.tex
Rewriting s_star according to Rankine-Hugoniot condition.
Add material on the rotationally-symmetric steady state solution.

@@
     David I. Ketcheson \and
-    Manuel Quezada de Luna
+    Manuel Quezada de Luna \and
+    David Batista
 }
@@
 \begin{abstract}
-We study numerical discretrizations of the shallow water equations
+We study numerical discretizations of the shallow water equations
 for the simulation of the circular hydraulic jump formed
 when a vertical fluid jet impacts a horizontal plate.  
-We focus on three numerical methods: finite volume Godunov-type
-methods using either a Roe or HLLE approximate Riemann solver, and
+We test a range of numerical methods, including finite volume Godunov-type
+methods with a variety of approximate Riemann solvers, and
 a continuous Galerkin finite element method based on flux-corrected transport (FCT).
-The finite volume methods are employed ona uniform cartesian grid
+The finite volume methods are employed on a uniform cartesian grid
 while the finite element method uses an unstructured triangular grid.
-The three methods give completely different results, even for
-highly-refined meshes.
+The methods give completely different results, even for highly-refined meshes.
 \end{abstract}
@@
 surface".  This phenomenon that is familiar in the everyday kitchen sink, is in
-fact highly nonlinear and unintuitive.  Near the jet, the flow is very shallow and
-fast, while further away it is deeper and slower.  The transition between these regions
-occurs in a very narrow region and takes the form of \emph{jump} or \emph{bore}
+fact highly nonlinear and complex.  Near the jet, the flow is shallow and
+supercritical, while further away it is deeper and subcritical.  The transition from supercritical
+to subcritical flow
+occurs in a very narrow region and takes the form of a \emph{jump} or \emph{bore}
 that is roughly circular if the surface is flat; we refer to it herein as a
@@
 
-Early experimental work the CHJ began some time later
+Early experimental work on the CHJ began some time later
 \cite{kurihara1946hydraulic,tani1949water,watson1964radial}.
-Watson \cite{watson1964radial} worked out the jump radius
-implied by Rayleigh's approach and then worked out the vertical
-velocity profile in the inner region, assuming a no-slip
+Watson \cite{watson1964radial} derived the jump radius
+implied by Rayleigh's approach and the vertical
+velocity profile in the supercritical region, assuming a no-slip
 boundary condition at the bottom.  He also studied the turbulent
@@
 
-While the jump is roughly circular, it may deviate from this shape and
-deforming rapidly in time.
+While the jump is roughly circular, under appropriate conditions it may deviate
+from this shape and deform rapidly and chaotically in time.
 Instability of the jump was observed from fairly early on \cite{craik1981circular}.
@@
 for a low-viscosity fluid like water the behavior is generally chaotic.
-The instability increases with the jet velocity and with the depth at the outside of the jump,
-and can be stable for sufficiently small velocities and depths.
+The strength of the instability increases with the jet velocity and with the
+depth at the outside of the jump.  The flow can be stable for sufficiently small
+velocities and depths.
+
+\subsection{Scope of this work}
+This work is intended to be more exploratory than explanatory; in other words,
+it is focused more on raising significant questions than on answering them.
+The central issue is the stability of the shallow water CHJ, from both a
+physical and a numerical perspective.
+The main result is our finding that a number of numerical methods that are
+in use give completely different solutions to the CHJ problem.
+These differences do not diminish when the grid is refined.
+Evidence suggests that, at least in some regimes, the CHJ is unstable
+and solutions are chaotic.  This implies that one should not expect pointwise
+convergence of solutions, so a traditional study of accuracy of the numerical
+methods is ruled out.
 
@@
 
-In \cite{peery1988blunt} a particular instability was observed to
-appear in numerical simulation of a bow shock, near the symmetry plane.
+In \cite{peery1988blunt} a numerical instability was observed to
+appear near the symmetry plane in the simulation of a bow shock.
 This phenomenon, known as a ``carbuncle", has been observed by many researchers
@@
 
+The carbuncle has not previously been observed in circular hydraulic jump
+simulations.  However, it is known to appear in the presence of steady shocks
+that are nearly aligned with the grid, so it is natural to expect it, and as we
+will see it does indeed arise.  Its presence is particularly interesting 
+in this circumstance since there is a known physical instability that affects
+the shock.  A natural question is whether the carbuncle in this instance is
+a purely numerical phenomenon or a manifestation of the physical instability.
+
 \section{The shallow water circular hydraulic jump}
 The shallow water equations in two (horizontal) dimensions are
+\begin{subequations} \label{eq:sw}
 \begin{align}
@@
 \end{align}
+\end{subequations}
 These can be written in vector form as
@@
 
-\subsection{Semi-analytical solution under rotational symmetry}
+\subsection{Semi-analytical steady solution under rotational symmetry}
+
+By assuming rotational symmetry in \eqref{eq:sw}, one obtains the
+system (in one spatial variable)
+\begin{subequations} \label{eq:rsw}
+\begin{align}
+    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
+    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
+\end{align}
+\end{subequations}
+Steady-state solutions of \eqref{eq:rsw} satisfy
+\begin{subequations}
+\begin{align}
+    rhu & = C \\
+    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
+\end{align}
+\end{subequations}
+for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
+We see that two types of steady profiles exist, depending on whether the flow
+is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
+include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.
+
+The steady, rotationally-symmetric circular hydraulic jump involves supercritical
+flow for $r<r_0$ and subcritical flow for $r>r_0$, where $r_0$ is the jump radius.
+The jump itself takes the form of a stationary shock wave.  The Rankine-Hugoniot jump
+conditions specify that for such a shock,
+\begin{align} \label{eq:RH}
+    h_+ - h_- & = \frac{-3h_- + \sqrt{h_-^2 + 8 h_- u_-^2/g}}{2} = \frac{3h_-}{2}\left(\sqrt{1+\frac{8}{9}(F_-^2-1)}-1\right),
+\end{align}
+where the subscripts $+, -$ denote states just inside or outside the jump radius, respectively.
+
+A steady-state, rotationally symmetric solution can be given for an annular region as follows:
+
+\begin{enumerate}
+    \item Specify the depth and velocity at the inner boundary (near the jet) and outer boundary.
+    \item Integrate \eqref{eq:dh} from both boundaries.
+    \item Find a radius $r_0$ where the matching condition \eqref{eq:RH} is satisfied.
+\end{enumerate}
+Due to the nature of solutions of \eqref{eq:dh}, it can be shown that the required jump
+radius $r_0$ always exists if the prescribed flow is supercritical at the inner boundary
+and subcritical at the outer boundary.
+
+Numerical tests suggest that the solution obtained in this way is a stable equilibrium
+of the reduced one-dimensional shallow water equations \eqref{eq:rsw}.  This
+solution provides a useful starting point for testing the stability of the CHJ
+as a solution of the full two-dimensional shallow water equations \eqref{eq:sw}.
 
 \section{Numerical methods}
+
+Potential methods to try:
+\begin{itemize}
+    \item Galerkin methods on cartesian grids (to see if carbuncles appear)
+    \item FV methods on perturbed grids (to avoid grid alignment)
+    \item FV methods on annular grids (with grid alignment everywhere)
+    \item Various Riemann solvers intended to avoid carbuncles
+\end{itemize}
 
@@
 $$
-    s^* = \frac{h_r v_r - h_l v_l}{h_r u_r v_r - h_l u_l v_l}.
+    s^* = \frac{h_r u_r v_r - h_l u_l v_l}{h_r v_r - h_l v_l}.
 $$

### 2020-07-07 rpn2swq-hllemcc.f90
Indentation and subroutine ending
Check indentation and add appropriate subroutine ending

@@ -84,6 +84,7 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     do i=2-mbc,mx+mbc
         alf(i) = auxl(inx,i)
         beta(i) = auxl(iny,i)
+
         unorl(i) = alf(i)*ql(2,i) + beta(i)*ql(3,i)
         unorr(i-1) = alf(i)*qr(2,i-1) + beta(i)*qr(3,i-1)
         utanl(i) = -beta(i)*ql(2,i) + alf(i)*ql(i,3)
@@ -95,13 +96,16 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     ! later used in routine rpt2 to do the transverse wave splitting.
 
     do i = 2-mbc, mx+mbc
-        h(i) = (qr(1,i-1)+ql(1,i))*0.50d0
+        h(i) = (qr(1,i-1)+ql(1,i))*0.50d0 ! = h_hat
         hsqrtl = dsqrt(qr(1,i-1))
         hsqrtr = dsqrt(ql(1,i))
         hsq2 = hsqrtl + hsqrtr
+        ! unorr(i-1)/hsqrtl is fine because unorr = momentum 
         u(i) = (unorr(i-1)/hsqrtl + unorl(i)/hsqrtr) / hsq2
         v(i) = (utanr(i-1)/hsqrtl + utanl(i)/hsqrtr) / hsq2
-        a(i) = dsqrt(grav*h(i))
+        ! u(i) = u_hat
+        ! v(i) = v_hat
+        a(i) = dsqrt(grav*h(i)) ! = c_hat
     enddo
 
     ! now split the jump in q at each interface into waves
@@ -144,92 +148,92 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     ! Incorporate entropy fix by adding a modified fraction of wave
     ! if s should change sign.
 
-    do i=2-mbc,mx+mbc
+        do i=2-mbc,mx+mbc
 
-        ! modified wave speeds to use HLLEM-entropy fix
+            ! modified wave speeds to use HLLEM-entropy fix
 
-        him1 = qr(1,i-1)
-        uim1 = unorr(i-1)
[…]

### 2020-07-07 numerics_paper/numerical_chj.tex
Update numerical_chj.tex
Add formulas for the HLLEMCC solver, specifically, the indicator function and the new absolute-value,

@@
 $$
-Here $\hat{v}, \hat{c}$ are the Roe-average normal velocity and sound speed, respectively.
+Here $\hat{v}, \hat{c}$ are the Roe-average normal velocity and sound speed, respectively, whereas the function $\phi$ is given by
+
+\begin{equation}\label{if}
+\phi(\theta)=min\left\lbrace 1,\, \left(\varepsilon\,\theta\,max\left\lbrace0,\,1-Fr_{u}^{\alpha} \right\rbrace \right)   \right\rbrace^{\beta}
+\end{equation}
+
+where $Fr_{u}$ is the Froude number perpendicular to the cell face and $\varepsilon$, $\alpha$, and $\beta$ are parameters that have to be chosen appropriately.
+For the shallow water equations, authors in \cite{kemm2014note} propose to use $\varepsilon=10^{-3}$ and $\alpha=\beta=\frac{1}{3}$.
+The value of $\theta$ is calculated as the residual in the Rankine-Hugoniot condition for the shear wave.\\
+
+The HLLEMCC solver follows the ideas of the Harten-Hyman entropy fix by splitting the shear wave with speed $\hat{u}$, into two waves moving with speeds $\hat{u}-\phi\hat{c}$ and $\hat{u}+\phi\hat{c}$ (see figure \ref{HLLEMCC}).
+Its implementation, in terms of the fluctuations, is performed by using the values $\hat{u}^{\pm}$ calculated with an absolute-value function defined as
+
+\begin{equation}\nonumber
+\varphi(\lambda)=\frac{\vert \lambda-\phi(\theta)\hat{c}\,\vert+\vert \lambda+\phi(\theta)\hat{c}\,\vert}{2}
+\end{equation}
 
@@
     \caption{Structure of the HLLEMCC approximate Riemann solution, with four waves.}
+    \label{HLLEMCC}
 \end{figure}

### 2020-07-12 rpn2swq_hllemcc.f90
Notebook to test HLLEMCC
This notebook has:
.- Original HLLEMCC code (Kemm's code): rpn2swq_hllemcc.f90
.- A modified HLLEMCC code: rpn2swq_hllemccroef.f90
.- Roe with entropy fix and GeoClaw solvers for comparison.

@@ -1,0 +1,240 @@
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+
+! Roe-solver for the 2D shallow water equations on a quadrilateral grid
+! 
+! Solve Riemann problems along one slice of data.
+! 
+! On input, ql contains the state vector at the left edge of each cell
+!           qr contains the state vector at the right edge of each cell
+!     
+! This data is along a slice in the x-direction if ixy=1
+!                            or the y-direction if ixy=2.
+! On output, wave contains the waves, s the speeds,
+! and amdq, apdq the decomposition of the flux difference
+!   f(qr(i-1)) - f(ql(i))
+! into leftgoing and rightgoing parts respectively.
+! With the Roe solver we have
+!    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
+! where A is the Roe matrix.  An entropy fix can also be incorporated
+! into the flux differences.
+! 
+! Note that the i'th Riemann problem has left state qr(i-1,:)
+!                                     and right state ql(i,:)
+! From the basic clawpack routines, this routine is called with ql = qr
+
+    implicit double precision (a-h,o-z)
+
+    dimension wave(meqn, mwaves, 1-mbc:maxm+mbc)
+    dimension    s(mwaves, 1-mbc:maxm+mbc)
+    dimension   ql(meqn, 1-mbc:maxm+mbc)
+    dimension   qr(meqn, 1-mbc:maxm+mbc)
+    dimension  apdq(meqn, 1-mbc:maxm+mbc)
+    dimension  amdq(meqn, 1-mbc:maxm+mbc)
+    dimension auxl(maux, 1-mbc:maxm+mbc)
+    dimension auxr(maux, 1-mbc:maxm+mbc)
+
+    ! local arrays -- common block comroe is passed to rpt2sh
+    ! ------------
+
+    dimension splus(mwaves)
+    dimension sminus(mwaves)
+
+    parameter (maxm2 = 1602)  !# assumes at most 1000x1000 grid with mbc=2
+    dimension delta(3)
+    logical efix, cfix
+    dimension unorl(-1:maxm2), unorr(-1:maxm2)
+    dimension utanl(-1:maxm2), utanr(-1:maxm2)
[…]

### 2020-07-12 rpn2swq_hllemccroef.f90

@@ -1,0 +1,283 @@
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! Modification of Kemm's code.
+! The entropy fix for the 1- and 3-wave is performed as in rpn2_shallow_roe_with_efix.f90, 
+! whereas the one for the 2-wave is the same as in Kemm's code.
+
+
+! Roe-solver for the 2D shallow water equations on a quadrilateral grid
+! 
+! Solve Riemann problems along one slice of data.
+! 
+! On input, ql contains the state vector at the left edge of each cell
+!           qr contains the state vector at the right edge of each cell
+!     
+! This data is along a slice in the x-direction if ixy=1
+!                            or the y-direction if ixy=2.
+! On output, wave contains the waves, s the speeds,
+! and amdq, apdq the decomposition of the flux difference
+!   f(qr(i-1)) - f(ql(i))
+! into leftgoing and rightgoing parts respectively.
+! With the Roe solver we have
+!    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
+! where A is the Roe matrix.  An entropy fix can also be incorporated
+! into the flux differences.
+! 
+! Note that the i'th Riemann problem has left state qr(i-1,:)
+!                                     and right state ql(i,:)
+! From the basic clawpack routines, this routine is called with ql = qr
+
+    implicit double precision (a-h,o-z)
+
+    dimension wave(meqn, mwaves, 1-mbc:maxm+mbc)
+    dimension    s(mwaves, 1-mbc:maxm+mbc)
+    dimension   ql(meqn, 1-mbc:maxm+mbc)
+    dimension   qr(meqn, 1-mbc:maxm+mbc)
+    dimension  apdq(meqn, 1-mbc:maxm+mbc)
+    dimension  amdq(meqn, 1-mbc:maxm+mbc)
+    dimension auxl(maux, 1-mbc:maxm+mbc)
+    dimension auxr(maux, 1-mbc:maxm+mbc)
+
+    ! local arrays -- common block comroe is passed to rpt2sh
+    ! ------------
+
+    dimension splus(mwaves)
+    dimension sminus(mwaves)
+
[…]

### 2020-07-13 hydraulic_jump_2D.py
Add new solvers
HLLEMCC
HLLEMCC modified
Roe (for comparison)

@@ -18,9 +18,9 @@ from matplotlib import animation
 import matplotlib.pyplot as plt
 from clawpack import pyclaw
 import numpy as np
-# If plotting in a Jupyter notebook, do this:
-# from clawpack.visclaw.JSAnimation import IPython_display
 
+# If plotting in a Jupyter notebook, do this:
+from clawpack.visclaw.JSAnimation import IPython_display
 
 def plot_surface(claw, save_plots=False, frames=101, val='surface',
                  vmin=0., vmax=0.5, clim=None, bathymetry=False):
@@ -47,7 +47,7 @@ def plot_surface(claw, save_plots=False, frames=101, val='surface',
         qq = np.maximum(b,h+b)
     elif val=='u':
         hu = frame.q[1,:,:]
-        qq = hu/h
+        qq = hu #/h
         
 
     x, y = frame.state.grid.p_centers    
@@ -68,7 +68,7 @@ def plot_surface(claw, save_plots=False, frames=101, val='surface',
             qq = np.maximum(b,h+b)
         elif val=='u':
             hu = frame.q[1,:,:]
-            qq = hu/h
+            qq = hu #/h
         im.set_data(qq.T)
         if clim:
             im.set_clim(*clim)
@@ -77,7 +77,7 @@ def plot_surface(claw, save_plots=False, frames=101, val='surface',
             fig.savefig(fname)   
         return im,
 
-    return animation.FuncAnimation(fig, fplot, frames=frames, interval=40, repeat=False)
+    return animation.FuncAnimation(fig, fplot, frames=frames, interval=40, repeat=True)
 
 def set_jet_values(_, state):
     """
@@ -162,7 +162,9 @@ def setup(h0, u0, r0=0.1, h_inf=0.15, g=1., num_cells=100, tfinal=1, solver_type
     
     from clawpack import riemann
     from clawpack import pyclaw
-    import shallow_hllc_2D
+    #import shallow_hllc_2D
+    import shallow_hllemcc_2D
+    import shallow_hllemccRoEF_2D
 
     riemann_solver = riemann_solver.lower()
 
@@ -170,8 +172,21 @@ def setup(h0, u0, r0=0.1, h_inf=0.15, g=1., num_cells=100, tfinal=1, solver_type
         if riemann_solver == 'hlle':
             solver = pyclaw.ClawSolver2D(riemann.shallow_hlle_2D)
             solver.fwave = False
+        if riemann_solver == 'roe':
[…]

### 2020-09-14 numerics_paper/numerical_chj.tex
Update numerical_chj.tex
Change the phrase "The value of $\theta$ is calculated as the residual" by the correct one "The value of $\theta$ is calculated as the 2-norm of the residual"
Add PyClaw simulation results.

@@
 \usepackage{graphicx}
+\usepackage{geometry}
+\usepackage{float}
+\usepackage{subfig}
 
@@
     David I. Ketcheson \and
-    Manuel Quezada de Luna \and
-    David Batista
+    Manuel Quezada de Luna
 }
@@
 For the shallow water equations, authors in \cite{kemm2014note} propose to use $\varepsilon=10^{-3}$ and $\alpha=\beta=\frac{1}{3}$.
-The value of $\theta$ is calculated as the residual in the Rankine-Hugoniot condition for the shear wave.\\
+The value of $\theta$ is calculated as the 2-norm of the residual in the Rankine-Hugoniot condition for the shear wave.\\
 
@@
 
+To demonstrate the range of qualitative solutions that can arise when using
+different numerical methods, we apply several schemes to the following setup.
+We take
+\begin{align*}
+    r_0 & = 0.1 & h_0 & = 0.5 & u_0 & = 0.75 \\
+    g & = 1 & h_\infty & = 0.15 & F_\infty & = 0.1
+\end{align*}
+where $r_0, h_0, u_0$ are the jet radius and the depth and velocity at the outer
+edge of the jet, respectively; $h_\infty$ and $F_\infty$ are the depth
+and Froude number (infinitely) far away from the jet.  The domain is $[-1,1]^2$
+and boundary conditions are imposed based on the 1D steady state solution with
+the chosen jet and far-field conditions.  The bathymetry is flat and friction is
+neglected.  The initial data is chosen to be far from the equilibrium and
+consists of a uniform depth of $h=0.15$ and zero velocity.
+Results using a $1000\times 1000$ grid are shown in Figure \ref{fig:hlle_vs_roe}.  
+We see that by $t=4$, the HLLE solution has already reached the equilibrium
+and remains there indefinitely.  In contrast, the Roe solution is highly
+unsteady with a very non-circular jump containing cusps that tend to form
+near the four points where the jump is aligned with a grid axis.
+
+Based on numerous tests, we have found that results for other initial data
+are very similar to those shown here, although of course the details
+of the Roe solution differ because the solution is chaotic.
+
+\begin{figure}
+  \begin{centering}
+    \subfloat[HLLE Riemann solver, t=4]{
+      \includegraphics[width=0.45\textwidth]{figures/hlle_1000_t4.eps} 
+    }
+    \subfloat[Roe Riemann solver with entropy fix, t=4]{
+      \includegraphics[width=0.45\textwidth]{figures/roe_1000_t4.eps}
+    }
+
+    \subfloat[HLLE Riemann solver, t=20\label{fig:hlle}]{
+      \includegraphics[width=0.45\textwidth]{figures/hlle_1000_t20.eps} 
+    }
+    \subfloat[Roe Riemann solver with entropy fix, t=20\label{fig:roe}]{
+      \includegraphics[width=0.45\textwidth]{figures/roe_1000_t20.eps}
+    }
+  \end{centering}
+  \caption{
+    Comparison of PyClaw solutions using the classic Clawpack discretization
+    with HLLE and Roe Riemann solvers.\label{fig:hlle_vs_roe}}
+\end{figure}
+
+Next we apply Kemm's HLLEMCC solver to the same scenario; results are shown
+in Figure \ref{fig:hllemcc}.  Results at early times are very similar to
+those for the Roe solver.  The cusps that appear have a stronger tendency
+to dissipate and then reform, as indicated in the figure at $t=20$.
+Nevertheless, they continue to appear indefinitely, based on longer-time
+simulations.
+
+\begin{figure}
+  \begin{centering}
+    \subfloat[HLLEMCC Riemann solver, t=4]{
+      \includegraphics[width=0.45\textwidth]{figures/hllemcc_1000_t4.eps} 
+    }
+    \subfloat[HLLEMCC Riemann solver, t=20]{
+      \includegraphics[width=0.45\textwidth]{figures/hllemcc_1000_t20.eps}
+    }
+  \end{centering}
+  \caption{
+    PyClaw solutions using the classic Clawpack discretization
+    with HLLEMCC Riemann solver.\label{fig:hllemcc}}
+\end{figure}
+
 \subsection{Stable jump}

### 2020-09-14 hydraulic_jump_2D.py
Make HLLEMCC parameters changeable at run-time.
Kemm's alpha, beta, and epsilon (see Kemm 2014).

@@ -164,9 +164,11 @@ def step_friction(solver, state, dt):
     q[1,:,:] = q[1,:,:] - dt*cf*u/h
     q[2,:,:] = q[2,:,:] - dt*cf*v/h
         
-def setup(h0=0.5, u0=0.75, r0=0.1, h_inf=0.15, g=1., num_cells=100, tfinal=1, solver_type='classic',
-          num_output_times=10, riemann_solver='hlle', boundary='subcritical', outdir='./_output',
-          friction=False, friction_coeff=0.01, F_bdy=0.1, use_petsc=False):
+def setup(h0=0.5, u0=0.75, r0=0.1, h_inf=0.15, g=1., num_cells=100, tfinal=1,
+          solver_type='classic', num_output_times=10, riemann_solver='hlle',
+          boundary='subcritical', outdir='./_output', friction=False,
+          friction_coeff=0.01, F_bdy=0.1, use_petsc=False, 
+          kalpha=1./3, kbeta=1.3, kepsilon=1.e-3):
     
     from clawpack import riemann
     if use_petsc:
@@ -287,6 +289,9 @@ def setup(h0=0.5, u0=0.75, r0=0.1, h_inf=0.15, g=1., num_cells=100, tfinal=1, so
     state.problem_data['h0'] = h0
     state.problem_data['u0'] = u0
     state.problem_data['grav'] = g   # Gravitational force
+    state.problem_data['kalpha'] = kalpha   # Kemm's alpha
+    state.problem_data['kbeta'] = kbeta   # Kemm's beta
+    state.problem_data['kepsilon'] = kepsilon   # Kemm's epsilon
     state.problem_data['F_bdy'] = F_bdy
     state.problem_data['cf'] = friction_coeff
 

### 2020-09-14 rpn2_sw_hllemcc.f90

@@ -34,9 +34,10 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
     double precision :: grav
     double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
-    double precision :: renore, rere1, reremu, reremv, rhind, rhparam
+    double precision :: renore, rere1, reremu, reremv, rhind
+    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
 
-    common /cparam/ grav
+    common /cparam/ grav, kalpha, kbeta, kepsilon
     common /comroe/ u, v, a, h
 
     data efix /.true./        !# use entropy fix for transonic rarefactions
@@ -153,18 +154,16 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
                 reremv = wave(mv,3,i)-wave(mv,1,i)
                 ! norm of relative residual
                 renore = dsqrt(rere1*rere1 + reremu*reremu + reremv*reremv)
-                ! Indicator for Rankine-Hugoniot condition, original:0.01d0 
-                rhparam = 0.001d0
-                ! rhind = a(i) * max(rhparam*renore,1.)
+                ! rhind = a(i) * max(kepsilon*renore,1.)
                 zfroude = dsqrt((u(i)*u(i))/(a(i)*a(i)))
-                ffroude  = dmax1(0.0d0,(1.0d0 - zfroude**(1.0d0/3.0d0)))
+                ffroude  = dmax1(0.0d0,(1.0d0 - zfroude**kalpha))
                 ! Alternative if cbrt for cubic root is available:
                 ! ffroude  = dmax1(0.0d0,(1.0d0 - dcbrt(zfroude)))
                
                 rhind = a(i)
-                rhind = rhind*dmin1(rhparam*renore*ffroude,1.0d0)**(1.0d0/3.0d0) 
+                rhind = rhind*dmin1(kepsilon*renore*ffroude,1.0d0)**kbeta 
                 ! Alternative if cbrt for cubic root is available:
[…]

### 2020-09-16 numerics_paper/numerical_chj.tex
More PyClaw results in paper.

@@
 
+The results above are computed with schemes that include second-order
+correction terms.  In Figure \ref{fig:first-order} we compare the corresponding
+first-order schemes.  Even by increasing the value of $\epsilon$ to $0.1$,
+the carbuncles are not suppressed by the HLLEMCC solver.  Indeed, the Roe
+and HLLEMCC solutions look very similar.
+
+\begin{figure}
+  \begin{centering}
+    \subfloat[Roe Riemann solver, 1st order, t=20]{
+      \includegraphics[width=0.45\textwidth]{figures/_1_roe_1000_t20.eps} 
+    }
+    \subfloat[HLLEMCC Riemann solver, $\epsilon=0.1$, 1st order, t=20]{
+      \includegraphics[width=0.45\textwidth]{figures/_1_hllemcc_1000_eps01_t20.eps}
+    }
+  \end{centering}
+  \caption{
+    PyClaw solutions using a 1st-order discretization
+    with different Riemann solvers.\label{fig:hllemcc}}
+\end{figure}
+
 \subsection{Stable jump}

### 2020-10-14 numerics_paper/version2/numerical_chj.tex
add outline for the paper

\documentclass[preprint, 11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[hmargin=0.9in, vmargin=1.25in]{geometry}
\usepackage[babel,german=quotes]{csquotes}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{caption}
\usepackage{amsmath}          % writing mathematical formulas
\usepackage{amsthm}           % writing mathematical theorems 
\usepackage{amssymb}          % writing mathematical symbols
\usepackage{bm, amsfonts}     % writing bold mathematical symbols
\usepackage{xcolor}
\usepackage{fixmath}
\usepackage{tikz}
\usepackage{bm}
\usepackage{makecell}
\usepackage{mathdots}
\usepackage{algpseudocode}
\usepackage{algorithm}

\newcommand{\W}{{\mathcal W}}
\newcommand{\A}{{\mathcal A}}

\title{Numerical simulation and entropy dissipative cure of the 
  carbuncle instability for the shallow water circular hydraulic jump}

\author{
    David I. Ketcheson \and
    Manuel Quezada de Luna
}
\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\subsection{The circular hydraulic jump}

\begin{itemize}
  \item Describe the formation of the circular hydraulic jump, some properties, etc. 
  \item Close this section stating that for some flow regimes, the CHJ becmes unstable and its 
    shape deviates from a circular shape. For high viscosity flows polygonal shapes are observed 
    and for low viscosity flows the shape appears to be chaotic. 
  \item We can use section 1.1 in the original version of the paper.
\end{itemize}

\subsection{Numerical shock instabilities}

\begin{itemize}
  \item Describe the carbuncle instability starting with Euler and going to SWEs. 
  \item Mention that we observe the cc instability for the CHJ as well. 
  \item Describe that in this case, the situation is more complicated than in other problems since we 
    expect the correct solution to have physical instabilities triggered by round off error. 
  \item Based on existing literature we know that we can use highly dissipative and robust methods to 
    avoid the formation of carbuncle instabilities. However, in this case, these methods also tend to 
    dissipate other important features of the solution unless the mesh is highly refined. 
  \item We are interested in methods that prevent the formation of cc instability in the CHJ 
    but preserve other features and physical instabilities. 
  \item There are multiple ideas and methods proposed to fix the cc insitability for the Euler 
    and recently for the SWEs. 
  \item It is believed that the cc instability occurs due to lack or reduced viscosity in the shear contact waves. 
  \item Some methods also propose to improve the entropy stability properties of the method. In particular, to induce or guarantee that the entropy is dissipated at shocks. 
  \item Based on some of these ideas we start with Rusanov's method using the cell averages of the solution. 
    This is a first-order method that is entropy stable and positivity preserving. We enhance its robustness by 
    overestimating the wave speed when strong shear is present in the solution. The result is a robust solver 
    that do not develop cc instabilities. However, the method is highly dissipative and, as a result, 
    many other important features of the solution are dissipated. 
  \item We then blend this highly dissipative method with a FV method based on Roe's average. 
  \item After this we introduce the second order corrections from Law-Wendroff's method. 
\end{itemize}

\subsection{Our contribution}
\begin{itemize}
\item We present a new problem where the cc instability occurs. In our opinion, this is a more challenging 
  problem than the standard bow problem since we expect the solution to be non-steady. 
  We aim to preserve instabilities that are inherently present in the CHJ for certain regimes. And, 
  at the same time, we aim to prevent the cc instability. 
\item We explore the use of entropy dissipation and stability as cure of the cc instability for the SWEs. 
\item We propose specific methods to blend the highly dissipative and robust Rusanov's RP and the 
  accurate 3-wave Roe's solver for the SWEs. In particular we propose methods that induce dissipation of entropy
  near the shocks and other methods that guarantee entropy stability. 
\item We demonstrate that these schemes prevent the formation of cc instabilities not only 
  in the well known bow shock problem, but also in the CHJ while preserving other important features of 
  the solution. 
\end{itemize}

\section{The shallow water circular hydraulic jump}
The shallow water equations in two (horizontal) dimensions are
\begin{subequations} \label{eq:sw}
\begin{align}
    h_t + (hu)_x + (hv)_y & = 0 \\
    (hu)_t + \left(hu^2 + \frac{1}{2}gh^2\right)_x + (huv)_y & = 0 \\
    (hv)_t + (huv)_x + \left(hv^2 + \frac{1}{2}gh^2\right)_y & = 0.
\end{align}
\end{subequations}
These can be written in vector form as
\begin{align}
    q_t + f(q)_x + g(q)_y & = 0.
\end{align}

\subsection{Semi-analytical steady solution under rotational symmetry}

By assuming rotational symmetry in \eqref{eq:sw}, one obtains the
system (in one spatial variable)
\begin{subequations} \label{eq:rsw}
\begin{align}
    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
\end{align}
\end{subequations}
Steady-state solutions of \eqref{eq:rsw} satisfy
\begin{subequations}
\begin{align}
    rhu & = C \\
    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
\end{align}
\end{subequations}
for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
We see that two types of steady profiles exist, depending on whether the flow
is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.

The steady, rotationally-symmetric circular hydraulic jump involves supercritical
flow for $r<r_0$ and subcritical flow for $r>r_0$, where $r_0$ is the jump radius.
The jump itself takes the form of a stationary shock wave.  The Rankine-Hugoniot jump
conditions specify that for such a shock,
\begin{align} \label{eq:RH}
    h_+ - h_- & = \frac{-3h_- + \sqrt{h_-^2 + 8 h_- u_-^2/g}}{2} = \frac{3h_-}{2}\left(\sqrt{1+\frac{8}{9}(F_-^2-1)}-1\right),
\end{align}
where the subscripts $+, -$ denote states just inside or outside the jump radius, respectively.

A steady-state, rotationally symmetric solution can be given for an annular region as follows:

\begin{enumerate}
    \item Specify the depth and velocity at the inner boundary (near the jet) and outer boundary.
    \item Integrate \eqref{eq:dh} from both boundaries.
    \item Find a radius $r_0$ where the matching condition \eqref{eq:RH} is satisfied.
\end{enumerate}
Due to the nature of solutions of \eqref{eq:dh}, it can be shown that the required jump
radius $r_0$ always exists if the prescribed flow is supercritical at the inner boundary
and subcritical at the outer boundary.

Numerical tests suggest that the solution obtained in this way is a stable equilibrium
of the reduced one-dimensional shallow water equations \eqref{eq:rsw}.  This
solution provides a useful starting point for testing the stability of the CHJ
as a solution of the full two-dimensional shallow water equations \eqref{eq:sw}.

\subsection{Formation of the CHJ}

\section{Numerical methods}
We consider the implementation of the Lax-Wendroff method in Clawpack. 
For a one-dimensional problem, the method can be written as follows:
\begin{align*}
  U_i^{n+1}=U_i^n-\frac{\Delta t}{\Delta x}\left[\A\Delta Q_{i-1/2}]
\end{align*}
where XX. In this section we focus on the first order method; i.e., we consider

XX

and propose different Riemann solvers. Once the Riemann solvers are defined for the low-order method, 
the second order corrections are incorporated following the same steps for all cases. Therefore, we focus our 
attention to design of the Riemann solvers.

\subsection{Roe's Riemann solver}

\subsection{Rusanov's Riemann solver}

\subsection{Blended Rusanov's and Roe's solver based on entropy residual}

\subsection{Blended Rusanov's and Roe's solver based on Tadmor's entropy stability condition}

\subsection{Entropy stable enhancement of Roe's Riemann solver}

\subsection{Other alternatives}

\section{Numerical results}

We consider two regimes. In the first regime, the physical instabilities are mild. 
By increasing the strength of the jump and other parameters, we consider a second 
and more strongly unstable regime. In both cases, we generate the CHJ as described in 
Section XX; that is, we impose supersonic inflow boundary conditions for the jet and 
subsonic outflow boundary conditions.

We consider two types of domain and two corresponding computational grids. 
The two types of domains are two-dimensional squares and circular domains. 
For both cases we use structured grids. See an example of these domains and grids in Figures XX and XX, 
respectively.

\subsection{Regime 1: midly unstable flow}

Show the results with Roe's and Rusanov's solver. 
Show the results with the three proposed methodologies.

\subsection{Regime 2: strongly unstable flow}

Show the results with Roe's and Rusanov's solver. 
Show the results with the three proposed methodologies.

\section{Conclusions}

\appendix
\section{Bow shock problem for the shallow water equations}

\section{Other numerical studies}

\bibliographystyle{apalike}
\bibliography{refs}

\end{document}

### 2020-11-02 numerics_paper/version2/numerical_chj.tex
Add missing \right.
tmp update, not ready to be reviewed

@@
 \newcommand{\A}{{\mathcal A}}
+\newcommand{\bff}{{\bf f}}
+\newcommand{\bfu}{{\bf u}}
+\newcommand{\bfv}{{\bf v}}
+\newcommand{\bfq}{{\bf q}}
+\newcommand{\bfx}{{\bf x}}
 
@@
 \section{Numerical methods}
-We consider the implementation of the Lax-Wendroff method in Clawpack. 
-For a one-dimensional problem, the method can be written as follows:
+In this work, we consider Riemann-based finite volume methods.
+Considering a one dimensional hyperbolic conservation law,
+the cell average of the numerical solution given by a finite volume scheme (with forward Euler) 
+can be written as follows:
+\begin{align}\label{first-order_FV}
+  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[F_{i+1/2}-F_{i-1/2}\right],
+\end{align}
+where $F_{i+1/2}=F_{i+1/2}(u^n)$ is a numerical flux; i.e., a discretization of the flux. 
+This numerical flux is usually is composed of two parts, a consistent term that approximates 
+the physical flux at the interface between cells $i$ and $i+1$, and a stabilization term. 
+For instance, the well known Rusanov's numerical flux is given by 
+\begin{align}\label{rusanov_num_flux}
+  F_{i+1/2}(z)=\frac{\bff(z_{i+1})+\bff(z_i)}{2} - \lambda_{i+1/2}(z_{i+1}-z_i),
+\end{align}
+where $\lambda_{i+1/2}$ is an upper bound on the wave speed associated with the Riemann problem 
+composed of the states $u_i$ and $u_{i+1}$. 
+If $z=u$ in \eqref{rusanov_num_flux}
+%; i.e., if the Rusanov's numerical flux is obtained considering 
+%the cell averages of the solution 
+and if the time step is small enough, 
+then the method \eqref{first-order_FV} is positivity preserving and entropy stable, see for instance XX. 
+However, the accuracy of the numerical flux, and hence of the method, would be 
+limited to first-order. To improve the accuracy properties of Rusanov's method, 
+two alternatives are common. 
+The numerical flux $F_{i+1/2}$ can be approximated based on polynomial 
+reconstructions around the interface $i+1/2$. 
+For smooth solutions, this strategy leads to arbitrarily high-order
+methods. Alternatively, second-order accuracy can be obtained by removing some of the 
+numerical dissipation introduced by the stabilization component of the numerical flux. 
+The second-order Lax-Wendroff method XX, which we use in this work, 
+is a popular example of this approach. 
+Both high-order corrections to Rusanov's method require careful treatment to guarantee 
+the solution is still positivity preserving and/or entropy stable.
+
+Our main objective is to study the instabilities in the CHJ and to propose a Riemann solver
+that is carbunclue free but do not dissipate important features of the solution. 
+To do that we combine the dissipative Rusanov's numerical flux with a Roe's average based
+approximate Riemann solver. A similar idea, to blend Rusanov's and a Roe's average 
+Riemann solver has been explored before in XX to solve the Euler equations. The blending 
+function in the aformentioned work is based on XX....
+
+After introducing the blended Riemann solver, we use it with the first order method 
+\eqref{first-order_FV} and then as building block of the Lax-Wendroff method.
+
+\subsection{Lax-Wendroff-LeVeque method}
+
+\section{Riemann solvers}
+
+\subsection{Two- versus three-wave solvers}
+The general finite volume method \eqref{first-order_FV} can be written in terms of left- 
+and right-going waves emanating from the interfaces XX. The net contribution of these waves 
+is called fluctuations. Based on fluctuations, method \eqref{first-order_FV} can be written as 
+\begin{align}\label{FV_via_fluct}
+  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right],
+\end{align}
+where $\A^+\Delta Q_{i-1/2}$ and $\A^-\Delta Q_{i+1/2}$ are the right- and left-going fluctuations 
+emanating from the interfaces $i-1/2$ and $i+1/2$, respectively.
+The waves and their speed of propagation (to compute the fluctuations) are tightly related to 
+Riemann solvers. For the shallow water equations in two-dimensions, the structure of the exact (and some 
+approximated) Riemann solvers contains three waves. This is the case, for example, of the well known 
+Roe's Riemann solver, see for instance XX. Simpler approaches are also possible. For instance, 
+method XX with the Rusanov's numerical flux XX can be rewritten in the form XX by considering 
+only two waves that propagate with the same speed in opposite directions.
+
+%\clearpage
+%We consider the implementation of the Lax-Wendroff method in Clawpack. 
+%Following XX, and assuming a one-dimensional problem, the method can be written as follows:
+%\begin{align*}
+%  U_i^{n+1}=U_i^n
+%  -\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right]
+%  -\frac{\Delta t}{\Delta x}\left[\tilde F_{i+1/2}-\tilde F_{i-1/2}\right],
+%\end{align*}
+%where XX and XX are right- and left-going fluctuations emanating from the interfaces XX and XX, respectively. 
+%The terms XX and XX are flux corrections that improve the accuracy to second order. 
+%. In this section we focus on the first order method; i.e., we consider 
+%XX 
+%
+%and propose different Riemann solvers. Once the Riemann solvers are defined for the low-order method, 
+%the second order corrections are incorporated following the same steps for all cases. Therefore, we focus our 
+%attention to design of the Riemann solvers.
+
+%\subsection{Roe's Riemann solver}
 \begin{align*}
-  U_i^{n+1}=U_i^n-\frac{\Delta t}{\Delta x}\left[\A\Delta Q_{i-1/2}]
+  F_{i+1/2}=\frac{\bff(u_{i+1})+\bff(u_i)}{2}+\frac{1}{2}|\hat A_{i+1/2}|(u_{i+1}-u_i)
 \end{align*}
-where XX. In this section we focus on the first order method; i.e., we consider
 
-XX
+%\subsection{Rusanov's Riemann solver}
 
-and propose different Riemann solvers. Once the Riemann solvers are defined for the low-order method, 
-the second order corrections are incorporated following the same steps for all cases. Therefore, we focus our 
-attention to design of the Riemann solvers.
+\clearpage
+\section{Entropy dissipative Riemann solver}
+Given a hyperbolic system of conservation laws 
+\begin{align}\label{cons_law}
+  \frac{\partial \bfu(\bfx)}{\partial t} + \nabla\cdot \bff(\bfu(\bfx))=0, \quad \bfx\in\Omega,
+\end{align}
+a convex function $\eta(\bfu)$ is called an entropy if there exists a function $\bfq(\bfu)$, called 
+entropy flux, such that XX. A weak solution of XX is called an entropy solution if 
+\begin{align}
+  \frac{\partial \eta(\bfu(\bfx))}{\partial t} + \nabla\cdot \bfq(\bfu(\bfx))\leq 0, \quad \bfx\in\Omega,
+\end{align}
+holds, in a distribution sense, for any entropy pair $(\eta(\bfu),\bfq(\bfu))$ of \eqref{cons_law}.
+It is desirable for a numerical method to satisfy a discrete version of the entropy inequality XX. 
+This is, however, not true in general. 
+In general, an entropy stable scheme introduces some type of nonlinear stabilization, which is 
+particularly important for solutions with strong discontinuities that might converge to non-physical 
+weak solutions. 
+In this work, we propose an entropy stable approximate Riemann solver for the one dimensional 
+shallow water equations. We use this Riemann solver as building block within the Lax-Wendroff 
+method, as presented in XX and implemented in Clawpack XX.
 
-\subsection{Roe's Riemann solver}
+\subsection{Entropy residual}
 
-\subsection{Rusanov's Riemann solver}
+Let $S_i$ be the set of cells containing cell $i$ and those that share a face with it. 
+Based on XX, we consider 
+\begin{align}\label{ent_residual}
+  \int_{S_i} \bfv(\bfu) \cdot \left[ \frac{\partial \bfu}{\partial t} + \nabla\cdot \bff(\bfu)\right]d\bfx
+  =\int_{S_i} \left[\frac{\partial \eta(\bfu)}{\partial t} + \bfv(\bfu) \cdot \nabla\cdot \bff(\bfu)\right] d\bfx
+\end{align}
+as a measurement of the entropy production in $S_i$. To avoid approximating the time derivative of the entropy, 
+we follow XX, XX and use XX to approximate XX=XX, which holds in smooth regions. Then we define 
+\begin{align}\label{Ri}
+  R_i := \frac{N_i}{D_i}, 
+\end{align}
+with
+\begin{align*}
+  N_i=
+  \left|\int_{S_i} \left[\bfv(\bfu) \cdot \nabla\cdot \bff(\bfu) - \nabla\cdot\bfq(\bfu) \right] d\bfx\right|,
+\end{align*}
+and $D_i$ being an upper bound of $N_i$ that acts as normalization so that $0\leq R_i\leq 1$.
+Note that $N_i=0$ if $\bfu$ is smooth in $S_i$. 
+In our implementation, we use
+\begin{align*}
+  N_i\approx \left|\bfv(\bfu_i)\cdot \int_{S_i}\nabla\cdot \bff(\bfu)d\bfx 
+  -\int_{S_i}\nabla\cdot\bfq(\bfu) d\bfx\right|,
+  \quad 
+  D_i = ||\bfv(\bfu_i)||_{\ell^2}\left|\left|\int_{S_i}\nabla\cdot\bff(\bfu)d\bfx\right|\right|_{\ell^2}
+  +\left|\int_{S_i}\nabla\cdot\bfq(\bfu)d\bfx\right|.
+\end{align*}
+In the next section we use $R_i$ to blend the Rusanov's and the Roe's Riemann solvers.
 
-\subsection{Blended Rusanov's and Roe's solver based on entropy residual}
+\subsection{Blended Riemann solver}
+In this section, we use \eqref{Ri}, which is a normalized measurement of the entropy residual \eqref{ent_residual}, 
+to blend the Rusanov's and the Roe's Riemann solvers. To do this, we define the following numerical flux:
+\begin{align}\label{num_flux_ev}
+  F_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} 
+  - \frac{1}{2} \left( R_{i+1/2}\lambda_{i+1/2}^{\max} + (1-R_{i+1/2})|\hat A_{i+1/2}| \right)(\bfu_{i+1}-\bfu_{i}),
+\end{align}
+and similarly for $F_{i-1/2}$. Here $R_{i+1/2}=\max\{R_{i+1},R_i\}$, $\lambda_{i+1/2}^{\max}$ is the 
+upper bound on the wave speed used in the Rusanov's Riemann solver, see \S XX, and $\hat A_{i+1/2}$
+is the Roe's averaged flux Jacobian, see \S XX.
 
-\subsection{Blended Rusanov's and Roe's solver based on Tadmor's entropy stability condition}
+%In terms of fluctuations, 
+The first-order method is given simply by \eqref{first-order_FV} 
+with the numerical flux $F_{i+1/2}$ given by \eqref{num_flux_ev}. 
+In smooth regions $R_{i+1/2}\approx 0$ so the numerical flux is essentially The Roe's numerical flux. 
+Near discontinuities or steep gradients, $R_{i+1/2}$ becomes large. If, in particular, $R_{i+1/2}\approx 1$, 
+then the numerical flux is essentially the Rusanov's numerical flux.  
+%
+The first-order method can also be written in terms of fluctuations. 
+To do this we consider the waves ($\W^p_{i\pm 1/2}$, with $p=1,2,3$) 
+from the Roe's solver and define 
+\begin{align}
+  \lambda_{i+1/2}^p := R_{i+1/2}\lambda_{i+1/2}^{\max} + (1-R_{i+1/2})|\hat \lambda_{i+1/2}^p|,
+\end{align}
+where $\hat\lambda_{i+1/2}^p$ is the $p$-th eigenvalue of the Roe's solver corresponding to the 
+interface $i+1/2$. The fluctuations are then given by 
+\begin{subequations}\label{ev_fluctuations}
+\begin{align}
+  \A^+\Delta Q_{i-1/2}&=\frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^p, \\
+  \A^-\Delta Q_{i+1/2}&=\frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^p.
+\end{align}
+\end{subequations}
 
-\subsection{Entropy stable enhancement of Roe's Riemann solver}
+To implement the second-order Lax-Wendroff type method in Clawpack, the Riemann solver must provide 
+the fluctuations, the waves and their speed. The fluctuations are simply \eqref{ev_fluctuations} 
+and the waves are the Roe's waves $\W_{i+1/2}^p$, with $p=1,2,3$. 
+In the framework of LeVeque, for a given interface $i+1/2$, 
+we would need to consider a wave speed $s_{i+1/2}^p$ so that 
+\begin{align*}
+  \hat \lambda_{i+1/2}^p+\lambda_{i+1/2}^p = s_{i+1/2}^p+|s_{i+1/2}^p|, \qquad 
+  \hat \lambda_{i+1/2}^p-\lambda_{i+1/2}^p = s_{i+1/2}^p-|s_{i+1/2}^p|,
+\end{align*}
+which is not possible to obtain in general (since the problem is overdetermined). Instead, we consider 
+simply $s_{i+1/2}^p=\hat\lambda_{i+1/2}^p$ and study the consequences of this choice 
+once the second-order corrections are applied.
+The second-order Lax-Wendroff-LeVeque method for the one-dimensional problem is given by 
+\begin{align}\label{second-order}
+  u_i^{n+1}=u_i^n
+  -\frac{\Delta t}{\Delta x}\left[\A^-\Delta Q_{i+1/2}+\A^+\Delta Q_{i-1/2}\right]
+  -\frac{\Delta t}{\Delta x}\left[\tilde F_{i+1/2}-\tilde F_{i-1/2}\right],
+\end{align}
+where 
+\begin{align*}
+  \tilde F_{i+1/2} = \frac{1}{2}\sum_p|\hat\lambda_{i+1/2}^p|
+  \left(1-\frac{\Delta t}{\Delta x}|\hat\lambda_{i+1/2}^p|\right)\tilde\W_{i+1/2}^p,
+\end{align*}
+and similarly for $\tilde F_{i-1/2}$, are the flux corrections with $\tilde\W_{i+1/2}^p$ 
+being a TVD-limited version of $\W_{i+1/2}^p$. 
+In this work, we consider the min-mod limiters, we refer to XX for more details.
 
-\subsection{Other alternatives}
+To understand the effect of choosing $s_{i+1/2}^p=\hat\lambda_{i+1/2}^p$ as the speed of $\W_{i+1/2}^p$, 
+let us assume that no limiters are applied; i.e., $\tilde\W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$.
+By doing this, equation XX can be written as 
+\begin{align*}
+  u_i^{n+1}=u_i^n
+  &-\frac{\Delta t}{2 \Delta x}
+  \sum_p
+  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}|\hat\lambda_{i+1/2}^p|^2 \right)\W_{i+1/2}^p
+  +
+  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}|\hat\lambda_{i-1/2}^p|^2 \right)\W_{i-1/2}^p\right] \\
+  &-\frac{\Delta t}{2\Delta x}
+  \sum_p 
+  \left[-\left(\lambda_{i+1/2}^{\max}-|\hat\lambda_{i+1/2}^p|\right)R_{i+1/2}\W_{i+1/2}^p
+    +\left(\lambda_{i-1/2}^{\max}-|\hat\lambda_{i-1/2}^p|\right)R_{i-1/2}\W_{i-1/2}^p\right],
+\end{align*}
+where the first row corresponds to the standard Lax-Wendroff method written in terms of 
+fluctuation-like quantities. 
+Since $\lambda_{i\pm1/2}^{\max}\geq |\hat\lambda_{i\pm1/2}^p|$, 
+the second row adds extra stabilization.
+Note that if $R_{i\pm 1/2}=0$, we recover the standard second-order Lax-Wendroff method based on 
+Roe's Riemann solver. 
+In the numerical experiments in \S XX, we obtain second-order experimental 
+order of convergence (EOC) for problems with smooth solution, which is expected since 
+$R_{i\pm 1/2}\approx 0$ in smooth regions.
+Also note that if $R_{i\pm 1/2}\approx 1$, which we expect to happen near steep gradients, 
+we do not recover the Lax-Wendroff method based on Rusanov's Riemann solver. 
+Instead, we obtain the first-order method XX with a flux correction. Assuming 
+the CFL-like condition $\Delta t/\Delta x|\hat\lambda_{i+1/2}^p|-1\leq 0$, the flux correction is anti-diffusive, 
+which slightly improves the accuracy of the first-order Rusanov's method. 
+In our numerical experiments 
+with smooth solutions, we obtain first-order EOC if we set $R_{i\pm1/2}=1$, $\forall i$. 
+%
+Therefore, the accuracy properties of method XX depend upon the behavior of $R_{i\pm 1/2}$. 
+In smooth regions, where $R_{i\pm 1/2}\approx 0$, we recover second-order convergence;
+near discontinuities, where $R_{i\pm 1/2}\approx 1$, we add extra dissipation that decreases the convergence 
+to first-order. It is this extra dissipation that reduces the occurance of carbuncle instabilities 
+in our numerical experiments. In smooth regions, the accuracy of the underlying Lax-Wendroff-LeVeque method 
+based on Roe's Riemann solver is preserved.
+
+\subsection{Entropy stable Riemann solver}
+
+\clearpage
 
 \section{Numerical results}
+Description of the methods. We consider either the first-order method XX or the second-order Lax-Wendroff
+type method XX. For each of these methods we use one of the following three Riemann solvers: 
+\begin{itemize}
+  \item LLF: Rusanov's or Local Lax Friederichs 2-wave Riemann solver reviewed in \S XX.
+  \item Roe: standard Roe's average based 3-wave Riemann solver reviewed in \S XX. 
+  \item Roe-fix: Roe's solver with entropy fix based on XX. 
+  \item EV:  entropy dissipative 3-wave Riemann solver described in \S XX.
+\end{itemize}
 
+\subsection{Dam break problem on a dry bed}
+
+\subsection{Dam break problem on a wet bed}
+We consider a one-dimensional dam break problem on a wet domain. 
+We follow the setup in XX. 
+The domain is $\Omega=(0,L)$ with $L=10$, the initial condition is given by 
+$hu(x,0)=0$ and 
+\begin{align*}
+  h(x,0) = 
+  \begin{cases}
+    h_l & \mbox{ if } x\in(0,x_0], \\
+    h_r & \mbox{ if } x\in(x_0,L),
+  \end{cases}
+\end{align*}
+with $x_0=5$, $h_l=0.005$ and $h_r=0.001$. 
+The exact solution, which can be found in XX and references therein, is given by 
+\begin{align*}
+  h(x,t) = 
+  \begin{cases}
+    h_l, \\
+    \frac{4}{9g}\left(\sqrt{gh_l}-\frac{x-x_0}{2t}\right)^2, \\
+    \frac{c_m^2}{g}, \\
+    h_r, 
+  \end{cases}
+\quad 
+  u(x,t) = 
+  \begin{cases}
+    0, &\mbox{ if } x\leq x_A(t), \\
+    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
+    2(\sqrt{gh_l}-c_m), & \mbox{ if } x_B(t)<x\leq x_C(t), \\
+    0, &\mbox{ if } x\leq x_C(t),
+  \end{cases}
+\end{align*}
+where $x_A(t)=x_0-t\sqrt{gh_l}$, $x_B(t)=x_0+t\left(2\sqrt{gh_l}-3c_m\right)$, 
+$x_C(t)=x_0+t\frac{2c_m^2\left(\sqrt{gh_l}-c_m\right)}{c_m^2-gh_r}$ and 
+$c_m$ is the solution of 
+$-8gh_rc_m^2\left(\sqrt{gh_l}-c_m\right)^2+\left(c_m^2-gh_r\right)^2\left(c_m^2+gh_r\right)=0$.
+We use $g=1$ and solve the problem up to the final time $T=5$ considering the 2nd-order Lax-Wendroff 
+type method XX with the LLF, Roe and EV Riemann solvers. The solution with different refinement levels 
+and each Riemann solver is shown in Figure XX. In Table XX, we summarize the results of a convergence test. 
+Since the solution is non-smooth, we expect no more than first order convergence rates. Note that the 
+results with the entropy dissipative solver EV are comparable against the results using Roe's solver. 
+That is, inducing entropy dissipation via the EV Riemann solver does not degrade the high-order accuracy 
+properties of the Roe's solver. 
+In contrast, the accuracy and convergence rates using the LLF solver are clearly degraded.
+
+\subsection{The circular hydraulic jump in one dimension}
+
+\subsection{Steady outflow}
+Let us consider a problem similar to that in the previous section, but with an outflow boundary condition. 
+That is, we consider the one-dimensional domain $\Omega=(0.1,1)$, the initial condition is 
+$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$, 
+and the right boundary condition is set to outflow, see the details in XX. 
+If the flow is supercritical ($|F|>1$), the exact solution converges to a 
+steady state profile given by the solution of the ODE XX. 
+Depending on the initial condition, shocks might develop before the steady state is reached. 
+Nevertheless, the steady state profile is smooth. 
+In Figure XX, we show the solution at different times based on the secnd-order method XX with 
+the EV Riemann solver. Additionally, in Table XX we summarize the result of a convergence study 
+based on the first- and second-order methods XX and XX, respectively. 
+We consider the Riemann solver in \S XX with Ri=0, Ri=1 and Ri computed via XX.
+
+\subsection{The circular hydraulic jump in two dimensions}
 We consider two regimes. In the first regime, the physical instabilities are mild. 
@@
 
+\clearpage
 \section{Conclusions}

### 2020-11-04 numerics_paper/version2/numerical_chj.tex
tmp update

@@
 
-%\subsection{Roe's Riemann solver}
+\subsection*{Roe's Riemann solver}
 \begin{align*}
@@
   \sum_p 
-  \left[-\left(\lambda_{i+1/2}^{\max}-|\hat\lambda_{i+1/2}^p|\right)R_{i+1/2}\W_{i+1/2}^p
-    +\left(\lambda_{i-1/2}^{\max}-|\hat\lambda_{i-1/2}^p|\right)R_{i-1/2}\W_{i-1/2}^p\right],
+  \left[-R_{i+1/2}\left(\lambda_{i+1/2}^{\max}-|\hat\lambda_{i+1/2}^p|\right)\W_{i+1/2}^p
+    +R_{i-1/2}\left(\lambda_{i-1/2}^{\max}-|\hat\lambda_{i-1/2}^p|\right)\W_{i-1/2}^p\right],
 \end{align*}
@@
 Instead, we obtain the first-order method XX with a flux correction. Assuming 
-the CFL-like condition $\Delta t/\Delta x|\hat\lambda_{i+1/2}^p|-1\leq 0$, the flux correction is anti-diffusive, 
+the CFL-like condition $\Delta t/\Delta x|\hat\lambda_{i\pm1/2}^p|-1\leq 0$, the flux correction is anti-diffusive, 
 which slightly improves the accuracy of the first-order Rusanov's method. 
@@
 to first-order. It is this extra dissipation that reduces the occurance of carbuncle instabilities 
-in our numerical experiments. In smooth regions, the accuracy of the underlying Lax-Wendroff-LeVeque method 
-based on Roe's Riemann solver is preserved.
+in our numerical experiments.
+%In smooth regions, the accuracy of the underlying Lax-Wendroff-LeVeque method 
+%based on Roe's Riemann solver is preserved.
 

### 2020-11-12 numerics_paper/version2/numerical_chj.tex
update

--- 2020-10-14 numerics_paper/version2/numerical_chj.tex
@@
 \usepackage[hmargin=0.9in, vmargin=1.25in]{geometry}
-\usepackage[babel,german=quotes]{csquotes}
+\usepackage[babel]{csquotes}
 \usepackage{subfig}
@@
 \newcommand{\A}{{\mathcal A}}
+\newcommand{\bff}{{\bf f}}
+\newcommand{\bfr}{{\bf r}}
+\newcommand{\bfF}{{\bf F}}
+\newcommand{\bfu}{{\bf u}}
+\newcommand{\bfv}{{\bf v}}
+\newcommand{\bfq}{{\bf q}}
+\newcommand{\bfx}{{\bf x}}
+\newcommand{\sgn}{{\rm sgn}}
+\newcommand{\Rus}{{\rm Rus}}
+\newcommand{\Roe}{{\rm Roe}}
+
+%\newdefinition{rmk}{Remark}
+\newtheorem{remark}{Remark}
 
@@
 \item We demonstrate that these schemes prevent the formation of cc instabilities not only 
-  in the well known bow shock problem, but also in the CHJ while preserving other important features of 
+  in the well-known bow shock problem, but also in the CHJ while preserving other important features of 
   the solution. 
@@
 
+\clearpage
 \section{Numerical methods}
-We consider the implementation of the Lax-Wendroff method in Clawpack. 
-For a one-dimensional problem, the method can be written as follows:
+In this work, we consider Riemann-based finite volume methods.
+Considering a one dimensional hyperbolic conservation law,
+the cell average of the numerical solution given by a 
+finite volume scheme (with forward Euler) can be written as follows:
+\begin{align}\label{first-order_FV}
+  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
+\end{align}
+where $\bfF_{i+1/2}=\bfF_{i+1/2}(u^n)$ is a numerical flux; i.e., $\bfF_{i+1/2}$ is 
+a discretization of the flux $\bff(\bfu(x_{i+1/2},t^n))$. 
+This numerical flux is composed of two parts, a consistent term that approximates 
+the physical flux at the interface between cells $i$ and $i+1$, and a stabilization term. 
+For instance, the well-known Rusanov's numerical flux is given by 
+\begin{align}\label{rusanov_num_flux}
+  \bfF_{i+1/2}(\bfu)=\frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} - \lambda_{i+1/2}(\bfu_{i+1}-\bfu_i),
+\end{align}
+where $\lambda_{i+1/2}$ is an upper bound on the wave speed associated with the Riemann problem 
+composed of the states $\bfu_i$ and $\bfu_{i+1}$. 
+%If $z=u$ in \eqref{rusanov_num_flux}
+%; i.e., if the Rusanov's numerical flux is obtained considering 
+%the cell averages of the solution 
+%and if the time step is small enough, 
+Method \eqref{first-order_FV} with Rusanov's flux \eqref{rusanov_num_flux}
+is positivity preserving and entropy stable, see for instance 
+\cite{perthame1996positivity,tadmor2003entropy,guermond2016invariant}
+However, the accuracy of the numerical flux, and hence of the method, is 
+limited to first-order. 
+Two alternatives are common to improve the accuracy properties of Rusanov's method.
+The numerical flux $\bfF_{i+1/2}$ can be approximated based on high-order polynomial 
+reconstructions (or combinations of them) around the interface $i+1/2$. 
+For smooth solutions, this strategy leads to arbitrarily high-order
+methods. Alternatively, second-order accuracy can be obtained by removing some of the 
+numerical dissipation introduced by the stabilization component of the numerical flux. 
+The second-order Lax-Wendroff method \cite{lax1959systems} is a popular example of this approach. 
+In this work we consider an adaptation of Lax-Wendroff method,
+sometimes known as Lax-Wendroff-LeVeque \cite{leveque1997wave,leveque2002finite}, 
+that incorporates limiters into the second-order correction. 
+These high-order corrections to Rusanov's method require careful treatment to guarantee 
+the solution is still positivity preserving and/or entropy stable.
+
+The accuracy of a given first- or second-order method depends heavily on the Riemann solver. 
+In particular, the Rusanov's Riemann solver is highly dissipative. In contrast, the well-known 
+Roe's Riemann solver produces more accurate solutions. 
+%
+As we discussed in the introduction, finite volume methods with 
+Roe's Riemann solvers are prone to the carbuncle instability for the CHJ. 
+In contrast, if we use the Rusanov's 
+Riemann solver, no carbuncles appear; however, many other important and physical features of 
+the solution are dissipated, unless the grid is highly refined. 
+Our main objective is to study the CHJ and to propose a Riemann solver
+that is carbuncle free but do not dissipate important features of the solution. 
+To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers. 
+Our blending criteria is based on the residual of the entropy (in)equality. 
+Similar ideas have been explored before to solve the Euler equations XX 
+and the shallow water equations XX. In these references, the authors use different 
+blending criteria.
+
+\subsection{Lax-Wendroff-LeVeque method}
+Let us now review the second-order correction to method \eqref{first-order_FV} implemented in 
+Clawpack \cite{clawpack} and PyClaw \cite{pyclaw-sisc}.
+We refer to \cite{leveque1997wave, leveque2002finite} for a detailed description of the method. 
+We consider only a one-dimensional problem; in the 
+numerical experiments in \S\ref{sec:num} we use second-order Strang's splitting \cite{strang1968construction} 
+to extend the method to multiple dimensions.
+
+The finite volume method \eqref{first-order_FV} can be written in terms of right- 
+and left-going waves emanating from the interfaces $i-1/2$ and $i+1/2$, respectively. 
+The net contribution of these waves is called right- and left-going fluctuations. 
+Upon defining $z^+:=\frac{1}{2}(z+|z|)$ and $z^-:=\frac{1}{2}(z-|z|)$, 
+the fluctuations are given by
+\begin{align}\label{fluct}
+  %\A^+\Delta Q_{i-1/2} := \frac{1}{2}\sum_p\left(s_{i-1/2}^p+|s_{i-1/2}^p|\right)\W_{i-1/2}^p, \quad
+  %\A^-\Delta Q_{i+1/2} := \frac{1}{2}\sum_p\left(s_{i+1/2}^p-|s_{i+1/2}^p|\right)\W_{i+1/2}^p,
+  \A^+\Delta Q_{i-1/2} := \sum_p\left(s_{i-1/2}^p\right)^+\W_{i-1/2}^p, \qquad
+  \A^-\Delta Q_{i+1/2} := \sum_p\left(s_{i+1/2}^p\right)^-\W_{i+1/2}^p,
+\end{align}
+where $\W_{i\pm 1/2}^p$ and $s_{i\pm 1/2}^2$ are the waves and their corresponding speeds, respectively. 
+The waves and their speed of propagation (to compute the fluctuations) are tightly related to Riemann solvers. 
+In \S\ref{sec:riemann_solvers} we review the definition of the waves and speeds associated with 
+Rusanov's and Roe's Riemann solvers.
+Based on fluctuations, method \eqref{first-order_FV} can be written as 
+\begin{align}\label{first-order_via_fluct}
+  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right].
+\end{align}
+
+The second-order Lax-Wendroff-LeVeque method removes some of the numerical dissipation in 
+\eqref{first-order_via_fluct} via a flux correction at each interface. The method is written as follows:
+\begin{align}\label{second-order_via_fluct}
+  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right]
+  -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
+\end{align}
+where 
 \begin{align*}
-  U_i^{n+1}=U_i^n-\frac{\Delta t}{\Delta x}\left[\A\Delta Q_{i-1/2}]
+  \tilde{F}_{i\pm 1/2}=\frac{1}{2}\sum_p|s_{i\pm 1/2}^p|\left(1-\frac{\Delta t}{\Delta x}|s_{i\pm 1/2}^p|\right)\tilde\W_{i\pm 1/2}^p
 \end{align*}
-where XX. In this section we focus on the first order method; i.e., we consider
+are the flux corrections at the interfaces $i\pm 1/2$. Here $\tilde{\W}_{i\pm 1/2}^p$ is a 
+TVD-limited version of $\W_{i\pm 1/2}^p$. 
+In this work, we consider the min-mod limiters, we refer to \cite{leveque1997wave} for more details. 
+If no limiters are used 
+($\tilde \W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$), \eqref{second-order_via_fluct} becomes
+the standard second-order Lax-Wendroff scheme, which in terms of fluctuation-like quantities 
+is given by
+\begin{align*}
+  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}
+\left[
+  \frac{1}{2}\left(s_{i-1/2}^p+\frac{\Delta t}{\Delta x}|s_{i-1/2}^p|^2\right)\W_{i-1/2}^p
+  +
+  \frac{1}{2}\left(s_{i+1/2}^p-\frac{\Delta t}{\Delta x}|s_{i+1/2}^p|^2\right)\W_{i+1/2}^p
+\right].
+\end{align*}
+To implement method \eqref{second-order_via_fluct}, 
+we need to choose a Riemann solver. Although the scheme is second-order 
+regardless of the Riemann solver, the accuracy and robustness depends upon the 
+Riemann solver. The approximate Riemann solver based on Rusanov's flux is robust and known to be carbuncle free. 
+However, it adds large amounts of numerical diffusion that, for relative coarse grids, 
+might dissipate other physical and important features of the solution. If the apprroximate Riemann solver 
+is based on the Roe's average, the amount of numerical diffusion is highly reduced; however, 
+such an approach is prone to develop the carbuncle instability for certain problems. 
+Many carbuncle fixes to the Roe's solver have been proposed for the Euler equations. 
+In the context of the shallow water equations, we review in \S\ref{sec:riemann_solvers} 
+the Riemann solver proposed in \cite{kemm2014note}.
+
+\section{Riemann solvers}\label{sec:riemann_solvers}
+
+For the shallow water equations in two-dimensions, the structure of the exact (and some 
+approximated) Riemann solvers contains three waves. This is the case, for example, of the well-known 
+Roe's Riemann solver, see for instance \cite{roe1981approximate} and 
+\cite[\S 15.3.3 and \S 21.7]{leveque2002finite}. 
+Simpler approaches are also possible. 
+For instance, methods \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}
+with the Rusanov's numerical flux are based on only two waves that 
+propagate with the same speed in opposite directions.
+We refer to \cite{ketcheson2020riemann} and references therein for 
+a full and detailed review of these two Riemann solvers.
+
+\subsection{Roe's average}\label{sec:roe}
+The Roe's Riemann solver is based on evaluating the flux Jacobian using the Roe's average
+\begin{align}\label{roe_average}
+  \bar h_{i-1/2}=\frac{1}{2}(h_{i-1}+h_i), \qquad 
+  \hat u_{i-1/2}=\frac{\sqrt{h_{i-1}}u_{i-1}+\sqrt{h_i}u_i}{\sqrt{h_{i-1}}+\sqrt{h_i}}, \qquad
+  \hat v_{i-1/2}=\frac{\sqrt{h_{i-1}}v_{i-1}+\sqrt{h_i}v_i}{\sqrt{h_{i-1}}+\sqrt{h_i}}.
+\end{align}
+Let 
+\begin{align*}
+\hat A_{i-1/2} =
+\begin{bmatrix}
+  0 & 1 & 0 \\
+  -\hat u_{i-1/2}^2+g\bar h_{i-1/2} & 2\hat u_{i-1/2} & 0 \\
+  -\hat u_{i-1/2} \hat v_{i-1/2} & \hat v_{i-1/2} & \hat u_{i-1/2}
+\end{bmatrix}
+\end{align*}
+be the flux Jacobian evaluated using the Roe's average at the interface $i-1/2$.
+The average \eqref{roe_average} is designed to guarantee: 
+i)  consistency of the approximate flux Jacobian 
+($\hat A_{i-1/2}\rightarrow \bff^\prime(\bfu(x_{i-1/2}))$ as $\bfu_i, \bfu_{i-1}\rightarrow \bfu(x_{i-1/2})$); 
+ii) hyperbolicity ($\hat A_{i-1/2}$ is diagonalizable with real eigenvalues);
+and iii) conservation ($\hat A_{i-1/2}(\bfu_i-\bfu_{i-1})=\bff(\bfu_i)-\bff(\bfu_{i-1})$). 
+Let
+\begin{align*}
+  \hat \bfr^1_{i-1/2} = 
+  \begin{bmatrix}
+    1 \\ 
+    \hat u_{i-1/2}-\sqrt{g\bar h_{i-1/2}}\\
+    \hat v_{i-1/2}
+  \end{bmatrix},
+  \qquad 
+  \hat \bfr^2_{i-1/2} = 
+  \begin{bmatrix}
+    0\\ 
+    0\\ 
+    1
+  \end{bmatrix},
+  \qquad
+  \hat \bfr^3_{i-1/2} = 
+  \begin{bmatrix}
+    1 \\ 
+    \hat u_{i-1/2}+\sqrt{g\bar h_{i-1/2}}\\
+    \hat v_{i-1/2}
+  \end{bmatrix}
+\end{align*}
+be the eigenvectors of $\hat A_{i-1/2}$.
+The waves for the two-dimensional Riemann solver based on the Roe's average are 
+$\W_{i-1/2}^p=\alpha_{i-1/2}^p\hat r_{i-1/2}^p$, with $p=1,2,3$. Here $\alpha_{i-1/2}^p$ is given by
+\begin{align}\label{system_for_alphas}
+  \left[\hat \bfr^1_{i-1/2} ~\hat \bfr^2_{i-1/2} ~\hat \bfr^3_{i-1/2}\right]
+  \begin{bmatrix}
+    \alpha^1_{i-1/2} \\
+    \alpha^2_{i-1/2} \\
+    \alpha^3_{i-1/2}
+  \end{bmatrix}
+  =\Delta Q_{i-1/2}:=\bfu_i-\bfu_{i-1}.
+\end{align} 
+%where $\alpha_{i-1/2}^p$ is the $p$-th entry from the solution of 
+%\end{align*}
+The speed of the waves is given by the eigenvalues of $\hat A_{i-1/2}$
+\begin{align*}
+  \hat\lambda_{i-1/2}^1=\hat u_{i-1/2}-g\sqrt{g\bar h_{i-1/2}}, \qquad
+  \hat\lambda_{i-1/2}^2=\hat u_{i-1/2}, \qquad
+  \hat\lambda_{i-1/2}^3=\hat u_{i-1/2}+g\sqrt{g\bar h_{i-1/2}}. \\
+\end{align*}
+
+The waves and speeds at interface $i+1/2$ are computed similarly. 
+The fluctuations are given by \eqref{fluct} with $s_{i\pm 1/2}^p=\hat\lambda_{i\pm 1/2}^p$ 
+Finally, the first- and second-order methods based on Roe's Riemann solver are given by 
+\eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively. 
+We close this review by remarking that the conservation property 
+$\hat A_{i-1/2}(\bfu_i-\bfu_{i-1})=\bff(\bfu_i)-\bff(\bfu_{i-1})$ implies that 
+\begin{align}\label{cons_roe}
+  \A^-\Delta Q_{i-1/2}+\A^+\Delta Q_{i-1/2}=\sum_p \hat\lambda_{i-1/2}^p\W_{i-1/2}^p=\bff(\bfu_{i})-\bff(\bfu_{i-1}),
+\end{align}
+a property we will use in the next section.
+
+\subsection{Rusanov's Riemann solver}\label{sec:rusanov}
+The Rusanov's Riemann solver considers only two waves that emanate from each interface, 
+say for instance interface $i-1/2$, and travel with the same speed $\lambda_{i-1/2}^{\max}\geq 0$ 
+but in opposite directions. 
+The speed of propagation $\lambda_{i-1/2}^{\max}$ is an upper bound on the maximum wave speed
+associated with the corresponding Riemann problem between the states $\bfu_{i-1}$ and $\bfu_{i}$. 
+In this work, we define $\lambda_{i-1/2}^{\max}$ via the guaranteed upper bound from \cite[Prop. 3.7]{azerad2017well}. 
+The waves are given by 
+\begin{align}\label{rusanov_waves}
+  \W_{i-1/2}^{1,\Rus}=\bfu_{i-1/2}-\bfu_{i-1}, \qquad 
+  \W_{i-1/2}^{2,\Rus}=\bfu_{i}-\bfu_{i-1/2}, 
+\end{align}
+where $\bfu_{i-1/2}$ is a middle state in the solution of the Riemann problem. The middle state is obtained
+by imposing
+\begin{align}\label{rus_cons}
+  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{1,\Rus}+\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus} = \bff(\bfu_{i})-\bff(\bfu_{i-1}),
+\end{align}
+which guarantees the solver is conservative. 
+The fluctuations are given by \eqref{fluct} with $s_{i-1/2}^1=-\lambda_{i-1/2}^{\max}$
+and $s_{i-1/2}^2=\lambda_{i-1/2}^{\max}$, and the first- and second-order methods based 
+on Rusanov's solver are given by \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.
+
+We close this section by remarking that the fluctuations associated with Rusanov's solver 
+can be expressed in terms of waves and speeds from the Roe's solver. To see this, first consider \eqref{rus_cons}
+and rewrite the right-going fluctuation as follows:
+\begin{align}\label{rus_rs_aux}
+  \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+  = \bff(\bfu_{i})-\bff(\bfu_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
+  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}.
+  %\A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
+\end{align}
+From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get 
+$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$. 
+Using \eqref{cons_roe}, \eqref{rus_rs_aux} becomes 
+\begin{align*}
+  \lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+  = \sum_p \left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
+  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus},
+\end{align*}
+which implies that 
+\begin{subequations}\label{fluct_rus}
+\begin{align}
+  \A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p,
+\end{align}
+and similarly, 
+\begin{align}
+  \A^{-,\Rus}\Delta Q_{i+1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i+1/2}^p - \lambda_{i+1/2}^{\max}\right)\W_{i+1/2}^p.
+\end{align}
+\end{subequations}
+
+\subsubsection{HLLEMCC solver}
+The HLLEMCC approximate Riemann solver for the shallow water equations was proposed
+in \cite{kemm2014note} specifically to deal with carbuncles.  The idea is to adjust
+the amount of dissipation applied to the shear wave, adding dissipation in regions
+where unphysical carbuncles would appear.  The approximate solution in this
+case consists of four waves; the faster waves use the HLLE wave speeds and carry
+jumps only in the depth and normal momentum, while the slower waves carry jumps
+only in the transverse momentum.  By adjusting the value of $\phi$, this solver
+can behave like the HLLE solver (if $\phi=1$) or like the Roe solver (if $\phi=0$).
+
+The middle states $q^1_m, q^2_m$ are the same as those used for the HLLE solver.
+Requiring conservation of the transverse momentum gives
+\begin{align*}
+  q^3_m = \frac{f^3_l - f^3_r - (\hat{v}-\phi\hat{c})q^3_l + (\hat{v}+\phi\hat{c})q^3_r}{2\phi\hat{c}}.
+\end{align*}
+Here $\hat{v}, \hat{c}$ are the Roe-average normal velocity and sound speed, respectively, whereas the function $\phi$ is given by
+\begin{equation}\label{if}
+\phi(\theta)=\min\left\lbrace 1,\, \left(\varepsilon\,\theta\,\max\left\lbrace0,\,1-Fr_{u}^{\alpha} \right\rbrace \right)   \right\rbrace^{\beta}
+\end{equation}
+where $Fr_{u}$ is the Froude number perpendicular to the cell face and $\varepsilon$, $\alpha$, and $\beta$ are parameters that have to be chosen appropriately.
+For the shallow water equations, authors in \cite{kemm2014note} propose to use $\varepsilon=10^{-3}$ and $\alpha=\beta=\frac{1}{3}$.
+The value of $\theta$ is calculated as the 2-norm of the residual in the Rankine-Hugoniot condition for the shear wave.
+
+The HLLEMCC solver follows the ideas of the Harten-Hyman entropy fix by splitting the shear wave with speed $\hat{u}$, into two waves moving with speeds $\hat{u}-\phi\hat{c}$ and $\hat{u}+\phi\hat{c}$ (see figure \ref{HLLEMCC}).
+Its implementation, in terms of the fluctuations, is performed by using the values $\hat{u}^{\pm}$ calculated with an absolute-value function defined as
+
+\begin{align*}
+\varphi(\lambda)=\frac{\vert \lambda-\phi(\theta)\hat{c}\,\vert+\vert \lambda+\phi(\theta)\hat{c}\,\vert}{2}
+\end{align*}
+
+\begin{figure}
+    \center
+    \includegraphics[width=0.6\textwidth]{figures/hllemcc.pdf}
+    \caption{Structure of the HLLEMCC approximate Riemann solution, with four waves.}
+    \label{HLLEMCC}
+\end{figure}
+
+\section{Entropy dissipative Riemann solver}\label{sec:ent_diss_rs}
+Consider a hyperbolic system of conservation laws 
+\begin{align}\label{cons_law}
+  \frac{\partial \bfu(\bfx)}{\partial t} + \nabla\cdot \bff(\bfu(\bfx))=0, \quad \bfx\in\Omega,
+\end{align}
+with a en entropy pair $\left(\eta(\bfu), \bfq(\bfu)\right)$. 
+A weak solution of \eqref{cons_law} is called an entropy solution if 
+\begin{align}\label{ent_inequality}
+  \frac{\partial \eta(\bfu(\bfx))}{\partial t} + \nabla\cdot \bfq(\bfu(\bfx))\leq 0, \quad \bfx\in\Omega,
+\end{align}
+holds, in a distribution sense, for any entropy pair.
+It is desirable for a numerical method to satisfy a discrete version of the entropy inequality 
+\eqref{ent_inequality}. This is, however, not true for many standard numerical discretizations. 
+In general, an entropy stable scheme introduces nonlinear stabilization, which is 
+particularly important for solutions with strong discontinuities that might converge to non-physical 
+weak solutions. 
+In this work, we propose an entropy stable approximate Riemann solver for the one dimensional 
+shallow water equations. We use this Riemann solver as building block within the Lax-Wendroff-LeVeque 
+method, as described in \cite{leveque1997wave,leveque2002finite} and implemented in Clawpack.
+
+\subsection{Entropy residual}
+Let $S_i$ be the set of cells containing cell $i$ and those that share a face with it. 
+Also let $\bfv(\bfu)=\eta^\prime(\bfu)$ be the entropy variable. 
+Based on \cite{guermond2011entropy}, we consider 
+\begin{align}\label{ent_residual}
+  \int_{S_i} \bfv(\bfu) \cdot \left[ \frac{\partial \bfu}{\partial t} + \nabla\cdot \bff(\bfu)\right]d\bfx
+  =\int_{S_i} \left[\frac{\partial \eta(\bfu)}{\partial t} + \bfv(\bfu) \cdot \nabla\cdot \bff(\bfu)\right] d\bfx
+\end{align}
+as a measurement of the entropy production in $S_i$. 
+To avoid approximating the time derivative of the entropy, we follow 
+\cite{guermond2018second, guermond2018well} and use 
+$\int_{K_i} \frac{\partial\eta(\bfu)}{\partial t}d\bfx=-\int_{K_i} \nabla\cdot\bfq(\bfu) d\bfx$, 
+which holds in smooth regions. Then we define 
+\begin{align}\label{Ri}
+  R_i := \frac{N_i}{D_i}, 
+\end{align}
+with
+\begin{align*}
+  N_i=
+  \left|\int_{S_i} \left[\bfv(\bfu) \cdot \nabla\cdot \bff(\bfu) - \nabla\cdot\bfq(\bfu) \right] d\bfx\right|,
+\end{align*}
+and $D_i$ being an upper bound of $N_i$ that acts as normalization so that $0\leq R_i\leq 1$.
+Note that $N_i\approx 0$ if $\bfu$ is smooth in $S_i$. 
+In our implementation, we use
+\begin{align*}
+  N_i\approx \left|\bfv(\bfu_i)\cdot \int_{S_i}\nabla\cdot \bff(\bfu)d\bfx 
+  -\int_{S_i}\nabla\cdot\bfq(\bfu) d\bfx\right|,
+  \quad 
+  %D_i = ||\bfv(\bfu_i)||_{\ell^2}\left|\left|\int_{S_i}\nabla\cdot\bff(\bfu)d\bfx\right|\right|_{\ell^2}
+  %+\left|\int_{S_i}\nabla\cdot\bfq(\bfu)d\bfx\right|.
+  D_i = \sum_{k=1}^{d+1}|\bfv_k(\bfu_i)|\left|\int_{S_i}\left(\nabla\cdot\bff(\bfu)\right)_kd\bfx\right|
+  +\left|\int_{S_i}\nabla\cdot\bfq(\bfu)d\bfx\right|,
+\end{align*}
+where $({\bf z})_k$ denotes the $k$-th component of ${\bf z}\in\mathbb{R}^{d+1}$ and $d$ is the number of physical 
+dimensions. In the next section, we use $R_i$ to blend Rusanov's and Roe's Riemann solvers.
+
+\subsection{Blended Riemann solver}\label{sec:blended_rs}
+In this section, we use \eqref{Ri}, which is a normalized measurement of the entropy residual \eqref{ent_residual}, 
+to blend the Rusanov's and the Roe's Riemann solvers. To do this, we define the following numerical flux:
+\begin{align}\label{num_flux_ev}
+  F_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} 
+  - \frac{1}{2} \left( R_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-R_{i+1/2})|\hat A_{i+1/2}| \right)(\bfu_{i+1}-\bfu_{i}),
+\end{align}
+and similarly for $F_{i-1/2}$. Here $\mathbb{I}\in\mathbb{R}^{3\times 3}$ is the identity matrix,
+$R_{i+1/2}=\max\{R_{i+1},R_i\}$, $\lambda_{i+1/2}^{\max}$ is the 
+upper bound on the wave speed used in the Rusanov's Riemann solver, see \S \ref{sec:rusanov}, 
+and $\hat A_{i+1/2}$ is the Roe's averaged flux Jacobian, see \S \ref{sec:roe}.
+
+%In terms of fluctuations, 
+The first-order method is given simply by \eqref{first-order_FV} 
+with the numerical flux $F_{i+1/2}$ given by \eqref{num_flux_ev}. 
+In smooth regions $R_{i+1/2}\approx 0$ so the numerical flux is essentially the Roe's numerical flux. 
+Near discontinuities or steep gradients, $R_{i+1/2}$ becomes large. If, in particular, $R_{i+1/2}\approx 1$, 
+then the numerical flux is essentially the Rusanov's numerical flux.  
+%
+The first-order method can also be written in terms of fluctuations via \eqref{first-order_via_fluct}. 
+To do this we consider the waves ($\W^p_{i\pm 1/2}$, with $p=1,2,3$) 
+from the Roe's solver and define 
+\begin{align}\label{lambda_p}
+  \lambda_{i+1/2}^p := R_{i+1/2}\lambda_{i+1/2}^{\max} + (1-R_{i+1/2})|\hat \lambda_{i+1/2}^p|,
+\end{align}
+where $\hat\lambda_{i+1/2}^p$ is the $p$-th eigenvalue of the Roe's solver corresponding to the 
+interface $i+1/2$. Note that $\lambda_{i+1/2}^p\geq 0$.
+The fluctuations are then given by 
+\begin{subequations}\label{ev_fluctuations}
+\begin{align}
+  \A^+\Delta Q_{i-1/2}&=\frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^p, \\
+  \A^-\Delta Q_{i+1/2}&=\frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^p.
+\end{align}
+\end{subequations}
+
+To implement the second-order Lax-Wendroff-LeVeque type method in Clawpack, the Riemann solver must provide 
+the fluctuations, the waves and their speed. The fluctuations are simply \eqref{ev_fluctuations} 
+and the waves are the Roe's waves $\W_{i+1/2}^p$, with $p=1,2,3$. 
+In the framework of LeVeque, for a given interface $i+1/2$, 
+we would need to consider a wave speed $s_{i+1/2}^p$ so that 
+\begin{align*}
+  \hat \lambda_{i+1/2}^p+\lambda_{i+1/2}^p = s_{i+1/2}^p+|s_{i+1/2}^p|, \qquad 
+  \hat \lambda_{i+1/2}^p-\lambda_{i+1/2}^p = s_{i+1/2}^p-|s_{i+1/2}^p|,
+\end{align*}
+which is not possible to obtain in general (since the problem is overdetermined). 
+Instead, we consider 
+$s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p$ 
+ and study the consequences of this choice once the second-order corrections are applied.
+Here $\sgn(z)$ is the sign function of $z$. 
+The second-order Lax-Wendroff-LeVeque method for the one-dimensional problem is given by 
+\begin{align}\label{second-order}
+  u_i^{n+1}=u_i^n
+  -\frac{\Delta t}{\Delta x}\left[\A^-\Delta Q_{i+1/2}+\A^+\Delta Q_{i-1/2}\right]
+  -\frac{\Delta t}{\Delta x}\left[\tilde F_{i+1/2}-\tilde F_{i-1/2}\right],
+\end{align}
+where 
+\begin{align*}
+  \tilde F_{i+1/2} = \frac{1}{2}\sum_p\lambda_{i+1/2}^p
+  \left(1-\frac{\Delta t}{\Delta x}\lambda_{i+1/2}^p\right)\tilde\W_{i+1/2}^p,
+\end{align*}
+and similarly for $\tilde F_{i-1/2}$, are the flux corrections with $\tilde\W_{i+1/2}^p$ 
+being a TVD-limited version of $\W_{i+1/2}^p$.
+
+To understand the effect of choosing $s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p$ 
+as the speed of $\W_{i+1/2}^p$, 
+let us assume that no limiters are applied; i.e., $\tilde\W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$.
+By doing this, equation \eqref{second-order} can be written as 
+\begin{align*}
+  u_i^{n+1}=u_i^n
+  &-\frac{\Delta t}{2 \Delta x}
+  \sum_p
+  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}(\lambda_{i+1/2}^p)^2 \right)\W_{i+1/2}^p
+  +
+  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}(\lambda_{i-1/2}^p)^2 \right)\W_{i-1/2}^p\right],
+\end{align*}
+which corresponds to the standard Lax-Wendroff method written in terms of fluctuation-like quantities. 
+Note that if $R_{i\pm 1/2}=0$, we recover the standard second-order Lax-Wendroff method based on 
+Roe's Riemann solver. 
+In the numerical experiments in \S\ref{sec:num}, we obtain second-order experimental 
+order of convergence (EOC) for problems with smooth solution, which is expected since 
+$R_{i\pm 1/2}\approx 0$ in smooth regions.
+Similarly, if $R_{i\pm 1/2}=1$, we recover the Lax-Wendroff method based on Rusanov's Riemann solver, 
+which is clear by writting the fluctuations in Rusanov's method in terms of Roe's waves and speeds, 
+see \eqref{fluct_rus}.
+
+\subsection{Entropy stable Riemann solver}
+The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a
+measurement of the entropy residual. The objective is to induce dissipation of entropy 
+near discontinuities. However, we have no guarantee that methods \eqref{first-order_via_fluct} 
+or \eqref{second-order_via_fluct} based on this Riemann solver are (locally) entropy stable. 
+In this section, we consider the one-dimensional problem and increase the numerical dissipation 
+to guarantee entropy stability for \eqref{first-order_via_fluct}. 
+To this end we follow the work by Tadmor, see for instance \cite{tadmor1987numerical, tadmor2003entropy} 
+and references therein.
+
+Recall method \eqref{first-order_FV}
+\begin{align*}
+  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
+\end{align*}
+with the numerical fluxes given by 
+\begin{align*}
+  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} 
+  - \frac{1}{2} \left( R_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-R_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfu_{i+1}-\bfu_{i}),
+\end{align*}
+and similarly for $\bfF_{i-1/2}$. Here $\lambda_{i+1/2}^{\min}$ is yet to be defined.
+We can write the numerical fluxes in terms of $\lambda^p_{i\pm 1/2}$ \eqref{lambda_p}
+and the Roe's waves $\W_{i\pm 1/2}^p$.
+Doing so, we get
+\begin{align*}
+  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} 
+  - \frac{1}{2}\sum_p(\lambda_{i+1/2}^p+\lambda_{i+1/2}^{\min})\W_{i+1/2}^p,
+\end{align*}
+and similarly for $\bfF_{i-1/2}$.
+Let $\bfv_i=\bfv(\bfu_i)$ and $q_i=q(\bfu_i)$ denote the entropy variable and the 
+(one-dimensional) entropy potential at cell $i$, respectively. 
+Also let $\psi_i=\bfv(\bfu_i)\cdot \bff(\bfu_i)-q(\bfu_i)$ be the entropy potential at cell $i$.
+From \cite[\S 4]{tadmor1987numerical}, if the numerical fluxes satisfy
+\begin{align}\label{es_cond}
+(\bfv_{i}-\bfv_{i-1})\cdot \bfF_{i-1/2}\leq \psi_{i}-\psi_{i-1},
+  \qquad
+  (\bfv_{i+1}-\bfv_i)\cdot \bfF_{i+1/2}\leq \psi_{i+1}-\psi_i,
+\end{align}
+then the following entropy inequality holds 
+\begin{align}\label{ent_ineq}
+  \frac{d\eta(\bfu_i)}{dt}+\frac{1}{\Delta x}\left[G_{i+1/2}-G_{i-1/2}\right]\leq 0,
+\end{align}
+where 
+\begin{align*}
+    G_{i+1/2} &= \frac{1}{2}\left(\bfv_i+\bfv_{i+1}\right)\bfF_{i+1/2}-\frac{1}{2}(\psi_{i}+\psi_{i+1}),
+\end{align*}
+and similarly for $G_{i-1/2}$, are discretizations of the entropy flux.
+
+Based on \eqref{es_cond}, we choose
+\begin{align}\label{lambda_min}
+  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\bfv_{i+1}-\bfv_{i})\cdot\left[\bff(\bfu_{i+1})+\bff(\bfu_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\psi_{i+1}-\psi_i)}{\frac{1}{2}(\bfv_{i+1}-\bfv_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}
+\end{align}
+which guarantees \eqref{es_cond} and, therefore, \eqref{ent_ineq}. 
+Since the blended Riemann solver induces dissipation of entropy, 
+we expect condition \eqref{es_cond} to be fulfilled most of the times even with $\lambda_{i+1/2}^{\min}=0$. 
+Therefore, \eqref{lambda_min} is used as a safeguard to guarantee entropy stability
+for the one-dimensional problem using method \eqref{first-order_FV} or \eqref{first-order_via_fluct}. 
+Extra modifications are needed to guarantee entropy stability of Lax-Wendroff-LeVeque method 
+\eqref{second-order_via_fluct} and its multidimensional extension via Strang's splitting. 
+We do not pursue these modifications in this work.
+
+\begin{remark}
+In terms of fluctuations, the entropy stability condition \eqref{es_cond} can be written as 
+\begin{align*}
+  (\bfv_{i}-\bfv_{i-1})\cdot 
+  \Big(
+  \underbrace{\bff(\bfu_i)-\A^+\Delta Q_{i-1/2}}_{=\bfF_{i-1/2}}\Big)\leq \psi_{i}-\psi_{i-1},
+  \quad
+  (\bfv_{i+1}-\bfv_i)\cdot 
+  \Big(\underbrace{\bff(\bfu_i)+\A^-\Delta Q_{i+1/2}}_{=\bfF_{i+1/2}}\Big)\leq \psi_{i+1}-\psi_i.
+\end{align*}
+\end{remark}
+
+\clearpage
+\section{Numerical results}\label{sec:num}
+Description of the methods. We consider either the first-order method XX or the second-order Lax-Wendroff
+type method XX. For each of these methods we use one of the following three Riemann solvers: 
+\begin{itemize}
+  \item LLF: Rusanov's or Local Lax Friederichs 2-wave Riemann solver reviewed in \S XX.
+  \item Roe: standard Roe's average based 3-wave Riemann solver reviewed in \S XX. 
+  \item Roe-fix: Roe's solver with entropy fix based on XX. 
+  \item EV:  entropy dissipative 3-wave Riemann solver described in \S XX.
+\end{itemize}
+
+\subsection{Dam break problem on a dry bed}
+We start with the one-dimensional dam break problem on a dry bed. 
+This problem is a canonical example that demonstrates the `entropy glitch' of Roe's solver 
+at transonic rarefactions. Since the baseline Roe's solver we use (see \S XX) does not contain 
+an entropy fix, it is important to demonstrate that the blended Riemann solver fixes the entropy glitch, 
+at least for this benchmark. 
+We follow the setup in XX. The domain is given by XX, and the initial condition is XX and XX. 
+The exact solution, which can be found in XX and references therein, is
 
@@
 
-and propose different Riemann solvers. Once the Riemann solvers are defined for the low-order method, 
-the second order corrections are incorporated following the same steps for all cases. Therefore, we focus our 
-attention to design of the Riemann solvers.
+We use g=1 and solve the problem up to T=10. In Figure XX we show the solution using method XX 
+with Roe's solver, Rusanov's solver and the blended solver. The entropy glitch, which is manifested 
+as a non-physical show at x=XX, is clearly shown with Roe's solver. Using Rusanov's solver and the 
+blended solver fix the entropy glitch. Using the blended solver leads to more accurate results. 
+In Table XX, we summarize the results of a convergence test.
 
-\subsection{Roe's Riemann solver}
+\subsection{Dam break problem on a wet bed}
+We consider a one-dimensional dam break problem on a wet domain. 
+We follow the setup in XX. 
+The domain is $\Omega=(0,L)$ with $L=10$, the initial condition is given by 
+$hu(x,0)=0$ and 
+\begin{align*}
+  h(x,0) = 
+  \begin{cases}
+    h_l & \mbox{ if } x\in(0,x_0], \\
+    h_r & \mbox{ if } x\in(x_0,L),
+  \end{cases}
+\end{align*}
+with $x_0=5$, $h_l=0.005$ and $h_r=0.001$. 
+The exact solution, which can be found in XX and references therein, is given by 
+\begin{align*}
+  h(x,t) = 
+  \begin{cases}
+    h_l, \\
+    \frac{4}{9g}\left(\sqrt{gh_l}-\frac{x-x_0}{2t}\right)^2, \\
+    \frac{c_m^2}{g}, \\
+    h_r, 
+  \end{cases}
+\quad 
+  u(x,t) = 
+  \begin{cases}
+    0, &\mbox{ if } x\leq x_A(t), \\
+    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
+    2(\sqrt{gh_l}-c_m), & \mbox{ if } x_B(t)<x\leq x_C(t), \\
+    0, &\mbox{ if } x\leq x_C(t),
+  \end{cases}
+\end{align*}
+where $x_A(t)=x_0-t\sqrt{gh_l}$, $x_B(t)=x_0+t\left(2\sqrt{gh_l}-3c_m\right)$, 
+$x_C(t)=x_0+t\frac{2c_m^2\left(\sqrt{gh_l}-c_m\right)}{c_m^2-gh_r}$ and 
+$c_m$ is the solution of 
+$-8gh_rc_m^2\left(\sqrt{gh_l}-c_m\right)^2+\left(c_m^2-gh_r\right)^2\left(c_m^2+gh_r\right)=0$.
+We use $g=1$ and solve the problem up to the final time $T=5$ considering the 2nd-order Lax-Wendroff 
+type method XX with the LLF, Roe and EV Riemann solvers. The solution with different refinement levels 
+and each Riemann solver is shown in Figure XX. In Table XX, we summarize the results of a convergence test. 
+Since the solution is non-smooth, we expect no more than first order convergence rates. Note that the 
+results with the entropy dissipative solver EV are comparable against the results using Roe's solver. 
+That is, inducing entropy dissipation via the EV Riemann solver does not degrade the high-order accuracy 
+properties of the Roe's solver. 
+In contrast, the accuracy and convergence rates using the LLF solver are clearly degraded.
 
-\subsection{Rusanov's Riemann solver}
+\subsection{The circular hydraulic jump in one dimension}
 
-\subsection{Blended Rusanov's and Roe's solver based on entropy residual}
+\subsection{Steady outflow}
+Let us consider a problem similar to that in the previous section, but with an outflow boundary condition. 
+That is, we consider the one-dimensional domain $\Omega=(0.1,1)$, the initial condition is 
+$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$, 
+and the right boundary condition is set to outflow, see the details in XX. 
+If the flow is supercritical ($|F|>1$), the exact solution converges to a 
+steady state profile given by the solution of the ODE XX. 
+Depending on the initial condition, shocks might develop before the steady state is reached. 
+Nevertheless, the steady state profile is smooth. 
+In Figure XX, we show the solution at different times based on the secnd-order method XX with 
+the EV Riemann solver. Additionally, in Table XX we summarize the result of a convergence study 
+based on the first- and second-order methods XX and XX, respectively. 
+We consider the Riemann solver in \S XX with Ri=0, Ri=1 and Ri computed via XX.
 
-\subsection{Blended Rusanov's and Roe's solver based on Tadmor's entropy stability condition}
-
-\subsection{Entropy stable enhancement of Roe's Riemann solver}
-
-\subsection{Other alternatives}
-
-\section{Numerical results}
-
+\subsection{The circular hydraulic jump in two dimensions}
 We consider two regimes. In the first regime, the physical instabilities are mild. 
@@
 
-\bibliographystyle{apalike}
+%\bibliographystyle{apalike}
+\bibliographystyle{plain}
 \bibliography{refs}
@@
 \end{document}
+
+%Random perturbation only in the jet
+%Turn off the perturbation after some time
+%Reduce the amplitude of the perturbation 
+%Non-rotationally sym. IC 
+%Consider Kemm's solver.

### 2020-11-16 numerics_paper/version2/numerical_chj.tex
update paper

@@
 \item We present a new problem where the cc instability occurs. In our opinion, this is a more challenging 
-  problem than the standard bow problem since we expect the solution to be non-steady. 
+  problem than the standard bow shock problem since we expect the solution to be non-steady. 
   We aim to preserve instabilities that are inherently present in the CHJ for certain regimes. And, 
   at the same time, we aim to prevent the cc instability. 
-\item We explore the use of entropy dissipation and stability as cure of the cc instability for the SWEs. 
-\item We propose specific methods to blend the highly dissipative and robust Rusanov's RP and the 
-  accurate 3-wave Roe's solver for the SWEs. In particular we propose methods that induce dissipation of entropy
-  near the shocks and other methods that guarantee entropy stability. 
-\item We demonstrate that these schemes prevent the formation of cc instabilities not only 
+\item We explore the use of entropy dissipation and entropy stability as cure of the cc instability for the SWEs. 
+\item We propose a specific method to blend the highly dissipative and robust Rusanov's RP and the 
+  accurate 3-wave Roe's solver for the SWEs. 
+\item We demonstrate that this scheme prevents the formation of cc instabilities not only 
   in the well-known bow shock problem, but also in the CHJ while preserving other important features of 
@@
 
-\subsection{Semi-analytical steady solution under rotational symmetry}
+\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
 
@@
 Steady-state solutions of \eqref{eq:rsw} satisfy
-\begin{subequations}
+\begin{subequations}\label{steady}
 \begin{align}
@@
 
-\subsection{Formation of the CHJ}
+\subsection{Formation of the CHJ}\label{sec:formation}
 
@@
 To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers. 
-Our blending criteria is based on the residual of the entropy (in)equality. 
-Similar ideas have been explored before to solve the Euler equations XX 
-and the shallow water equations XX. In these references, the authors use different 
-blending criteria.
+Our blending criteria is based on the residual of the entropy (in)equality,
+which induces dissipation of entropy near discontinuities. 
+Blending these Riemann solvers has been proposed before
+to solve the Euler equations, see e.g.
+\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
+the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
+and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
+In these references, the authors use blending criteria different than ours.
+Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
+For example, in \cite{ismail2009affordable,ismail2009proposed},
+the authors developed carbuncle free solvers for the Euler equations
+via dissipation of entropy.
 
-\subsection{Lax-Wendroff-LeVeque method}
+\subsection{Lax-Wendroff-LeVeque method}\label{sec:lax-wendroff-leveque}
 Let us now review the second-order correction to method \eqref{first-order_FV} implemented in 
@@
 
-\subsection{Entropy stable Riemann solver}
+\subsection{Entropy stable Riemann solver}\label{sec:entropy_stable}
 The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a
@@
 
+\begin{remark}[About the effect of $\lambda_{i\pm 1/2}^{\min}$ for Roe's solver.]
+  If we set $R_{i\pm 1/2}=0, ~\forall i$ in \eqref{lambda_p}, then the Riemann solver becomes the Roe's solver.
+  In this case, it is well known that transonic rarefactions develop non-physical shocks,
+  a phenomenon known as `entropy-glitch'. To fix this problem, many entropy fixes to the Roe's
+  solver have been developed. Enhancing the dissipation via \eqref{lambda_min} behaves as an entropy fix for
+  the Roe's solver, which we demonstrate via a numerical experiment in \S \ref{sec:rp_dry_bed}. 
+\end{remark}
+
 \clearpage
 \section{Numerical results}\label{sec:num}
-Description of the methods. We consider either the first-order method XX or the second-order Lax-Wendroff
-type method XX. For each of these methods we use one of the following three Riemann solvers: 
-\begin{itemize}
-  \item LLF: Rusanov's or Local Lax Friederichs 2-wave Riemann solver reviewed in \S XX.
-  \item Roe: standard Roe's average based 3-wave Riemann solver reviewed in \S XX. 
-  \item Roe-fix: Roe's solver with entropy fix based on XX. 
-  \item EV:  entropy dissipative 3-wave Riemann solver described in \S XX.
-\end{itemize}
+In this section we present one- and two-dimensional numerical experiments to demonstrate the behavior of the
+blended Riemann solver from \S\ref{sec:blended_rs} with the extra entropy dissipation from \S\ref{sec:entropy_stable}. 
+We compare the behavior of the blended solver against the standard Roe's and Rusanov's solvers.
+In most of the experiments we use these Riemann solvers with the Lax-Wendroff-LeVeque method reviewed in \S\ref{sec:lax-wendroff-leveque}.
+When the exact solution is available, we report a convergence test based on the $L^1$-error for the water height
+\begin{align*}
+  E_1=\sum_i|K_i|~|h_i-h^e(\bfx_i)|,
+\end{align*}
+where $|K_i|$ is the length or area of cell $i$,
+$h_i$ is the cell average of the water height at cell $i$,
+and $h^e(\bfx_i)$ is the exact water height evaluated at the center of cell $i$.
 
-\subsection{Dam break problem on a dry bed}
+We consider a total of six problems. We start with a one-dimensional Riemann problem over a dry bed. 
+For this problem we use the first-order method \eqref{first-order_via_fluct},
+which helps preventing violations of positivity. Although, only the Rusanov's solver guarantees
+positivity, we do not get negative values for the water height with this method and the other Riemann solvers.
+In \S \ref{sec:steady_outflow} we consider a one-dimensional problem with a smooth steady state solution. We observe the expected
+second-order accuracy of method \eqref{second-order_via_fluct} with Roe's and the blended solvers. Using
+the Rusanov's Riemann solver degrades the accuracy to first-order. The overall accuracy of the blended solver
+is not degraded since the extra dissipation is not applied in smooth regions. 
+In \S\ref{sec:rp_wet_bed} and \S\ref{sec:1D_chj} we consider one-dimensional problems with strong shocks. We observe similar accuracy with the
+blended solver than with Roe's solver; that is, inducing entropy dissipation near the shocks do not degrade the accuracy
+of the underlying Roe's solver. This extra dissipation, however, prevents the formation of carbuncles in other experiments,
+which we demonstrate in \S \ref{sec:bow_shock} and \S\ref{sec:2D_chj}. 
+The main focus of this work is in the formation of carbuncle instabilities in the two-dimensional CHJ. We present
+an extensive set of experiments about this problem in \S \ref{sec:2D_chj}.
+We consider not only Roe's, Rusanov's and the blended solvers, but also the HLLE solver and a solver proposed in \cite{kemm2014note},
+which is designed to remove the carbuncle instabilities in the shallow water equations.
+
+\subsection{Dam break problem on a dry bed}\label{sec:rp_dry_bed}
 We start with the one-dimensional dam break problem on a dry bed. 
 This problem is a canonical example that demonstrates the `entropy glitch' of Roe's solver 
-at transonic rarefactions. Since the baseline Roe's solver we use (see \S XX) does not contain 
-an entropy fix, it is important to demonstrate that the blended Riemann solver fixes the entropy glitch, 
-at least for this benchmark. 
-We follow the setup in XX. The domain is given by XX, and the initial condition is XX and XX. 
-The exact solution, which can be found in XX and references therein, is
+at transonic rarefactions. Since the baseline Roe's solver we use (see \S\ref{sec:roe}) does not contain 
+an entropy fix, it is important to demonstrate that the blended Riemann solver fixes the entropy glitch.
+We follow the setup in \cite[\S 4.1.2]{delestre2013swashes}.
+The domain is given by $\Omega=(0,L=10)$, and the initial condition is $hu(x,0)=0$ and
+\begin{align*}
+  h(x,0)=
+  \begin{cases}
+    h_l, & \mbox{ if }, x\leq x_0=5 \\
+    h_r, &\mbox{ otherwise},
+  \end{cases}
+\end{align*}
+with $h_l=5\times 10^{-3}$ and $h_r=0$. In our experiments we use $h_r=1\times 10^{-15}$ to avoid division by zero.
+The exact solution, which can also be found in \cite{delestre2013swashes} and references therein, is 
+\begin{align*}
+  h(x,t) = 
+  \begin{cases}
+    h_l, \\
+    \frac{4}{9g}\left(\sqrt{gh_l}-\frac{x-x_0}{2t}\right)^2, \\
+    0,
+  \end{cases}
+\quad 
+  u(x,t) = 
+  \begin{cases}
+    0, &\mbox{ if } x\leq x_A(t), \\
+    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
+    0, &\mbox{ if } x_B(t)\leq x,
+  \end{cases}
+\end{align*}
+with $x_A(t)=x_0-t\sqrt{gh_l}$ and $x_B(t)=x_0+2t\sqrt{gh_l}$.
+We use $g=1$ and solve the problem up to $t=10$.
+In figures \ref{fig:rp_dry_bed_roe}-\ref{fig:rp_dry_bed_blended},
+we show the solution using method \eqref{first-order_via_fluct} 
+with Roe's solver, Rusanov's solver and the blended solver, respectively.
+The entropy glitch, which is manifested as a non-physical shock at $x=x_0$,
+is present with Roe's solver. Using Rusanov's and the blended solvers
+fixes the entropy glitch.
+In Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test. 
+Note that using the blended solver leads to more accurate results.
 
-XX
+\begin{figure}[!h]
+  {\scriptsize
+    \subfloat[Roe's Riemann solver.\label{fig:rp_dry_bed_roe}]{
+      \includegraphics[scale=0.36]{figures/rp_dry_roe.png}}
+    ~
+    \subfloat[Rusanov's Riemann solver.\label{fig:rp_dry_bed_llf}]{
+        \includegraphics[scale=0.36]{figures/rp_dry_llf.png}}
 
-We use g=1 and solve the problem up to T=10. In Figure XX we show the solution using method XX 
-with Roe's solver, Rusanov's solver and the blended solver. The entropy glitch, which is manifested 
-as a non-physical show at x=XX, is clearly shown with Roe's solver. Using Rusanov's solver and the 
-blended solver fix the entropy glitch. Using the blended solver leads to more accurate results. 
-In Table XX, we summarize the results of a convergence test.
+    \subfloat[Blended Riemann solver. \label{fig:rp_dry_bed_blended}]{
+      \includegraphics[scale=0.36]{figures/rp_dry_blended.png}}
+    ~
+    \subfloat[Roe's Riemann solver with $\lambda_{i\pm 1/2}^{\min}$.\label{fig:rp_dry_bed_roe_with_lmin}]{
+      \includegraphics[scale=0.36]{figures/rp_dry_roe_with_lmin.png}}    
+  }
+  \caption{
+    Dam break problem over a dry bed using method \eqref{first-order_via_fluct} with different
+    Riemann solvers. We show the numerical solution and the exact solution (in dashed black) at $t=10$.
+    We consider different refinements with
+    (cyan) $\Delta x=1/400$, (red) $\Delta x=1/800$, (blue) $\Delta x=1/1600$.
+    \label{fig:rp_dry_bed}}
+\end{figure}
 
-\subsection{Dam break problem on a wet bed}
-We consider a one-dimensional dam break problem on a wet domain. 
-We follow the setup in XX. 
+For this particular problem, using method \eqref{first-order_via_fluct} with the blended Riemann solver
+leads to $\lambda_{i\pm 1/2}^{\min}=0$ for all the experiments, but this is not true in general.
+Finally, let us demonstrate the effect of $\lambda_{i\pm 1/2}^{\min}$ by imposing $R_{i}=0$ in \eqref{lambda_p},
+which is equivalent to using Roe's solver with extra dissipation given by $\lambda_{i\pm 1/2}^{\min}$.
+In Figure \ref{fig:rp_dry_bed_roe_with_lmin}, we show the solution using different refinements,
+and in the last column of Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test.
+Note that the entropy glitch is highly reduced. 
+To remove completely the entropy glitch we could add high-order entropy dissipation following \cite{tadmor2003entropy}
+and references therein.
+
+\begin{table}[!ht]\scriptsize
+  \begin{center}
+    \begin{tabular}{||c||c|c||c|c||c|c||c|c||} \hline
+      & \multicolumn{2}{c||}{Roe's solver}  
+      &\multicolumn{2}{c||}{Rusanov's solver} 
+      &\multicolumn{2}{c||}{Blended solver}
+      &\multicolumn{2}{c||}{Roe's with $\lambda_{i\pm 1/2}^{\min}$} \\ \cline{2-9}
+      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
+      1/50   & 6.21E-04 &  --  & 6.95E-04 &   -- & 5.10E-04 &  --  & 5.95E-04 & --   \\
+      1/100  & 4.70E-04 & 0.40 & 5.27E-04 & 0.40 & 3.53E-04 & 0.52 & 4.06E-04 & 0.55 \\
+      1/200  & 3.91E-04 & 0.26 & 3.88E-04 & 0.44 & 2.38E-04 & 0.56 & 2.83E-04 & 0.52 \\
+      1/400  & 3.16E-04 & 0.30 & 2.64E-04 & 0.55 & 1.51E-04 & 0.65 & 1.89E-04 & 0.58 \\
+      1/800  & 2.35E-04 & 0.42 & 1.67E-04 & 0.66 & 9.43E-05 & 0.68 & 1.20E-04 & 0.65 \\
+      1/1600 & 2.00E-04 & 0.23 & 1.01E-04 & 0.72 & 5.70E-05 & 0.72 & 7.42E-05 & 0.68 \\ \hline
+    \end{tabular}
+    \caption{Grid convergence study for the dam break problem over a dry bed
+      using method \eqref{first-order_via_fluct} with different Riemann solvers.\label{table:rp_dry_bed}}
+  \end{center}
+\end{table}
+
+\subsection{Dam break problem on a wet bed}\label{sec:rp_wet_bed}
+We consider now a one-dimensional dam break problem on a wet domain. 
+We follow the setup in \cite[\S 4.1.1]{delestre2013swashes}. 
 The domain is $\Omega=(0,L)$ with $L=10$, the initial condition is given by 
@@
 with $x_0=5$, $h_l=0.005$ and $h_r=0.001$. 
-The exact solution, which can be found in XX and references therein, is given by 
+The exact solution, which can be found in \cite{delestre2013swashes} and references therein,
+is given by 
 \begin{align*}
@@
 $-8gh_rc_m^2\left(\sqrt{gh_l}-c_m\right)^2+\left(c_m^2-gh_r\right)^2\left(c_m^2+gh_r\right)=0$.
-We use $g=1$ and solve the problem up to the final time $T=5$ considering the 2nd-order Lax-Wendroff 
-type method XX with the LLF, Roe and EV Riemann solvers. The solution with different refinement levels 
-and each Riemann solver is shown in Figure XX. In Table XX, we summarize the results of a convergence test. 
+We use $g=1$ and solve the problem up to the final time $t=5$ considering method \eqref{second-order_via_fluct}
+with Roe's, Rusanov's and the blended solvers. The solution with different refinement levels 
+and each Riemann solver is shown in Figure \ref{fig:rp_wet_bed}.
+In Table \ref{table:rp_wet_bed}, we summarize the results of a convergence test. 
 Since the solution is non-smooth, we expect no more than first order convergence rates. Note that the 
-results with the entropy dissipative solver EV are comparable against the results using Roe's solver. 
-That is, inducing entropy dissipation via the EV Riemann solver does not degrade the high-order accuracy 
-properties of the Roe's solver. 
-In contrast, the accuracy and convergence rates using the LLF solver are clearly degraded.
+results with the entropy dissipative blended solver are comparable against the results using Roe's solver. 
+That is, inducing entropy dissipation via the blended Riemann solver does not degrade the high-order
+accuracy properties of the Roe's solver. 
+In contrast, the accuracy and convergence rates using Rusanov's solver are clearly degraded.
 
-\subsection{The circular hydraulic jump in one dimension}
+\begin{figure}[!h]
+  {\scriptsize
+    \subfloat[Roe's Riemann solver.]{
+      \includegraphics[scale=0.29]{figures/rp_wet_roe.png}}
+    \quad
+    \subfloat[Rusanov's Riemann solver.]{
+        \includegraphics[scale=0.29]{figures/rp_wet_llf.png}}
+    \quad
+    \subfloat[Blended Riemann solver.]{
+      \includegraphics[scale=0.29]{figures/rp_wet_blended.png}}
+  }
+  \caption{
+    Dam break problem over a wet bed using method \eqref{second-order_via_fluct} with different
+    Riemann solvers. We show the numerical solution and the exact solution (in dashed black) at $t=10$.
+    We consider different refinements with
+    (cyan) $\Delta x=1/400$, (red) $\Delta x=1/800$, (blue) $\Delta x=1/1600$.
+    \label{fig:rp_wet_bed}}
+\end{figure}
 
-\subsection{Steady outflow}
+\begin{table}[!ht]\scriptsize
+  \begin{center}
+    \begin{tabular}{||c||c|c||c|c||c|c||} \hline
+      & \multicolumn{2}{c||}{Roe's solver}  
+      &\multicolumn{2}{c||}{Rusanov's solver} 
+      &\multicolumn{2}{c||}{Blended solver} \\ \cline{2-7}
+      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
+      1/50   & 4.22E-04 &  --  & 5.73E-04 &  --  & 4.34E-04 & --   \\
+      1/100  & 2.00E-04 & 1.07 & 3.35E-04 & 0.78 & 2.08E-04 & 1.05 \\
+      1/200  & 1.11E-04 & 0.85 & 2.01E-04 & 0.74 & 1.13E-04 & 0.88 \\
+      1/400  & 5.05E-05 & 1.13 & 1.18E-04 & 0.77 & 5.26E-05 & 1.10 \\
+      1/800  & 2.60E-05 & 0.95 & 7.08E-05 & 0.74 & 2.73E-05 & 0.94 \\
+      1/1600 & 1.29E-05 & 1.01 & 3.79E-05 & 0.90 & 1.32E-05 & 1.04 \\ \hline
+    \end{tabular}
+    \caption{Grid convergence study for the dam break problem over a wet bed
+      using method \eqref{second-order_via_fluct} with different Riemann solvers.\label{table:rp_wet_bed}}
+  \end{center}
+\end{table}
+
+\subsection{The circular hydraulic jump in one-dimension}\label{sec:1D_chj}
+In this section we consider for the first time a CHJ. We start with a one-dimensional problem
+embedded into the two-dimensional domain $\Omega=(0.1,1)\times(0,0.1)$.
+A semi-analytical solution at steady state is available as described in \S \ref{sec:steady_chj};
+therefore, we can perform a convergence test. 
+We consider the initial condition $h(x,t=0)=0.1$, $hu(x,t=0)=0$ and $hv(x,t=0)=0$.
+At the left boundary, we impose $h(0.1,t)=0.3$, $hu(0.1,t)=0.225$ and $hv(0.1,t)=0$.
+At the right boundary we impose $h(1,t)=0.37387387318873766$, $hu(1,t)=\beta/h(1,t)$ and $hv(1,t)=0$ where $\beta=0.0225$.
+The values on the right boundary are chosen so that the CHJ is located at $x=0.3$, as described in \S \ref{sec:formation}.
+%
+In the first row of Figure \ref{fig:rp_wet_bed}, we show (slices of) the solution (at $y=0.05$)
+at different times using the second-order method \eqref{second-order_via_fluct} with Roe's, Rusanov's and the blended Riemann solvers. 
+In the second row of the same figure, we plot $R_i$ \eqref{Ri}, which is a (normalized) measurement of the
+entropy residual. Note that $R_i$ accurately locates the shocks and is (nearly) zero in smooth regions. 
+In Table \ref{table:chj_1D}, we summarize the results of a convergence study. 
+We remark that the convergence with Roe's solver is erratic.
+Moreover, for the finer grids, the solution appears to become worse.
+Using Rusanov's and the blended Riemann solvers delivers cleaner convergence rates. 
+In \cite[\S 5]{kemm2018heuristical} and \cite{bader2014carbuncle}, the authors consider one-dimensional problems embedded into
+two-dimensional domains and observe different instabilities for the
+Euler and the shallow water equations, respectively.
+The authors attributed the problem to a manifestation of the carbuncle instability
+that lead to an unstable location of the steady shock with certain Riemann solvers, including Roe's solver.
+
+\begin{figure}[!h]
+  {\scriptsize
+    \includegraphics[scale=0.29]{figures/chj_1D_t0p5.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/chj_1D_t1p3.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/chj_1D_t10p0.png}
+
+    \vspace{15pt}
+    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t0p5.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t1p3.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t10p0.png}
+  }
+  \caption{
+    One-dimensional CHJ using method \eqref{second-order_via_fluct} with 
+    (cyan) Roe's, (red) Rusanov's and (blue) the blended Riemann solvers.
+    In the second row we plot $R_i$, which is given by \eqref{Ri}.
+    In all simulations we consider $\Delta x=1/400$.
+    \label{fig:1D_chj}}
+\end{figure}
+
+\begin{table}[!ht]\scriptsize
+  \begin{center}
+    \begin{tabular}{||c||c|c||c|c||c|c||} \hline
+      & \multicolumn{2}{c||}{Roe's solver}  
+      &\multicolumn{2}{c||}{Rusanov's solver} 
+      &\multicolumn{2}{c||}{Blended solver} \\ \cline{2-7}
+      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
+      1/50  & 1.56E-04 &  --   & 8.80E-04 &  --  & 3.56E-04 &  --  \\
+      1/100 & 1.06E-04 & 0.55  & 5.09E-04 & 0.79 & 1.95E-04 & 0.86 \\
+      1/200 & 7.77E-05 & 0.45  & 2.96E-04 & 0.78 & 1.07E-04 & 0.87 \\
+      1/400 & 7.50E-06 & 3.37  & 1.49E-04 & 0.99 & 3.68E-05 & 1.53 \\
+      1/800 & 9.07E-06 & -0.27 & 7.98E-05 & 0.90 & 1.67E-05 & 1.14 \\ \hline
+    \end{tabular}
+    \caption{Grid convergence study for the one-dimensional CHJ at $t=10$
+      using method \eqref{second-order_via_fluct} with different Riemann solvers.\label{table:chj_1D}}
+  \end{center}
+\end{table}
+
+\subsection{Steady outflow}\label{sec:steady_outflow}
 Let us consider a problem similar to that in the previous section, but with an outflow boundary condition. 
-That is, we consider the one-dimensional domain $\Omega=(0.1,1)$, the initial condition is 
-$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$, 
-and the right boundary condition is set to outflow, see the details in XX. 
-If the flow is supercritical ($|F|>1$), the exact solution converges to a 
-steady state profile given by the solution of the ODE XX. 
+That is, we consider the one-dimensional domain $\Omega=(0.1,1)$, the initial condition 
+$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$, 
+and set the right boundary condition to outflow.
+We refer to \cite[\S 21.8.5]{leveque2002finite} for details about the outflow boundary condition. 
+If the flow is supercritical ($|F|>1$ where $F$ is the Froude number), the exact solution converges to a 
+steady state profile given by the solution of \eqref{steady}. 
 Depending on the initial condition, shocks might develop before the steady state is reached. 
 Nevertheless, the steady state profile is smooth. 
-In Figure XX, we show the solution at different times based on the secnd-order method XX with 
-the EV Riemann solver. Additionally, in Table XX we summarize the result of a convergence study 
-based on the first- and second-order methods XX and XX, respectively. 
-We consider the Riemann solver in \S XX with Ri=0, Ri=1 and Ri computed via XX.
+In Figure \ref{fig:steady_outflow}, we show the solution and $R_i$ at different times using
+the second-order method \eqref{second-order_via_fluct}
+with Roe's, Rusanov's and the blended Riemann solvers.
+Additionally, in Table \ref{table:steady_outflow}, we summarize the results of a convergence study 
+based on methods \eqref{first-order_via_fluct} and
+\eqref{second-order_via_fluct}, using the same Riemann solvers.
 
-\subsection{The circular hydraulic jump in two dimensions}
-We consider two regimes. In the first regime, the physical instabilities are mild. 
-By increasing the strength of the jump and other parameters, we consider a second 
-and more strongly unstable regime. In both cases, we generate the CHJ as described in 
-Section XX; that is, we impose supersonic inflow boundary conditions for the jet and 
-subsonic outflow boundary conditions.
+\begin{figure}[!h]
+  {\scriptsize
+    \includegraphics[scale=0.29]{figures/outflow_1D_t0p5.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/outflow_1D_t1p3.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/outflow_1D_t10p0.png}
 
-We consider two types of domain and two corresponding computational grids. 
-The two types of domains are two-dimensional squares and circular domains. 
-For both cases we use structured grids. See an example of these domains and grids in Figures XX and XX, 
-respectively.
+    \vspace{15pt}
+    \includegraphics[scale=0.29]{figures/outflow_1D_Ri_t0p5.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/outflow_1D_Ri_t1p3.png}
+    \qquad
+    \includegraphics[scale=0.29]{figures/outflow_1D_Ri_t10p0.png}
+  }
+  \caption{
+    Steady outflow problem using method \eqref{second-order_via_fluct} with 
+    (cyan) Roe's, (red) Rusanov's and (dashed blue) the blended Riemann solvers.
+    In the second row we plot $R_i$, which is given by \eqref{Ri}.
+    In all simulations we consider $\Delta x=1/400$.
+    \label{fig:steady_outflow}}
+\end{figure}
 
-\subsection{Regime 1: midly unstable flow}
+\begin{table}[!ht]\scriptsize
+  \begin{center}
+    \begin{tabular}{||c||c|c||c|c||c|c||c|c||c|c||c|c||} \hline
+      & \multicolumn{6}{c||}{First-order method \eqref{first-order_via_fluct}}
+      & \multicolumn{6}{c||}{Second-order method \eqref{second-order_via_fluct}} \\ \cline{2-13}
+      & \multicolumn{2}{c||}{Roe's solver}  
+      &\multicolumn{2}{c||}{Rusanov's solver} 
+      &\multicolumn{2}{c||}{Blended solver}
+      & \multicolumn{2}{c||}{Roe's solver}  
+      &\multicolumn{2}{c||}{Rusanov's solver} 
+      &\multicolumn{2}{c||}{Blended solver}
+      \\ \cline{2-13}
+      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate\\ \hline
+      1/50   & 2.16E-4 &  --  & 9.97E-4 &   -- & 2.20E-4 & --   & 4.13E-5 &   -- & 7.32E-4 &  --  & 4.97E-5 & --   \\
+      1/100  & 1.14E-4 & 0.92 & 6.38E-4 & 0.64 & 1.16E-4 & 0.92 & 1.47E-5 & 1.48 & 4.43E-4 & 0.73 & 1.89E-5 & 1.39 \\
+      1/200  & 5.86E-5 & 0.95 & 3.88E-4 & 0.72 & 5.93E-5 & 0.96 & 4.81E-6 & 1.61 & 2.57E-4 & 0.79 & 6.36E-6 & 1.57 \\
+      1/400  & 2.98E-5 & 0.97 & 2.26E-4 & 0.78 & 3.00E-5 & 0.98 & 1.42E-6 & 1.76 & 1.44E-4 & 0.83 & 1.92E-6 & 1.72 \\
+      1/800  & 1.50E-5 & 0.98 & 1.27E-4 & 0.83 & 1.51E-5 & 0.99 & 4.16E-7 & 1.76 & 7.81E-5 & 0.88 & 5.61E-7 & 1.77 \\
+      1/1600 & 7.53E-6 & 0.99 & 6.90E-5 & 0.88 & 7.55E-6 & 0.99 & 1.14E-7 & 1.86 & 4.12E-5 & 0.92 & 1.54E-7 & 1.86 \\ \hline
+    \end{tabular}
+    \caption{Grid convergence study for the steady outflow problem
+      using methods \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}
+      with different Riemann solvers.\label{table:steady_outflow}}
+  \end{center}
+\end{table}
 
-Show the results with Roe's and Rusanov's solver. 
-Show the results with the three proposed methodologies.
+%\clearpage
+\subsection{Steady shock around a cylinder}\label{sec:bow_shock}
+\subsection{The circular hydraulic jump in two dimensions}\label{sec:2D_chj}
+%We consider two regimes. In the first regime, the physical instabilities are mild. 
+%By increasing the strength of the jump and other parameters, we consider a second 
+%and more strongly unstable regime. In both cases, we generate the CHJ as described in 
+%Section XX; that is, we impose supersonic inflow boundary conditions for the jet and 
+%subsonic outflow boundary conditions.
 
-\subsection{Regime 2: strongly unstable flow}
+%We consider two types of domain and two corresponding computational grids. 
+%The two types of domains are two-dimensional squares and circular domains. 
+%For both cases we use structured grids. See an example of these domains and grids in Figures XX and XX, 
+%respectively.
 
-Show the results with Roe's and Rusanov's solver. 
-Show the results with the three proposed methodologies.
+%\subsection{Regime 1: midly unstable flow}
+
+%Show the results with Roe's and Rusanov's solver. 
+%Show the results with the three proposed methodologies.
+
+%\subsection{Regime 2: strongly unstable flow}
+
+%Show the results with Roe's and Rusanov's solver. 
+%Show the results with the three proposed methodologies.
 
 \section{Conclusions}
-
-\appendix
-\section{Bow shock problem for the shallow water equations}
-
-\section{Other numerical studies}
 

### 2020-11-17 numerics_paper/version2/numerical_chj.tex
Draft introduction.

@@
 \begin{abstract}
-
+We study numerical discretizations of the shallow water equations
+for the simulation of the circular hydraulic jump formed
+when a vertical fluid jet impacts a horizontal plate.  
+This is a challenging problem that includes a curved steady shock
+that can be physically unstable.  We show that numerical methods are prone to
+either suppress the instability completely or form artificial carbuncles.
+We test existing cures for the carbuncle as well as a new method,
+and show results that exhibit the shock instability without unphysical carbuncles.
 \end{abstract}
@@
 
-\begin{itemize}
-  \item Describe the formation of the circular hydraulic jump, some properties, etc. 
-  \item Close this section stating that for some flow regimes, the CHJ becmes unstable and its 
-    shape deviates from a circular shape. For high viscosity flows polygonal shapes are observed 
-    and for low viscosity flows the shape appears to be chaotic. 
-  \item We can use section 1.1 in the original version of the paper.
-\end{itemize}
+Perhaps the first reference to the observation of the circular hydraulic jump
+comes from Lord Rayleigh \cite{rayleigh1914theory}, who wrote that it
+``may usually be seen whenever a stream of water from a tap strikes a horizontal
+surface".  This phenomenon that is familiar in the everyday kitchen sink, is in
+fact highly nonlinear and complex.  Near the jet, the flow is shallow and
+supercritical, while further away it is deeper and subcritical.  The transition from supercritical
+to subcritical flow
+occurs in a very narrow region and takes the form of a \emph{jump} or \emph{bore}
+that is roughly circular if the surface is flat; we refer to it herein as a
+circular hydraulic jump (CHJ).
+
+Early experimental work on the CHJ began some time later
+\cite{kurihara1946hydraulic,tani1949water,watson1964radial}.
+Watson \cite{watson1964radial} derived the jump radius
+implied by Rayleigh's approach and the vertical
+velocity profile in the supercritical region, assuming a no-slip
+boundary condition at the bottom.  He also studied the turbulent
+flow case and performed experiments.
+More detailed experiments revealed different qualitative classification
+of jumps  \cite{ishigai1977heat,craik1981circular}.
+Although later work incorporated more physical details (such as surface tension) into the models
+\cite{bush2003influence}, Bohr et. al. showed that important properties of the jump
+(particularly its radius) could be reasonably predicted using a simple shallow water
+model \cite{bohr1993shallow}.
+
+While the jump is roughly circular, under appropriate conditions it may deviate
+from this shape and deform rapidly and chaotically in time.
+Instability of the jump was observed from fairly early on \cite{craik1981circular}.
+Under special circumstances with more viscous fluids, the jump instability may lead to
+the formation of curious shapes such as polygons \cite{ellegaard1998creating}, but
+for a low-viscosity fluid like water the behavior is generally chaotic.
+The strength of the instability increases with the jet velocity and with the
+depth at the outside of the jump.  For fluids with finite viscosity, the flow
+can also be completely steady for sufficiently small velocities and depths.
 
 \subsection{Numerical shock instabilities}
+
+In \cite{peery1988blunt} a numerical instability was observed to
+appear near the symmetry plane in the simulation of a bow shock.
+This phenomenon, subsequently dubbed ``carbuncle", has been observed by many researchers
+in similar numerical experiments for the Euler equations, and many remedies
+have been proposed, mainly in the form of additional numerical dissipation
+\cite{quirk1997contribution,pandolfi2001numerical,dumbser2004matrix,chauvat2005shock,ismail2009proposed,shen2014stability}.
+Most notably, the HLLE Riemann solver suppresses the carbuncle instability \cite{quirk1997contribution}.
+For a recent review of numerical shock instability and work to alleviate it,
+we refer to \cite[Section 2.5]{simonnumerical}.
+
+Given the similarity of structure between the Euler equations and the shallow
+water equations, it is not surprising that carbuncles appear in numerical
+solutions of the latter as well \cite{kemm2014note}.
+The shallow water carbuncle behaves similarly to the
+Euler carbuncle: it appears when the Roe solver is used, but not when
+the HLL solver is used, and can be suppressed by similar numerical methods
+\cite{kemm2014note,bader2014carbuncle}.
+
+As we will see, carbuncles also appear in the numerical solution of the shallow
+water circular hydraulic jump.  This is natural, since the solution involves a
+standing shock wave.  Dealing with the carbuncle in this context is particularly
+interesting and challenging, since this standing shock should (at least in an
+appropriate flow regime) be unstable, and some research has suggested that the
+carbuncle is the manifestation of a true physical
+instability \cite{moschetta2001carbuncle,elling2009carbuncle}.
 

### 2020-11-19 numerics_paper/version2/numerical_chj.tex
Editorial revisions.

@@
 \newcommand{\A}{{\mathcal A}}
+\newcommand{\apdq}{\A^+ \!\!{\Delta} Q}
+\newcommand{\amdq}{\A^- \!\!{\Delta} Q}
+\newcommand{\imh}{{i-1/2}}
+\newcommand{\iph}{{i+1/2}}
 \newcommand{\bff}{{\bf f}}
@@
 \section{The shallow water circular hydraulic jump}
-The shallow water equations in two (horizontal) dimensions are
+We consider the shallow water model in two horizontal dimensions:
 \begin{subequations} \label{eq:sw}
@@
 \end{subequations}
-These can be written in vector form as
+Here $h, u,$ and $v$ are respectively the depth and the $x$- and $y$-components of
+velocity, which are functions of the two spatial coordinates $(x,y)$ as well as time $t$.
+The gravitational force is proportional to $g$.
+%The system \eqref{eq:sw} can be written in vector form as
+%\begin{align}
+%    q_t + f(q)_x + g(q)_y & = 0.
+%\end{align}
+
+\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
+In this work, we consider the initial boundary value problem consisting of the
+shallow water model \eqref{eq:sw} in an annular domain
 \begin{align}
-    q_t + f(q)_x + g(q)_y & = 0.
-\end{align}
-
-\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
-
-By assuming rotational symmetry in \eqref{eq:sw}, one obtains the
-system (in one spatial variable)
+r_\textup{jet} \le r \le r_\infty
+\end{align}
+where $r = \sqrt{x^2+y^2}$, with prescribed inflow at $r=r_\textup{jet}$
+and prescribed outflow at $r=r_\infty$.
+The domain and boundary conditions are rotationally symmetric.
+By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
 \begin{subequations} \label{eq:rsw}
@@
 \end{subequations}
+where the depth $h$ and radial velocity $u$ are functions of the single spatial coordinate $r$.
 Steady-state solutions of \eqref{eq:rsw} satisfy
@@
 
-A steady-state, rotationally symmetric solution can be given for an annular region as follows:
+A steady-state, rotationally symmetric solution can be given for an annular region with prescribed
+flow at the inner and outer boundaries as follows:
 
@@
 
-Numerical tests suggest that the solution obtained in this way is a stable equilibrium
-of the reduced one-dimensional shallow water equations \eqref{eq:rsw}.  This
-solution provides a useful starting point for testing the stability of the CHJ
-as a solution of the full two-dimensional shallow water equations \eqref{eq:sw}.
-
-\subsection{Formation of the CHJ}\label{sec:formation}
-
-\clearpage
+Numerical tests suggest that the solution of \eqref{eq:rsw} tends rapidly approaches that
+just described under general initial conditions as long as the flow at the jet is
+supercritical and the outer outflow is subcritical.  This solution remains steady
+if the rate of inflow and outflow are matched.  The stability of this solution in
+the face of non-rotationally-symmetric perturbations is an important question not
+only in the shallow water context but for more realistic fluid models and physically.
+It will play an important role in the results we present later.
+The rotationally-symmetric steady state is a useful initial condition for studies
+of CHJ stability.
+
 \section{Numerical methods}
@@
 However, the accuracy of the numerical flux, and hence of the method, is 
-limited to first-order. 
+limited to first order. 
 Two alternatives are common to improve the accuracy properties of Rusanov's method.
@@
 To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers. 
-Our blending criteria is based on the residual of the entropy (in)equality,
+Our blending criterion is based on the residual of the entropy (in)equality,
 which induces dissipation of entropy near discontinuities. 
@@
   %\A^-\Delta Q_{i+1/2} := \frac{1}{2}\sum_p\left(s_{i+1/2}^p-|s_{i+1/2}^p|\right)\W_{i+1/2}^p,
-  \A^+\Delta Q_{i-1/2} := \sum_p\left(s_{i-1/2}^p\right)^+\W_{i-1/2}^p, \qquad
-  \A^-\Delta Q_{i+1/2} := \sum_p\left(s_{i+1/2}^p\right)^-\W_{i+1/2}^p,
+  \apdq_\imh := \sum_p\left(s_{i-1/2}^p\right)^+\W_{i-1/2}^p, \qquad
+  \amdq_\iph := \sum_p\left(s_{i+1/2}^p\right)^-\W_{i+1/2}^p, \\
 \end{align}
@@
 \begin{align}\label{first-order_via_fluct}
-  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right].
+  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right].
 \end{align}
@@
 If no limiters are used 
-($\tilde \W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$), \eqref{second-order_via_fluct} becomes
+($\tilde \W_{i\pm 1/2}^p = \W_{i\pm 1/2}^p$), \eqref{second-order_via_fluct} becomes
 the standard second-order Lax-Wendroff scheme, which in terms of fluctuation-like quantities 
@@
 Riemann solver. The approximate Riemann solver based on Rusanov's flux is robust and known to be carbuncle free. 
-However, it adds large amounts of numerical diffusion that, for relative coarse grids, 
-might dissipate other physical and important features of the solution. If the apprroximate Riemann solver 
+However, it adds large amounts of numerical diffusion that, for relatively coarse grids, 
+might dissipate other physical and important features of the solution. If the approximate Riemann solver 
 is based on the Roe's average, the amount of numerical diffusion is highly reduced; however, 
@@
 
-\section{Entropy dissipative Riemann solver}\label{sec:ent_diss_rs}
+\section{Entropy-dissipative Riemann solver}\label{sec:ent_diss_rs}
 Consider a hyperbolic system of conservation laws 
@@
 It is desirable for a numerical method to satisfy a discrete version of the entropy inequality 
-\eqref{ent_inequality}. This is, however, not true for many standard numerical discretizations. 
+\eqref{ent_inequality}, but many widely-used discretizations do not.
 In general, an entropy stable scheme introduces nonlinear stabilization, which is 
@@
 
-\subsection{Entropy stable Riemann solver}\label{sec:entropy_stable}
+\subsection{Entropy-stable Riemann solver}\label{sec:entropy_stable}
 The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a

### 2020-11-19 numerics_paper/version2/numerical_chj.tex
Some notation changes, wording, and remove trailing whitespace.

@@
 Method \eqref{first-order_FV} with Rusanov's flux \eqref{rusanov_num_flux}
-is positivity preserving and entropy stable, see for instance 
+is positivity preserving and entropy stable; see for instance
 \cite{perthame1996positivity,tadmor2003entropy,guermond2016invariant}
@@
 The accuracy of a given first- or second-order method depends heavily on the Riemann solver.
-In particular, the Rusanov's Riemann solver is highly dissipative. In contrast, the well-known 
+In particular, Rusanov's Riemann solver is highly dissipative. In contrast, the well-known
 Roe's Riemann solver produces more accurate solutions.
@@
 Roe's Riemann solvers are prone to the carbuncle instability for the CHJ.
-In contrast, if we use the Rusanov's 
+In contrast, if we use Rusanov's
 Riemann solver, no carbuncles appear; however, many other important and physical features of
@@
 might dissipate other physical and important features of the solution. If the approximate Riemann solver
-is based on the Roe's average, the amount of numerical diffusion is highly reduced; however, 
+is based on Roe's average, the amount of numerical diffusion is highly reduced; however,
 such an approach is prone to develop the carbuncle instability for certain problems.
-Many carbuncle fixes to the Roe's solver have been proposed for the Euler equations. 
+Many carbuncle fixes to Roe's solver have been proposed for the Euler equations.
 In the context of the shallow water equations, we review in \S\ref{sec:riemann_solvers}
@@
 For instance, methods \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}
-with the Rusanov's numerical flux are based on only two waves that 
+with Rusanov's numerical flux are based on only two waves that
 propagate with the same speed in opposite directions.
@@
 \subsection{Roe's average}\label{sec:roe}
-The Roe's Riemann solver is based on evaluating the flux Jacobian using the Roe's average
+The Roe's Riemann solver is based on evaluating the flux Jacobian using Roe's average
 \begin{align}\label{roe_average}
@@
 \end{align*}
-be the flux Jacobian evaluated using the Roe's average at the interface $i-1/2$.
+be the flux Jacobian evaluated using Roe's average at the interface $i-1/2$.
 The average \eqref{roe_average} is designed to guarantee:
@@
 be the eigenvectors of $\hat A_{i-1/2}$.
-The waves for the two-dimensional Riemann solver based on the Roe's average are 
+The waves for the two-dimensional Riemann solver based on Roe's average are
 $\W_{i-1/2}^p=\alpha_{i-1/2}^p\hat r_{i-1/2}^p$, with $p=1,2,3$. Here $\alpha_{i-1/2}^p$ is given by
@@
 We close this section by remarking that the fluctuations associated with Rusanov's solver
-can be expressed in terms of waves and speeds from the Roe's solver. To see this, first consider \eqref{rus_cons}
+can be expressed in terms of waves and speeds from Roe's solver. To see this, first consider \eqref{rus_cons}
 and rewrite the right-going fluctuation as follows:
@@
 \eqref{ent_inequality}, but many widely-used discretizations do not.
-In general, an entropy stable scheme introduces nonlinear stabilization, which is 
-particularly important for solutions with strong discontinuities that might converge to non-physical 
-weak solutions. 
 In this work, we propose an entropy stable approximate Riemann solver for the one dimensional
@@
 \cite{guermond2018second, guermond2018well} and use
-$\int_{K_i} \frac{\partial\eta(\bfu)}{\partial t}d\bfx=-\int_{K_i} \nabla\cdot\bfq(\bfu) d\bfx$, 
+$\int \frac{\partial\eta(\bfu)}{\partial t}d\bfx=-\int \nabla\cdot\bfq(\bfu) d\bfx$,
 which holds in smooth regions. Then we define
 \begin{align}\label{Ri}
-  R_i := \frac{N_i}{D_i}, 
+  \theta_i := \frac{R_i}{D_i},
 \end{align}
@@
 \begin{align*}
-  N_i=
+  R_i=
   \left|\int_{S_i} \left[\bfv(\bfu) \cdot \nabla\cdot \bff(\bfu) - \nabla\cdot\bfq(\bfu) \right] d\bfx\right|,
 \end{align*}
-and $D_i$ being an upper bound of $N_i$ that acts as normalization so that $0\leq R_i\leq 1$.
-Note that $N_i\approx 0$ if $\bfu$ is smooth in $S_i$. 
+and $D_i$ being an upper bound of $R_i$ that acts as normalization so that $0\leq \theta_i\leq 1$.
+Note that $R_i\approx 0$ if $\bfu$ is smooth in $S_i$.
 In our implementation, we use
 \begin{align*}
-  N_i\approx \left|\bfv(\bfu_i)\cdot \int_{S_i}\nabla\cdot \bff(\bfu)d\bfx 
+  R_i\approx \left|\bfv(\bfu_i)\cdot \int_{S_i}\nabla\cdot \bff(\bfu)d\bfx
   -\int_{S_i}\nabla\cdot\bfq(\bfu) d\bfx\right|,
@@
 where $({\bf z})_k$ denotes the $k$-th component of ${\bf z}\in\mathbb{R}^{d+1}$ and $d$ is the number of physical
-dimensions. In the next section, we use $R_i$ to blend Rusanov's and Roe's Riemann solvers.
+dimensions. In the next section, we use $\theta_i$ to blend Rusanov's and Roe's Riemann solvers.
 
@@
 In this section, we use \eqref{Ri}, which is a normalized measurement of the entropy residual \eqref{ent_residual},
-to blend the Rusanov's and the Roe's Riemann solvers. To do this, we define the following numerical flux:
+to blend Rusanov's and Roe's Riemann solvers. To do this, we define the following numerical flux:
 \begin{align}\label{num_flux_ev}
   F_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2}
-  - \frac{1}{2} \left( R_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-R_{i+1/2})|\hat A_{i+1/2}| \right)(\bfu_{i+1}-\bfu_{i}),
-\end{align}
-and similarly for $F_{i-1/2}$. Here $\mathbb{I}\in\mathbb{R}^{3\times 3}$ is the identity matrix,
-$R_{i+1/2}=\max\{R_{i+1},R_i\}$, $\lambda_{i+1/2}^{\max}$ is the 
-upper bound on the wave speed used in the Rusanov's Riemann solver, see \S \ref{sec:rusanov}, 
-and $\hat A_{i+1/2}$ is the Roe's averaged flux Jacobian, see \S \ref{sec:roe}.
-
-%In terms of fluctuations, 
+  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max} I + (1-\theta_{i+1/2})|\hat A_{i+1/2}| \right)(\bfu_{i+1}-\bfu_{i}),
+\end{align}
+and similarly for $F_{i-1/2}$. Here $I\in\mathbb{R}^{3\times 3}$ is the identity matrix,
+$\theta_{i+1/2}=\max\{\theta_{i+1},\theta_i\}$, $\lambda_{i+1/2}^{\max}$ is the
+upper bound on the wave speed used in Rusanov's Riemann solver (see \S \ref{sec:rusanov}),
+and $\hat A_{i+1/2}$ is Roe's averaged flux Jacobian (see \S \ref{sec:roe}).
+
 The first-order method is given simply by \eqref{first-order_FV}
 with the numerical flux $F_{i+1/2}$ given by \eqref{num_flux_ev}.
-In smooth regions $R_{i+1/2}\approx 0$ so the numerical flux is essentially the Roe's numerical flux. 
-Near discontinuities or steep gradients, $R_{i+1/2}$ becomes large. If, in particular, $R_{i+1/2}\approx 1$, 
-then the numerical flux is essentially the Rusanov's numerical flux.  
-%
+In smooth regions $\theta_{i+1/2}\approx 0$ so the numerical flux is essentially Roe's flux.
+Near discontinuities or steep gradients, $\theta_{i+1/2}$ becomes large. In the limit $\theta_{i+1/2}\approx 1$
+the numerical flux is essentially Rusanov's flux.
+
 The first-order method can also be written in terms of fluctuations via \eqref{first-order_via_fluct}.
 To do this we consider the waves ($\W^p_{i\pm 1/2}$, with $p=1,2,3$)
-from the Roe's solver and define 
+from Roe's solver and define
 \begin{align}\label{lambda_p}
-  \lambda_{i+1/2}^p := R_{i+1/2}\lambda_{i+1/2}^{\max} + (1-R_{i+1/2})|\hat \lambda_{i+1/2}^p|,
-\end{align}
-where $\hat\lambda_{i+1/2}^p$ is the $p$-th eigenvalue of the Roe's solver corresponding to the 
+  \lambda_{i+1/2}^p := \theta_{i+1/2}\lambda_{i+1/2}^{\max} + (1-\theta_{i+1/2})|\hat \lambda_{i+1/2}^p|,
+\end{align}
+where $\hat\lambda_{i+1/2}^p$ is the $p$-th eigenvalue of Roe's solver corresponding to the
 interface $i+1/2$. Note that $\lambda_{i+1/2}^p\geq 0$.
@@
 \begin{align}
-  \A^+\Delta Q_{i-1/2}&=\frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^p, \\
-  \A^-\Delta Q_{i+1/2}&=\frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^p.
+  \apdq_\imh & = \frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^p, \\
+  \amdq_\iph & = \frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^p.
 \end{align}
@@
 
-To implement the second-order Lax-Wendroff-LeVeque type method in Clawpack, the Riemann solver must provide 
-the fluctuations, the waves and their speed. The fluctuations are simply \eqref{ev_fluctuations} 
-and the waves are the Roe's waves $\W_{i+1/2}^p$, with $p=1,2,3$. 
-In the framework of LeVeque, for a given interface $i+1/2$, 
-we would need to consider a wave speed $s_{i+1/2}^p$ so that 
-\begin{align*}
-  \hat \lambda_{i+1/2}^p+\lambda_{i+1/2}^p = s_{i+1/2}^p+|s_{i+1/2}^p|, \qquad 
-  \hat \lambda_{i+1/2}^p-\lambda_{i+1/2}^p = s_{i+1/2}^p-|s_{i+1/2}^p|,
-\end{align*}
-which is not possible to obtain in general (since the problem is overdetermined). 
+To implement a second-order LWL scheme based on this solver, in addition to
+these fluctuations we must define a set of waves corresponding speeds.
+Using only the waves from the Roe solver, it is in general not possible
+to choose speeds that yield the fluctuations \eqref{ev_fluctuations}.
+%In the framework of LeVeque, for a given interface $i+1/2$,
+%we would need to consider a wave speed $s_{i+1/2}^p$ so that
+%\begin{align*}
+%  \hat \lambda_{i+1/2}^p+\lambda_{i+1/2}^p = s_{i+1/2}^p+|s_{i+1/2}^p|, \qquad
+%  \hat \lambda_{i+1/2}^p-\lambda_{i+1/2}^p = s_{i+1/2}^p-|s_{i+1/2}^p|,
+%\end{align*}
+%which is not possible to obtain in general (since the problem is overdetermined).
 Instead, we consider
@@
   \tilde F_{i+1/2} = \frac{1}{2}\sum_p\lambda_{i+1/2}^p
-  \left(1-\frac{\Delta t}{\Delta x}\lambda_{i+1/2}^p\right)\tilde\W_{i+1/2}^p,
-\end{align*}
-and similarly for $\tilde F_{i-1/2}$, are the flux corrections with $\tilde\W_{i+1/2}^p$ 
-being a TVD-limited version of $\W_{i+1/2}^p$.
+  \left(1-\frac{\Delta t}{\Delta x}\lambda_{i+1/2}^p\right)\tilde\W_{i+1/2}^p
+\end{align*}
+are the flux corrections with $\tilde\W_{i+1/2}^p$ being a limited version of $\W_{i+1/2}^p$.
 
@@
 which corresponds to the standard Lax-Wendroff method written in terms of fluctuation-like quantities.
-Note that if $R_{i\pm 1/2}=0$, we recover the standard second-order Lax-Wendroff method based on 
+Note that if $\theta_{i\pm 1/2}=0$, we recover the standard second-order Lax-Wendroff method based on
 Roe's Riemann solver.
@@
 order of convergence (EOC) for problems with smooth solution, which is expected since
-$R_{i\pm 1/2}\approx 0$ in smooth regions.
-Similarly, if $R_{i\pm 1/2}=1$, we recover the Lax-Wendroff method based on Rusanov's Riemann solver, 
-which is clear by writting the fluctuations in Rusanov's method in terms of Roe's waves and speeds, 
+$\theta_{i\pm 1/2}\approx 0$ in smooth regions.
+Similarly, if $\theta_{i\pm 1/2}=1$, we recover the Lax-Wendroff method based on Rusanov's Riemann solver,
+which is clear by writing the fluctuations in Rusanov's method in terms of Roe's waves and speeds;
 see \eqref{fluct_rus}.
@@
 \subsection{Entropy-stable Riemann solver}\label{sec:entropy_stable}
-The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a
-measurement of the entropy residual. The objective is to induce dissipation of entropy 
+The Riemann solver presented in the previous section blends the solvers of Rusanov and Roe based on a
+local measurement of the entropy residual. The objective is to induce dissipation of entropy
 near discontinuities. However, we have no guarantee that methods \eqref{first-order_via_fluct}
 or \eqref{second-order_via_fluct} based on this Riemann solver are (locally) entropy stable.
-In this section, we consider the one-dimensional problem and increase the numerical dissipation 
-to guarantee entropy stability for \eqref{first-order_via_fluct}. 
-To this end we follow the work by Tadmor, see for instance \cite{tadmor1987numerical, tadmor2003entropy} 
-and references therein.
-
-Recall method \eqref{first-order_FV}
-\begin{align*}
-  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
-\end{align*}
-with the numerical fluxes given by 
-\begin{align*}
-  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} 
-  - \frac{1}{2} \left( R_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-R_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfu_{i+1}-\bfu_{i}),
-\end{align*}
-and similarly for $\bfF_{i-1/2}$. Here $\lambda_{i+1/2}^{\min}$ is yet to be defined.
-We can write the numerical fluxes in terms of $\lambda^p_{i\pm 1/2}$ \eqref{lambda_p}
-and the Roe's waves $\W_{i\pm 1/2}^p$.
-Doing so, we get
-\begin{align*}
-  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} 
-  - \frac{1}{2}\sum_p(\lambda_{i+1/2}^p+\lambda_{i+1/2}^{\min})\W_{i+1/2}^p,
-\end{align*}
-and similarly for $\bfF_{i-1/2}$.
+In this section, we further modify the first-order method
+\eqref{first-order_via_fluct} in order to guarantee local entropy stability.
+
 Let $\bfv_i=\bfv(\bfu_i)$ and $q_i=q(\bfu_i)$ denote the entropy variable and the
@@
 \end{align*}
-and similarly for $G_{i-1/2}$, are discretizations of the entropy flux.
-
-Based on \eqref{es_cond}, we choose
+is the discrete entropy flux.
+
+To enforce \eqref{es_cond} we further modify the flux \eqref{num_flux_ev}:
+\begin{align*}
+  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2}
+  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-\theta_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfu_{i+1}-\bfu_{i}).
+\end{align*}
+Here $\lambda_{i+1/2}^{\min}$ is yet to be defined.
+We can write the numerical fluxes in terms of $\lambda^p_{i\pm 1/2}$ \eqref{lambda_p}
+and Roe's waves $\W_{i\pm 1/2}^p$.
+Doing so, we get
+\begin{align*}
+  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2}
+  - \frac{1}{2}\sum_p(\lambda_{i+1/2}^p+\lambda_{i+1/2}^{\min})\W_{i+1/2}^p.
+\end{align*}
+
+We set
 \begin{align}\label{lambda_min}
@@
 which guarantees \eqref{es_cond} and, therefore, \eqref{ent_ineq}.
-Since the blended Riemann solver induces dissipation of entropy, 
-we expect condition \eqref{es_cond} to be fulfilled most of the times even with $\lambda_{i+1/2}^{\min}=0$. 
-Therefore, \eqref{lambda_min} is used as a safeguard to guarantee entropy stability
-for the one-dimensional problem using method \eqref{first-order_FV} or \eqref{first-order_via_fluct}. 
-Extra modifications are needed to guarantee entropy stability of Lax-Wendroff-LeVeque method 
-\eqref{second-order_via_fluct} and its multidimensional extension via Strang's splitting. 
-We do not pursue these modifications in this work.
+
+Since the blended Riemann solver described in Section \ref{} already
+tends to introduce entropy dissipation,
+we expect condition \eqref{es_cond} to be fulfilled most of the time even with $\lambda_{i+1/2}^{\min}=0$.
+But \eqref{lambda_min} is used as a safeguard to guarantee entropy stability of the first-order method.
+Extra modifications would be needed to guarantee entropy stability of the second-order Lax-Wendroff-LeVeque method
+\eqref{second-order_via_fluct} and its multidimensional extension via Strang splitting.
+We do not pursue such modifications in this work.
 
@@
 \begin{remark}[About the effect of $\lambda_{i\pm 1/2}^{\min}$ for Roe's solver.]
-  If we set $R_{i\pm 1/2}=0, ~\forall i$ in \eqref{lambda_p}, then the Riemann solver becomes the Roe's solver.
+  If we set $\theta_{i\pm 1/2}=0, ~\forall i$ in \eqref{lambda_p}, then the Riemann solver becomes Roe's solver.
   In this case, it is well known that transonic rarefactions develop non-physical shocks,
-  a phenomenon known as `entropy-glitch'. To fix this problem, many entropy fixes to the Roe's
-  solver have been developed. Enhancing the dissipation via \eqref{lambda_min} behaves as an entropy fix for
-  the Roe's solver, which we demonstrate via a numerical experiment in \S \ref{sec:rp_dry_bed}. 
+  a phenomenon known as `entropy glitch'. To fix this problem, many entropy fixes to Roe's
+  solver have been developed. Enhancing the dissipation via \eqref{lambda_min}
+  also serves as a transonic entropy fix for Roe's solver, which we demonstrate
+  via a numerical experiment in \S \ref{sec:rp_dry_bed}.
 \end{remark}
@@
 For this problem we use the first-order method \eqref{first-order_via_fluct},
-which helps preventing violations of positivity. Although, only the Rusanov's solver guarantees
+which helps preventing violations of positivity. Although, only Rusanov's solver guarantees
 positivity, we do not get negative values for the water height with this method and the other Riemann solvers.
@@
 second-order accuracy of method \eqref{second-order_via_fluct} with Roe's and the blended solvers. Using
-the Rusanov's Riemann solver degrades the accuracy to first-order. The overall accuracy of the blended solver
+Rusanov's Riemann solver degrades the accuracy to first-order. The overall accuracy of the blended solver
 is not degraded since the extra dissipation is not applied in smooth regions.
@@
 leads to $\lambda_{i\pm 1/2}^{\min}=0$ for all the experiments, but this is not true in general.
-Finally, let us demonstrate the effect of $\lambda_{i\pm 1/2}^{\min}$ by imposing $R_{i}=0$ in \eqref{lambda_p},
+Finally, let us demonstrate the effect of $\lambda_{i\pm 1/2}^{\min}$ by imposing $\theta_{i}=0$ in \eqref{lambda_p},
 which is equivalent to using Roe's solver with extra dissipation given by $\lambda_{i\pm 1/2}^{\min}$.
@@
 That is, inducing entropy dissipation via the blended Riemann solver does not degrade the high-order
-accuracy properties of the Roe's solver. 
+accuracy properties of Roe's solver.
 In contrast, the accuracy and convergence rates using Rusanov's solver are clearly degraded.
@@
 at different times using the second-order method \eqref{second-order_via_fluct} with Roe's, Rusanov's and the blended Riemann solvers.
-In the second row of the same figure, we plot $R_i$ \eqref{Ri}, which is a (normalized) measurement of the
-entropy residual. Note that $R_i$ accurately locates the shocks and is (nearly) zero in smooth regions. 
+In the second row of the same figure, we plot $\theta_i$ \eqref{Ri}, which is a (normalized) measurement of the
+entropy residual. Note that $\theta_i$ accurately locates the shocks and is (nearly) zero in smooth regions.
 In Table \ref{table:chj_1D}, we summarize the results of a convergence study.
@@
     (cyan) Roe's, (red) Rusanov's and (blue) the blended Riemann solvers.
-    In the second row we plot $R_i$, which is given by \eqref{Ri}.
+    In the second row we plot $\theta_i$, which is given by \eqref{Ri}.
     In all simulations we consider $\Delta x=1/400$.
@@
 Nevertheless, the steady state profile is smooth.
-In Figure \ref{fig:steady_outflow}, we show the solution and $R_i$ at different times using
+In Figure \ref{fig:steady_outflow}, we show the solution and $\theta_i$ at different times using
 the second-order method \eqref{second-order_via_fluct}
@@
     (cyan) Roe's, (red) Rusanov's and (dashed blue) the blended Riemann solvers.
-    In the second row we plot $R_i$, which is given by \eqref{Ri}.
+    In the second row we plot $\theta_i$, which is given by \eqref{Ri}.
     In all simulations we consider $\Delta x=1/400$.

### 2020-11-19 numerics_paper/version2/numerical_chj.tex
Add missing reference.

@@
 
-Since the blended Riemann solver described in Section \ref{} already
+Since the blended Riemann solver described in Section \ref{sec:blended_rs} already
 tends to introduce entropy dissipation,

### 2020-11-19 rpn2_swq_hllemcc.f90
Fix a bug in HLLEMCC solver for mapped grids.

@@ -181,7 +181,7 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
                 ! Alternative if cbrt for cubic root is available:
                 ! ffroude  = dmax1(0.0d0,(1.0d0 - dcbrt(zfroude)))
                
-                rhind = a(i)
+                rhind = a(i) * auxl(ilenrat,i)
                 rhind = rhind*dmin1(kepsilon*renore*ffroude,1.0d0)**kbeta 
                 ! Alternative if cbrt for cubic root is available:
                 ! rhind = rhind*dcbrt(dmin1(kepsilon*renore*ffroude,1.0d0))

### 2020-11-22 tmp_llf/hydraulic_jump_2D.py
Introduce LWL at first use and use it consistently. Fix typo.

@@ -277,7 +277,7 @@ def setup(
         elif riemann_solver == 'llf':
             solver = pyclaw.ClawSolver2D(shallow_llf_2D)
             solver.num_eqn = 3
-            solver.num_waves = 3
+            solver.num_waves = 2
             solver.fwave = False
         else:
             raise Exception('Unrecognized Riemann solver') 
@@ -413,7 +413,7 @@ if __name__ == "__main__":
     grav=1.0
     
     # ***** numerics ***** #
-    num_cells = 500
+    num_cells = 200
     solver_type = 'classic'
     #riemann_solver = 'hlle'
     riemann_solver = 'llf'
@@ -469,7 +469,7 @@ if __name__ == "__main__":
         num_cells=num_cells,
         solver_type = solver_type,
         riemann_solver=riemann_solver,        
-        use_petsc=True)
+        use_petsc=False)
     #
     
     claw.run()

### 2020-11-22 numerics_paper/version2/numerical_chj.tex

@@
 In this work we consider an adaptation of Lax-Wendroff method,
-sometimes known as Lax-Wendroff-LeVeque \cite{leveque1997wave,leveque2002finite},
+sometimes known as Lax-Wendroff-LeVeque (LWL) \cite{leveque1997wave,leveque2002finite},
 that incorporates limiters into the second-order correction.
@@
 
-The second-order Lax-Wendroff-LeVeque method removes some of the numerical dissipation in
+The second-order LWL method removes some of the numerical dissipation in
 \eqref{first-order_via_fluct} via a flux correction at each interface. The method is written as follows:
@@
 In this work, we propose an entropy stable approximate Riemann solver for the one dimensional
-shallow water equations. We use this Riemann solver as building block within the Lax-Wendroff-LeVeque
+shallow water equations. We use this Riemann solver as building block within the LWL
 method, as described in \cite{leveque1997wave,leveque2002finite} and implemented in Clawpack.
@@
 To implement a second-order LWL scheme based on this solver, in addition to
-these fluctuations we must define a set of waves corresponding speeds.
+these fluctuations we must define a set of waves and corresponding speeds.
 Using only the waves from the Roe solver, it is in general not possible
@@
 Here $\sgn(z)$ is the sign function of $z$.
-The second-order Lax-Wendroff-LeVeque method for the one-dimensional problem is given by
+The second-order LWL method for the one-dimensional problem is given by
 \begin{align}\label{second-order}
@@
 \end{align}
-which guarantees \eqref{es_cond} and, therefore, \eqref{ent_ineq}.
+which guarantees \eqref{es_cond}, and therefore \eqref{ent_ineq}.
 
@@
 But \eqref{lambda_min} is used as a safeguard to guarantee entropy stability of the first-order method.
-Extra modifications would be needed to guarantee entropy stability of the second-order Lax-Wendroff-LeVeque method
+Extra modifications would be needed to guarantee entropy stability of the second-order LWL method
 \eqref{second-order_via_fluct} and its multidimensional extension via Strang splitting.
@@
 We compare the behavior of the blended solver against the standard Roe's and Rusanov's solvers.
-In most of the experiments we use these Riemann solvers with the Lax-Wendroff-LeVeque method reviewed in \S\ref{sec:lax-wendroff-leveque}.
+In most of the experiments we use these Riemann solvers with the LWL method reviewed in \S\ref{sec:lax-wendroff-leveque}.
 When the exact solution is available, we report a convergence test based on the $L^1$-error for the water height

### 2020-11-23 numerics_paper/version2/numerical_chj.tex
Shorten material on approximate Riemann solvers.

@@
 \newcommand{\amdq}{\A^- \!\!{\Delta} Q}
+\newcommand{\Ft}{\tilde{F}}
 \newcommand{\imh}{{i-1/2}}
@@
 \section{Numerical methods}
-In this work, we consider Riemann-based finite volume methods.
-Considering a one dimensional hyperbolic conservation law,
-the cell average of the numerical solution given by a
-finite volume scheme (with forward Euler) can be written as follows:
-\begin{align}\label{first-order_FV}
-  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
-\end{align}
-where $\bfF_{i+1/2}=\bfF_{i+1/2}(u^n)$ is a numerical flux; i.e., $\bfF_{i+1/2}$ is
-a discretization of the flux $\bff(\bfu(x_{i+1/2},t^n))$.
-This numerical flux is composed of two parts, a consistent term that approximates
-the physical flux at the interface between cells $i$ and $i+1$, and a stabilization term.
-For instance, the well-known Rusanov's numerical flux is given by
-\begin{align}\label{rusanov_num_flux}
-  \bfF_{i+1/2}(\bfu)=\frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2} - \lambda_{i+1/2}(\bfu_{i+1}-\bfu_i),
-\end{align}
-where $\lambda_{i+1/2}$ is an upper bound on the wave speed associated with the Riemann problem
-composed of the states $\bfu_i$ and $\bfu_{i+1}$.
-%If $z=u$ in \eqref{rusanov_num_flux}
-%; i.e., if the Rusanov's numerical flux is obtained considering
-%the cell averages of the solution
-%and if the time step is small enough,
-Method \eqref{first-order_FV} with Rusanov's flux \eqref{rusanov_num_flux}
-is positivity preserving and entropy stable; see for instance
-\cite{perthame1996positivity,tadmor2003entropy,guermond2016invariant}
-However, the accuracy of the numerical flux, and hence of the method, is
-limited to first order.
-Two alternatives are common to improve the accuracy properties of Rusanov's method.
-The numerical flux $\bfF_{i+1/2}$ can be approximated based on high-order polynomial
-reconstructions (or combinations of them) around the interface $i+1/2$.
-For smooth solutions, this strategy leads to arbitrarily high-order
-methods. Alternatively, second-order accuracy can be obtained by removing some of the
-numerical dissipation introduced by the stabilization component of the numerical flux.
-The second-order Lax-Wendroff method \cite{lax1959systems} is a popular example of this approach.
-In this work we consider an adaptation of Lax-Wendroff method,
-sometimes known as Lax-Wendroff-LeVeque (LWL) \cite{leveque1997wave,leveque2002finite},
-that incorporates limiters into the second-order correction.
-These high-order corrections to Rusanov's method require careful treatment to guarantee
-the solution is still positivity preserving and/or entropy stable.
-
-The accuracy of a given first- or second-order method depends heavily on the Riemann solver.
-In particular, Rusanov's Riemann solver is highly dissipative. In contrast, the well-known
-Roe's Riemann solver produces more accurate solutions.
-%
-As we discussed in the introduction, finite volume methods with
-Roe's Riemann solvers are prone to the carbuncle instability for the CHJ.
-In contrast, if we use Rusanov's
-Riemann solver, no carbuncles appear; however, many other important and physical features of
-the solution are dissipated, unless the grid is highly refined.
-Our main objective is to study the CHJ and to propose a Riemann solver
-that is carbuncle free but do not dissipate important features of the solution.
-To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers.
-Our blending criterion is based on the residual of the entropy (in)equality,
-which induces dissipation of entropy near discontinuities.
-Blending these Riemann solvers has been proposed before
-to solve the Euler equations, see e.g.
-\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
-the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
-and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
-In these references, the authors use blending criteria different than ours.
-Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
-For example, in \cite{ismail2009affordable,ismail2009proposed},
-the authors developed carbuncle free solvers for the Euler equations
-via dissipation of entropy.
-
-\subsection{Lax-Wendroff-LeVeque method}\label{sec:lax-wendroff-leveque}
-Let us now review the second-order correction to method \eqref{first-order_FV} implemented in
-Clawpack \cite{clawpack} and PyClaw \cite{pyclaw-sisc}.
-We refer to \cite{leveque1997wave, leveque2002finite} for a detailed description of the method.
-We consider only a one-dimensional problem; in the
-numerical experiments in \S\ref{sec:num} we use second-order Strang's splitting \cite{strang1968construction}
-to extend the method to multiple dimensions.
-
-The finite volume method \eqref{first-order_FV} can be written in terms of right-
-and left-going waves emanating from the interfaces $i-1/2$ and $i+1/2$, respectively.
-The net contribution of these waves is called right- and left-going fluctuations.
+In this work we study the behavior of certain shock-capturing finite volume
+methods based on the use of Riemann solvers.  For simplicity, we discuss these
+solvers in the context of a 1-dimensional mesh.  
+In the numerical experiments in \S\ref{sec:num} we use second-order Strang
+splitting \cite{strang1968construction} to extend the method to multiple dimensions.
+
+The numerical solution $u_i$
+represents the average value over cell $i$, which extends from $x_\imh$ to $x_\iph$.
+At each time step and at each interface $x_\imh$, we approximately solve the
+Riemann problem with initial states $(u^n_{i-1},u^n_i)$.  The approximate
+solution consists of a set of traveling discontinuities or waves $\W^p_\imh$
+and corresponding speeds $s^p_\imh$.
+We use the wave propagation framework developed by LeVeque \cite{leveque1997wave, leveque2002finite}
+and employed in the Clawpack software \cite{clawpack,pyclaw-sisc}, which implements the LWL scheme
+\begin{align}\label{second-order_via_fluct}
+  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
+  -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
+\end{align}
 Upon defining $z^+:=\frac{1}{2}(z+|z|)$ and $z^-:=\frac{1}{2}(z-|z|)$,
@@
 \begin{align}\label{fluct}
-  %\A^+\Delta Q_{i-1/2} := \frac{1}{2}\sum_p\left(s_{i-1/2}^p+|s_{i-1/2}^p|\right)\W_{i-1/2}^p, \quad
-  %\A^-\Delta Q_{i+1/2} := \frac{1}{2}\sum_p\left(s_{i+1/2}^p-|s_{i+1/2}^p|\right)\W_{i+1/2}^p,
   \apdq_\imh := \sum_p\left(s_{i-1/2}^p\right)^+\W_{i-1/2}^p, \qquad
@@
 \end{align}
-where $\W_{i\pm 1/2}^p$ and $s_{i\pm 1/2}^2$ are the waves and their corresponding speeds, respectively.
-The waves and their speed of propagation (to compute the fluctuations) are tightly related to Riemann solvers.
-In \S\ref{sec:riemann_solvers} we review the definition of the waves and speeds associated with
-Rusanov's and Roe's Riemann solvers.
-Based on fluctuations, method \eqref{first-order_FV} can be written as
+and represent the effect (to first
+order accuracy) of waves coming from the solution of the Riemann problem at
+each of the neighboring interfaces.  Meanwhile, $\Ft_{i\pm 1/2}$ are correction
+fluxes that make the method second-order accurate:
+\begin{align*}
+  \tilde{F}_{i\pm 1/2}=\frac{1}{2}\sum_p|s_{i\pm 1/2}^p|\left(1-\frac{\Delta t}{\Delta x}|s_{i\pm 1/2}^p|\right)\tilde\W_{i\pm 1/2}^p.
+\end{align*}
+Here $\tilde{\W}_{i\pm 1/2}^p$ is a TVD-limited version of $\W_{i\pm 1/2}^p$.
+We will sometimes work with
+the first-order method obtained by setting the correction fluxes to zero:
 \begin{align}\label{first-order_via_fluct}
-  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right].
-\end{align}
-
-The second-order LWL method removes some of the numerical dissipation in
-\eqref{first-order_via_fluct} via a flux correction at each interface. The method is written as follows:
-\begin{align}\label{second-order_via_fluct}
-  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\A^+\Delta Q_{i-1/2}+\A^-\Delta Q_{i+1/2}\right]
-  -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
-\end{align}
-where
-\begin{align*}
-  \tilde{F}_{i\pm 1/2}=\frac{1}{2}\sum_p|s_{i\pm 1/2}^p|\left(1-\frac{\Delta t}{\Delta x}|s_{i\pm 1/2}^p|\right)\tilde\W_{i\pm 1/2}^p
-\end{align*}
-are the flux corrections at the interfaces $i\pm 1/2$. Here $\tilde{\W}_{i\pm 1/2}^p$ is a
-TVD-limited version of $\W_{i\pm 1/2}^p$.
-In this work, we consider the min-mod limiters, we refer to \cite{leveque1997wave} for more details.
-If no limiters are used
-($\tilde \W_{i\pm 1/2}^p = \W_{i\pm 1/2}^p$), \eqref{second-order_via_fluct} becomes
-the standard second-order Lax-Wendroff scheme, which in terms of fluctuation-like quantities
-is given by
-\begin{align*}
-  u_i^{n+1}=u_i^n-\frac{\Delta t}{\Delta x}
-\left[
-  \frac{1}{2}\left(s_{i-1/2}^p+\frac{\Delta t}{\Delta x}|s_{i-1/2}^p|^2\right)\W_{i-1/2}^p
-  +
-  \frac{1}{2}\left(s_{i+1/2}^p-\frac{\Delta t}{\Delta x}|s_{i+1/2}^p|^2\right)\W_{i+1/2}^p
-\right].
-\end{align*}
-To implement method \eqref{second-order_via_fluct},
-we need to choose a Riemann solver. Although the scheme is second-order
-regardless of the Riemann solver, the accuracy and robustness depends upon the
-Riemann solver. The approximate Riemann solver based on Rusanov's flux is robust and known to be carbuncle free.
-However, it adds large amounts of numerical diffusion that, for relatively coarse grids,
-might dissipate other physical and important features of the solution. If the approximate Riemann solver
-is based on Roe's average, the amount of numerical diffusion is highly reduced; however,
-such an approach is prone to develop the carbuncle instability for certain problems.
-Many carbuncle fixes to Roe's solver have been proposed for the Euler equations.
-In the context of the shallow water equations, we review in \S\ref{sec:riemann_solvers}
-the Riemann solver proposed in \cite{kemm2014note}.
-
-\section{Riemann solvers}\label{sec:riemann_solvers}
-
-For the shallow water equations in two-dimensions, the structure of the exact (and some
-approximated) Riemann solvers contains three waves. This is the case, for example, of the well-known
-Roe's Riemann solver, see for instance \cite{roe1981approximate} and
-\cite[\S 15.3.3 and \S 21.7]{leveque2002finite}.
-Simpler approaches are also possible.
-For instance, methods \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}
-with Rusanov's numerical flux are based on only two waves that
-propagate with the same speed in opposite directions.
+  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
+\end{align}
+which, for the conservative Riemann solvers we use herein, can also be written
+as a first-order flux-differencing scheme:
+\begin{align}\label{first-order_FV}
+  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
+\end{align}
+where $\bfF_{i+1/2}$ is a numerical approximation of the flux at $x_\iph$.
+
+Next we briefly review two commonly-used Riemann solvers: those of Rusanov and
+Roe.  Neither of these solvers deals with the carbuncle instability in fully
+satisfactory way.  Rusanov's solver suppresses the carbuncle but is known to
+also suppress related real instabilities, while Roe's solver exhibits carbuncles.
+Later we will combine these two solvers in order to better deal with shock instability.
 We refer to \cite{ketcheson2020riemann} and references therein for
-a full and detailed review of these two Riemann solvers.
-
-\subsection{Roe's average}\label{sec:roe}
-The Roe's Riemann solver is based on evaluating the flux Jacobian using Roe's average
+a full and detailed description of these two Riemann solvers in the context of
+the shallow water equations.
+
+\subsection{Rusanov's Riemann solver}\label{sec:rusanov}
+The Rusanov solver approximates the Riemann solution by with a single wave traveling in each
+direction.  Both waves are assumed to travel with speed $\lambda^{\max}_\imh$, which is an
+upper bound on the wave speeds appearing in the true solution of the Riemann problem.
+In this work, we take $\lambda_{i-1/2}^{\max}$ to be the upper bound
+from \cite[Prop. 3.7]{azerad2017well}.
+The waves are then given by
+\begin{align}\label{rusanov_waves}
+  \W_{i-1/2}^{1,\Rus}=\bfu_{i-1/2}-\bfu_{i-1}, \qquad
+  \W_{i-1/2}^{2,\Rus}=\bfu_{i}-\bfu_{i-1/2},
+\end{align}
+where the intermediate state $\bfu_{i-1/2}$ is determined by imposing conservation:
+\begin{align}\label{rus_cons}
+  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{1,\Rus}+\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus} = \bff(\bfu_{i})-\bff(\bfu_{i-1}).
+\end{align}
+The fluctuations are given by \eqref{fluct} with $s_{i-1/2}^1=-\lambda_{i-1/2}^{\max}$
+and $s_{i-1/2}^2=\lambda_{i-1/2}^{\max}$, and the first- and second-order methods based
+on Rusanov's solver are given by \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.
+
+\subsection{Roe's Riemann solver} \label{sec:roe}
+The Roe Riemann solver is based on evaluating the flux Jacobian using Roe's average
 \begin{align}\label{roe_average}
@@
 \end{align}
-Let
-\begin{align*}
-\hat A_{i-1/2} =
-\begin{bmatrix}
-  0 & 1 & 0 \\
-  -\hat u_{i-1/2}^2+g\bar h_{i-1/2} & 2\hat u_{i-1/2} & 0 \\
-  -\hat u_{i-1/2} \hat v_{i-1/2} & \hat v_{i-1/2} & \hat u_{i-1/2}
-\end{bmatrix}
-\end{align*}
-be the flux Jacobian evaluated using Roe's average at the interface $i-1/2$.
-The average \eqref{roe_average} is designed to guarantee:
-i)  consistency of the approximate flux Jacobian
-($\hat A_{i-1/2}\rightarrow \bff^\prime(\bfu(x_{i-1/2}))$ as $\bfu_i, \bfu_{i-1}\rightarrow \bfu(x_{i-1/2})$);
-ii) hyperbolicity ($\hat A_{i-1/2}$ is diagonalizable with real eigenvalues);
-and iii) conservation ($\hat A_{i-1/2}(\bfu_i-\bfu_{i-1})=\bff(\bfu_i)-\bff(\bfu_{i-1})$).
-Let
+The waves and speeds are given by the eigenvectors and eigenvalues of the approximate
+flux Jacobian obtained using these averages, resulting in
+\begin{align*}
+  \hat\lambda_{i-1/2}^1=\hat u_{i-1/2}-g\sqrt{g\bar h_{i-1/2}}, \qquad
+  \hat\lambda_{i-1/2}^2=\hat u_{i-1/2}, \qquad
+  \hat\lambda_{i-1/2}^3=\hat u_{i-1/2}+g\sqrt{g\bar h_{i-1/2}}. \\
+\end{align*}
+and $\W_{i-1/2}^p=\alpha_{i-1/2}^p\hat r_{i-1/2}^p$, with
 \begin{align*}
@@
 \end{align*}
-be the eigenvectors of $\hat A_{i-1/2}$.
-The waves for the two-dimensional Riemann solver based on Roe's average are
-$\W_{i-1/2}^p=\alpha_{i-1/2}^p\hat r_{i-1/2}^p$, with $p=1,2,3$. Here $\alpha_{i-1/2}^p$ is given by
+and the factors $\alpha^p_\imh$ obtained by solving
 \begin{align}\label{system_for_alphas}
@@
 \end{align}
-%where $\alpha_{i-1/2}^p$ is the $p$-th entry from the solution of
+
+\section{An entropy-based blending of Rusanov and Roe}
+
+As we discussed in the introduction, finite volume methods with
+Roe's Riemann solvers are prone to the carbuncle instability for the CHJ.
+In contrast, if we use Rusanov's
+Riemann solver, no carbuncles appear; however, many other important and physical features of
+the solution are dissipated, unless the grid is highly refined.
+Our main objective is to study the CHJ and to propose a Riemann solver
+that is carbuncle-free but does not dissipate important features of the solution.
+To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers.
+Our blending criterion is based on the residual of the entropy (in)equality,
+which induces dissipation of entropy near discontinuities.
+Blending these Riemann solvers has been proposed before
+to solve the Euler equations, see e.g.
+\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
+the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
+and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
+In these references, the authors use blending criteria different than ours.
+Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
+For example, in \cite{ismail2009affordable,ismail2009proposed},
+the authors developed carbuncle free solvers for the Euler equations
+via dissipation of entropy.
+
+%We close this section by remarking that the fluctuations associated with Rusanov's solver
+%can be expressed in terms of waves and speeds from Roe's solver. To see this, first consider \eqref{rus_cons}
+%and rewrite the right-going fluctuation as follows:
+%\begin{align}\label{rus_rs_aux}
+%  \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+%  = \bff(\bfu_{i})-\bff(\bfu_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
+%  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}.
+%  %\A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
+%\end{align}
+%From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get
+%$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$.
+%Using \eqref{cons_roe}, \eqref{rus_rs_aux} becomes
+%\begin{align*}
+%  \lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+%  = \sum_p \left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
+%  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus},
 %\end{align*}
-The speed of the waves is given by the eigenvalues of $\hat A_{i-1/2}$
-\begin{align*}
-  \hat\lambda_{i-1/2}^1=\hat u_{i-1/2}-g\sqrt{g\bar h_{i-1/2}}, \qquad
-  \hat\lambda_{i-1/2}^2=\hat u_{i-1/2}, \qquad
-  \hat\lambda_{i-1/2}^3=\hat u_{i-1/2}+g\sqrt{g\bar h_{i-1/2}}. \\
-\end{align*}
-
-The waves and speeds at interface $i+1/2$ are computed similarly.
-The fluctuations are given by \eqref{fluct} with $s_{i\pm 1/2}^p=\hat\lambda_{i\pm 1/2}^p$
-Finally, the first- and second-order methods based on Roe's Riemann solver are given by
-\eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.
-We close this review by remarking that the conservation property
-$\hat A_{i-1/2}(\bfu_i-\bfu_{i-1})=\bff(\bfu_i)-\bff(\bfu_{i-1})$ implies that
-\begin{align}\label{cons_roe}
-  \A^-\Delta Q_{i-1/2}+\A^+\Delta Q_{i-1/2}=\sum_p \hat\lambda_{i-1/2}^p\W_{i-1/2}^p=\bff(\bfu_{i})-\bff(\bfu_{i-1}),
-\end{align}
-a property we will use in the next section.
-
-\subsection{Rusanov's Riemann solver}\label{sec:rusanov}
-The Rusanov's Riemann solver considers only two waves that emanate from each interface,
-say for instance interface $i-1/2$, and travel with the same speed $\lambda_{i-1/2}^{\max}\geq 0$
-but in opposite directions.
-The speed of propagation $\lambda_{i-1/2}^{\max}$ is an upper bound on the maximum wave speed
-associated with the corresponding Riemann problem between the states $\bfu_{i-1}$ and $\bfu_{i}$.
-In this work, we define $\lambda_{i-1/2}^{\max}$ via the guaranteed upper bound from \cite[Prop. 3.7]{azerad2017well}.
-The waves are given by
-\begin{align}\label{rusanov_waves}
-  \W_{i-1/2}^{1,\Rus}=\bfu_{i-1/2}-\bfu_{i-1}, \qquad
-  \W_{i-1/2}^{2,\Rus}=\bfu_{i}-\bfu_{i-1/2},
-\end{align}
-where $\bfu_{i-1/2}$ is a middle state in the solution of the Riemann problem. The middle state is obtained
-by imposing
-\begin{align}\label{rus_cons}
-  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{1,\Rus}+\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus} = \bff(\bfu_{i})-\bff(\bfu_{i-1}),
-\end{align}
-which guarantees the solver is conservative.
-The fluctuations are given by \eqref{fluct} with $s_{i-1/2}^1=-\lambda_{i-1/2}^{\max}$
-and $s_{i-1/2}^2=\lambda_{i-1/2}^{\max}$, and the first- and second-order methods based
-on Rusanov's solver are given by \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.
-
-We close this section by remarking that the fluctuations associated with Rusanov's solver
-can be expressed in terms of waves and speeds from Roe's solver. To see this, first consider \eqref{rus_cons}
-and rewrite the right-going fluctuation as follows:
-\begin{align}\label{rus_rs_aux}
-  \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
-  = \bff(\bfu_{i})-\bff(\bfu_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
-  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}.
-  %\A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
-\end{align}
-From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get
-$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$.
-Using \eqref{cons_roe}, \eqref{rus_rs_aux} becomes
-\begin{align*}
-  \lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
-  = \sum_p \left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
-  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus},
-\end{align*}
-which implies that
-\begin{subequations}\label{fluct_rus}
-\begin{align}
-  \A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p,
-\end{align}
-and similarly,
-\begin{align}
-  \A^{-,\Rus}\Delta Q_{i+1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i+1/2}^p - \lambda_{i+1/2}^{\max}\right)\W_{i+1/2}^p.
-\end{align}
-\end{subequations}
-
-\subsubsection{HLLEMCC solver}
-The HLLEMCC approximate Riemann solver for the shallow water equations was proposed
-in \cite{kemm2014note} specifically to deal with carbuncles.  The idea is to adjust
-the amount of dissipation applied to the shear wave, adding dissipation in regions
-where unphysical carbuncles would appear.  The approximate solution in this
-case consists of four waves; the faster waves use the HLLE wave speeds and carry
-jumps only in the depth and normal momentum, while the slower waves carry jumps
-only in the transverse momentum.  By adjusting the value of $\phi$, this solver
-can behave like the HLLE solver (if $\phi=1$) or like the Roe solver (if $\phi=0$).
-
-The middle states $q^1_m, q^2_m$ are the same as those used for the HLLE solver.
-Requiring conservation of the transverse momentum gives
-\begin{align*}
-  q^3_m = \frac{f^3_l - f^3_r - (\hat{v}-\phi\hat{c})q^3_l + (\hat{v}+\phi\hat{c})q^3_r}{2\phi\hat{c}}.
-\end{align*}
-Here $\hat{v}, \hat{c}$ are the Roe-average normal velocity and sound speed, respectively, whereas the function $\phi$ is given by
-\begin{equation}\label{if}
-\phi(\theta)=\min\left\lbrace 1,\, \left(\varepsilon\,\theta\,\max\left\lbrace0,\,1-Fr_{u}^{\alpha} \right\rbrace \right)   \right\rbrace^{\beta}
-\end{equation}
-where $Fr_{u}$ is the Froude number perpendicular to the cell face and $\varepsilon$, $\alpha$, and $\beta$ are parameters that have to be chosen appropriately.
+%which implies that
+%\begin{subequations}\label{fluct_rus}
+%\begin{align}
+%  \A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p,
+%\end{align}
+%and similarly,
+%\begin{align}
+%  \A^{-,\Rus}\Delta Q_{i+1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i+1/2}^p - \lambda_{i+1/2}^{\max}\right)\W_{i+1/2}^p.
+%\end{align}
+%\end{subequations}
+
+%\subsubsection{HLLEMCC solver}
+%The HLLEMCC approximate Riemann solver for the shallow water equations was proposed
+%in \cite{kemm2014note} specifically to deal with carbuncles.  The idea is to adjust
+%the amount of dissipation applied to the shear wave, adding dissipation in regions
+%where unphysical carbuncles would appear.  The approximate solution in this
+%case consists of four waves; the faster waves use the HLLE wave speeds and carry
+%jumps only in the depth and normal momentum, while the slower waves carry jumps
+%only in the transverse momentum.  By adjusting the value of $\phi$, this solver
+%can behave like the HLLE solver (if $\phi=1$) or like the Roe solver (if $\phi=0$).
+%
+%The middle states $q^1_m, q^2_m$ are the same as those used for the HLLE solver.
+%Requiring conservation of the transverse momentum gives
+%\begin{align*}
+%  q^3_m = \frac{f^3_l - f^3_r - (\hat{v}-\phi\hat{c})q^3_l + (\hat{v}+\phi\hat{c})q^3_r}{2\phi\hat{c}}.
+%\end{align*}
+%Here $\hat{v}, \hat{c}$ are the Roe-average normal velocity and sound speed, respectively, whereas the function $\phi$ is given by
+%\begin{equation}\label{if}
+%\phi(\theta)=\min\left\lbrace 1,\, \left(\varepsilon\,\theta\,\max\left\lbrace0,\,1-Fr_{u}^{\alpha} \right\rbrace \right)   \right\rbrace^{\beta}
+%\end{equation}
+%where $Fr_{u}$ is the Froude number perpendicular to the cell face and $\varepsilon$, $\alpha$, and $\beta$ are parameters that have to be chosen appropriately.
 For the shallow water equations, authors in \cite{kemm2014note} propose to use $\varepsilon=10^{-3}$ and $\alpha=\beta=\frac{1}{3}$.
 The value of $\theta$ is calculated as the 2-norm of the residual in the Rankine-Hugoniot condition for the shear wave.
-
-The HLLEMCC solver follows the ideas of the Harten-Hyman entropy fix by splitting the shear wave with speed $\hat{u}$, into two waves moving with speeds $\hat{u}-\phi\hat{c}$ and $\hat{u}+\phi\hat{c}$ (see figure \ref{HLLEMCC}).
+%
+%The HLLEMCC solver follows the ideas of the Harten-Hyman entropy fix by splitting the shear wave with speed $\hat{u}$, into two waves moving with speeds $\hat{u}-\phi\hat{c}$ and $\hat{u}+\phi\hat{c}$ (see figure \ref{HLLEMCC}).
 Its implementation, in terms of the fluctuations, is performed by using the values $\hat{u}^{\pm}$ calculated with an absolute-value function defined as
-
-\begin{align*}
-\varphi(\lambda)=\frac{\vert \lambda-\phi(\theta)\hat{c}\,\vert+\vert \lambda+\phi(\theta)\hat{c}\,\vert}{2}
-\end{align*}
-
-\begin{figure}
-    \center
-    \includegraphics[width=0.6\textwidth]{figures/hllemcc.pdf}
-    \caption{Structure of the HLLEMCC approximate Riemann solution, with four waves.}
-    \label{HLLEMCC}
-\end{figure}
-
-\section{Entropy-dissipative Riemann solver}\label{sec:ent_diss_rs}
-Consider a hyperbolic system of conservation laws
-\begin{align}\label{cons_law}
-  \frac{\partial \bfu(\bfx)}{\partial t} + \nabla\cdot \bff(\bfu(\bfx))=0, \quad \bfx\in\Omega,
-\end{align}
-with a en entropy pair $\left(\eta(\bfu), \bfq(\bfu)\right)$.
-A weak solution of \eqref{cons_law} is called an entropy solution if
-\begin{align}\label{ent_inequality}
-  \frac{\partial \eta(\bfu(\bfx))}{\partial t} + \nabla\cdot \bfq(\bfu(\bfx))\leq 0, \quad \bfx\in\Omega,
-\end{align}
-holds, in a distribution sense, for any entropy pair.
-It is desirable for a numerical method to satisfy a discrete version of the entropy inequality
-\eqref{ent_inequality}, but many widely-used discretizations do not.
-In this work, we propose an entropy stable approximate Riemann solver for the one dimensional
-shallow water equations. We use this Riemann solver as building block within the LWL
-method, as described in \cite{leveque1997wave,leveque2002finite} and implemented in Clawpack.
-
-\subsection{Entropy residual}
+%
+%\begin{align*}
+%\varphi(\lambda)=\frac{\vert \lambda-\phi(\theta)\hat{c}\,\vert+\vert \lambda+\phi(\theta)\hat{c}\,\vert}{2}
+%\end{align*}
+%
+%\begin{figure}
+%    \center
+%    \includegraphics[width=0.6\textwidth]{figures/hllemcc.pdf}
+%    \caption{Structure of the HLLEMCC approximate Riemann solution, with four waves.}
+%    \label{HLLEMCC}
+%\end{figure}
+
+\subsection{The entropy residual}
 Let $S_i$ be the set of cells containing cell $i$ and those that share a face with it.
@@
 as a measurement of the entropy production in $S_i$.
-To avoid approximating the time derivative of the entropy, we follow
+To avoid the need to compute the time derivative of the entropy, we follow
 \cite{guermond2018second, guermond2018well} and use
@@
 In this section, we use \eqref{Ri}, which is a normalized measurement of the entropy residual \eqref{ent_residual},
-to blend Rusanov's and Roe's Riemann solvers. To do this, we define the following numerical flux:
+to blend Rusanov's and Roe's Riemann solvers. To do this, we define the blended numerical flux:
 \begin{align}\label{num_flux_ev}
   F_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2}
-  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max} I + (1-\theta_{i+1/2})|\hat A_{i+1/2}| \right)(\bfu_{i+1}-\bfu_{i}),
-\end{align}
-and similarly for $F_{i-1/2}$. Here $I\in\mathbb{R}^{3\times 3}$ is the identity matrix,
+  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max} I + (1-\theta_{i+1/2})|\hat A_{i+1/2}| \right)(\bfu_{i+1}-\bfu_{i}).
+\end{align}
+Here $I\in\mathbb{R}^{3\times 3}$ is the identity matrix,
 $\theta_{i+1/2}=\max\{\theta_{i+1},\theta_i\}$, $\lambda_{i+1/2}^{\max}$ is the

### 2020-11-23 numerics_paper/version2/numerical_chj.tex
Put back description of Rusanov in terms of Roe.

@@
 Similarly, if $\theta_{i\pm 1/2}=1$, we recover the Lax-Wendroff method based on Rusanov's Riemann solver,
-which is clear by writing the fluctuations in Rusanov's method in terms of Roe's waves and speeds;
-see \eqref{fluct_rus}.
+which can be seen by writing the fluctuations in Rusanov's method in terms of Roe's waves and speeds.
+
+To see this, first consider \eqref{rus_cons} and rewrite the right-going fluctuation as follows:
+\begin{align}\label{rus_rs_aux}
+  \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+  = \bff(\bfu_{i})-\bff(\bfu_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
+  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}.
+  %\A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
+\end{align}
+From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get
+$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$.
+Using \eqref{cons_roe}, \eqref{rus_rs_aux} becomes
+\begin{align*}
+  \lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+  = \sum_p \left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
+  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus},
+\end{align*}
+which implies that
+\begin{subequations}\label{fluct_rus}
+\begin{align}
+  \A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p,
+\end{align}
+and similarly,
+\begin{align}
+  \A^{-,\Rus}\Delta Q_{i+1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i+1/2}^p - \lambda_{i+1/2}^{\max}\right)\W_{i+1/2}^p.
+\end{align}
+\end{subequations}
 

### 2020-11-23 numerics_paper/version2/numerical_chj.tex

@@
 
+\subsection{Wave propagation methods}\label{sec:waveprop}
 The numerical solution $u_i$
@@
 a full and detailed description of these two Riemann solvers in the context of
-the shallow water equations.
+the shallow water equations.  Both of these solvers satisfy the conservation
+property
+\begin{align} \label{rs_conservation}
+    \amdq_\imh + \apdq_\imh = \bff(\bfu_i) - \bff(\bfu_{i-1}).
+\end{align}
 
@@
 %$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$.
-%Using \eqref{cons_roe}, \eqref{rus_rs_aux} becomes
+%Using \eqref{rs_conservation}, \eqref{rus_rs_aux} becomes
 %\begin{align*}
@@
 $\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$.
-Using \eqref{cons_roe}, \eqref{rus_rs_aux} becomes
+Using \eqref{rs_conservation}, \eqref{rus_rs_aux} becomes
 \begin{align*}
@@
 We compare the behavior of the blended solver against the standard Roe's and Rusanov's solvers.
-In most of the experiments we use these Riemann solvers with the LWL method reviewed in \S\ref{sec:lax-wendroff-leveque}.
+In most of the experiments we use these Riemann solvers with the LWL method reviewed in \S\ref{sec:waveprop}.
 When the exact solution is available, we report a convergence test based on the $L^1$-error for the water height

### 2020-11-25 numerics_paper/version2/numerical_chj.tex
Change notation to avoid potential conflicts.
u -> q
q -> \entflux (\psi)
v -> \entvar (\eta')
\psi -> \efp (entropy flux potential, \Psi for now)

@@
 \newcommand{\bfu}{{\bf u}}
-\newcommand{\bfv}{{\bf v}}
+\newcommand{\entvar}{\eta'}
 \newcommand{\bfq}{{\bf q}}
@@
 \newcommand{\Roe}{{\rm Roe}}
+\newcommand{\efp}{\Psi}
+\newcommand{\entflux}{\psi}
 
@@
 \subsection{Wave propagation methods}\label{sec:waveprop}
-The numerical solution $u_i$
-represents the average value over cell $i$, which extends from $x_\imh$ to $x_\iph$.
+The numerical solution $q_i$ represents the average value of the set of
+conserved quantities over cell $i$, which extends from $x_\imh$ to $x_\iph$.
 At each time step and at each interface $x_\imh$, we approximately solve the
-Riemann problem with initial states $(u^n_{i-1},u^n_i)$.  The approximate
+Riemann problem with initial states $(q^n_{i-1},q^n_i)$.  The approximate
 solution consists of a set of traveling discontinuities or waves $\W^p_\imh$
@@
 \begin{align}\label{second-order_via_fluct}
-  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
+  q_i^{n+1} = q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
   -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
@@
 \begin{align}\label{first-order_via_fluct}
-  u_i^{n+1} = u_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
+  q_i^{n+1} = q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
 \end{align}
@@
 \begin{align}\label{first-order_FV}
-  \bfu_i^{n+1}=\bfu_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
+  q_i^{n+1}=q_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
 \end{align}
@@
 \begin{align} \label{rs_conservation}
-    \amdq_\imh + \apdq_\imh = \bff(\bfu_i) - \bff(\bfu_{i-1}).
+    \amdq_\imh + \apdq_\imh = \bff(\bfq_i) - \bff(\bfq_{i-1}).
 \end{align}
@@
 \begin{align}\label{rusanov_waves}
-  \W_{i-1/2}^{1,\Rus}=\bfu_{i-1/2}-\bfu_{i-1}, \qquad
-  \W_{i-1/2}^{2,\Rus}=\bfu_{i}-\bfu_{i-1/2},
+  \W_{i-1/2}^{1,\Rus}=\bfq_{i-1/2}-\bfq_{i-1}, \qquad
+  \W_{i-1/2}^{2,\Rus}=\bfq_{i}-\bfq_{i-1/2},
 \end{align}
@@
 \begin{align}\label{rus_cons}
-  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{1,\Rus}+\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus} = \bff(\bfu_{i})-\bff(\bfu_{i-1}).
+  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{1,\Rus}+\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus} = \bff(\bfq_{i})-\bff(\bfq_{i-1}).
 \end{align}
@@
   \end{bmatrix}
-  =\Delta Q_{i-1/2}:=\bfu_i-\bfu_{i-1}.
+  =\Delta Q_{i-1/2}:=q_i-q_{i-1}.
 \end{align}
@@
 %  \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
-%  = \bff(\bfu_{i})-\bff(\bfu_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
+%  = \bff(q_{i})-\bff(q_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
 %  -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}.
@@
 %From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get
-%$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$.
+%$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=q_i-q_{i-1}=\sum_p\W_{i-1/2}^p$.
 %Using \eqref{rs_conservation}, \eqref{rus_rs_aux} becomes
@@
 \subsection{The entropy residual}
-Let $S_i$ be the set of cells containing cell $i$ and those that share a face with it.
-Also let $\bfv(\bfu)=\eta^\prime(\bfu)$ be the entropy variable.
+Let $\entvar(\bfq)$ and $\entflux(\bfq)$ denote the entropy variable and entropy flux.
 Based on \cite{guermond2011entropy}, we consider
 \begin{align}\label{ent_residual}
-  \int_{S_i} \bfv(\bfu) \cdot \left[ \frac{\partial \bfu}{\partial t} + \nabla\cdot \bff(\bfu)\right]d\bfx
-  =\int_{S_i} \left[\frac{\partial \eta(\bfu)}{\partial t} + \bfv(\bfu) \cdot \nabla\cdot \bff(\bfu)\right] d\bfx
-\end{align}
-as a measurement of the entropy production in $S_i$.
+  \int_{S_i} \entvar(\bfq) \cdot \left[ \frac{\partial \bfq}{\partial t} + \nabla\cdot \bff(\bfq)\right]d\bfx
+  =\int_{S_i} \left[\frac{\partial \eta(\bfq)}{\partial t} + \entvar(\bfq) \cdot \nabla\cdot \bff(\bfq)\right] d\bfx
+\end{align}
+as a measurement of the entropy production in cell $i$ (here $S_i=[x_\imh,x_\iph]$.
 To avoid the need to compute the time derivative of the entropy, we follow
 \cite{guermond2018second, guermond2018well} and use
-$\int \frac{\partial\eta(\bfu)}{\partial t}d\bfx=-\int \nabla\cdot\bfq(\bfu) d\bfx$,
+$\int \frac{\partial\eta(\bfq)}{\partial t}d\bfx=-\int \nabla\cdot\entflux(\bfq) d\bfx$,
 which holds in smooth regions. Then we define
@@
   R_i=
-  \left|\int_{S_i} \left[\bfv(\bfu) \cdot \nabla\cdot \bff(\bfu) - \nabla\cdot\bfq(\bfu) \right] d\bfx\right|,
+  \left|\int_{S_i} \left[\entvar(\bfq) \cdot \nabla\cdot \bff(\bfq) - \nabla\cdot\entflux(\bfq) \right] d\bfx\right|,
 \end{align*}
 and $D_i$ being an upper bound of $R_i$ that acts as normalization so that $0\leq \theta_i\leq 1$.
-Note that $R_i\approx 0$ if $\bfu$ is smooth in $S_i$.
+Note that $R_i\approx 0$ if $\bfq$ is smooth in $S_i$.
 In our implementation, we use
 \begin{align*}
-  R_i\approx \left|\bfv(\bfu_i)\cdot \int_{S_i}\nabla\cdot \bff(\bfu)d\bfx
-  -\int_{S_i}\nabla\cdot\bfq(\bfu) d\bfx\right|,
+  R_i\approx \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(\bfq)d\bfx
+  -\int_{S_i}\nabla\cdot\entflux(\bfq) d\bfx\right|,
   \quad
-  %D_i = ||\bfv(\bfu_i)||_{\ell^2}\left|\left|\int_{S_i}\nabla\cdot\bff(\bfu)d\bfx\right|\right|_{\ell^2}
-  %+\left|\int_{S_i}\nabla\cdot\bfq(\bfu)d\bfx\right|.
-  D_i = \sum_{k=1}^{d+1}|\bfv_k(\bfu_i)|\left|\int_{S_i}\left(\nabla\cdot\bff(\bfu)\right)_kd\bfx\right|
-  +\left|\int_{S_i}\nabla\cdot\bfq(\bfu)d\bfx\right|,
+  %D_i = ||\entvar(\bfq_i)||_{\ell^2}\left|\left|\int_{S_i}\nabla\cdot\bff(\bfq)d\bfx\right|\right|_{\ell^2}
+  %+\left|\int_{S_i}\nabla\cdot\bfq(\bfq)d\bfx\right|.
+  D_i = \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\int_{S_i}\left(\nabla\cdot\bff(\bfq)\right)_kd\bfx\right|
+  +\left|\int_{S_i}\nabla\cdot\entflux(\bfq)d\bfx\right|,
 \end{align*}
@@
 \begin{align}\label{num_flux_ev}
-  F_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2}
-  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max} I + (1-\theta_{i+1/2})|\hat A_{i+1/2}| \right)(\bfu_{i+1}-\bfu_{i}).
+  F_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
+  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max} I + (1-\theta_{i+1/2})|\hat A_{i+1/2}| \right)(\bfq_{i+1}-\bfq_{i}).
 \end{align}
@@
 \begin{align}\label{second-order}
-  u_i^{n+1}=u_i^n
+  q_i^{n+1}=q_i^n
   -\frac{\Delta t}{\Delta x}\left[\A^-\Delta Q_{i+1/2}+\A^+\Delta Q_{i-1/2}\right]
@@
 \begin{align*}
-  u_i^{n+1}=u_i^n
+  q_i^{n+1}=q_i^n
   &-\frac{\Delta t}{2 \Delta x}
@@
   \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
-  = \bff(\bfu_{i})-\bff(\bfu_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
+  = \bff(\bfq_{i})-\bff(\bfq_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
   -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}.
@@
 From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get
-$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfu_i-\bfu_{i-1}=\sum_p\W_{i-1/2}^p$.
+$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfq_i-\bfq_{i-1}=\sum_p\W_{i-1/2}^p$.
 Using \eqref{rs_conservation}, \eqref{rus_rs_aux} becomes
@@
 
-Let $\bfv_i=\bfv(\bfu_i)$ and $q_i=q(\bfu_i)$ denote the entropy variable and the
-(one-dimensional) entropy potential at cell $i$, respectively.
-Also let $\psi_i=\bfv(\bfu_i)\cdot \bff(\bfu_i)-q(\bfu_i)$ be the entropy potential at cell $i$.
+Let $\entvar_i=\entvar(\bfq_i)$ and $\entflux_i=\entflux(\bfq_i)$ denote the entropy variable and the
+(one-dimensional) entropy flux at cell $i$, respectively.
+Also let $\efp=\entvar(\bfq_i)\cdot \bff(\bfq_i)-\entflux(\bfq_i)$ be the entropy potential at cell $i$.
 From \cite[\S 4]{tadmor1987numerical}, if the numerical fluxes satisfy
 \begin{align}\label{es_cond}
-(\bfv_{i}-\bfv_{i-1})\cdot \bfF_{i-1/2}\leq \psi_{i}-\psi_{i-1},
+(\entvar_{i}-\entvar_{i-1})\cdot \bfF_{i-1/2}\leq \efp{i}-\efp{i-1},
   \qquad
-  (\bfv_{i+1}-\bfv_i)\cdot \bfF_{i+1/2}\leq \psi_{i+1}-\psi_i,
+  (\entvar_{i+1}-\entvar_i)\cdot \bfF_{i+1/2}\leq \efp{i+1}-\efp,
 \end{align}
@@
 \begin{align}\label{ent_ineq}
-  \frac{d\eta(\bfu_i)}{dt}+\frac{1}{\Delta x}\left[G_{i+1/2}-G_{i-1/2}\right]\leq 0,
+  \frac{d\eta(\bfq_i)}{dt}+\frac{1}{\Delta x}\left[G_{i+1/2}-G_{i-1/2}\right]\leq 0,
 \end{align}
@@
 \begin{align*}
-    G_{i+1/2} &= \frac{1}{2}\left(\bfv_i+\bfv_{i+1}\right)\bfF_{i+1/2}-\frac{1}{2}(\psi_{i}+\psi_{i+1}),
+    G_{i+1/2} &= \frac{1}{2}\left(\entvar_i+\entvar_{i+1}\right)\bfF_{i+1/2}-\frac{1}{2}(\efp{i}+\efp{i+1}),
 \end{align*}
@@
 \begin{align*}
-  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2}
-  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-\theta_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfu_{i+1}-\bfu_{i}).
+  \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
+  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-\theta_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfq_{i+1}-\bfq_{i}).
 \end{align*}
@@
 \begin{align*}
-  \bfF_{i+1/2} = \frac{\bff(\bfu_{i+1})+\bff(\bfu_i)}{2}
+  \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
   - \frac{1}{2}\sum_p(\lambda_{i+1/2}^p+\lambda_{i+1/2}^{\min})\W_{i+1/2}^p.
@@
 \begin{align}\label{lambda_min}
-  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\bfv_{i+1}-\bfv_{i})\cdot\left[\bff(\bfu_{i+1})+\bff(\bfu_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\psi_{i+1}-\psi_i)}{\frac{1}{2}(\bfv_{i+1}-\bfv_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}
+  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\left[\bff(\bfq_{i+1})+\bff(\bfq_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\efp{i+1}-\efp)}{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}
 \end{align}
@@
 \begin{align*}
-  (\bfv_{i}-\bfv_{i-1})\cdot
+  (\entvar_{i}-\entvar_{i-1})\cdot
   \Big(
-  \underbrace{\bff(\bfu_i)-\A^+\Delta Q_{i-1/2}}_{=\bfF_{i-1/2}}\Big)\leq \psi_{i}-\psi_{i-1},
+  \underbrace{\bff(\bfq_i)-\A^+\Delta Q_{i-1/2}}_{=\bfF_{i-1/2}}\Big)\leq \efp{i}-\efp{i-1},
   \quad
-  (\bfv_{i+1}-\bfv_i)\cdot
-  \Big(\underbrace{\bff(\bfu_i)+\A^-\Delta Q_{i+1/2}}_{=\bfF_{i+1/2}}\Big)\leq \psi_{i+1}-\psi_i.
+  (\entvar_{i+1}-\entvar_i)\cdot
+  \Big(\underbrace{\bff(\bfq_i)+\A^-\Delta Q_{i+1/2}}_{=\bfF_{i+1/2}}\Big)\leq \efp{i+1}-\efp.
 \end{align*}
@@
 and $h^e(\bfx_i)$ is the exact water height evaluated at the center of cell $i$.
+Since the methods under consideration are at most second order accurate, comparison
+of cell averages with centered point values is a sufficient way to test their accuracy.
 
@@
 For this problem we use the first-order method \eqref{first-order_via_fluct},
-which helps preventing violations of positivity. Although, only Rusanov's solver guarantees
-positivity, we do not get negative values for the water height with this method and the other Riemann solvers.
+to avoid negative depth values. Although only Rusanov's solver is proven to guarantee
+positivity, we do not get negative values for the water height with any of the first-order methods.
 In \S \ref{sec:steady_outflow} we consider a one-dimensional problem with a smooth steady state solution. We observe the expected

### 2020-11-25 numerics_paper/version2/numerical_chj.tex
Remove boldface characters.
tmp branch to update numerical results

--- 2020-11-23 numerics_paper/version2/numerical_chj.tex
@@
 \newcommand{\Roe}{{\rm Roe}}
+\newcommand{\Fr}{{\rm Fr}}
 
@@
       $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
-      1/50   & 6.21E-04 &  --  & 6.95E-04 &   -- & 5.10E-04 &  --  & 5.95E-04 & --   \\
-      1/100  & 4.70E-04 & 0.40 & 5.27E-04 & 0.40 & 3.53E-04 & 0.52 & 4.06E-04 & 0.55 \\
-      1/200  & 3.91E-04 & 0.26 & 3.88E-04 & 0.44 & 2.38E-04 & 0.56 & 2.83E-04 & 0.52 \\
-      1/400  & 3.16E-04 & 0.30 & 2.64E-04 & 0.55 & 1.51E-04 & 0.65 & 1.89E-04 & 0.58 \\
-      1/800  & 2.35E-04 & 0.42 & 1.67E-04 & 0.66 & 9.43E-05 & 0.68 & 1.20E-04 & 0.65 \\
-      1/1600 & 2.00E-04 & 0.23 & 1.01E-04 & 0.72 & 5.70E-05 & 0.72 & 7.42E-05 & 0.68 \\ \hline
+      1/50   & 6.21E-04 &  --  & 6.95E-04 &   -- & 5.08E-04 & --   & 5.95E-04 & --   \\
+      1/100  & 4.70E-04 & 0.40 & 5.27E-04 & 0.40 & 3.39E-04 & 0.58 & 4.06E-04 & 0.55 \\
+      1/200  & 3.91E-04 & 0.26 & 3.88E-04 & 0.44 & 2.30E-04 & 0.56 & 2.83E-04 & 0.52 \\
+      1/400  & 3.16E-04 & 0.30 & 2.64E-04 & 0.55 & 1.49E-04 & 0.62 & 1.89E-04 & 0.58 \\
+      1/800  & 2.35E-04 & 0.42 & 1.67E-04 & 0.66 & 9.24E-05 & 0.68 & 1.20E-04 & 0.65 \\
+      1/1600 & 2.00E-04 & 0.23 & 1.01E-04 & 0.72 & 5.66E-05 & 0.70 & 7.42E-05 & 0.68 \\ \hline
     \end{tabular}
@@
       $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
-      1/50   & 4.22E-04 &  --  & 5.73E-04 &  --  & 4.34E-04 & --   \\
-      1/100  & 2.00E-04 & 1.07 & 3.35E-04 & 0.78 & 2.08E-04 & 1.05 \\
-      1/200  & 1.11E-04 & 0.85 & 2.01E-04 & 0.74 & 1.13E-04 & 0.88 \\
-      1/400  & 5.05E-05 & 1.13 & 1.18E-04 & 0.77 & 5.26E-05 & 1.10 \\
-      1/800  & 2.60E-05 & 0.95 & 7.08E-05 & 0.74 & 2.73E-05 & 0.94 \\
-      1/1600 & 1.29E-05 & 1.01 & 3.79E-05 & 0.90 & 1.32E-05 & 1.04 \\ \hline
+      1/50   & 4.22E-04 &  --  & 5.73E-04 &  --  & 4.24E-04 &  --  \\ 
+      1/100  & 2.00E-04 & 1.07 & 3.35E-04 & 0.78 & 1.99E-04 & 1.08 \\
+      1/200  & 1.11E-04 & 0.85 & 2.01E-04 & 0.74 & 1.10E-04 & 0.84 \\
+      1/400  & 5.05E-05 & 1.13 & 1.18E-04 & 0.77 & 5.07E-05 & 1.12 \\
+      1/800  & 2.60E-05 & 0.95 & 7.08E-05 & 0.74 & 2.61E-05 & 0.95 \\
+      1/1600 & 1.29E-05 & 1.01 & 3.79E-05 & 0.90 & 1.29E-05 & 1.01 \\ \hline
     \end{tabular}
@@
 
-\subsection{The circular hydraulic jump in one-dimension}\label{sec:1D_chj}
-In this section we consider for the first time a CHJ. We start with a one-dimensional problem
-embedded into the two-dimensional domain $\Omega=(0.1,1)\times(0,0.1)$.
-A semi-analytical solution at steady state is available as described in \S \ref{sec:steady_chj};
-therefore, we can perform a convergence test.
-We consider the initial condition $h(x,t=0)=0.1$, $hu(x,t=0)=0$ and $hv(x,t=0)=0$.
-At the left boundary, we impose $h(0.1,t)=0.3$, $hu(0.1,t)=0.225$ and $hv(0.1,t)=0$.
-At the right boundary we impose $h(1,t)=0.37387387318873766$, $hu(1,t)=\beta/h(1,t)$ and $hv(1,t)=0$ where $\beta=0.0225$.
-The values on the right boundary are chosen so that the CHJ is located at $x=0.3$, as described in \S \ref{sec:formation}.
-%
-In the first row of Figure \ref{fig:rp_wet_bed}, we show (slices of) the solution (at $y=0.05$)
-at different times using the second-order method \eqref{second-order_via_fluct} with Roe's, Rusanov's and the blended Riemann solvers.
-In the second row of the same figure, we plot $\theta_i$ \eqref{Ri}, which is a (normalized) measurement of the
-entropy residual. Note that $\theta_i$ accurately locates the shocks and is (nearly) zero in smooth regions.
-In Table \ref{table:chj_1D}, we summarize the results of a convergence study.
-We remark that the convergence with Roe's solver is erratic.
-Moreover, for the finer grids, the solution appears to become worse.
-Using Rusanov's and the blended Riemann solvers delivers cleaner convergence rates.
-In \cite[\S 5]{kemm2018heuristical} and \cite{bader2014carbuncle}, the authors consider one-dimensional problems embedded into
-two-dimensional domains and observe different instabilities for the
-Euler and the shallow water equations, respectively.
-The authors attributed the problem to a manifestation of the carbuncle instability
-that lead to an unstable location of the steady shock with certain Riemann solvers, including Roe's solver.
-
-\begin{figure}[!h]
-  {\scriptsize
-    \includegraphics[scale=0.29]{figures/chj_1D_t0p5.png}
-    \qquad
-    \includegraphics[scale=0.29]{figures/chj_1D_t1p3.png}
-    \qquad
-    \includegraphics[scale=0.29]{figures/chj_1D_t10p0.png}
-
-    \vspace{15pt}
-    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t0p5.png}
-    \qquad
-    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t1p3.png}
-    \qquad
-    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t10p0.png}
-  }
-  \caption{
-    One-dimensional CHJ using method \eqref{second-order_via_fluct} with
-    (cyan) Roe's, (red) Rusanov's and (blue) the blended Riemann solvers.
-    In the second row we plot $\theta_i$, which is given by \eqref{Ri}.
-    In all simulations we consider $\Delta x=1/400$.
-    \label{fig:1D_chj}}
-\end{figure}
-
-\begin{table}[!ht]\scriptsize
-  \begin{center}
-    \begin{tabular}{||c||c|c||c|c||c|c||} \hline
-      & \multicolumn{2}{c||}{Roe's solver}
-      &\multicolumn{2}{c||}{Rusanov's solver}
-      &\multicolumn{2}{c||}{Blended solver} \\ \cline{2-7}
-      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
-      1/50  & 1.56E-04 &  --   & 8.80E-04 &  --  & 3.56E-04 &  --  \\
-      1/100 & 1.06E-04 & 0.55  & 5.09E-04 & 0.79 & 1.95E-04 & 0.86 \\
-      1/200 & 7.77E-05 & 0.45  & 2.96E-04 & 0.78 & 1.07E-04 & 0.87 \\
-      1/400 & 7.50E-06 & 3.37  & 1.49E-04 & 0.99 & 3.68E-05 & 1.53 \\
-      1/800 & 9.07E-06 & -0.27 & 7.98E-05 & 0.90 & 1.67E-05 & 1.14 \\ \hline
-    \end{tabular}
-    \caption{Grid convergence study for the one-dimensional CHJ at $t=10$
-      using method \eqref{second-order_via_fluct} with different Riemann solvers.\label{table:chj_1D}}
-  \end{center}
-\end{table}
+%% \subsection{The circular hydraulic jump in one-dimension}\label{sec:1D_chj}
+%% In this section we consider for the first time a CHJ. We start with a one-dimensional problem
+%% embedded into the two-dimensional domain $\Omega=(0.1,1)\times(0,0.1)$.
+%% A semi-analytical solution at steady state is available as described in \S \ref{sec:steady_chj};
+%% therefore, we can perform a convergence test.
+%% We consider the initial condition $h(x,t=0)=0.1$, $hu(x,t=0)=0$ and $hv(x,t=0)=0$.
+%% At the left boundary, we impose $h(0.1,t)=0.3$, $hu(0.1,t)=0.225$ and $hv(0.1,t)=0$.
+%% At the right boundary we impose $h(1,t)=0.37387387318873766$, $hu(1,t)=\beta/h(1,t)$ and $hv(1,t)=0$ where $\beta=0.0225$.
+%% The values on the right boundary are chosen so that the CHJ is located at $x=0.3$, as described in \S \ref{sec:formation}.
+%% %
+%% In the first row of Figure \ref{fig:rp_wet_bed}, we show (slices of) the solution (at $y=0.05$)
+%% at different times using the second-order method \eqref{second-order_via_fluct} with Roe's, Rusanov's and the blended Riemann solvers.
+%% In the second row of the same figure, we plot $\theta_i$ \eqref{Ri}, which is a (normalized) measurement of the
+%% entropy residual. Note that $\theta_i$ accurately locates the shocks and is (nearly) zero in smooth regions.
+%% In Table \ref{table:chj_1D}, we summarize the results of a convergence study.
+%% We remark that the convergence with Roe's solver is erratic.
+%% Moreover, for the finer grids, the solution appears to become worse.
+%% Using Rusanov's and the blended Riemann solvers delivers cleaner convergence rates.
+%% In \cite[\S 5]{kemm2018heuristical} and \cite{bader2014carbuncle}, the authors consider one-dimensional problems embedded into
+%% two-dimensional domains and observe different instabilities for the
+%% Euler and the shallow water equations, respectively.
+%% The authors attributed the problem to a manifestation of the carbuncle instability
+%% that lead to an unstable location of the steady shock with certain Riemann solvers, including Roe's solver.
+
+%% \begin{figure}[!h]
+%%   {\scriptsize
+%%     \includegraphics[scale=0.29]{figures/chj_1D_t0p5.png}
+%%     \qquad
+%%     \includegraphics[scale=0.29]{figures/chj_1D_t1p3.png}
+%%     \qquad
+%%     \includegraphics[scale=0.29]{figures/chj_1D_t10p0.png}
+
+%%     \vspace{15pt}
+%%     \includegraphics[scale=0.29]{figures/chj_1D_Ri_t0p5.png}
+%%     \qquad
+%%     \includegraphics[scale=0.29]{figures/chj_1D_Ri_t1p3.png}
+%%     \qquad
+%%     \includegraphics[scale=0.29]{figures/chj_1D_Ri_t10p0.png}
+%%   }
+%%   \caption{
+%%     One-dimensional CHJ using method \eqref{second-order_via_fluct} with
+%%     (cyan) Roe's, (red) Rusanov's and (blue) the blended Riemann solvers.
+%%     In the second row we plot $\theta_i$, which is given by \eqref{Ri}.
+%%     In all simulations we consider $\Delta x=1/400$.
+%%     \label{fig:1D_chj}}
+%% \end{figure}
+
+%% \begin{table}[!ht]\scriptsize
+%%   \begin{center}
+%%     \begin{tabular}{||c||c|c||c|c||c|c||} \hline
+%%       & \multicolumn{2}{c||}{Roe's solver}
+%%       &\multicolumn{2}{c||}{Rusanov's solver}
+%%       &\multicolumn{2}{c||}{Blended solver} \\ \cline{2-7}
+%%       $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
+%%       1/50  & 1.56E-04 &  --   & 8.80E-04 &  --  & 2.48E-04 & -- \\
+%%       1/100 & 1.06E-04 & 0.55  & 5.09E-04 & 0.79 & 1.70E-04 & 0.54 \\
+%%       1/200 & 7.77E-05 & 0.45  & 2.96E-04 & 0.78 & 9.24E-05 & 0.87 \\
+%%       1/400 & 7.50E-06 & 3.37  & 1.49E-04 & 0.99 & 1.70E-05 & 2.44 \\
+%%       1/800 & 9.07E-06 & -0.27 & 7.98E-05 & 0.90 & 1.28E-05 & 0.40 \\ \hline
+%%     \end{tabular}
+%%     \caption{Grid convergence study for the one-dimensional CHJ at $t=10$
+%%       using method \eqref{second-order_via_fluct} with different Riemann solvers.\label{table:chj_1D}}
+%%   \end{center}
+%% \end{table}
 
 \subsection{Steady outflow}\label{sec:steady_outflow}
-Let us consider a problem similar to that in the previous section, but with an outflow boundary condition.
-That is, we consider the one-dimensional domain $\Omega=(0.1,1)$, the initial condition
-$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$,
-and set the right boundary condition to outflow.
+Let us consider a one-dimensional problem with a constant inflow and a steady outflow.
+The domain is given by $\Omega=(0.1,1)$, the initial condition is
+$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$,
+and the right boundary condition is set to outflow.
 We refer to \cite[\S 21.8.5]{leveque2002finite} for details about the outflow boundary condition.
-If the flow is supercritical ($|F|>1$ where $F$ is the Froude number), the exact solution converges to a
+If the flow is supercritical ($|\Fr|>1$ where $\Fr$ is the Froude number), the exact solution converges to a
 steady state profile given by the solution of \eqref{steady}.
@@
       $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate\\ \hline
-      1/50   & 2.16E-4 &  --  & 9.97E-4 &   -- & 2.20E-4 & --   & 4.13E-5 &   -- & 7.32E-4 &  --  & 4.97E-5 & --   \\
-      1/100  & 1.14E-4 & 0.92 & 6.38E-4 & 0.64 & 1.16E-4 & 0.92 & 1.47E-5 & 1.48 & 4.43E-4 & 0.73 & 1.89E-5 & 1.39 \\
-      1/200  & 5.86E-5 & 0.95 & 3.88E-4 & 0.72 & 5.93E-5 & 0.96 & 4.81E-6 & 1.61 & 2.57E-4 & 0.79 & 6.36E-6 & 1.57 \\
-      1/400  & 2.98E-5 & 0.97 & 2.26E-4 & 0.78 & 3.00E-5 & 0.98 & 1.42E-6 & 1.76 & 1.44E-4 & 0.83 & 1.92E-6 & 1.72 \\
-      1/800  & 1.50E-5 & 0.98 & 1.27E-4 & 0.83 & 1.51E-5 & 0.99 & 4.16E-7 & 1.76 & 7.81E-5 & 0.88 & 5.61E-7 & 1.77 \\
-      1/1600 & 7.53E-6 & 0.99 & 6.90E-5 & 0.88 & 7.55E-6 & 0.99 & 1.14E-7 & 1.86 & 4.12E-5 & 0.92 & 1.54E-7 & 1.86 \\ \hline
+      1/50   & 2.16E-4 &  --  & 9.97E-4 &   -- & 2.16E-4 & --   & 4.13E-5 &   -- & 7.32E-4 &  --  & 4.15E-5 & --   \\
+      1/100  & 1.14E-4 & 0.92 & 6.38E-4 & 0.64 & 1.14E-4 & 0.92 & 1.47E-5 & 1.48 & 4.43E-4 & 0.73 & 1.48E-5 & 1.48 \\
+      1/200  & 5.86E-5 & 0.95 & 3.88E-4 & 0.72 & 5.87E-5 & 0.95 & 4.81E-6 & 1.61 & 2.57E-4 & 0.79 & 4.83E-6 & 1.61 \\
+      1/400  & 2.98E-5 & 0.97 & 2.26E-4 & 0.78 & 2.98E-5 & 0.97 & 1.42E-6 & 1.76 & 1.44E-4 & 0.83 & 1.42E-6 & 1.76 \\
+      1/800  & 1.50E-5 & 0.98 & 1.27E-4 & 0.83 & 1.50E-5 & 0.98 & 4.16E-7 & 1.76 & 7.81E-5 & 0.88 & 4.18E-7 & 1.76 \\
+      1/1600 & 7.53E-6 & 0.99 & 6.90E-5 & 0.88 & 7.53E-6 & 0.99 & 1.14E-7 & 1.86 & 4.12E-5 & 0.92 & 1.15E-7 & 1.86 \\ \hline
     \end{tabular}
@@
 
-%\clearpage
 \subsection{Steady shock around a cylinder}\label{sec:bow_shock}
+
 \subsection{The circular hydraulic jump in two dimensions}\label{sec:2D_chj}
-%We consider two regimes. In the first regime, the physical instabilities are mild.
-%By increasing the strength of the jump and other parameters, we consider a second
-%and more strongly unstable regime. In both cases, we generate the CHJ as described in
-%Section XX; that is, we impose supersonic inflow boundary conditions for the jet and
-%subsonic outflow boundary conditions.
-
-%We consider two types of domain and two corresponding computational grids.
-%The two types of domains are two-dimensional squares and circular domains.
-%For both cases we use structured grids. See an example of these domains and grids in Figures XX and XX,
-%respectively.
+Let us finally consider the numerical experiments for the CHJ in two-dimensions.
+The domain is given by the annulus $\Omega=\{(x,y)~|~ 0.1\leq\sqrt{x^2+y^2}\leq 1\}$.
+The computational mesh is a logically quadrilateral mesh, see Figure \ref{fig:mesh_chj}.
+In all the experiments in this section, the mesh consists of $10^6$ cells. 
+Let
+\begin{align*}
+  \partial\Omega_I:=\{(x,y)\in\Omega ~|~ \sqrt{x^2+y^2}=0.1\}, \qquad \mbox{ and } \qquad
+  \partial\Omega_o:=\{(x,y)\in\Omega ~|~ \sqrt{x^2+y^2}=1\},
+\end{align*}
+denote the inflow and outflow boundaries, respectively. 
+The boundary conditions are 
+\begin{subequations}\label{bcs_chj}
+  \begin{align}
+  h(x,y,t)=h_I, \quad
+  u(x,y,t)=|\bfu_I| \left(\frac{x}{r_I}\right), \quad
+  v(x,y,t)=|\bfu_I| \left(\frac{y}{r_I}\right), \qquad \forall (x,y)\in\partial\Omega_I, \\
+  %
+  h(x,y,t)=h_o, \quad 
+  u(x,y,t)= \frac{\beta }{r_o h}\left(\frac{x}{r_o}\right), \quad
+  v(x,y,t)= \frac{\beta }{r_o h}\left(\frac{y}{r_o}\right), \qquad \forall (x,y)\in\partial\Omega_o, \label{bcs_chj_outflow}
+  \end{align}
+\end{subequations}
+where $h_I=0.3$, $r_I=0.1$, $r_o=1$ and $\beta=r_Ih_I|\bfu_I|$.
+We choose $|\bfu_I|$ and $h_o$ to generate a radially symmetric shock at $r_s:=\sqrt{x^2+y^2}=0.3$, as described in \S \ref{sec:steady_chj}.
+By selecting the boundary conditions $|\bfu_I|$ and $h_o$ we can move through different regimes.
+In particular, we consider two cases.
+For the first case, the Froude number at the inflow boundary is $\Fr\approx 1.3693$.
+In this case, the CHJ appears to be nearly stable.
+However, using Roe's solver leads to the formation of carbuncle instabilities.
+%
+The second regime is a more aggresive regime with a Froude number of $\Fr\approx 27.3861$ at the inflow boundary. 
+In this regime, using Roe's solver also leads to carbuncle instabilities. In addition, 
+due to the large Froude number, random perturbations might trigger physical instabilities. 
+If the random perturbations model or mimic real disturbances in a physical experiment, it is then
+desirable to properly simulate the unstable behavior of the flow without developing carbuncle instabilities. 
+Therefore, the numerical method must introduce a carbuncle fix without dissipating important features of the
+solution that could lead to physical instabilities. 
+For this reason, we believe this problem is more challenging than the standard bow shock problem.
+
+\begin{figure}[!h]
+  \centering 
+  \includegraphics[scale=0.15]{figures/mesh.png}
+  \caption{
+    Example of a grid for the problem with the two-dimensional CHJ.
+    \label{fig:mesh_chj}}
+\end{figure}
+
+\subsubsection{Regime I}
+In this case, the boundary conditions are given by \eqref{bcs_chj} with
+\begin{align}\label{bcs_chj_r1}
+  |\bfu_I|=0.75, \qquad h_o=0.37387387318873766. 
+\end{align}
+The initial condition is a symmetric CHJ located at $\sqrt{x^2+y^2}=0.3$, generated as described in \S \ref{sec:steady_chj}.
+In Figure \ref{fig:chj_r1}, we show the magnitude of the momentum at different times using
+method \eqref{second-order_via_fluct} with Roe's Riemann solver.
+The solution clearly develops carbuncle instabilities.
+We could fix these instabilities using Rusanov's, Kemm's or the blended Riemann solvers. 
+In Figure \ref{fig:chj_r1_diff}, we show the corresponding solutions at $t=3$.
+For this test case, all solvers give qualitatively similar solutions;
+however, we expect Rusanov's solver to be less accurate.
+
+\begin{figure}[!h]
+  \centering 
+  \includegraphics[scale=0.28]{figures/chj_r1_roe_p1.png} \quad
+  \includegraphics[scale=0.28]{figures/chj_r1_roe_p2.png} \quad
+  \includegraphics[scale=0.28]{figures/chj_r1_roe_p3.png} \quad
+  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
+    \eqref{bcs_chj} and \eqref{bcs_chj_r1}. 
+    We consider method \eqref{second-order_via_fluct} with Roe's solver.    
+    The color scale in the plots go from (blue) $0.0746$ to (red) $0.3738$,
+    which corresponds to the minimum and maximum values of the momentum at $t=0$.
+    \label{fig:chj_r1}}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering 
+  \includegraphics[scale=0.28]{figures/chj_r1_llf_p3.png} \quad
+  \includegraphics[scale=0.5]{figures/cross.png} \quad
+  \includegraphics[scale=0.28]{figures/chj_r1_blended_p3.png} \quad
+  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
+    \eqref{bcs_chj} and \eqref{bcs_chj_r1}. 
+    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
+    (middle) Kemm's and (right) the blended solvers. 
+    The color scale in the plots go from (blue) $0.0746$ to (red) $0.3738$,
+    which corresponds to the minimum and maximum values of the momentum at $t=0$.
+    \label{fig:chj_r1_diff}}
+\end{figure}
+
+\subsubsection{Regime II}
+We now consider a more aggresive regime. The boundary conditions are given by \eqref{bcs_chj} with 
+\begin{align}\label{bcs_chj_r2}
+  |\bfu_I|=15, \qquad h_o=6.6845019298155357.
+\end{align}
+Similar than before, we follow \S \ref{sec:steady_chj} to generate an initial condition with a symmetric CHJ located at
+$\sqrt{x^2+y^2}=0.3$. In Figure XX, we show the magnitude of the momentum at different times using method
+\eqref{second-order_via_fluct} with Roe's Riemann solver. Just like before, the solution develops
+carbuncle instabilities. Let us now use Rusanov's, Kemm's and the blended solvers to fix the
+carbuncle instabilities. In Figure XX, we show the corresponding solutions at $t=XX$.
+{\color{red} TODO: finish this part depending on the results with Kemm's solver.}
+%
+Opposed to what we observed in the first test case, in this situation the Froude number is
+high enough that numerical perturbations start to trigger a response around the CHJ.
+For the resolution that we consider, this behavior is dissipated by the extra numerical viscosity introduced by
+Rusanov's solver. In Figure XX, we show the solution at a later time considering the same solvers as before.
+Clearly, using the Rusanov's solver dissipated the physical response to these numerical perturbations.
+
+\begin{figure}[!h]
+  \centering 
+  \includegraphics[scale=0.28]{figures/chj_r2_roe_p1.png} \quad
+  \includegraphics[scale=0.28]{figures/chj_r2_roe_p2.png} \quad
+  \includegraphics[scale=0.28]{figures/chj_r2_roe_p3.png}
+  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
+    \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
+    We consider method \eqref{second-order_via_fluct} with Roe's solver.
+    The color scale in the plots go from (blue) $0.4502$ to (red) $4.4798$,
+    which corresponds to the minimum and maximum values of the momentum at $t=0$.
+    \label{fig:chj_r2}}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering 
+  \includegraphics[scale=0.28]{figures/chj_r2_llf_t0p1.png} \quad
+  \includegraphics[scale=0.5]{figures/cross.png} \quad
+  \includegraphics[scale=0.28]{figures/chj_r2_blended_t0p1.png} 
+  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
+    \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
+    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
+    (middle) Kemm's and (right) the blended solvers. 
+    The color scale in the plots go from (blue) $0.4502$ to (red) $4.4798$,
+    which corresponds to the minimum and maximum values of the momentum at $t=0$.
+    \label{fig:chj_r2}}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering 
+  \includegraphics[scale=0.5]{figures/cross.png} \quad
+  \includegraphics[scale=0.5]{figures/cross.png} \quad
+  \includegraphics[scale=0.28]{figures/chj_r2_blended_t5p0.png} 
+  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
+    \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
+    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
+    (middle) Kemm's and (right) the blended solvers. 
+    The color scale in the plots go from (blue) $0.4502$ to (red) $4.4798$,
+    which corresponds to the minimum and maximum values of the momentum at $t=0$.
+    \label{fig:chj_r2}}
+\end{figure}
+
+\subsubsection{Regime II with random perturbations}
+Finally, we introduce a physical random perturbation to the inflow boundary condition.
+Let $\epsilon(x,y,t)\in[-0.01,0.01]$ be a random number for a given location $(x,y)$,
+at a given time $t$. 
+The inflow boundary condition is now given by
+\begin{subequations}
+  \begin{align}
+    h(x,y,t)=\tilde h_I, \quad
+    u(x,y,t)=|\tilde \bfu_I| \left(\frac{x}{r_I}\right), \quad
+    v(x,y,t)=|\tilde \bfu_I| \left(\frac{y}{r_I}\right), \qquad \forall (x,y)\in\partial\Omega_I,
+  \end{align}
+\end{subequations}
+with $\tilde h_I = \frac{h_I}{1+\epsilon(x,y)}$ and $|\tilde\bfu_I|=|\bfu_I|(1+\epsilon(x,y))$.
+Here, just as before, $h_I=0.3$ and $|\bfu_I|=15$. 
+Note that $\beta=r_I\tilde h|\tilde\bfu_I|=r_Ih_I\bfu_I$.
+The outflow boundary condition is given by
+\eqref{bcs_chj_outflow} with $h_o=6.6845019298155357$.
+
+\begin{figure}[!h]
+  \centering
+  \begin{tabular}{cccc}
+    \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t1p0.png} &
+    \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t2p0.png} &
+    \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t4p0.png} &
+    \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t6p0.png}
+  \end{tabular}
+  \caption{XX}
+\end{figure}
 

### 2020-11-27 cylinder.py
add code for cylinder in main folder

@@ -1,0 +1,218 @@
+"""
+Options:
+    - Riemann solver: GeoClaw, Roe, HLLE, HLLEMCC
+"""
+
+from matplotlib import animation
+import matplotlib.pyplot as plt
+from clawpack import pyclaw
+import numpy as np
+from ent_residual import *
+
+tol = 1E-7
+
+# If plotting in a Jupyter notebook, do this:
+#from clawpack.visclaw.JSAnimation import IPython_display
+
+# *********************************************************** #
+# ********** FUNCTIONS TO COMPUTE ENTROPY RESIDUAL ********** #
+# *********************************************************** #
+def b4step(self, state, old_state=None):
+    """
+    Set zero normal boundary inside the cylinder
+    """
+    # get entropy residual #
+    state.get_qbc_from_q(2,self.qbc)
+    Ri = get_ent_residual(state,self.qbc)
+
+    set_Ri = state.problem_data['set_Ri']
+    if set_Ri is None:
+        state.aux[state.num_aux-1,:,:] = Ri[2:-2,2:-2]
+    else:
+        state.aux[state.num_aux-1,:,:] = set_Ri
+    #
+
+    # set values at the cylinder #
+    xc, yc = state.p_centers
+    rc = np.sqrt((xc-40)**2 + (yc-50)**2)
+    in_solid = np.where((rc <= 20.0))
+
+    nx = (xc-40)/(rc+1e-7)
+    ny = (yc-50)/(rc+1e-7)
+    hu=state.q[1,:,:]
+    hv=state.q[2,:,:]
+    
+    hu_tan = hu*ny - hv*nx
+    state.q[1,in_solid[0],in_solid[1]] =  ny[in_solid[0],in_solid[1]] * hu_tan[in_solid[0],in_solid[1]]
+    state.q[2,in_solid[0],in_solid[1]] = -nx[in_solid[0],in_solid[1]] * hu_tan[in_solid[0],in_solid[1]]
+#
+
+def inflow_bcs(state, dim, _, qbc, __, num_ghost):
+    qbc[0,:num_ghost,:] = state.problem_data['h0']
+    qbc[1,:num_ghost,:] = state.problem_data['u0']
+    qbc[2,:num_ghost,:] = 0.0
+#
+
+def setup(
+        use_dmin_blended = 1.0,
+        set_Ri = None,
+        h0=2.0,
+        u0=20.0,
+        g=1.,
+        x_num_cells=160,
+        y_num_cells=400,
+        tfinal=10,
[…]

### 2020-11-27 plot.py

@@ -1,0 +1,125 @@
+from clawpack.petclaw.solution import Solution
+#from petclaw.io.petsc import read_petsc
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pylab as pl
+from matplotlib import rc
+#rc('text', usetex=True)
+import numpy as np
+import os
+
+def plot_q(frame,
+           bathymetry=False,
+           file_prefix='claw',
+           path='./_output/',
+           plot_ent_residual=False,
+           plot_pcolor=True,
+           hLim=None,
+           momMagLim=None,
+           plot_slices=True):
+    import sys
+    sys.path.append('.')
+
+    sol=Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
+    
+    x=sol.state.grid.x.centers; y=sol.state.grid.y.centers
+    mx=len(x); my=len(y)
+
+    h=sol.state.q[0,:,:]
+    q1=sol.state.q[1,:,:]
+    q2=sol.state.q[2,:,:]
+    qMag = np.sqrt(q1**2 + q2**2)
+
+    yy,xx = np.meshgrid(y,x)
+
+    if frame < 10:
+        str_frame = "000"+str(frame)
+    elif frame < 100:
+        str_frame = "00"+str(frame)
+    elif frame < 1000:
+        str_frame = "0"+str(frame)
+    else:
+        str_frame = str(frame)
+    #
+
+    grad_h = np.gradient(h)
+    abs_grad_h = np.sqrt(grad_h[0]**2 + grad_h[1]**2)
+    
+    if plot_pcolor:
+        fig = pl.figure(figsize=(4,8))
+        # ***** Water depth ***** #
+        #pl.subplot(121)
+        #pl.pcolormesh(xx,yy,h)
+        #pl.axis('equal')
+        #cb = pl.colorbar();
+        #if hLim is not None:
+        #    pl.clim(hLim[0],hLim[1])
+        #
+        # ***** Magnitude of momentum ***** #
+        #pl.subplot(122)
+        import clawpack.visclaw.colormaps as cm
+        scm = cm.schlieren_colormap()
+        pl.pcolormesh(xx,yy,abs_grad_h,cmap=scm)
+        #pl.pcolormesh(xx,yy,qMag)
+        pl.axis('equal')
+        #cb = pl.colorbar();
+        #pl.yticks([])
+        #if momMagLim is not None:
+        #    pl.clim(momMagLim[0],momMagLim[1])
+        #
[…]

### 2020-11-29 numerics_paper/version2/numerical_chj.tex
Put back missing subscripts on entropy potential.

@@
 The computational mesh is a logically quadrilateral mesh, see Figure \ref{fig:mesh_chj}.
-In all the experiments in this section, the mesh consists of $10^6$ cells. 
+For all the experiments in this section, the mesh consists of $10^6$ cells. 
 Let
@@
 
+For most of the experiments we show a Schlieren plot for the water height.
+That is, we plot $||\nabla h||_{\ell^2}$ with a white-to-black logarithm based colormap.
+For some experiments we also plot the magnitude of the momentum $|h\bfu|=h\sqrt{u^2+v^2}$.
+
 \begin{figure}[!h]
@@
 The initial condition is a symmetric CHJ located at $\sqrt{x^2+y^2}=0.3$, generated as described in \S \ref{sec:steady_chj}.
-In Figure \ref{fig:chj_r1}, we show the magnitude of the momentum at different times using
+In Figure \ref{fig:chj_r1_roe}, we show the solution at different times using
 method \eqref{second-order_via_fluct} with Roe's Riemann solver.
@@
 In Figure \ref{fig:chj_r1_diff}, we show the corresponding solutions at $t=3$.
-For this test case, all solvers give qualitatively similar solutions;
+For this test case, all carbuncle free solvers give qualitatively similar solutions;
 however, we expect Rusanov's solver to be less accurate.
@@
   \centering
-  \includegraphics[scale=0.28]{figures/chj_r1_roe_p1.png} \quad
-  \includegraphics[scale=0.28]{figures/chj_r1_roe_p2.png} \quad
-  \includegraphics[scale=0.28]{figures/chj_r1_roe_p3.png} \quad
-  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
-    \eqref{bcs_chj} and \eqref{bcs_chj_r1}. 
-    We consider method \eqref{second-order_via_fluct} with Roe's solver.    
-    The color scale in the plots go from (blue) $0.0746$ to (red) $0.3738$,
-    which corresponds to the minimum and maximum values of the momentum at $t=0$.
+  \subfloat[Solution at (left) $t=1$, (middle) $t=1.5$ and (right) $t=3$ using Roe's Riemann solver.\label{fig:chj_r1_roe}]{
+    \begin{tabular}{ccc} 
+      \includegraphics[scale=0.3]{figures/chj_r1_roe_p1.png} &
+      \includegraphics[scale=0.3]{figures/chj_r1_roe_p2.png} &
+      \includegraphics[scale=0.3]{figures/chj_r1_roe_p3.png}
+    \end{tabular}
+  }
+
+  \subfloat[Solution at $t=3$ using (left) Rusanov's, (middle) Kemm's and (right) the blended solvers.\label{fig:chj_r1_diff}]{
+    \begin{tabular}{ccc}
+      \includegraphics[scale=0.3]{figures/chj_r1_llf_p3.png} &
+      \includegraphics[scale=0.5]{figures/cross.png} &
+      \includegraphics[scale=0.3]{figures/chj_r1_blended_p3.png}
+    \end{tabular}
+    }
+    \caption{CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
+      We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
+      In all cases we show the Schlieren plot for the water height $h$.
+    %The color scale in the plots go from (blue) $0.0746$ to (red) $0.3738$,
+    %which corresponds to the minimum and maximum values of the momentum at $t=0$.
     \label{fig:chj_r1}}
-\end{figure}
-
-\begin{figure}[!h]
-  \centering 
-  \includegraphics[scale=0.28]{figures/chj_r1_llf_p3.png} \quad
-  \includegraphics[scale=0.5]{figures/cross.png} \quad
-  \includegraphics[scale=0.28]{figures/chj_r1_blended_p3.png} \quad
-  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
-    \eqref{bcs_chj} and \eqref{bcs_chj_r1}. 
-    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
-    (middle) Kemm's and (right) the blended solvers. 
-    The color scale in the plots go from (blue) $0.0746$ to (red) $0.3738$,
-    which corresponds to the minimum and maximum values of the momentum at $t=0$.
-    \label{fig:chj_r1_diff}}
 \end{figure}
@@
 Similar than before, we follow \S \ref{sec:steady_chj} to generate an initial condition with a symmetric CHJ located at
-$\sqrt{x^2+y^2}=0.3$. In Figure XX, we show the magnitude of the momentum at different times using method
+$\sqrt{x^2+y^2}=0.3$. In Figure \ref{fig:chj_r2_roe}, we show the solution at different times using method
 \eqref{second-order_via_fluct} with Roe's Riemann solver. Just like before, the solution develops
 carbuncle instabilities. Let us now use Rusanov's, Kemm's and the blended solvers to fix the
-carbuncle instabilities. In Figure XX, we show the corresponding solutions at $t=XX$.
+carbuncle instabilities. In Figure \ref{fig:chj_r2_diff}, we show the corresponding solutions at $t=0.1$.
 {\color{red} TODO: finish this part depending on the results with Kemm's solver.}
@@
 Opposed to what we observed in the first test case, in this situation the Froude number is
-high enough that numerical perturbations start to trigger a response around the CHJ.
+high enough that numerical perturbations start to trigger a response around the CHJ
+(see the middle and right panels of Figure \ref{fig:chj_r2_diff}).
 For the resolution that we consider, this behavior is dissipated by the extra numerical viscosity introduced by
-Rusanov's solver. In Figure XX, we show the solution at a later time considering the same solvers as before.
-Clearly, using the Rusanov's solver dissipated the physical response to these numerical perturbations.
+Rusanov's solver (see the left panel of Figure \ref{fig:chj_r2_diff}).
+In Figure \ref{fig:chj_r2_later}, we show the solution at a later time considering the same solvers as before.
+This time, we plot not only the Schlieren plot for the water height, but also the magnitude of the momentum. 
+Clearly, using Rusanov's solver dissipates the physical response to these numerical perturbations.
 
@@
   \centering
-  \includegraphics[scale=0.28]{figures/chj_r2_roe_p1.png} \quad
-  \includegraphics[scale=0.28]{figures/chj_r2_roe_p2.png} \quad
-  \includegraphics[scale=0.28]{figures/chj_r2_roe_p3.png}
-  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
-    \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
-    We consider method \eqref{second-order_via_fluct} with Roe's solver.
-    The color scale in the plots go from (blue) $0.4502$ to (red) $4.4798$,
-    which corresponds to the minimum and maximum values of the momentum at $t=0$.
-    \label{fig:chj_r2}}
+  \subfloat[Solution at (left) $t=0.09$, (middle) $t=0.1$ and (right) $t=0.11$ using Roe's Riemann solver.\label{fig:chj_r2_roe}]{
+    \begin{tabular}{ccc}
+      \includegraphics[scale=0.3]{figures/chj_r2_roe_p1.png} &
+      \quad\includegraphics[scale=0.3]{figures/chj_r2_roe_p2.png} \quad &
+      \includegraphics[scale=0.3]{figures/chj_r2_roe_p3.png}
+    \end{tabular}
+  }
+
+  \vspace{10pt}
+  \subfloat[Solution at $t=0.1$ using (left) Rusanov's, (middle) Kemm's and (right) the blended solvers.\label{fig:chj_r2_diff}]{
+    \begin{tabular}{ccc}
+      \includegraphics[scale=0.3]{figures/chj_r2_llf_t0p1.png} &
+      \quad \includegraphics[scale=0.5]{figures/cross.png} \quad &
+      \includegraphics[scale=0.3]{figures/chj_r2_blended_t0p1.png}
+    \end{tabular}
+    }
+  \caption{CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
+    We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
+    In all cases we show the Schlieren plot for the water height $h$.
+    %The color scale in the plots go from (blue) $0.45$ to (red) $4.5$,
+    %which corresponds to the minimum and maximum values of the momentum at $t=0$.
+  }
 \end{figure}
@@
   \centering
-  \includegraphics[scale=0.28]{figures/chj_r2_llf_t0p1.png} \quad
+  \subfloat[$t=5$]{
+    \includegraphics[scale=0.3]{figures/chj_r2_rusanov_t5p0.png} \quad
     \includegraphics[scale=0.5]{figures/cross.png} \quad
-  \includegraphics[scale=0.28]{figures/chj_r2_blended_t0p1.png} 
-  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
-    \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
+    \includegraphics[scale=0.3]{figures/chj_r2_blended_t5p0.png}
+  }
+  \caption{CHJ at $t=5$ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
     We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
     (middle) Kemm's and (right) the blended solvers. 
-    The color scale in the plots go from (blue) $0.4502$ to (red) $4.4798$,
-    which corresponds to the minimum and maximum values of the momentum at $t=0$.
-    \label{fig:chj_r2}}
-\end{figure}
-
-\begin{figure}[!h]
-  \centering 
-  \includegraphics[scale=0.5]{figures/cross.png} \quad
-  \includegraphics[scale=0.5]{figures/cross.png} \quad
-  \includegraphics[scale=0.28]{figures/chj_r2_blended_t5p0.png} 
-  \caption{Magnitude of the momentum for a CHJ with boundary conditions given by
-    \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
-    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
-    (middle) Kemm's and (right) the blended solvers. 
-    The color scale in the plots go from (blue) $0.4502$ to (red) $4.4798$,
-    which corresponds to the minimum and maximum values of the momentum at $t=0$.
-    \label{fig:chj_r2}}
+    In all cases we superimpose the Schlieren plot for the water height $h$
+    an the magnitude of the momentum with a uniform (white-to-blue) colormap. 
+    The scale for colormap is (white) $0.45$ to (blue) $4.5$, 
+    which corresponds to the minimum and maximum values at $t=0$.
+    \label{fig:chj_r2_later}}
 \end{figure}
@@
 \end{subequations}
-with $\tilde h_I = \frac{h_I}{1+\epsilon(x,y)}$ and $|\tilde\bfu_I|=|\bfu_I|(1+\epsilon(x,y))$.
+with $\tilde h_I = \frac{h_I}{1+\epsilon(x,y,t)}$ and $|\tilde\bfu_I|=|\bfu_I|(1+\epsilon(x,y,t))$.
 Here, just as before, $h_I=0.3$ and $|\bfu_I|=15$. 
@@
 \eqref{bcs_chj_outflow} with $h_o=6.6845019298155357$.
+In this case, due to the large Froude number, the random perturbations in the inflow
+might trigger physical instabilities, which we want to preserve. At the same time, we
+do not want the solution to develop carbuncle instabilities.
+In figures \ref{fig:chj_r2wp_rusanov} and \ref{fig:chj_r2wp_blended}, we show the solution
+at different times considering method \eqref{second-order_via_fluct} with Rusanov's and
+the blended Riemann solvers, respectively.
 
@@
   \centering
+  \subfloat[Rusanov's Riemann solver\label{fig:chj_r2wp_rusanov}]{
     \begin{tabular}{cccc}
-    \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t1p0.png} &
+      $t=2$ & $t=3$ & $t=4$ & $t=5$ \\
+      \includegraphics[scale=0.24]{figures/chj_r2wp_rusanov_t2p0.png} &
+      \includegraphics[scale=0.24]{figures/chj_r2wp_rusanov_t3p0.png} &
+      \includegraphics[scale=0.24]{figures/chj_r2wp_rusanov_t4p0.png} &
+      \includegraphics[scale=0.24]{figures/chj_r2wp_rusanov_t5p0.png}
+    \end{tabular}
+  }
+
+  \subfloat[Blended Riemann solver\label{fig:chj_r2wp_blended}]{
+    \begin{tabular}{cccc}
       \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t2p0.png} &
+      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t3p0.png} &
       \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t4p0.png} &
-    \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t6p0.png}
+      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t5p0.png}
     \end{tabular}
-  \caption{XX}
+  }
+  \caption{CHJ at different times with a random perturbation at the inflow boundary.
+    We consider method \eqref{second-order_via_fluct} with (a) Roe's and (b) the blended
+    Riemann solvers. In all cases we show the Schlieren plot for the water height $h$.}
 \end{figure}

### 2020-11-30 cylinder.py
fix small prob with inflow_bcs

@@ -49,7 +49,7 @@ def b4step(self, state, old_state=None):
 
 def inflow_bcs(state, dim, _, qbc, __, num_ghost):
     qbc[0,:num_ghost,:] = state.problem_data['h0']
-    qbc[1,:num_ghost,:] = state.problem_data['u0']
+    qbc[1,:num_ghost,:] = state.problem_data['h0']*state.problem_data['u0']
     qbc[2,:num_ghost,:] = 0.0
 #
 

### 2020-11-30 cylinder.py
quick fixes to not use petsc

@@ -22,12 +22,19 @@ def b4step(self, state, old_state=None):
     Set zero normal boundary inside the cylinder
     """
     # get entropy residual #
-    state.get_qbc_from_q(2,self.qbc)
-    Ri = get_ent_residual(state,self.qbc)
-
+    if use_petsc:
+        state.get_qbc_from_q(2,self.qbc)
+        Ri = get_ent_residual(state,self.qbc)
+    else:
+        Ri = get_ent_residual(state)
+    #
+    
     set_Ri = state.problem_data['set_Ri']
     if set_Ri is None:
-        state.aux[state.num_aux-1,:,:] = Ri[2:-2,2:-2]
+        if use_petsc:
+            state.aux[state.num_aux-1,:,:] = Ri[2:-2,2:-2]
+        else:
+            state.aux[state.num_aux-1,:,:] = Ri
     else:
         state.aux[state.num_aux-1,:,:] = set_Ri
     #
@@ -71,7 +78,7 @@ def setup(
         file_prefix='claw',
         kalpha=1./3,
         kbeta=1.3,
-        kepsilon=1.e-2):
+        kepsilon=1.e-3):
     
     from clawpack import riemann
     if use_petsc:
@@ -154,6 +161,7 @@ def setup(
 
         state.problem_data['use_dmin_blended'] = use_dmin_blended
         state.problem_data['set_Ri'] = set_Ri
+        state.problem_data['use_petsc'] = use_petsc
         state.problem_data['grav'] = g   # Gravitational force
         state.problem_data['h0'] = h0
         state.problem_data['u0'] = u0
@@ -189,18 +197,20 @@ if __name__ == "__main__":
     # ***** numerics ***** #
     # ******************** #
     # riemann solver
-    riemann_solver = 'es'
-    #riemann_solver = 'hllemcc'
+    #riemann_solver = 'es'
+    riemann_solver = 'hllemcc'
     
     use_dmin_blended = 1.0
     set_Ri = None
+    use_petsc = False
     
     # mesh
-    refn = 1
+    refn = 0
     mx=int((2**refn)*160)
     my=int((2**refn)*400)
     
     claw = setup(
+        use_petsc=use_petsc,
         use_dmin_blended = use_dmin_blended,
         set_Ri = set_Ri,
         # parameters

### 2020-11-30 plot.py

@@ -1,4 +1,5 @@
-from clawpack.petclaw.solution import Solution
+from clawpack.petclaw.solution import Solution as petclaw_Solution
+from clawpack.pyclaw.solution import Solution as pyclaw_Solution
 #from petclaw.io.petsc import read_petsc
 import matplotlib
 matplotlib.use('Agg')
@@ -9,6 +10,7 @@ import numpy as np
 import os
 
 def plot_q(frame,
+           use_petsc=True,
            bathymetry=False,
            file_prefix='claw',
            path='./_output/',
@@ -20,8 +22,12 @@ def plot_q(frame,
     import sys
     sys.path.append('.')
 
-    sol=Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
-    
+    if use_petsc:
+        sol=petclaw_Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
+    else:
+        file_prefix='fort'
+        sol=pyclaw_Solution(frame,read_aux=False,path=path,file_prefix=file_prefix)
+    #    
     x=sol.state.grid.x.centers; y=sol.state.grid.y.centers
     mx=len(x); my=len(y)
 
@@ -100,14 +106,16 @@ if __name__== "__main__":
     to_frame   = 1000
     frames=xrange(from_frame,to_frame+1)
 
+    use_petsc = False
+    
     ##########################################
     # Get min and max values for color plots #
-    sol=Solution(0,file_format='petsc',read_aux=False,path='./_output/',file_prefix='claw')
-    hMin = sol.state.q[0,:,:].min(); hMax = sol.state.q[0,:,:].max()
-    print (hMin, hMax)
-    q1=sol.state.q[1,:,:]; q2=sol.state.q[2,:,:]
-    momMag = np.sqrt(q1**2 + q2**2)
-    momMagMin = momMag.min(); momMagMax = momMag.max()
+    #sol=Solution(0,file_format='petsc',read_aux=False,path='./_output/',file_prefix='claw')
+    #hMin = sol.state.q[0,:,:].min(); hMax = sol.state.q[0,:,:].max()
+    #print (hMin, hMax)
+    #q1=sol.state.q[1,:,:]; q2=sol.state.q[2,:,:]
+    #momMag = np.sqrt(q1**2 + q2**2)
+    #momMagMin = momMag.min(); momMagMax = momMag.max()
     ###########################################
     
[…]

### 2020-11-30 cylinder.py
Use transverse solver for HLLEMCC.
Also make code run in Python 3.
Initialize velocity in cylinder.py.

@@ -96,18 +96,19 @@ def setup(
             solver.num_eqn = 3
             solver.num_waves = 3
             solver.fwave = False
+            solver.dimensional_split=True
         elif riemann_solver == 'hllemcc':
             solver = pyclaw.ClawSolver2D(shallow_hllemcc_2D)
             solver.num_eqn = 3
             solver.num_waves = 3
             solver.fwave = False
+            solver.dimensional_split=False
+            solver.transverse_waves = 1
         else:
             raise Exception('Unrecognized Riemann solver')
-        solver.dimensional_split=True
         solver.limiters = pyclaw.limiters.tvd.minmod
         solver.cfl_max     = 0.46
         solver.cfl_desired = 0.45
-        solver.transverse_waves = 2
         solver.order = 2
     elif solver_type == 'sharpclaw':
         raise Exception('use solver_type classic for now')
@@ -172,7 +173,7 @@ def setup(
         # INITIAL CONDITIONS #
         nx=xc/(rc+1.e-7); ny=yc/(rc+1.e-7)
         state.q[0,:,:] = state.problem_data['h0']
-        state.q[1,:,:] = 0.0
+        state.q[1,:,:] = state.problem_data['u0']
         state.q[2,:,:] = 0.0
 
         # create clawpack solution

### 2020-11-30 plot.py

@@ -104,7 +104,7 @@ if __name__== "__main__":
     if not os.path.exists('./_plots/slices'): os.mkdir('_plots/slices')
     from_frame = 0
     to_frame   = 1000
-    frames=xrange(from_frame,to_frame+1)
+    frames=range(from_frame,to_frame+1)
 
     use_petsc = False
     

### 2020-11-30 rpt2_shallow_roe_with_efix.f90
Add transverse Roe solver (for HLLEMCC).

@@ -1,0 +1,82 @@
+! =====================================================
+subroutine rpt2(ixy,imp,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,aux1,aux2,aux3,asdq,bmasdq,bpasdq)
+! =====================================================
+    implicit double precision (a-h,o-z)
+
+!     # Riemann solver in the transverse direction for the shallow water
+!     equations .
+!     # Split asdq (= A^* \Delta q, where * = + or -)
+!     # into down-going flux difference bmasdq (= B^- A^* \Delta q)
+!     #    and up-going flux difference bpasdq (= B^+ A^* \Delta q)
+
+    dimension     ql(meqn, 1-mbc:maxm+mbc)
+    dimension     qr(meqn, 1-mbc:maxm+mbc)
+    dimension   asdq(meqn, 1-mbc:maxm+mbc)
+    dimension bmasdq(meqn, 1-mbc:maxm+mbc)
+    dimension bpasdq(meqn, 1-mbc:maxm+mbc)
+
+    dimension waveb(3,3),sb(3)
+
+!   # grav must be set elsewhere
+    common /cparam/ grav
+
+!   # Roe averages quantities of each interface
+    parameter (maxm2 = 1800)
+    double precision u(-6:maxm2+7),v(-6:maxm2+7),a(-6:maxm2+7), &
+                     h(-6:maxm2+7)
+
+    if (ixy == 1) then
+        mu = 2
+        mv = 3
+    else
+        mu = 3
+        mv = 2
+    endif
+            
+    do 10 i = 2-mbc, mx+mbc
+        h(i) = (qr(1,i-1)+ql(1,i))*0.50d0
+        hsqrtl = dsqrt(qr(1,i-1))
+        hsqrtr = dsqrt(ql(1,i))
+        hsq2 = hsqrtl + hsqrtr
+        u(i) = (qr(mu,i-1)/hsqrtl + ql(mu,i)/hsqrtr) / hsq2
+        v(i) = (qr(mv,i-1)/hsqrtl + ql(mv,i)/hsqrtr) / hsq2
+        a(i) = dsqrt(grav*h(i))
+    10 END DO
+
+
+    do 20 i = 2-mbc, mx+mbc
+        a1 = (0.50d0/a(i))*((v(i)+a(i))*asdq(1,i)-asdq(mv,i))
+        a2 = asdq(mu,i) - u(i)*asdq(1,i)
[…]

### 2020-11-30 cylinder.py
Set initial momentum correctly.

@@ -173,7 +173,7 @@ def setup(
         # INITIAL CONDITIONS #
         nx=xc/(rc+1.e-7); ny=yc/(rc+1.e-7)
         state.q[0,:,:] = state.problem_data['h0']
-        state.q[1,:,:] = state.problem_data['u0']
+        state.q[1,:,:] = state.problem_data['h0']*state.problem_data['u0']
         state.q[2,:,:] = 0.0
 
         # create clawpack solution

### 2020-11-30 rpn2_swq_hllemcc.f90
Add transverse HLLEMCC solver for mapped grids.

@@ -1,6 +1,6 @@
 subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
 
-! Roe-solver for the 2D shallow water equations on a mapped 
+! Roe solver for the 2D shallow water equations on a mapped 
 ! (logically quadrilateral) grid
 
 ! waves: 3
@@ -20,8 +20,7 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
     double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
 
-    ! local arrays -- common block comroe is passed to rpt2sh
-    ! ------------
+    ! local arrays
 
     double precision, dimension(mwaves) :: splus, sminus
     integer :: depth
@@ -40,7 +39,6 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
 
     common /cparam/ grav, kalpha, kbeta, kepsilon
-    common /comroe/ u, v, a, h
 
     data efix /.true./        !# use entropy fix for transonic rarefactions
     data cfix /.true./        !# additionally use carbuncle fix
@@ -84,8 +82,6 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     enddo
 
     ! compute the Roe-averaged variables needed in the Roe solver.
-    ! These are stored in the common block comroe since they are
-    ! later used in routine rpt2 to do the transverse wave splitting.
 
     do i = 2-mbc, mx+mbc
         hsqrt_l = dsqrt(qr(depth,i-1))

### 2020-11-30 rpt2_shallow_roe_mapped.f90

@@ -1,0 +1,102 @@
+subroutine rpt2(ixy,imp,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,aux1,aux2,aux3,asdq,bmasdq,bpasdq)
+
+! Transverse Roe solver for the 2D shallow water equations
+! on a mapped grid.
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!   1 depth
+!   2 x-momentum
+!   3 y-momentum
+
+    implicit none
+
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(in) :: ql, qr, asdq
+    integer, intent(in) :: ixy, imp, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(maux, 1-mbc:maxm+mbc), intent(in) :: aux1, aux2, aux3
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(out) :: bmasdq, bpasdq
+
+    ! Local variables
+    double precision, dimension(3,3) :: waveb
+    double precision, dimension(3) :: sb
+    integer :: mu, mv, i, m, mw, inx, iny, ilenrat
+    double precision :: hsqrt_l, hsqrt_r, hsq2, a1, a2, a3, grav, u, v, a, h, alpha, beta
+    double precision :: asdqh, asdqn, asdqt
+
+!   # grav must be set elsewhere
+    common /cparam/ grav
+
+!   # Roe average quantities of each interface
+    integer, parameter :: maxm2 = 1800
+    double precision, dimension(-6:maxm2+7) :: unorl, unorr, utanl, utanr
+
+    if (ixy == 1) then
+        inx = 1
+        iny = 2
+        ilenrat = 3
+    else
+        inx = 4
+        iny = 5
+        ilenrat = 6
+    endif
+            
+    do i=2-mbc, mx+mbc
+        alpha = aux2(inx,i)
+        beta = aux2(iny,i)
+        unorl = alpha*ql(2,i) + beta*ql(3,i)
+        unorr = alpha*qr(2,i-1) + beta*qr(3,i-1)
+        utanl = -beta*ql(2,i) + alpha*ql(3,i)
+        utanr = -beta*qr(2,i-1) + alpha*qr(3,i-1)
+    enddo
+
[…]

### 2020-12-01 rpt2_shallow_roe_mapped.f90
Correction to transverse mapped solver.

@@ -33,13 +33,13 @@ subroutine rpt2(ixy,imp,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,aux1,aux2,aux3,asdq,b
     double precision, dimension(-6:maxm2+7) :: unorl, unorr, utanl, utanr
 
     if (ixy == 1) then
-        inx = 1
-        iny = 2
-        ilenrat = 3
-    else
         inx = 4
         iny = 5
         ilenrat = 6
+    else
+        inx = 1
+        iny = 2
+        ilenrat = 3
     endif
             
     do i=2-mbc, mx+mbc

### 2020-12-03 rpn2_swq_hllemcc.f90
Transverse mapped solver based on Clawpack 4 code.

@@ -1,224 +1,0 @@
-subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-
-! Roe solver for the 2D shallow water equations on a mapped 
-! (logically quadrilateral) grid
-
-! waves: 3
-! equations: 3
-
-! Conserved quantities:
-!       1 depth
-!       2 x-momentum
-!       3 y-momentum
-
-    implicit none
-
-    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
-    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
-    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
-    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
-
-    ! local arrays
-
-    double precision, dimension(mwaves) :: splus, sminus
-    integer :: depth
-    integer :: i, m, mw
-    integer :: inx, iny, ilenrat
-    double precision :: ffroude, a1, a2, a3
-
-    integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
-    double precision, dimension(3) :: delta
-    logical :: efix, cfix
-    double precision, dimension(-1:maxm2) :: u, v, a, h, unorl, unorr, utanl, utanr, alpha, beta
-    double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
-    double precision :: grav
-    double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
-    double precision :: renore, rere1, reremu, reremv, rhind
-    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
-
-    common /cparam/ grav, kalpha, kbeta, kepsilon
-
-    data efix /.true./        !# use entropy fix for transonic rarefactions
-    data cfix /.true./        !# additionally use carbuncle fix
-
[…]

### 2020-12-03 rpn2_swq_hllemcc.f90
Put back solver that I accidentally deleted.

@@ -1,0 +1,228 @@
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+
+! Roe-solver for the 2D shallow water equations on a mapped 
+! (logically quadrilateral) grid
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x-momentum
+!       3 y-momentum
+
+    implicit none
+
+    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
+    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
+    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
+
+    ! local arrays -- common block comroe is passed to rpt2sh
+    ! ------------
+
+    double precision, dimension(mwaves) :: splus, sminus
+    integer :: depth
+    integer :: i, m, mw
+    integer :: inx, iny, ilenrat
+    double precision :: ffroude, a1, a2, a3
+
+    integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
+    double precision, dimension(3) :: delta
+    logical :: efix, cfix
+    double precision, dimension(-1:maxm2) :: u, v, a, h, unorl, hunorr, hutanl, hutanr, alpha, beta
+    double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
+    double precision :: grav
+    double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
+    double precision :: renore, rere1, reremu, reremv, rhind
+    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
+
+    common /cparam/ grav, kalpha, kbeta, kepsilon
+    common /comroe/ u, v, a, h
+
+    data efix /.true./        !# use entropy fix for transonic rarefactions
[…]

### 2020-12-03 rpn2_swq_hllemcc.f90
Fix typo.

@@ -32,7 +32,7 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
     integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
     double precision, dimension(3) :: delta
     logical :: efix, cfix
-    double precision, dimension(-1:maxm2) :: u, v, a, h, unorl, hunorr, hutanl, hutanr, alpha, beta
+    double precision, dimension(-1:maxm2) :: u, v, a, h, hunorl, hunorr, hutanl, hutanr, alpha, beta
     double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
     double precision :: grav
     double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude

### 2020-12-03 numerics_paper/version2/numerical_chj.tex

@@
 \newcommand{\iph}{{i+1/2}}
-\newcommand{\bff}{{\bf f}}
-\newcommand{\bfr}{{\bf r}}
-\newcommand{\bfF}{{\bf F}}
-\newcommand{\bfu}{{\bf u}}
-\newcommand{\bfv}{{\bf v}}
-\newcommand{\bfq}{{\bf q}}
-\newcommand{\bfx}{{\bf x}}
+\newcommand{\bff}{{f}}
+\newcommand{\bfr}{{r}}
+\newcommand{\bfF}{{F}}
+\newcommand{\bfu}{{u}}
+\newcommand{\entvar}{\eta'}
+\newcommand{\bfq}{{q}}
+\newcommand{\bfx}{{x}}
 \newcommand{\sgn}{{\rm sgn}}
@@
 \newcommand{\Roe}{{\rm Roe}}
+\newcommand{\efp}{\psi}
+\newcommand{\entflux}{g}
 \newcommand{\Fr}{{\rm Fr}}
@@
 
-Let $\bfv_i=\bfv(\bfu_i)$ and $q_i=q(\bfu_i)$ denote the entropy variable and the
-(one-dimensional) entropy potential at cell $i$, respectively.
-Also let $\psi_i=\bfv(\bfu_i)\cdot \bff(\bfu_i)-q(\bfu_i)$ be the entropy potential at cell $i$.
+Let $\entvar_i=\entvar(\bfq_i)$ and $\entflux_i=\entflux(\bfq_i)$ denote the entropy variable and the
+(one-dimensional) entropy flux at cell $i$, respectively.
+Also let $\efp=\entvar(\bfq_i)\cdot \bff(\bfq_i)-\entflux(\bfq_i)$ be the entropy potential at cell $i$.
 From \cite[\S 4]{tadmor1987numerical}, if the numerical fluxes satisfy
 \begin{align}\label{es_cond}
-(\bfv_{i}-\bfv_{i-1})\cdot \bfF_{i-1/2}\leq \psi_{i}-\psi_{i-1},
+(\entvar_{i}-\entvar_{i-1})\cdot \bfF_{i-1/2}\leq \efp_{i}-\efp_{i-1},
   \qquad
-  (\bfv_{i+1}-\bfv_i)\cdot \bfF_{i+1/2}\leq \psi_{i+1}-\psi_i,
+  (\entvar_{i+1}-\entvar_i)\cdot \bfF_{i+1/2}\leq \efp_{i+1}-\efp_i,
 \end{align}
@@
 \begin{align*}
-    G_{i+1/2} &= \frac{1}{2}\left(\bfv_i+\bfv_{i+1}\right)\bfF_{i+1/2}-\frac{1}{2}(\psi_{i}+\psi_{i+1}),
+    G_{i+1/2} &= \frac{1}{2}\left(\entvar_i+\entvar_{i+1}\right)\bfF_{i+1/2}-\frac{1}{2}(\efp_{i}+\efp_{i+1}),
 \end{align*}
@@
 \begin{align}\label{lambda_min}
-  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\bfv_{i+1}-\bfv_{i})\cdot\left[\bff(\bfu_{i+1})+\bff(\bfu_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\psi_{i+1}-\psi_i)}{\frac{1}{2}(\bfv_{i+1}-\bfv_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}
+  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\left[\bff(\bfq_{i+1})+\bff(\bfq_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\efp_{i+1}-\efp_i)}{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}
 \end{align}
@@
 \begin{align*}
-  (\bfv_{i}-\bfv_{i-1})\cdot
+  (\entvar_{i}-\entvar_{i-1})\cdot
   \Big(
-  \underbrace{\bff(\bfu_i)-\A^+\Delta Q_{i-1/2}}_{=\bfF_{i-1/2}}\Big)\leq \psi_{i}-\psi_{i-1},
+  \underbrace{\bff(\bfq_i)-\A^+\Delta Q_{i-1/2}}_{=\bfF_{i-1/2}}\Big)\leq \efp_{i}-\efp_{i-1},
   \quad
-  (\bfv_{i+1}-\bfv_i)\cdot
-  \Big(\underbrace{\bff(\bfu_i)+\A^-\Delta Q_{i+1/2}}_{=\bfF_{i+1/2}}\Big)\leq \psi_{i+1}-\psi_i.
+  (\entvar_{i+1}-\entvar_i)\cdot
+  \Big(\underbrace{\bff(\bfq_i)+\A^-\Delta Q_{i+1/2}}_{=\bfF_{i+1/2}}\Big)\leq \efp_{i+1}-\efp_i.
 \end{align*}

### 2020-12-03 mql/FV/cylinder/cylinder.py
add cleaned code for cylinder

@@ -1,0 +1,230 @@
+"""
+Options:
+    - Riemann solver: GeoClaw, Roe, HLLE, HLLEMCC
+"""
+
+from matplotlib import animation
+import matplotlib.pyplot as plt
+from clawpack import pyclaw
+import numpy as np
+from ent_residual import *
+
+tol = 1E-7
+
+# If plotting in a Jupyter notebook, do this:
+#from clawpack.visclaw.JSAnimation import IPython_display
+
+# *********************************************************** #
+# ********** FUNCTIONS TO COMPUTE ENTROPY RESIDUAL ********** #
+# *********************************************************** #
+def b4step(self, state, old_state=None):
+    """
+    Set zero normal boundary inside the cylinder
+    """
+    # get entropy residual #
+    if use_petsc:
+        state.get_qbc_from_q(2,self.qbc)
+        Ri = get_ent_residual(state,self.qbc)
+    else:
+        Ri = get_ent_residual(state)
+    #
+    
+    set_Ri = state.problem_data['set_Ri']
+    if set_Ri is None:
+        if use_petsc:
+            state.aux[state.num_aux-1,:,:] = Ri[2:-2,2:-2]
+        else:
+            state.aux[state.num_aux-1,:,:] = Ri
+    else:
+        state.aux[state.num_aux-1,:,:] = set_Ri
+    #
+
+    # set values at the cylinder #
+    xc, yc = state.p_centers
+    rc = np.sqrt((xc-40)**2 + (yc-50)**2)
+    in_solid = np.where((rc <= 20.0))
+
+    nx = (xc-40)/(rc+1e-7)
+    ny = (yc-50)/(rc+1e-7)
+    hu=state.q[1,:,:]
+    hv=state.q[2,:,:]
+    
+    hu_tan = hu*ny - hv*nx
+    state.q[1,in_solid[0],in_solid[1]] =  ny[in_solid[0],in_solid[1]] * hu_tan[in_solid[0],in_solid[1]]
+    state.q[2,in_solid[0],in_solid[1]] = -nx[in_solid[0],in_solid[1]] * hu_tan[in_solid[0],in_solid[1]]
+#
+
+def inflow_bcs(state, dim, _, qbc, __, num_ghost):
+    qbc[0,:num_ghost,:] = state.problem_data['h0']
+    qbc[1,:num_ghost,:] = state.problem_data['h0']*state.problem_data['u0']
[…]

### 2020-12-03 mql/FV/cylinder/plot.py

@@ -1,0 +1,117 @@
+from clawpack.petclaw.solution import Solution as petclaw_Solution
+from clawpack.pyclaw.solution import Solution as pyclaw_Solution
+#from petclaw.io.petsc import read_petsc
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pylab as pl
+from matplotlib import rc
+#rc('text', usetex=True)
+import numpy as np
+import os
+
+def plot_q(frame,
+           plot_schlieren=True,
+           use_petsc=True,
+           bathymetry=False,
+           file_prefix='claw',
+           path='./_output/',
+           plot_ent_residual=False,
+           plot_pcolor=True,
+           momMagLim=None,
+           plot_slices=True):
+    import sys
+    sys.path.append('.')
+
+    if use_petsc:
+        sol=petclaw_Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
+    else:
+        file_prefix='fort'
+        sol=pyclaw_Solution(frame,read_aux=False,path=path,file_prefix=file_prefix)
+    #    
+    x=sol.state.grid.x.centers; y=sol.state.grid.y.centers
+    mx=len(x); my=len(y)
+
+    h=sol.state.q[0,:,:]
+    q1=sol.state.q[1,:,:]
+    q2=sol.state.q[2,:,:]
+    qMag = np.sqrt(q1**2 + q2**2)
+
+    yy,xx = np.meshgrid(y,x)
+
+    if frame < 10:
+        str_frame = "000"+str(frame)
+    elif frame < 100:
+        str_frame = "00"+str(frame)
+    elif frame < 1000:
+        str_frame = "0"+str(frame)
+    else:
+        str_frame = str(frame)
+    #
+
+    grad_h = np.gradient(h)
+    abs_grad_h = np.sqrt(grad_h[0]**2 + grad_h[1]**2)
+    
+    if plot_pcolor:
+        fig = pl.figure(figsize=(4,8))
+        if plot_schlieren:
+            import clawpack.visclaw.colormaps as cm
+            scm = cm.schlieren_colormap()
+            pl.pcolormesh(xx,yy,abs_grad_h,cmap=scm)                
+        #
+        else:
+            pl.pcolormesh(xx,yy,qMag)
+            if momMagLim is not None:
+                pl.clim(momMagLim[0],momMagLim[1])
+            #
+        #
+        pl.axis('equal')
[…]

### 2020-12-03 mql/FV/cylinder/rpn2_shallow_es.f90

@@ -1,0 +1,397 @@
+subroutine get_ent_variables(grav,dh,dhu,dw1,dw2)
+  double precision grav,dh,dhu,dw1,dw2,du
+  du = dhu/dh
+  dw1 = grav*dh-.5d0*du**2.d0
+  dw2 = du
+end subroutine get_ent_variables
+
+subroutine get_ent_flux(grav,dh,dhu,dEntFlux)
+  double precision grav,dh,dhu,dEntFlux,du
+  du = dhu/dh
+  dEntFlux = (.5d0*dh*du*du+grav*dh*dh)*du
+end subroutine get_ent_flux
+
+subroutine get_ent_pot(dw1,dw2,df1,df2,dEntFlux,dEntPot)
+  double precision dw1,dw2,df1,df2,dEntFlux,dEntPot
+  dEntPot = dw1*df1+dw2*df2-dEntFlux
+end subroutine get_ent_pot
+
+subroutine get_fZ(grav,dh,dhZ,fZ)
+  double precision grav,dh,dhZ,fZ
+  if (dh<= dhZ) then
+     fZ = 2.d0*(dsqrt(grav*dh)-dsqrt(grav*dhZ))
+  else
+     fZ = (dh-dhZ)*dsqrt(0.5d0*grav*(dh+dhZ)/dh/dhZ)
+  endif
+  return
+end subroutine get_fZ
+
+subroutine get_f(grav,dh,dhL,dhR,velL,velR,f)
+  double precision grav,dh,dhL,dhR,velL,velR,dfL,dfR,f
+  call get_fZ(grav,dh,dhL,dfL)
+  call get_fZ(grav,dh,dhR,dfR) 
+  f = dfL + dfR + velR - velL
+  return
+end subroutine get_f
+
+subroutine get_lambda_llf(grav,dhL,dhR,velL,velR,dmax)
+  double precision grav,dhL,dhR,velL,velR,dmax
+  double precision dhMin,dhMax,x0,dfMin,dfMax,dhStar,dlambda1,dlambda3
+  dhMin = min(dhL,dhR)
+  dhMax = max(dhL,dhR)            
+  x0 = (2.d0*dsqrt(2.d0)-1.d0)**2.d0
+  call get_f(grav,x0*dhMin,dhL,dhR,velL,velR,dfMin)
+  call get_f(grav,x0*dhMax,dhL,dhR,velL,velR,dfMax)
[…]

### 2020-12-03 mql/FV/cylinder/rpn2_sw_hllemcc.f90

@@ -1,0 +1,207 @@
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+
+! Roe-solver for the 2D shallow water equations on a Cartesian grid
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x-momentum
+!       3 y-momentum
+
+    implicit none
+
+    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
+    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
+    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
+
+    ! local arrays -- common block comroe is passed to rpt2sh
+    ! ------------
+
+    double precision, dimension(mwaves) :: splus, sminus
+    integer :: depth, mu, mv
+    integer :: i, m, mw
+    double precision :: ffroude, a1, a2, a3
+
+    integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
+    double precision, dimension(3) :: delta
+    logical :: efix, cfix
+    double precision, dimension(-1:maxm2) :: u, v, a, h
+    double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
+    double precision :: grav
+    double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
+    double precision :: renore, rere1, reremu, reremv, rhind
+    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
+
+    common /cparam/ grav, kalpha, kbeta, kepsilon
+    common /comroe/ u, v, a, h
+
+    data efix /.true./        !# use entropy fix for transonic rarefactions
+    data cfix /.true./        !# additionally use carbuncle fix
+
[…]

### 2020-12-03 mql/FV/chj/annulus/hydraulic_jump_2D_annulus.py
add cleaned code for chj in the annulus

@@ -1,0 +1,744 @@
+"""
+Code to simulate 2D circular hydraulic jump in PyClaw.
+
+This version uses a mapped (annular) grid.
+
+In order to get a jump, you must either introduce a bathymetric barrier (circular ring),
+include bottom friction, or impose a subcritical flow at the outer boundary.
+Bathymetry is not enabled for the mapped grid solver at present.
+"""
+
+from matplotlib import animation
+import matplotlib.pyplot as plt
+from clawpack import pyclaw
+import numpy as np
+from mapc2p import mapc2p, mapc2p_annulus
+
+# If plotting in a Jupyter notebook, do this:
+#from clawpack.visclaw.JSAnimation import IPython_display
+
+# Mapped grid parameters
+theta_lower = 0.0
+theta_upper = 2*np.pi
+tol=1E-15
+
+# ENTROPY AND ITS DERIVATIVES
+def get_entropy(g,h,hu,hv):
+    return 0.5*g*h**2 + 0.5*(hu**2+hv**2)/(h+tol)
+def dEntropy_dh(g,h,hu,hv):
+    return g*h - 0.5*(hu**2+hv**2)/(h**2+tol)
+def dEntropy_dhu(g,h,hu,hv):
+    return hu/(h+tol)
+def dEntropy_dhv(g,h,hu,hv):
+    return hv/(h+tol)
+# FLUXES OF THE SHALLOW WATER EQUATIONS
+def xFlux1(g,h,hu,hv):
+    return hu
+def xFlux2(g,h,hu,hv):
+    return hu**2/(h+tol) + 0.5*g*h**2
+def xFlux3(g,h,hu,hv):
+    return hu*hv/(h+tol)
+def yFlux1(g,h,hu,hv):
+    return hv
+def yFlux2(g,h,hu,hv):
+    return hu*hv/(h+tol)
+def yFlux3(g,h,hu,hv):
+    return hv**2/(h+tol) + 0.5*g*h**2
+# ENTROPY FLUX 
+def xEntFlux(g,h,hu,hv):
+    return (0.5*(hu**2+hv**2)/(h+tol) + g*h**2)*hu/(h+tol) 
+def yEntFlux(g,h,hu,hv):
[…]

### 2020-12-03 mql/FV/chj/annulus/mapc2p.py

@@ -1,0 +1,26 @@
+import numpy as np
+
+def mapc2p(xc,yc):
+    xp = xc*np.cos(yc)
+    yp = xc*np.sin(yc)
+    return xp, yp
+
+def mapc2p_annulus(xc, yc):
+    """
+    Specifies the mapping to curvilinear coordinates.
+
+    Inputs: c_centers = Computational cell centers
+                 [array ([Xc1, Xc2, ...]), array([Yc1, Yc2, ...])]
+
+    Output: p_centers = Physical cell centers
+                 [array ([Xp1, Xp2, ...]), array([Yp1, Yp2, ...])]
+    """  
+    p_centers = []
+
+    # Polar coordinates (first coordinate = radius,  second coordinate = theta)
+    p_centers.append(xc[:]*np.cos(yc[:]))
+    p_centers.append(xc[:]*np.sin(yc[:]))
+    
+    return p_centers
+
+

### 2020-12-03 mql/FV/chj/annulus/plot.py

@@ -1,0 +1,178 @@
+from clawpack.petclaw.solution import Solution
+#from petclaw.io.petsc import read_petsc
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pylab as pl
+from matplotlib import rc
+#rc('text', usetex=True)
+import numpy as np
+import os
+
+tol=1E-7
+
+def plot_q(frame,
+           plot_jump=True,
+           plot_schlieren=True,
+           plot_mesh=False,
+           clim=None,
+           xlim=None,
+           ylim=None,
+           file_prefix='claw',
+           path='./_output/',
+           plot_pcolor=True,
+           plot_ent_residual=False,
+           plot_slices=True):
+    import sys
+    sys.path.append('.')
+
+    import mapc2p
+    sol=Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
+    
+    xc_centers, yc_centers = sol.state.grid.c_centers    
+    xc = np.zeros((xc_centers.shape[0],xc_centers.shape[1]+1))
+    yc = np.zeros((yc_centers.shape[0],yc_centers.shape[1]+1))
+    xc[:,:-1] = xc_centers
+    xc[:,-1]  = xc_centers[:,0]
+    yc[:,:-1] = yc_centers
+    yc[:,-1]  = yc_centers[:,0]    
+    
+    xp,yp = mapc2p.mapc2p(xc,yc)
+    
+    # plot mesh
+    if plot_mesh:
+        pl.figure(figsize=(20,20))
+        xc_edges, yc_edges = sol.state.grid.c_nodes
+        xp_edges, yp_edges = mapc2p.mapc2p(xc_edges,yc_edges)
+        pl.plot(xp_edges,yp_edges,'-k',lw=1.25)
+        pl.plot(xp_edges.transpose(),yp_edges.transpose(),'-k',lw=1.25)
+        pl.axis('equal')
+        pl.xticks(fontsize=20)
+        pl.yticks(fontsize=20)
+        pl.xlim([-1,1])
+        pl.ylim([-1,1])
+        pl.savefig('mesh.png')
+    #
+    
+    h=sol.state.q[0,:,:]
+    q1=sol.state.q[1,:,:]
+    q2=sol.state.q[2,:,:]
+    qMag = np.sqrt(q1**2 + q2**2)
+
+    grad_h = np.gradient(h)
[…]

### 2020-12-03 mql/FV/chj/annulus/rpn2_sw_annular_es.f90

@@ -1,0 +1,364 @@
+subroutine get_ent_variables(grav,dh,dhu,dw1,dw2)
+  double precision grav,dh,dhu,dw1,dw2,du
+  du = dhu/dh
+  dw1 = grav*dh-.5d0*du**2.d0
+  dw2 = du
+end subroutine get_ent_variables
+
+subroutine get_ent_flux(grav,dh,dhu,dEntFlux)
+  double precision grav,dh,dhu,dEntFlux,du
+  du = dhu/dh
+  dEntFlux = (.5d0*dh*du*du+grav*dh*dh)*du
+end subroutine get_ent_flux
+
+subroutine get_ent_pot(dw1,dw2,df1,df2,dEntFlux,dEntPot)
+  double precision dw1,dw2,df1,df2,dEntFlux,dEntPot
+  dEntPot = dw1*df1+dw2*df2-dEntFlux
+end subroutine get_ent_pot
+
+subroutine get_fZ(grav,dh,dhZ,fZ)
+  double precision grav,dh,dhZ,fZ
+  if (dh<= dhZ) then
+     fZ = 2.d0*(dsqrt(grav*dh)-dsqrt(grav*dhZ))
+  else
+     fZ = (dh-dhZ)*dsqrt(0.5d0*grav*(dh+dhZ)/dh/dhZ)
+  endif
+  return
+end subroutine get_fZ
+
+subroutine get_f(grav,dh,dhL,dhR,velL,velR,f)
+  double precision grav,dh,dhL,dhR,velL,velR,dfL,dfR,f
+  call get_fZ(grav,dh,dhL,dfL)
+  call get_fZ(grav,dh,dhR,dfR) 
+  f = dfL + dfR + velR - velL
+  return
+end subroutine get_f
+
+subroutine get_lambda_llf(grav,dhL,dhR,velL,velR,dmax)
+  double precision grav,dhL,dhR,velL,velR,dmax
+  double precision dhMin,dhMax,x0,dfMin,dfMax,dhStar,dlambda1,dlambda3
+  dhMin = min(dhL,dhR)
+  dhMax = max(dhL,dhR)            
+  x0 = (2.d0*dsqrt(2.d0)-1.d0)**2.d0
+  call get_f(grav,x0*dhMin,dhL,dhR,velL,velR,dfMin)
+  call get_f(grav,x0*dhMax,dhL,dhR,velL,velR,dfMax)
[…]

### 2020-12-03 mql/FV/chj/annulus/rpn2_sw_annular_hllemcc.f90

@@ -1,0 +1,228 @@
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+
+! Roe-solver for the 2D shallow water equations on a mapped 
+! (logically quadrilateral) grid
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x-momentum
+!       3 y-momentum
+
+    implicit none
+
+    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
+    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
+    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
+
+    ! local arrays -- common block comroe is passed to rpt2sh
+    ! ------------
+
+    double precision, dimension(mwaves) :: splus, sminus
+    integer :: depth
+    integer :: i, m, mw
+    integer :: inx, iny, ilenrat
+    double precision :: ffroude, a1, a2, a3
+
+    integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
+    double precision, dimension(3) :: delta
+    logical :: efix, cfix
+    double precision, dimension(-1:maxm2) :: u, v, a, h, unorl, unorr, utanl, utanr, alpha, beta
+    double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
+    double precision :: grav
+    double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
+    double precision :: renore, rere1, reremu, reremv, rhind
+    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
+
+    common /cparam/ grav, kalpha, kbeta, kepsilon
+    common /comroe/ u, v, a, h
+
+    data efix /.true./        !# use entropy fix for transonic rarefactions
[…]

### 2020-12-03 mql/FV/chj/annulus/rpt2_shallow_roe_with_efix_annulus.f90

@@ -1,0 +1,131 @@
+! =====================================================
+subroutine rpt2(ixy,imp,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,aux1,aux2,aux3,asdq,bmasdq,bpasdq)
+! =====================================================
+    implicit none !double precision (a-h,o-z)
+
+!     # Riemann solver in the transverse direction for the shallow water
+!     equations .
+!     # Split asdq (= A^* \Delta q, where * = + or -)
+!     # into down-going flux difference bmasdq (= B^- A^* \Delta q)
+!     #    and up-going flux difference bpasdq (= B^+ A^* \Delta q)
+
+    integer, intent(in) :: ixy, imp, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(in) :: ql,qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: aux1, aux2, aux3
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(in) :: asdq
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(out) :: bmasdq, bpasdq
+
+    double precision, dimension (3,3) :: wave
+    double precision, dimension (3) :: s, delta
+    double precision :: grav
+
+    integer :: mw,m,inx,iny,ilenrat,i
+    double precision :: hsqrt_r,hsqrt_l,hsq2,a1,a2,a3
+    double precision :: w11,w12,w13,w21,w22,w23,w31,w32,w33,d1,d2,d3
+    
+!   # grav must be set elsewhere
+    common /cparam/ grav
+
+!   # Roe averages quantities of each interface
+    integer, parameter :: maxm2 = 1800
+    double precision, dimension (-6:maxm2+7) :: u, v, a, h, hunor_r, hunor_l, hutan_r, hutan_l, alpha, beta
+
+    if (ixy.eq.1) then
+        inx = 1
+        iny = 2
+        ilenrat = 3
+    else
+        inx = 4
+        iny = 5
+        ilenrat = 6
+     endif
+     
+    ! Compute the rotation matrix:
+    ! [ alpha beta ]
[…]

### 2020-12-03 mql/FV/chj/hydraulic_jump_2D.py
remove some old code

@@ -1,452 +1,0 @@
-"""
-Code to simulate 2D circular hydraulic jump in PyClaw.
-
-Currently uses a Cartesian grid.
-
-In order to get a jump, you must either introduce a bathymetric barrier (circular ring),
-include bottom friction, or impose a subcritical flow at the outer boundary.
-To work with bathymetry, you have to use a Riemann solver that handles variable
-bathymetry; for now this means the GeoClaw solver.
-
-We could also get a jump by introducing bottom friction; we haven't tried this yet.
-
-Options:
-    - Riemann solver: GeoClaw, Roe, HLLE, HLLEMCC
-"""
-
-from matplotlib import animation
-import matplotlib.pyplot as plt
-from clawpack import pyclaw
-import numpy as np
-
-# If plotting in a Jupyter notebook, do this:
-#from clawpack.visclaw.JSAnimation import IPython_display
-
-def plot_surface(claw, make_anim=True, save_plots=False, frames=101, val='surface',
-                 vmin=0., vmax=0.5, clim=None, bathymetry=False, plotdir='./_plots'):
-    """
-    Plot results of 2D shallow water simulation as a pcolor plot and animate
-    (intended for Jupyter notebook).
-
-    If variable bathymetry is used, set bathymetry=True.
-    """
-    print('plotting')
-    fig = plt.figure(figsize=[10,10])
-    ax1 = fig.add_subplot(111)
-    try:
-        # Use solution in memory
-        frame = claw.frames[0]
-    except:
-        # Read solution from disk
-        frame = pyclaw.Solution(0,path=claw,read_aux=bathymetry)
-    if bathymetry:
-        b = frame.aux[0,:,:]
-    else:
-        b = 0.
-    h = frame.q[0,:,:]
-    if val == 'surface':
-        qq = np.maximum(b,h+b)
-    elif val=='u':
-        hu = frame.q[1,:,:]
-        qq = hu #/h
-        
-
-    x, y = frame.state.grid.p_centers    
-
-    im = ax1.imshow(qq.T, #cmap='Blues',
-                    extent=[x.min(), x.max(), y.min(), y.max()],
-                    vmin=vmin, vmax=vmax,
[…]

### 2020-12-03 mql/FV/chj/plot.py

@@ -1,195 +1,0 @@
-from clawpack.petclaw.solution import Solution
-#from petclaw.io.petsc import read_petsc
-import matplotlib
-matplotlib.use('Agg')
-import matplotlib.pylab as pl
-from matplotlib import rc
-#rc('text', usetex=True)
-import numpy as np
-import os
-
-def reconstruct_solution(h):
-    x_hL = np.zeros_like(h)
-    x_hR = np.zeros_like(h)
-    x_hL[2:-2,:] = (-3.*h[:-4,:] + 27*h[1:-3,:] + 47*h[2:-2,:] - 13*h[3:-1,:] + 2*h[4:,:])/60.
-    x_hR[2:-2,:] = (2.*h[:-4,:] - 13*h[1:-3,:] + 47*h[2:-2,:] + 27*h[3:-1,:] - 3*h[4:,:])/60.
-
-    y_hL = np.zeros_like(h)
-    y_hR = np.zeros_like(h)
-    y_hL[:,2:-2] = (-3.*h[:,:-4] + 27*h[:,1:-3] + 47*h[:,2:-2] - 13*h[:,3:-1] + 2*h[:,4:])/60.
-    y_hR[:,2:-2] = (2.*h[:,:-4] - 13*h[:,1:-3] + 47*h[:,2:-2] + 27*h[:,3:-1] - 3*h[:,4:])/60.
-
-    return x_hL, x_hR, y_hL, y_hR
-#
-    
-def get_entropy(g,h,hu,hv):
-    return 0.5*g*h**2 + 0.5*(hu**2+hv**2)/(h+1E-7)
-#
-
-def xFlux(g,h,hu,hv):
-    tol=1E-7
-    return (0.5*(hu**2+hv**2)/(h+tol) + g*h**2)*hu/(h+tol) 
-#
-def yFlux(g,h,hu,hv):
-    tol=1E-7
-    return (0.5*(hu**2+hv**2)/(h+tol) + g*h**2)*hv/(h+tol) 
-#
-
-def plot_q(frame,
-           bathymetry=False,
-           file_prefix='claw',
-           path='./_output/',
-           plot_pcolor=True,
-           plot_ent_residual=False,
-           hLim=None,
-           momMagLim=None,
-           plot_slices=True):
-    import sys
-    sys.path.append('.')
-
-    sol=Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
[…]

### 2020-12-03 mql/FV/chj/rpn2_shallow_hllc.f90

@@ -1,165 +1,0 @@
-! =====================================================
-subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-! =====================================================
-
-! HLLC solver for the 2D shallow water equations.
-
-! waves: 3
-! equations: 3
-
-! Conserved quantities:
-!       1 depth
-!       2 x-momentum
-!       3 y-momentum
-
-    implicit none
-
-    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
-    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
-    
-    double precision, dimension(meqn, mwaves, 1-mbc:maxm+mbc), intent(out) :: wave
-    double precision, dimension(mwaves, 1-mbc:maxm+mbc), intent(out) :: s
-    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(out) :: amdq, apdq
-    
-    double precision, dimension(2,1) :: q_hat
-    double precision :: u_l, u_r, c_l, c_r, u_hat, c_hat, v_l, v_r, v_star, v_hat, h_hat
-    double precision :: h_l, h_r, hsqrt_l, hsqrt_r, hsq2
-    double precision :: grav, a, b, c, aprime, bprime, root0, root
-    double precision :: h_ml, hu_ml, hv_ml, h_mr, hu_mr, hv_mr, s1, s2, s_star, u_ml, u_mr
-    integer :: depth, mu, mv
-    integer :: i, m, mw, j, maxiter
-    
-
-    common /cparam/  grav
-
-!     # set mu to point to  the component of the system that corresponds
-!     # to momentum in the direction of this slice, mv to the orthogonal
-!     # momentum:
-
-    ! max num of iteration for the root-finding algorithm
-    maxiter = 20   
-    depth = 1
-    if (ixy.eq.1) then
-        mu = 2
-        mv = 3
-    else
-        mu = 3
-        mv = 2
-    endif
-
-    do i=2-mbc,mx+mbc
-        h_l = qr(depth,i-1)
-        h_r = ql(depth,i)
[…]

### 2020-12-03 mql/FV/chj/rpn2_sw_hllemcc.f90

@@ -1,207 +1,0 @@
-subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-
-! Roe-solver for the 2D shallow water equations on a Cartesian grid
-
-! waves: 3
-! equations: 3
-
-! Conserved quantities:
-!       1 depth
-!       2 x-momentum
-!       3 y-momentum
-
-    implicit none
-
-    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
-    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
-    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
-    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
-
-    ! local arrays -- common block comroe is passed to rpt2sh
-    ! ------------
-
-    double precision, dimension(mwaves) :: splus, sminus
-    integer :: depth, mu, mv
-    integer :: i, m, mw
-    double precision :: ffroude, a1, a2, a3
-
-    integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
-    double precision, dimension(3) :: delta
-    logical :: efix, cfix
-    double precision, dimension(-1:maxm2) :: u, v, a, h
-    double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
-    double precision :: grav
-    double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
-    double precision :: renore, rere1, reremu, reremv, rhind
-    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
-
-    common /cparam/ grav, kalpha, kbeta, kepsilon
-    common /comroe/ u, v, a, h
-
-    data efix /.true./        !# use entropy fix for transonic rarefactions
-    data cfix /.true./        !# additionally use carbuncle fix
-
[…]

### 2020-12-03 mql/FV/chj/rpn2_swq_hllemcc.f90

@@ -1,235 +1,0 @@
-subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-
-! Roe-solver for the 2D shallow water equations on a mapped 
-! (logically quadrilateral) grid
-
-! waves: 3
-! equations: 3
-
-! Conserved quantities:
-!       1 depth
-!       2 x-momentum
-!       3 y-momentum
-
-    implicit none
-
-    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
-    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
-    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
-    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
-
-    ! local arrays -- common block comroe is passed to rpt2sh
-    ! ------------
-
-    double precision, dimension(mwaves) :: splus, sminus
-    integer :: depth, mu, mv
-    integer :: i, m, mw
-    integer :: inx, iny, ilenrat
-    double precision :: ffroude, a1, a2, a3
-
-    integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
-    double precision, dimension(3) :: delta
-    logical :: efix, cfix
-    double precision, dimension(-1:maxm2) :: u, v, a, h, unorl, unorr, utanl, utanr, alf, beta
-    double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
-    double precision :: grav
-    double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
-    double precision :: renore, rere1, reremu, reremv, rhind
-    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
-
-    common /cparam/ grav, kalpha, kbeta, kepsilon
-    common /comroe/ u, v, a, h
-
[…]

### 2020-12-03 mql/FV/chj/rpn2swq-hllemcc.f

@@ -1,274 +1,0 @@
-
-c
-c
-c     =====================================================
-      subroutine rpn2(ixy,maxm,meqn,mwaves,mbc,mx,ql,qr,auxl,auxr,
-     &     wave,s,amdq,apdq)
-c     =====================================================
-c     
-c     # Roe-solver for the 2D shallow water equations
-c     #  on a quadrilateral grid
-c     
-c     # solve Riemann problems along one slice of data.
-c     
-c     # On input, ql contains the state vector at the left edge of each
-c     cell
-c     #           qr contains the state vector at the right edge of each
-c     cell
-c     
-c     # This data is along a slice in the x-direction if ixy=1
-c     #                            or the y-direction if ixy=2.
-c     # On output, wave contains the waves, s the speeds,
-c     # and amdq, apdq the decomposition of the flux difference
-c     #   f(qr(i-1)) - f(ql(i))
-c     # into leftgoing and rightgoing parts respectively.
-c     # With the Roe solver we have
-c     #    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
-c     # where A is the Roe matrix.  An entropy fix can also be
-c     incorporated
-c     # into the flux differences.
-c     
-c     # Note that the i'th Riemann problem has left state qr(i-1,:)
-c     #                                    and right state ql(i,:)
-c     # From the basic clawpack routines, this routine is called with ql
-c     = qr
-c     
-c     
-      implicit double precision (a-h,o-z)
-c     
-      dimension wave(1-mbc:maxm+mbc, meqn, mwaves)
-      dimension    s(1-mbc:maxm+mbc, mwaves)
-      dimension   ql(1-mbc:maxm+mbc, meqn)
-      dimension   qr(1-mbc:maxm+mbc, meqn)
-      dimension  apdq(1-mbc:maxm+mbc, meqn)
-      dimension  amdq(1-mbc:maxm+mbc, meqn)
-      dimension auxl(1-mbc:maxm+mbc, 7)
-      dimension auxr(1-mbc:maxm+mbc, 7)
-c     
-c     local arrays -- common block comroe is passed to rpt2sh
-c     ------------
-c     
[…]

### 2020-12-03 mql/FV/chj/rpn2swq-hllemcc.f90

@@ -1,236 +1,0 @@
-subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-
-! Roe-solver for the 2D shallow water equations on a quadrilateral grid
-! 
-! waves: 3
-! equations: 3
-! 
-! On input, ql contains the state vector at the left edge of each cell
-!           qr contains the state vector at the right edge of each cell
-!     
-! This data is along a slice in the x-direction if ixy=1
-!                            or the y-direction if ixy=2.
-! On output, wave contains the waves, s the speeds,
-! and amdq, apdq the decomposition of the flux difference
-!   f(qr(i-1)) - f(ql(i))
-! into leftgoing and rightgoing parts respectively.
-! With the Roe solver we have
-!    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
-! where A is the Roe matrix.  An entropy fix can also be incorporated
-! into the flux differences.
-! 
-! Note that the i'th Riemann problem has left state qr(i-1,:)
-!                                     and right state ql(i,:)
-! From the basic clawpack routines, this routine is called with ql = qr
-
-    implicit none
-
-    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
-    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
-    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
-    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
-    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
-
-    ! local arrays -- common block comroe is passed to rpt2sh
-    ! ------------
-
-    double precision, dimension(mwaves) :: splus, sminus
-
-    integer parameter (maxm2 = 1602)  !# assumes at most 1000x1000 grid with mbc=2
-    double precision, dimension(3) :: delta
-    logical :: efix, cfix
[…]

### 2020-12-03 mql/FV/chj/rpn2swq_hllemcc.f90

@@ -1,240 +1,0 @@
-subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-
-! Roe-solver for the 2D shallow water equations on a quadrilateral grid
-! 
-! Solve Riemann problems along one slice of data.
-! 
-! On input, ql contains the state vector at the left edge of each cell
-!           qr contains the state vector at the right edge of each cell
-!     
-! This data is along a slice in the x-direction if ixy=1
-!                            or the y-direction if ixy=2.
-! On output, wave contains the waves, s the speeds,
-! and amdq, apdq the decomposition of the flux difference
-!   f(qr(i-1)) - f(ql(i))
-! into leftgoing and rightgoing parts respectively.
-! With the Roe solver we have
-!    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
-! where A is the Roe matrix.  An entropy fix can also be incorporated
-! into the flux differences.
-! 
-! Note that the i'th Riemann problem has left state qr(i-1,:)
-!                                     and right state ql(i,:)
-! From the basic clawpack routines, this routine is called with ql = qr
-
-    implicit double precision (a-h,o-z)
-
-    dimension wave(meqn, mwaves, 1-mbc:maxm+mbc)
-    dimension    s(mwaves, 1-mbc:maxm+mbc)
-    dimension   ql(meqn, 1-mbc:maxm+mbc)
-    dimension   qr(meqn, 1-mbc:maxm+mbc)
-    dimension  apdq(meqn, 1-mbc:maxm+mbc)
-    dimension  amdq(meqn, 1-mbc:maxm+mbc)
-    dimension auxl(maux, 1-mbc:maxm+mbc)
-    dimension auxr(maux, 1-mbc:maxm+mbc)
-
-    ! local arrays -- common block comroe is passed to rpt2sh
-    ! ------------
-
-    dimension splus(mwaves)
-    dimension sminus(mwaves)
-
-    parameter (maxm2 = 1602)  !# assumes at most 1000x1000 grid with mbc=2
-    dimension delta(3)
-    logical efix, cfix
-    dimension unorl(-1:maxm2), unorr(-1:maxm2)
-    dimension utanl(-1:maxm2), utanr(-1:maxm2)
[…]

### 2020-12-03 mql/FV/chj/rpn2swq_hllemccroef.f90

@@ -1,283 +1,0 @@
-subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
-! Modification of Kemm's code.
-! The entropy fix for the 1- and 3-wave is performed as in rpn2_shallow_roe_with_efix.f90, 
-! whereas the one for the 2-wave is the same as in Kemm's code.
-
-
-! Roe-solver for the 2D shallow water equations on a quadrilateral grid
-! 
-! Solve Riemann problems along one slice of data.
-! 
-! On input, ql contains the state vector at the left edge of each cell
-!           qr contains the state vector at the right edge of each cell
-!     
-! This data is along a slice in the x-direction if ixy=1
-!                            or the y-direction if ixy=2.
-! On output, wave contains the waves, s the speeds,
-! and amdq, apdq the decomposition of the flux difference
-!   f(qr(i-1)) - f(ql(i))
-! into leftgoing and rightgoing parts respectively.
-! With the Roe solver we have
-!    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
-! where A is the Roe matrix.  An entropy fix can also be incorporated
-! into the flux differences.
-! 
-! Note that the i'th Riemann problem has left state qr(i-1,:)
-!                                     and right state ql(i,:)
-! From the basic clawpack routines, this routine is called with ql = qr
-
-    implicit double precision (a-h,o-z)
-
-    dimension wave(meqn, mwaves, 1-mbc:maxm+mbc)
-    dimension    s(mwaves, 1-mbc:maxm+mbc)
-    dimension   ql(meqn, 1-mbc:maxm+mbc)
-    dimension   qr(meqn, 1-mbc:maxm+mbc)
-    dimension  apdq(meqn, 1-mbc:maxm+mbc)
-    dimension  amdq(meqn, 1-mbc:maxm+mbc)
-    dimension auxl(maux, 1-mbc:maxm+mbc)
-    dimension auxr(maux, 1-mbc:maxm+mbc)
-
-    ! local arrays -- common block comroe is passed to rpt2sh
-    ! ------------
-
-    dimension splus(mwaves)
-    dimension sminus(mwaves)
-
[…]

### 2020-12-03 mql/FV/chj/rpt2_dummy.f90

@@ -1,10 +1,0 @@
-! =====================================================
-subroutine rpt2(ixy,imp,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,aux1,aux2,aux3,asdq,bmasdq,bpasdq)
-! =====================================================
-    implicit double precision (a-h,o-z)
-
-!     # Dummy transverse Riemann solver, for use in dimensionally-split algorithm.
-
-    write(*,*) 'Error: Dummy transverse Riemann solver called!'
-    return
-    end subroutine rpt2

### 2021-01-13 numerics_paper/version2/numerical_chj.tex
Add section on flow past a cylinder.
Also some minor improvements to code.

@@
 
-\subsection{Steady shock around a cylinder}\label{sec:bow_shock}
+\subsection{Flow past a cylinder}\label{sec:bow_shock}
+In this section we consider the formation of a bow shock when a
+uniform flow encounters a cylindrical obstacle.  This problem has
+been studied previously in the context of carbuncle formation in
+several works for the Euler equations and in
+\cite{kemm2014note,bader2014carbuncle} for the shallow water equations.
+We present results for existing solvers as a form of verification,
+along with results for the new blended solver.  The main question of
+interest is the formation of carbuncles.  It is known, for instance,
+that the Roe solver incorrectly generates a carbuncle at the center
+of the bow shock.
+
+We model only the flow on the upstream side of the cylinder, since
+our interest is in the resolution of the bow shock.  We take the domain
+$[0,40]\times[0,100]$ with a cylinder of radius 20 centered at $(40,50)$.
+Reflecting boundary conditions are imposed at the surface of the cylinder,
+along with outflow conditions along the rest of the right edge of the domain.
+The depth and velocity are set initially and (for all time) at the other
+boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.  We use a $160 \times 400$
+uniform Cartesian grid.
+
+We show results at $t=80$, after the flow has reached
+a steady state.  With a Roe solver, negative values are generated at an early
+time, leading to failure of the solver.  Therefore, we show results only for
+HLLE, HLLEMCC, and the blended solver.  These three methods give very similar
+results.
+
+\begin{figure}[!h]
+  {\scriptsize
+    \includegraphics[scale=0.3]{figures/cylinder_HLLE.png}
+    \qquad \qquad \qquad
+    \includegraphics[scale=0.3]{figures/cylinder_HLLEMCC.png}
+    \qquad \qquad \qquad
+    \includegraphics[scale=0.3]{figures/cylinder_ES.png}
+  }
+  \caption{
+    Flow past a cylinder.  Results for HLLE, HLLEMCC, and the blended solver.
+    \label{fig:steady_outflow}}
+\end{figure}
+
+We have also conducted a more challenging version of this test, in which
+the velocity within the domain is initially zero.  In that case, the HLLE
+and blended solvers give carbuncle-free results similar to Figure \ref{fig:steady_outflow},
+but HLLEMCC exhibits a carbuncle.
 

### 2021-01-13 cylinder.py

@@ -93,19 +93,23 @@ def setup(
     if solver_type == 'classic':
         if riemann_solver == 'es':
             solver = pyclaw.ClawSolver2D(shallow_es_2D)
-            solver.num_eqn = 3
+            solver.dimensional_split=True
             solver.num_waves = 3
-            solver.fwave = False
+        elif riemann_solver == 'roe':
+            solver = pyclaw.ClawSolver2D(riemann.shallow_roe_with_efix_2D)
+            solver.dimensional_split=True
+        elif riemann_solver == 'hlle':
+            solver = pyclaw.ClawSolver2D(riemann.shallow_hlle_2D)
             solver.dimensional_split=True
         elif riemann_solver == 'hllemcc':
             solver = pyclaw.ClawSolver2D(shallow_hllemcc_2D)
-            solver.num_eqn = 3
-            solver.num_waves = 3
-            solver.fwave = False
             solver.dimensional_split=False
             solver.transverse_waves = 1
+            solver.num_waves = 3
         else:
             raise Exception('Unrecognized Riemann solver')
+        solver.num_eqn = 3
+        solver.fwave = False
         solver.limiters = pyclaw.limiters.tvd.minmod
         solver.cfl_max     = 0.46
         solver.cfl_desired = 0.45
@@ -188,7 +192,7 @@ if __name__ == "__main__":
     # ********************** #
     # ***** parameters ***** #
     # ********************** #
-    tfinal = 100
+    tfinal = 80
     nDTOut = 1
     h0 = 1.0
     u0 = 5.0

### 2021-01-13 hydraulic_jump_2D.py

@@ -240,34 +240,28 @@ def setup(initialConditionType=0,
     if solver_type == 'classic':
         if riemann_solver == 'hlle':
             solver = pyclaw.ClawSolver2D(riemann.shallow_hlle_2D)
-            solver.fwave = False
         elif riemann_solver == 'roe':
             solver = pyclaw.ClawSolver2D(riemann.shallow_roe_with_efix_2D)
-            solver.fwave = False
         elif riemann_solver == 'hllc':
             solver = pyclaw.ClawSolver2D(shallow_hllc_2D)
             solver.num_eqn = 3
             solver.num_waves = 3
-            solver.fwave = False
         elif riemann_solver == 'hllemcc':
             solver = pyclaw.ClawSolver2D(shallow_hllemcc_2D)
             solver.num_eqn = 3
             solver.num_waves = 3
-            solver.fwave = False
         elif riemann_solver == 'hllemccroef':
             solver = pyclaw.ClawSolver2D(shallow_hllemccRoEF_2D)
             solver.num_eqn = 3
             solver.num_waves = 3
-            solver.fwave = False
         elif riemann_solver == 'hllem':
             # Not yet implemented
             solver = pyclaw.ClawSolver2D(riemann.shallow_hllem_2D)
-            solver.fwave = False
         elif riemann_solver == 'geoclaw':
             solver = pyclaw.ClawSolver2D(riemann.sw_aug_2D)
-            solver.fwave = True
         else:
             raise Exception('Unrecognized Riemann solver') 
+        solver.fwave = True
         #solver.dimensional_split=True
         #solver.limiters = pyclaw.limiters.tvd.minmod
         solver.cfl_max     = 0.9

### 2021-01-13 plot.py

@@ -75,10 +75,12 @@ def plot_q(frame,
         #
         pl.xticks(fontsize=20)
         pl.yticks(fontsize=20)
-        pl.tight_layout()
         pl.xlim([0,40])
         pl.ylim([0,100])
+        pl.axis('image')
         pl.title("t= "+str(sol.state.t),fontsize=20)
+        pl.tight_layout()
+        pl.xlim([0,40])
 
         
         pl.savefig('./_plots/pcolor/'+str_frame+'.png')

### 2021-01-13 numerics_paper/version2/numerical_chj.tex
Revise discussion of previous work on carbuncle-free solvers.
Also fix capitalization in the bibliography.

@@
 Next we briefly review two commonly-used Riemann solvers: those of Rusanov and
-Roe.  Neither of these solvers deals with the carbuncle instability in fully
+Roe.  We refer to \cite{ketcheson2020riemann} and references therein for
+a full and detailed description of these two Riemann solvers in the context of
+the shallow water equations.
+Neither of these solvers deals with the carbuncle instability in a fully
 satisfactory way.  Rusanov's solver suppresses the carbuncle but is known to
@@
 Later we will combine these two solvers in order to better deal with shock instability.
-We refer to \cite{ketcheson2020riemann} and references therein for
-a full and detailed description of these two Riemann solvers in the context of
-the shallow water equations.  Both of these solvers satisfy the conservation
-property
+Both of these solvers satisfy the conservation property
 \begin{align} \label{rs_conservation}
@@
 
-As we discussed in the introduction, finite volume methods with
-Roe's Riemann solvers are prone to the carbuncle instability for the CHJ.
-In contrast, if we use Rusanov's
-Riemann solver, no carbuncles appear; however, many other important and physical features of
-the solution are dissipated, unless the grid is highly refined.
+As we discussed in the introduction, finite volume methods that use
+a Roe Riemann solver are prone to the carbuncle instability for the CHJ.
+In contrast, methods that use a Rusanov Riemann solver exhibit no carbuncles. 
+However, such methods tend to dissipate other important physical features of
+the solution, unless the grid is highly refined.
 Our main objective is to study the CHJ and to propose a Riemann solver
@@
 Blending these Riemann solvers has been proposed before
-to solve the Euler equations, see e.g.
-\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
-the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
-and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
-In these references, the authors use blending criteria different than ours.
-Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
-For example, in \cite{ismail2009affordable,ismail2009proposed},
-the authors developed carbuncle free solvers for the Euler equations
-via dissipation of entropy.
+to solve the Euler equations (see
+\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy}),
+the shallow water equations (see \cite{bader2014carbuncle,kemm2014note}),
+and even the Navier-Stokes equations (see \cite{nishikawa2008very,ohwada2018simple}).
+Our approach differs from those just cited in that it is based on the local entropy;
+for a related approach in the context of the Euler equations, see
+\cite{ismail2009affordable,ismail2009proposed}.
+
+It is important to recognize also that it is not correct to simply suppress
+the carbuncle phenomenon in all situations, since it is related to a physical
+instability in the Euler equations; see \cite{moschetta2001carbuncle,elling2009carbuncle}.
+A test problem was contrived in \cite{elling2009carbuncle} to illustrate this.
+In the present work, we provide a test problem that possesses a genuine instability
+that leads to carbuncles in many numerical approximations.  This is an ideal
+test for assessing numerical methods, since neither the presence of carbuncles
+nor the complete absence of instability represents the correct behavior.
 
@@
     Flow past a cylinder.  Results for HLLE, HLLEMCC, and the blended solver.
-    \label{fig:steady_outflow}}
+    \label{fig:cylinder_flow}}
 \end{figure}

### 2021-01-14 numerics_paper/version2/numerical_chj.tex
Minor edits to sections 4-5.

@@
 fluxes that make the method second-order accurate:
-\begin{align*}
+\begin{align}\label{correction-fluxes}
   \tilde{F}_{i\pm 1/2}=\frac{1}{2}\sum_p|s_{i\pm 1/2}^p|\left(1-\frac{\Delta t}{\Delta x}|s_{i\pm 1/2}^p|\right)\tilde\W_{i\pm 1/2}^p.
-\end{align*}
+\end{align}
 Here $\tilde{\W}_{i\pm 1/2}^p$ is a TVD-limited version of $\W_{i\pm 1/2}^p$.
@@
 
+In what follows, we use $\W^{p,\Rus}$ to denote the waves in the Rusanov
+solver and simply $\W^p$ to denote the waves in the Roe solver.
 %We close this section by remarking that the fluctuations associated with Rusanov's solver
@@
 \end{align*}
-where $({\bf z})_k$ denotes the $k$-th component of ${\bf z}\in\mathbb{R}^{d+1}$ and $d$ is the number of physical
+where $(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$ and $d$ is the number of physical
 dimensions. In the next section, we use $\theta_i$ to blend Rusanov's and Roe's Riemann solvers.
@@
 Here $\sgn(z)$ is the sign function of $z$.
+
 The second-order LWL method for the one-dimensional problem is given by
-\begin{align}\label{second-order}
-  q_i^{n+1}=q_i^n
-  -\frac{\Delta t}{\Delta x}\left[\A^-\Delta Q_{i+1/2}+\A^+\Delta Q_{i-1/2}\right]
-  -\frac{\Delta t}{\Delta x}\left[\tilde F_{i+1/2}-\tilde F_{i-1/2}\right],
-\end{align}
-where
-\begin{align*}
-  \tilde F_{i+1/2} = \frac{1}{2}\sum_p\lambda_{i+1/2}^p
-  \left(1-\frac{\Delta t}{\Delta x}\lambda_{i+1/2}^p\right)\tilde\W_{i+1/2}^p
-\end{align*}
-are the flux corrections with $\tilde\W_{i+1/2}^p$ being a limited version of $\W_{i+1/2}^p$.
-
+\eqref{second-order_via_fluct} and \eqref{correction-fluxes}.
 To understand the effect of choosing $s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p$
 as the speed of $\W_{i+1/2}^p$,
-let us assume that no limiters are applied; i.e., $\tilde\W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$.
-By doing this, equation \eqref{second-order} can be written as
+let us assume that no limiter is applied; i.e., $\tilde\W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$.
+Then \eqref{second-order_via_fluct} can be written as
 \begin{align*}
@@
 \end{align*}
-which corresponds to the standard Lax-Wendroff method written in terms of fluctuation-like quantities.
-Note that if $\theta_{i\pm 1/2}=0$, we recover the standard second-order Lax-Wendroff method based on
+which is just the standard Lax-Wendroff method written wave propagation form.
+Now taking $\theta_{i\pm 1/2}=0$ corresponds to using a Roe solver while taking
+$\theta_{i\pm 1/2}=1$ corresponds to using a Rusanov solver.
+
+Note that if $\theta_{i\pm 1/2}=0$ we recover the standard second-order Lax-Wendroff method based on
 Roe's Riemann solver.
@@
 \begin{align}\label{rus_rs_aux}
-  \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+  \A^{+,\Rus}\Delta Q_{i-1/2}:=\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}
   = \bff(\bfq_{i})-\bff(\bfq_{i-1}) + \lambda_{i-1/2}^{\max}\left(\W_{i-1/2}^{1,\Rus}+\W_{i-1/2}^{2,\Rus}\right)
@@
 \begin{align*}
-  \lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\max}
+  \lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}
   = \sum_p \left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
@@
 
-We consider a total of six problems. We start with a one-dimensional Riemann problem over a dry bed.
-For this problem we use the first-order method \eqref{first-order_via_fluct},
+We consider a total of six problems. We start in \S \ref{sec:rp_dry_bed} with a
+one-dimensional Riemann problem over a dry bed.
+For this problem we use the first-order methods \eqref{first-order_via_fluct},
 to avoid negative depth values. Although only Rusanov's solver is proven to guarantee
 positivity, we do not get negative values for the water height with any of the first-order methods.
+In \S \ref{sec:rp_wet_bed} we apply the second-order methods to a dam-break problem
+with a wet bed.
 In \S \ref{sec:steady_outflow} we consider a one-dimensional problem with a smooth steady state solution. We observe the expected
@@
 In \S\ref{sec:rp_wet_bed} and \S\ref{sec:1D_chj} we consider one-dimensional problems with strong shocks. We observe similar accuracy with the
-blended solver than with Roe's solver; that is, inducing entropy dissipation near the shocks do not degrade the accuracy
-of the underlying Roe's solver. This extra dissipation, however, prevents the formation of carbuncles in other experiments,
+blended solver or the Roe solver; this suggests that the numerical entropy
+dissipation near the shocks does not degrade the accuracy
+of the underlying Roe solver. This extra dissipation, however, prevents the formation of carbuncles in other experiments,
 which we demonstrate in \S \ref{sec:bow_shock} and \S\ref{sec:2D_chj}.
 The main focus of this work is in the formation of carbuncle instabilities in the two-dimensional CHJ. We present
-an extensive set of experiments about this problem in \S \ref{sec:2D_chj}.
+an extensive set of experiments for this problem in \S \ref{sec:2D_chj}.
 We consider not only Roe's, Rusanov's and the blended solvers, but also the HLLE solver and a solver proposed in \cite{kemm2014note},
@@
 This problem is a canonical example that demonstrates the `entropy glitch' of Roe's solver
-at transonic rarefactions. Since the baseline Roe's solver we use (see \S\ref{sec:roe}) does not contain
+at transonic rarefactions. Since the baseline Roe solver we use (see \S\ref{sec:roe}) does not contain
 an entropy fix, it is important to demonstrate that the blended Riemann solver fixes the entropy glitch.
@@
 with $x_A(t)=x_0-t\sqrt{gh_l}$ and $x_B(t)=x_0+2t\sqrt{gh_l}$.
-We use $g=1$ and solve the problem up to $t=10$.
-In figures \ref{fig:rp_dry_bed_roe}-\ref{fig:rp_dry_bed_blended},
+We take $g=1$ and solve the problem up to $t=10$.
+In Figures \ref{fig:rp_dry_bed_roe}-\ref{fig:rp_dry_bed_blended},
 we show the solution using method \eqref{first-order_via_fluct}
@@
 and in the last column of Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test.
-Note that the entropy glitch is highly reduced.
+Note that the entropy glitch is greatly reduced.
 To remove completely the entropy glitch we could add high-order entropy dissipation following \cite{tadmor2003entropy}
@@
 Since the solution is non-smooth, we expect no more than first order convergence rates. Note that the
-results with the entropy dissipative blended solver are comparable against the results using Roe's solver.
+results with the entropy dissipative blended solver are comparable to the results using Roe's solver.
 That is, inducing entropy dissipation via the blended Riemann solver does not degrade the high-order
@@
 We refer to \cite[\S 21.8.5]{leveque2002finite} for details about the outflow boundary condition.
-If the flow is supercritical ($|\Fr|>1$ where $\Fr$ is the Froude number), the exact solution converges to a
-steady state profile given by the solution of \eqref{steady}.
-Depending on the initial condition, shocks might develop before the steady state is reached.
-Nevertheless, the steady state profile is smooth.
+Regardless of the initial condition, the exact solution converges to a
+steady state profile given by one of the two solutions of \eqref{steady},
+corresponding to subcritical or supercritical flow.  In the present case we
+have imposed a supercritical inflow.
 In Figure \ref{fig:steady_outflow}, we show the solution and $\theta_i$ at different times using
@@
 \eqref{second-order_via_fluct}, using the same Riemann solvers.
+Although the chosen initial
+condition leads to shock formation, the steady state is smooth and close to
+second order convergence is observed for the Roe and blended solvers.
 
@@
   {\scriptsize
-    \includegraphics[scale=0.3]{figures/cylinder_HLLE.png}
-    \qquad \qquad \qquad
-    \includegraphics[scale=0.3]{figures/cylinder_HLLEMCC.png}
-    \qquad \qquad \qquad
-    \includegraphics[scale=0.3]{figures/cylinder_ES.png}
+    \subfloat[HLLE Solver.]{\includegraphics[scale=0.4]{figures/cylinder_HLLE.png}}
+    \quad \quad \quad \quad \quad \quad
+    \subfloat[HLLEMCC Solver.]{\includegraphics[scale=0.4]{figures/cylinder_HLLEMCC.png}}
+    \quad \quad \quad \quad \quad \quad
+    \subfloat[Blended solver.]{\includegraphics[scale=0.4]{figures/cylinder_ES.png}}
   }
@@
 In this regime, using Roe's solver also leads to carbuncle instabilities. In addition, 
-due to the large Froude number, random perturbations might trigger physical instabilities. 
+due to the large Froude number, random perturbations may trigger physical instabilities. 
 If the random perturbations model or mimic real disturbances in a physical experiment, it is then
@@
 In this case, due to the large Froude number, the random perturbations in the inflow
-might trigger physical instabilities, which we want to preserve. At the same time, we
+may trigger physical instabilities, which we want to preserve. At the same time, we
 do not want the solution to develop carbuncle instabilities.

### 2021-01-14 numerics_paper/version2/numerical_chj.tex
More editing of Section 5.

@@
 \subsection{The circular hydraulic jump in two dimensions}\label{sec:2D_chj}
-Let us finally consider the numerical experiments for the CHJ in two-dimensions.
+Let us finally consider the numerical experiments for the CHJ in two dimensions.
 The domain is given by the annulus $\Omega=\{(x,y)~|~ 0.1\leq\sqrt{x^2+y^2}\leq 1\}$.
-The computational mesh is a logically quadrilateral mesh, see Figure \ref{fig:mesh_chj}.
-For all the experiments in this section, the mesh consists of $10^6$ cells. 
+The computational mesh is logically quadrilateral, of the type shown in Figure \ref{fig:mesh_chj}.
+For all the experiments in this section, we use a mesh with 1000 cells in each (radial
+and angular) direction.
 Let
@@
 We choose $|\bfu_I|$ and $h_o$ to generate a radially symmetric shock at $r_s:=\sqrt{x^2+y^2}=0.3$, as described in \S \ref{sec:steady_chj}.
-By selecting the boundary conditions $|\bfu_I|$ and $h_o$ we can move through different regimes.
-In particular, we consider two cases.
-For the first case, the Froude number at the inflow boundary is $\Fr\approx 1.3693$.
-In this case, the CHJ appears to be nearly stable.
-However, using Roe's solver leads to the formation of carbuncle instabilities.
-%
-The second regime is a more aggresive regime with a Froude number of $\Fr\approx 27.3861$ at the inflow boundary. 
-In this regime, using Roe's solver also leads to carbuncle instabilities. In addition, 
-due to the large Froude number, random perturbations may trigger physical instabilities. 
-If the random perturbations model or mimic real disturbances in a physical experiment, it is then
-desirable to properly simulate the unstable behavior of the flow without developing carbuncle instabilities. 
-Therefore, the numerical method must introduce a carbuncle fix without dissipating important features of the
-solution that could lead to physical instabilities. 
-For this reason, we believe this problem is more challenging than the standard bow shock problem.
+By adjusting the boundary conditions $|\bfu_I|$ and $h_o$ we can study different flow regimes.
+We focus on two cases.
+%For the first case, the Froude number at the inflow boundary is $\Fr\approx 1.3693$.
+%In this case, the CHJ appears to be nearly stable.
+%However, using Roe's solver leads to the formation of carbuncle instabilities.
+%%
+%The second regime is a more aggresive regime with a Froude number of $\Fr\approx 27.3861$ at the inflow boundary. 
+%In this regime, using Roe's solver also leads to carbuncle instabilities. In addition, 
+%due to the large Froude number, random perturbations may trigger physical instabilities. 
+%If the random perturbations model or mimic real disturbances in a physical experiment, it is then
+%desirable to properly simulate the unstable behavior of the flow without developing carbuncle instabilities. 
+%Therefore, the numerical method must introduce a carbuncle fix without dissipating important features of the
+%solution that could lead to physical instabilities. 
+%For this reason, we believe this problem is more challenging than the standard bow shock problem.
 
 For most of the experiments we show a Schlieren plot for the water height.
-That is, we plot $||\nabla h||_{\ell^2}$ with a white-to-black logarithm based colormap.
+That is, we plot $||\nabla h||_{\ell^2}$ with a greyscale logarithmic colormap.
 For some experiments we also plot the magnitude of the momentum $|h\bfu|=h\sqrt{u^2+v^2}$.
@@
 The solution clearly develops carbuncle instabilities.
-We could fix these instabilities using Rusanov's, Kemm's or the blended Riemann solvers. 
-In Figure \ref{fig:chj_r1_diff}, we show the corresponding solutions at $t=3$.
-For this test case, all carbuncle free solvers give qualitatively similar solutions;
-however, we expect Rusanov's solver to be less accurate.
+In Figure \ref{fig:chj_r1_diff}, we show solutions at $t=3$ for
+Rusanov's, Kemm's, and the blended Riemann solver.
+For this test case, these three solvers give qualitatively similar solutions,
+all of which are free from carbuncles.
+%however, we expect Rusanov's solver to be less accurate.
 
@@
 \end{align}
-Similar than before, we follow \S \ref{sec:steady_chj} to generate an initial condition with a symmetric CHJ located at
+Again we follow \S \ref{sec:steady_chj} to generate an initial condition with a symmetric CHJ located at
 $\sqrt{x^2+y^2}=0.3$. In Figure \ref{fig:chj_r2_roe}, we show the solution at different times using method
-\eqref{second-order_via_fluct} with Roe's Riemann solver. Just like before, the solution develops
-carbuncle instabilities. Let us now use Rusanov's, Kemm's and the blended solvers to fix the
-carbuncle instabilities. In Figure \ref{fig:chj_r2_diff}, we show the corresponding solutions at $t=0.1$.
+\eqref{second-order_via_fluct} with Roe's Riemann solver. Again the solution develops
+carbuncle instabilities.  In Figure \ref{fig:chj_r2_diff}, we 
+see that with Rusanov's, Kemm's and the blended solver, the carbuncles are avoided.
 {\color{red} TODO: finish this part depending on the results with Kemm's solver.}
 %
-Opposed to what we observed in the first test case, in this situation the Froude number is
+In contrast to what we observed in the first test case, in this situation the Froude number is
 high enough that numerical perturbations start to trigger a response around the CHJ
@@
 Finally, we introduce a physical random perturbation to the inflow boundary condition.
-Let $\epsilon(x,y,t)\in[-0.01,0.01]$ be a random number for a given location $(x,y)$,
-at a given time $t$. 
-The inflow boundary condition is now given by
+We use the same mean values \eqref{bcs_chj_r2}, but now we set
 \begin{subequations}
@@
 with $\tilde h_I = \frac{h_I}{1+\epsilon(x,y,t)}$ and $|\tilde\bfu_I|=|\bfu_I|(1+\epsilon(x,y,t))$.
-Here, just as before, $h_I=0.3$ and $|\bfu_I|=15$. 
-Note that $\beta=r_I\tilde h|\tilde\bfu_I|=r_Ih_I\bfu_I$.
-The outflow boundary condition is given by
-\eqref{bcs_chj_outflow} with $h_o=6.6845019298155357$.
+Here $\epsilon(x,y,t)$ is chosen at each inflow boundary ghost cell and at each
+time step as an i.i.d. random variable from the uniform distribution $[-0.01,0.01]$.
 In this case, due to the large Froude number, the random perturbations in the inflow

### 2021-01-25 numerics_paper/version2/numerical_chj.tex
Cleanup and reorganization of the paper.

--- 2021-01-13 numerics_paper/version2/numerical_chj.tex
@@
 \begin{abstract}
-We study numerical discretizations of the shallow water equations
-for the simulation of the circular hydraulic jump formed
-when a vertical fluid jet impacts a horizontal plate.
-This is a challenging problem that includes a curved steady shock
-that can be physically unstable.  We show that numerical methods are prone to
-either suppress the instability completely or form artificial carbuncles.
+The numerical artifact known as a carbuncle in the context of the Euler
+equations has been shown to appear also in shallow water simulations.
+We study how to deal with carbuncles in the shallow water context,
+with an emphasis on a new test problem involving a steady circular hydraulic
+jump that can be physically unstable.  We show that numerical methods are prone to
+either suppress the instability completely or form carbuncles.
 We test existing cures for the carbuncle as well as a new method,
@@
 \section{Introduction}
+
+\subsection{Numerical shock instabilities}
+
+In \cite{peery1988blunt} a numerical instability was observed to
+appear near the symmetry plane in the simulation of a bow shock.
+This phenomenon, subsequently dubbed ``carbuncle", has been observed by many researchers
+in similar numerical experiments for the Euler equations, and many remedies
+have been proposed, mainly in the form of additional numerical dissipation
+\cite{quirk1997contribution,pandolfi2001numerical,dumbser2004matrix,chauvat2005shock,ismail2009proposed,shen2014stability}.
+Most notably, the HLLE Riemann solver suppresses the carbuncle instability \cite{quirk1997contribution}.
+For a recent review of numerical shock instability and work to alleviate it,
+we refer to \cite[Section 2.5]{simonnumerical}.
+
+Given the similarity of structure between the Euler equations and the shallow
+water equations, it is not surprising that carbuncles appear in numerical
+solutions of the latter as well \cite{kemm2014note}.
+The shallow water carbuncle behaves similarly to the
+Euler carbuncle: it appears when the Roe solver is used, but not when
+the HLL solver is used, and can be suppressed by similar numerical methods
+\cite{kemm2014note,bader2014carbuncle}.
+
+The most common test problems used to investigate carbuncle formation are
+that of bow shock formation or a steady, grid-aligned planar shock.
+In both of these problems, the correct behavior is the formation of a stable
+shock profile, which is achieved easily by simple methods with excessive
+diffusion.  This means these problems are not adequate on their own to
+evaluate methods for practical calculations.
+
+Elling \cite{elling2009carbuncle} proposed instead a problem specifically
+designed to feature a carbuncle as the physical solution.  This has
+been used as a test problem to identify methods that impose excessive dissipation.
+
+Herein we introduce a new and more exacting test problem.  Like some
+of the problems above, it includes an equilibrium solution consisting of a
+steady shock.  Similar to the Elling problem, the equilibrium is unstable.
+However, the correct manifestation of the instability is different from
+the carbuncle.  This allows us to distinguish schemes that yield correct
+behavior from both those that are too dissipative and those that generate
+carbuncles.
 
@@
 
-\subsection{Numerical shock instabilities}
-
-In \cite{peery1988blunt} a numerical instability was observed to
-appear near the symmetry plane in the simulation of a bow shock.
-This phenomenon, subsequently dubbed ``carbuncle", has been observed by many researchers
-in similar numerical experiments for the Euler equations, and many remedies
-have been proposed, mainly in the form of additional numerical dissipation
-\cite{quirk1997contribution,pandolfi2001numerical,dumbser2004matrix,chauvat2005shock,ismail2009proposed,shen2014stability}.
-Most notably, the HLLE Riemann solver suppresses the carbuncle instability \cite{quirk1997contribution}.
-For a recent review of numerical shock instability and work to alleviate it,
-we refer to \cite[Section 2.5]{simonnumerical}.
-
-Given the similarity of structure between the Euler equations and the shallow
-water equations, it is not surprising that carbuncles appear in numerical
-solutions of the latter as well \cite{kemm2014note}.
-The shallow water carbuncle behaves similarly to the
-Euler carbuncle: it appears when the Roe solver is used, but not when
-the HLL solver is used, and can be suppressed by similar numerical methods
-\cite{kemm2014note,bader2014carbuncle}.
-
-As we will see, carbuncles also appear in the numerical solution of the shallow
+As we will see, carbuncles can appear in the numerical solution of the shallow
 water circular hydraulic jump.  This is natural, since the solution involves a
@@
 
-\begin{itemize}
-  \item Describe the carbuncle instability starting with Euler and going to SWEs.
-  \item Mention that we observe the cc instability for the CHJ as well.
-  \item Describe that in this case, the situation is more complicated than in other problems since we
-    expect the correct solution to have physical instabilities triggered by round off error.
-  \item Based on existing literature we know that we can use highly dissipative and robust methods to
-    avoid the formation of carbuncle instabilities. However, in this case, these methods also tend to
-    dissipate other important features of the solution unless the mesh is highly refined.
-  \item We are interested in methods that prevent the formation of cc instability in the CHJ
-    but preserve other features and physical instabilities.
-  \item There are multiple ideas and methods proposed to fix the cc insitability for the Euler
-    and recently for the SWEs.
-  \item It is believed that the cc instability occurs due to lack or reduced viscosity in the shear contact waves.
-  \item Some methods also propose to improve the entropy stability properties of the method. In particular, to induce or guarantee that the entropy is dissipated at shocks.
-  \item Based on some of these ideas we start with Rusanov's method using the cell averages of the solution.
-    This is a first-order method that is entropy stable and positivity preserving. We enhance its robustness by
-    overestimating the wave speed when strong shear is present in the solution. The result is a robust solver
-    that do not develop cc instabilities. However, the method is highly dissipative and, as a result,
-    many other important features of the solution are dissipated.
-  \item We then blend this highly dissipative method with a FV method based on Roe's average.
-  \item After this we introduce the second order corrections from Law-Wendroff's method.
-\end{itemize}
-
-\subsection{Our contribution}
-\begin{itemize}
-\item We present a new problem where the cc instability occurs. In our opinion, this is a more challenging
-  problem than the standard bow shock problem since we expect the solution to be non-steady.
-  We aim to preserve instabilities that are inherently present in the CHJ for certain regimes. And,
-  at the same time, we aim to prevent the cc instability.
-\item We explore the use of entropy dissipation and entropy stability as cure of the cc instability for the SWEs.
-\item We propose a specific method to blend the highly dissipative and robust Rusanov's RP and the
-  accurate 3-wave Roe's solver for the SWEs.
-\item We demonstrate that this scheme prevents the formation of cc instabilities not only
-  in the well-known bow shock problem, but also in the CHJ while preserving other important features of
-  the solution.
-\end{itemize}
-
-\section{The shallow water circular hydraulic jump}
+\subsection{Outline}
+In this work, we describe and study the circular hydraulic jump as a new
+test problem for shallow water discretizations.  We find that although some
+existing methods behave acceptably on previous test problems, they
+are not capable of providing accurate solutions for the circular hydraulic jump
+across the range of flow regimes we study.  We develop a new entropy-stable
+Riemann solver for the shallow water equations that generates solutions with
+the correct behavior.
+
+\section{Numerical methods for the shallow water equations}
 We consider the shallow water model in two horizontal dimensions:
@@
 The gravitational force is proportional to $g$.
-%The system \eqref{eq:sw} can be written in vector form as
-%\begin{align}
-%    q_t + f(q)_x + g(q)_y & = 0.
-%\end{align}
-
-\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
-In this work, we consider the initial boundary value problem consisting of the
-shallow water model \eqref{eq:sw} in an annular domain
+The system \eqref{eq:sw} can be written in vector form as
 \begin{align}
-r_\textup{jet} \le r \le r_\infty
-\end{align}
-where $r = \sqrt{x^2+y^2}$, with prescribed inflow at $r=r_\textup{jet}$
-and prescribed outflow at $r=r_\infty$.
-The domain and boundary conditions are rotationally symmetric.
-By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
-\begin{subequations} \label{eq:rsw}
-\begin{align}
-    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
-    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
-\end{align}
-\end{subequations}
-where the depth $h$ and radial velocity $u$ are functions of the single spatial coordinate $r$.
-Steady-state solutions of \eqref{eq:rsw} satisfy
-\begin{subequations}\label{steady}
-\begin{align}
-    rhu & = C \\
-    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
-\end{align}
-\end{subequations}
-for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
-We see that two types of steady profiles exist, depending on whether the flow
-is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
-include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.
-
-The steady, rotationally-symmetric circular hydraulic jump involves supercritical
-flow for $r<r_0$ and subcritical flow for $r>r_0$, where $r_0$ is the jump radius.
-The jump itself takes the form of a stationary shock wave.  The Rankine-Hugoniot jump
-conditions specify that for such a shock,
-\begin{align} \label{eq:RH}
-    h_+ - h_- & = \frac{-3h_- + \sqrt{h_-^2 + 8 h_- u_-^2/g}}{2} = \frac{3h_-}{2}\left(\sqrt{1+\frac{8}{9}(F_-^2-1)}-1\right),
-\end{align}
-where the subscripts $+, -$ denote states just inside or outside the jump radius, respectively.
-
-A steady-state, rotationally symmetric solution can be given for an annular region with prescribed
-flow at the inner and outer boundaries as follows:
-
-\begin{enumerate}
-    \item Specify the depth and velocity at the inner boundary (near the jet) and outer boundary.
-    \item Integrate \eqref{eq:dh} from both boundaries.
-    \item Find a radius $r_0$ where the matching condition \eqref{eq:RH} is satisfied.
-\end{enumerate}
-Due to the nature of solutions of \eqref{eq:dh}, it can be shown that the required jump
-radius $r_0$ always exists if the prescribed flow is supercritical at the inner boundary
-and subcritical at the outer boundary.
-
-Numerical tests suggest that the solution of \eqref{eq:rsw} tends rapidly approaches that
-just described under general initial conditions as long as the flow at the jet is
-supercritical and the outer outflow is subcritical.  This solution remains steady
-if the rate of inflow and outflow are matched.  The stability of this solution in
-the face of non-rotationally-symmetric perturbations is an important question not
-only in the shallow water context but for more realistic fluid models and physically.
-It will play an important role in the results we present later.
-The rotationally-symmetric steady state is a useful initial condition for studies
-of CHJ stability.
-
-\section{Numerical methods}
+    q_t + f(q)_x + g(q)_y & = 0.
+\end{align}
+
 In this work we study the behavior of certain shock-capturing finite volume
 methods based on the use of Riemann solvers.  For simplicity, we discuss these
-solvers in the context of a 1-dimensional mesh.  
+solvers in the context of a 1-dimensional problem and mesh.  
 In the numerical experiments in \S\ref{sec:num} we use second-order Strang
@@
 \subsection{Wave propagation methods}\label{sec:waveprop}
-The numerical solution $q_i$ represents the average value of the set of
+Let $Q_i$ represent the average value of the set of
 conserved quantities over cell $i$, which extends from $x_\imh$ to $x_\iph$.
 At each time step and at each interface $x_\imh$, we approximately solve the
-Riemann problem with initial states $(q^n_{i-1},q^n_i)$.  The approximate
+Riemann problem with initial states $(Q^n_{i-1},Q^n_i)$.  The approximate
 solution consists of a set of traveling discontinuities or waves $\W^p_\imh$
@@
 We use the wave propagation framework developed by LeVeque \cite{leveque1997wave, leveque2002finite}
-and employed in the Clawpack software \cite{clawpack,pyclaw-sisc}, which implements the LWL scheme
+and employed in the Clawpack software \cite{clawpack,pyclaw-sisc}, which implements the 
+Lax-Wendroff-LeVeque (LWL) scheme
 \begin{align}\label{second-order_via_fluct}
-  q_i^{n+1} = q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
+  Q_i^{n+1} = Q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
   -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
@@
 \begin{align}\label{first-order_via_fluct}
-  q_i^{n+1} = q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
+  Q_i^{n+1} = Q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
 \end{align}
@@
 \begin{align}\label{first-order_FV}
-  q_i^{n+1}=q_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
+  Q_i^{n+1}=Q_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
 \end{align}
@@
 
-Next we briefly review two commonly-used Riemann solvers: those of Rusanov and
-Roe.  We refer to \cite{ketcheson2020riemann} and references therein for
-a full and detailed description of these two Riemann solvers in the context of
-the shallow water equations.
-Neither of these solvers deals with the carbuncle instability in a fully
+Next we briefly review three existing Riemann solvers for the shallow water equations.
+We start with the methods of Rusanov and of Roe.  Neither of these solvers
+deals with the carbuncle instability in a fully
 satisfactory way.  Rusanov's solver suppresses the carbuncle but is known to
@@
 Later we will combine these two solvers in order to better deal with shock instability.
-Both of these solvers satisfy the conservation property
+We refer to \cite{ketcheson2020riemann} and references therein for
+a full and detailed description of these two Riemann solvers in the context of
+the shallow water equations.  Both of these solvers satisfy the conservation
+property
 \begin{align} \label{rs_conservation}
@@
 
+\subsection{HLLEMCC Riemann solver} \label{sec:kemm}
+Several specialized Riemann solvers have been proposed in order to deal
+with carbuncles.  Among them, only that proposed by Kemm has been implemented
+and tested on the shallow water equations \cite{kemm2014note,bader2014carbuncle}.
+The idea behind that solver, known as the HLLEMCC solver, is to combine
+the HLLE and Roe solvers in such a way that the resulting method behaves
+like Roe's when away from potential carbuncle instabilities, and like HLLE
+where the potential for such instabilities arises.   This is done using
+an indicator function based on the local residual of the Rankine-Hugoniot
+condition for the shear wave.  The method requires some parameters that
+may need to be tuned for a specific problem.  We will consider this method
+in the numerical tests below.
+
 \section{An entropy-based blending of Rusanov and Roe}
 
-As we discussed in the introduction, finite volume methods that use
-a Roe Riemann solver are prone to the carbuncle instability for the CHJ.
-In contrast, methods that use a Rusanov Riemann solver exhibit no carbuncles. 
-However, such methods tend to dissipate other important physical features of
-the solution, unless the grid is highly refined.
-Our main objective is to study the CHJ and to propose a Riemann solver
+As we discussed in the introduction, finite volume methods that use a
+Roe Riemann solver are prone to exhibit the carbuncle instability.
+Rusanov solver, like the HLLE solver, suppresses these instabilities but may also
+dissipate other important features of the solution, unless the grid is highly refined.
+In this section, we propose a Riemann solver
 that is carbuncle-free but does not dissipate important features of the solution.
@@
 Blending these Riemann solvers has been proposed before
-to solve the Euler equations (see
-\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy}),
-the shallow water equations (see \cite{bader2014carbuncle,kemm2014note}),
-and even the Navier-Stokes equations (see \cite{nishikawa2008very,ohwada2018simple}).
-Our approach differs from those just cited in that it is based on the local entropy;
-for a related approach in the context of the Euler equations, see
-\cite{ismail2009affordable,ismail2009proposed}.
-
-It is important to recognize also that it is not correct to simply suppress
-the carbuncle phenomenon in all situations, since it is related to a physical
-instability in the Euler equations; see \cite{moschetta2001carbuncle,elling2009carbuncle}.
-A test problem was contrived in \cite{elling2009carbuncle} to illustrate this.
-In the present work, we provide a test problem that possesses a genuine instability
-that leads to carbuncles in many numerical approximations.  This is an ideal
-test for assessing numerical methods, since neither the presence of carbuncles
-nor the complete absence of instability represents the correct behavior.
+to solve the Euler equations, see e.g.
+\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
+the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
+and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
+In these references, the authors use blending criteria different than ours.
+Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
+For example, in \cite{ismail2009affordable,ismail2009proposed},
+the authors developed carbuncle free solvers for the Euler equations
+via dissipation of entropy.
 
@@
 
-\subsection{Flow past a cylinder}\label{sec:bow_shock}
-In this section we consider the formation of a bow shock when a
-uniform flow encounters a cylindrical obstacle.  This problem has
-been studied previously in the context of carbuncle formation in
-several works for the Euler equations and in
-\cite{kemm2014note,bader2014carbuncle} for the shallow water equations.
-We present results for existing solvers as a form of verification,
-along with results for the new blended solver.  The main question of
-interest is the formation of carbuncles.  It is known, for instance,
-that the Roe solver incorrectly generates a carbuncle at the center
-of the bow shock.
-
-We model only the flow on the upstream side of the cylinder, since
-our interest is in the resolution of the bow shock.  We take the domain
-$[0,40]\times[0,100]$ with a cylinder of radius 20 centered at $(40,50)$.
-Reflecting boundary conditions are imposed at the surface of the cylinder,
-along with outflow conditions along the rest of the right edge of the domain.
-The depth and velocity are set initially and (for all time) at the other
-boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.  We use a $160 \times 400$
-uniform Cartesian grid.
-
-We show results at $t=80$, after the flow has reached
-a steady state.  With a Roe solver, negative values are generated at an early
-time, leading to failure of the solver.  Therefore, we show results only for
-HLLE, HLLEMCC, and the blended solver.  These three methods give very similar
-results.
-
-\begin{figure}[!h]
-  {\scriptsize
-    \includegraphics[scale=0.3]{figures/cylinder_HLLE.png}
-    \qquad \qquad \qquad
-    \includegraphics[scale=0.3]{figures/cylinder_HLLEMCC.png}
-    \qquad \qquad \qquad
-    \includegraphics[scale=0.3]{figures/cylinder_ES.png}
-  }
-  \caption{
-    Flow past a cylinder.  Results for HLLE, HLLEMCC, and the blended solver.
-    \label{fig:cylinder_flow}}
-\end{figure}
-
-We have also conducted a more challenging version of this test, in which
-the velocity within the domain is initially zero.  In that case, the HLLE
-and blended solvers give carbuncle-free results similar to Figure \ref{fig:steady_outflow},
-but HLLEMCC exhibits a carbuncle.
+\subsection{Steady shock around a cylinder}\label{sec:bow_shock}
+
+\section{Numerical study of the circular hydraulic jump}
+
+\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
+In this section, we consider the initial boundary value problem consisting of the
+shallow water model \eqref{eq:sw} in an annular domain
+\begin{align}
+r_\textup{jet} \le r \le r_\infty
+\end{align}
+where $r = \sqrt{x^2+y^2}$, with prescribed inflow at $r=r_\textup{jet}$
+and prescribed outflow at $r=r_\infty$.
+The domain and boundary conditions are rotationally symmetric.
+By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
+\begin{subequations} \label{eq:rsw}
+\begin{align}
+    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
+    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
+\end{align}
+\end{subequations}
+where the depth $h$ and radial velocity $u$ are functions of the single spatial coordinate $r$.
+Steady-state solutions of \eqref{eq:rsw} satisfy
+\begin{subequations}\label{steady}
+\begin{align}
+    rhu & = C \\
+    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
+\end{align}
+\end{subequations}
+for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
+We see that two types of steady profiles exist, depending on whether the flow
+is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
+include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.
+
+The steady, rotationally-symmetric circular hydraulic jump involves supercritical
+flow for $r<r_0$ and subcritical flow for $r>r_0$, where $r_0$ is the jump radius.
+The jump itself takes the form of a stationary shock wave.  The Rankine-Hugoniot jump
+conditions specify that for such a shock,
+\begin{align} \label{eq:RH}
+    h_+ - h_- & = \frac{-3h_- + \sqrt{h_-^2 + 8 h_- u_-^2/g}}{2} = \frac{3h_-}{2}\left(\sqrt{1+\frac{8}{9}(F_-^2-1)}-1\right),
+\end{align}
+where the subscripts $+, -$ denote states just inside or outside the jump radius, respectively.
+
+A steady-state, rotationally symmetric solution can be given for an annular region with prescribed
+flow at the inner and outer boundaries as follows:
+
+\begin{enumerate}
+    \item Specify the depth and velocity at the inner boundary (near the jet) and outer boundary.
+    \item Integrate \eqref{eq:dh} from both boundaries.
+    \item Find a radius $r_0$ where the matching condition \eqref{eq:RH} is satisfied.
+\end{enumerate}
+Due to the nature of solutions of \eqref{eq:dh}, it can be shown that the required jump
+radius $r_0$ always exists if the prescribed flow is supercritical at the inner boundary
+and subcritical at the outer boundary.
+
+Numerical tests suggest that the solution of \eqref{eq:rsw} tends rapidly approaches that
+just described under general initial conditions as long as the flow at the jet is
+supercritical and the outer outflow is subcritical.  This solution remains steady
+if the rate of inflow and outflow are matched.  The stability of this solution in
+the face of non-rotationally-symmetric perturbations is an important question not
+only in the shallow water context but for more realistic fluid models and physically.
+It will play an important role in the results we present later.
+The rotationally-symmetric steady state is a useful initial condition for studies
+of CHJ stability.
 

### 2021-02-09 numerics_paper/version2/numerical_chj.tex

--- 2021-01-14 numerics_paper/version2/numerical_chj.tex
@@
 
+\subsection{HLLEMCC Riemann solver} \label{sec:kemm}
+Several specialized Riemann solvers have been proposed in order to deal
+with carbuncles.  Among them, only that proposed by Kemm has been implemented
+and tested on the shallow water equations \cite{kemm2014note,bader2014carbuncle}.
+The idea behind that solver, known as the HLLEMCC solver, is to combine
+the HLLE and Roe solvers in such a way that the resulting method behaves
+like Roe's when away from potential carbuncle instabilities, and like HLLE
+where the potential for such instabilities arises.   This is done using
+an indicator function based on the local residual of the Rankine-Hugoniot
+condition for the shear wave.  The method requires some parameters that
+may need to be tuned for a specific problem.  We will consider this method
+in the numerical tests below.
+
 \section{An entropy-based blending of Rusanov and Roe}
 
-As we discussed in the introduction, finite volume methods that use
-a Roe Riemann solver are prone to the carbuncle instability for the CHJ.
+As we discussed in the introduction, finite volume methods that use a
+Roe Riemann solver are prone to exhibit the carbuncle instability.
 In contrast, methods that use a Rusanov Riemann solver exhibit no carbuncles. 
@@
 
+\section{Numerical study of the circular hydraulic jump}
+
+\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
+In this section, we consider the initial boundary value problem consisting of the
+shallow water model \eqref{eq:sw} in an annular domain
+\begin{align}
+r_\textup{jet} \le r \le r_\infty
+\end{align}
+where $r = \sqrt{x^2+y^2}$, with prescribed inflow at $r=r_\textup{jet}$
+and prescribed outflow at $r=r_\infty$.
+The domain and boundary conditions are rotationally symmetric.
+By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
+\begin{subequations} \label{eq:rsw}
+\begin{align}
+    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
+    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
+\end{align}
+\end{subequations}
+where the depth $h$ and radial velocity $u$ are functions of the single spatial coordinate $r$.
+Steady-state solutions of \eqref{eq:rsw} satisfy
+\begin{subequations}\label{steady}
+\begin{align}
+    rhu & = C \\
+    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
+\end{align}
+\end{subequations}
+for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
+We see that two types of steady profiles exist, depending on whether the flow
+is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
+include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.
+
+The steady, rotationally-symmetric circular hydraulic jump involves supercritical
+flow for $r<r_0$ and subcritical flow for $r>r_0$, where $r_0$ is the jump radius.
+The jump itself takes the form of a stationary shock wave.  The Rankine-Hugoniot jump
+conditions specify that for such a shock,
+\begin{align} \label{eq:RH}
+    h_+ - h_- & = \frac{-3h_- + \sqrt{h_-^2 + 8 h_- u_-^2/g}}{2} = \frac{3h_-}{2}\left(\sqrt{1+\frac{8}{9}(F_-^2-1)}-1\right),
+\end{align}
+where the subscripts $+, -$ denote states just inside or outside the jump radius, respectively.
+
+A steady-state, rotationally symmetric solution can be given for an annular region with prescribed
+flow at the inner and outer boundaries as follows:
+
+\begin{enumerate}
+    \item Specify the depth and velocity at the inner boundary (near the jet) and outer boundary.
+    \item Integrate \eqref{eq:dh} from both boundaries.
+    \item Find a radius $r_0$ where the matching condition \eqref{eq:RH} is satisfied.
+\end{enumerate}
+Due to the nature of solutions of \eqref{eq:dh}, it can be shown that the required jump
+radius $r_0$ always exists if the prescribed flow is supercritical at the inner boundary
+and subcritical at the outer boundary.
+
+Numerical tests suggest that the solution of \eqref{eq:rsw} tends rapidly approaches that
+just described under general initial conditions as long as the flow at the jet is
+supercritical and the outer outflow is subcritical.  This solution remains steady
+if the rate of inflow and outflow are matched.  The stability of this solution in
+the face of non-rotationally-symmetric perturbations is an important question not
+only in the shallow water context but for more realistic fluid models and physically.
+It will play an important role in the results we present later.
+The rotationally-symmetric steady state is a useful initial condition for studies
+of CHJ stability.
+
 \subsection{The circular hydraulic jump in two dimensions}\label{sec:2D_chj}

### 2021-02-09 numerics_paper/version2/numerical_chj.tex
Write out full blended scheme.

@@
 \end{align*}
-and $\W_{i-1/2}^p=\alpha_{i-1/2}^p\hat r_{i-1/2}^p$, with
+and $\W_{i-1/2}^{p,\Roe}=\alpha_{i-1/2}^p\hat r_{i-1/2}^p$, with
 \begin{align*}
@@
 
-In what follows, we use $\W^{p,\Rus}$ to denote the waves in the Rusanov
-solver and simply $\W^p$ to denote the waves in the Roe solver.
 %We close this section by remarking that the fluctuations associated with Rusanov's solver
@@
 \subsection{Blended Riemann solver}\label{sec:blended_rs}
+In what follows, we use $\W^{p,\Rus}$ to denote the waves in the Rusanov
+solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.  We propose
+a new scheme consisting of the second-order LWL method \eqref{second-order_via_fluct} using
+the first-order fluxes
+\begin{align}
+  \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
+  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-\theta_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfq_{i+1}-\bfq_{i})
+\end{align}
+and correction fluxes \eqref{correction-fluxes} where
+\begin{align} \label{s_p}
+    s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p
+\end{align}
+with
+\begin{align}\label{lambda_p}
+  \lambda_{i+1/2}^p := \theta_{i+1/2}\lambda_{i+1/2}^{\max} + (1-\theta_{i+1/2})|\hat \lambda_{i+1/2}^p| + \lambda_\iph^{\min}.
+\end{align}
+The term $\lambda_\iph^{\min}$ is used to ensure that the scheme would be entropy stable
+if the second-order corrections were neglected.  The normalized entropy
+residual $\theta$ controls whether the scheme behaves more like that of Rusanov
+or Roe.  Specifically, if $\lambda_\iph^{\min}=0$ and $\theta_\iph=0$, this scheme
+is equivalent to that of Rusanov, while if $\lambda_\iph^{\min}=0$ and $\theta_\iph=1$, 
+this scheme is equivalent to that of Roe.
+
 In this section, we use \eqref{Ri}, which is a normalized measurement of the entropy residual \eqref{ent_residual},
@@
 and $\hat A_{i+1/2}$ is Roe's averaged flux Jacobian (see \S \ref{sec:roe}).
-
-The first-order method is given simply by \eqref{first-order_FV}
-with the numerical flux $F_{i+1/2}$ given by \eqref{num_flux_ev}.
-In smooth regions $\theta_{i+1/2}\approx 0$ so the numerical flux is essentially Roe's flux.
-Near discontinuities or steep gradients, $\theta_{i+1/2}$ becomes large. In the limit $\theta_{i+1/2}\approx 1$
-the numerical flux is essentially Rusanov's flux.
+This flux is equal to the Rusanov flux if $\theta_\iph=0$ and the Roe flux if $\theta_\iph=1$.
+We expect that $\theta_\iph\approx 0 $ in smooth regions, while $\theta_\iph\approx 1$ near
+shocks.
+
+%The first-order method is given simply by \eqref{first-order_FV}
+%with the numerical flux $F_{i+1/2}$ given by \eqref{num_flux_ev}.
 
 The first-order method can also be written in terms of fluctuations via \eqref{first-order_via_fluct}.
-To do this we consider the waves ($\W^p_{i\pm 1/2}$, with $p=1,2,3$)
-from Roe's solver and define
+To do this we define the blended wave speed
 \begin{align}\label{lambda_p}
@@
 \begin{align}
-  \apdq_\imh & = \frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^p, \\
-  \amdq_\iph & = \frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^p.
+  \apdq_\imh & = \frac{1}{2}\sum_p \left(\hat\lambda_{i-1/2}^p + \lambda_{i-1/2}^p\right)\W_{i-1/2}^{p,\Roe}, \\
+  \amdq_\iph & = \frac{1}{2}\sum_p \left(\hat\lambda_{i+1/2}^p - \lambda_{i+1/2}^p\right)\W_{i+1/2}^{p,\Roe}.
 \end{align}
 \end{subequations}
-
-To implement a second-order LWL scheme based on this solver, in addition to
-these fluctuations we must define a set of waves and corresponding speeds.
+These can be used in \eqref{first-order_via_fluct} to obtain a first-order scheme.
+To implement the correction fluxes required for the second-order scheme \eqref{second-order_via_fluct},
+we must also define a set of waves and corresponding speeds.
 Using only the waves from the Roe solver, it is in general not possible
@@
 %which is not possible to obtain in general (since the problem is overdetermined).
-Instead, we consider
-$s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p$
- and study the consequences of this choice once the second-order corrections are applied.
-Here $\sgn(z)$ is the sign function of $z$.
-
-The second-order LWL method for the one-dimensional problem is given by
-\eqref{second-order_via_fluct} and \eqref{correction-fluxes}.
-To understand the effect of choosing $s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p$
-as the speed of $\W_{i+1/2}^p$,
-let us assume that no limiter is applied; i.e., $\tilde\W_{i\pm 1/2}^p\equiv \W_{i\pm 1/2}^p$.
-Then \eqref{second-order_via_fluct} can be written as
+Instead, we set
+\begin{align} \label{s_p}
+    s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p
+\end{align}
+where $\sgn(z)$ is the sign function of $z$.
+Using this in \eqref{correction-fluxes} and \eqref{ev_fluctuations} in 
+\eqref{second-order_via_fluct}, we obtain (in the absence of limiting)
+the second-order scheme
 \begin{align*}
@@
 \end{align*}
-which is just the standard Lax-Wendroff method written wave propagation form.
-Now taking $\theta_{i\pm 1/2}=0$ corresponds to using a Roe solver while taking
-$\theta_{i\pm 1/2}=1$ corresponds to using a Rusanov solver.
-
 Note that if $\theta_{i\pm 1/2}=0$ we recover the standard second-order Lax-Wendroff method based on
 Roe's Riemann solver.
-In the numerical experiments in \S\ref{sec:num}, we obtain second-order experimental
-order of convergence (EOC) for problems with smooth solution, which is expected since
-$\theta_{i\pm 1/2}\approx 0$ in smooth regions.
 Similarly, if $\theta_{i\pm 1/2}=1$, we recover the Lax-Wendroff method based on Rusanov's Riemann solver,

### 2021-02-10 numerics_paper/version2/numerical_chj.tex
Reorganize section 3.

@@
 \end{align}
-which, for the conservative Riemann solvers we use herein, can also be written
-as a first-order flux-differencing scheme:
-\begin{align}\label{first-order_FV}
-  Q_i^{n+1}=Q_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right],
-\end{align}
-where $\bfF_{i+1/2}$ is a numerical approximation of the flux at $x_\iph$.
+We note that, for the conservative Riemann solvers we use herein, the LWL scheme
+\eqref{second-order_via_fluct} can also be written purely in flux-differencing
+form:
+\begin{align}\label{flux-differencing-form}
+  Q_i^{n+1}=Q_i^n-\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right]
+ -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
+\end{align}
+with appropriately chosen first order fluxes $\bfF_{i+1/2}$ and correction
+fluxes $\tilde{F}_{i-1/2}$.
 
@@
 the solution, unless the grid is highly refined.
-Our main objective is to study the CHJ and to propose a Riemann solver
-that is carbuncle-free but does not dissipate important features of the solution.
-To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers.
-Our blending criterion is based on the residual of the entropy (in)equality,
-which induces dissipation of entropy near discontinuities.
-Blending these Riemann solvers has been proposed before
-to solve the Euler equations (see
+In this section we propose a Riemann solver
+that is carbuncle-free but avoids dissipating important features of the solution.
+To do so we first use a combination of Rusanov's and Roe's solvers, switching
+between them based on a local measure of the entropy residual.
+Previous works have also proposed blending Riemann solvers with different
+amounts of dissipation in the context of the Euler equations (see
 \cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy}),
@@
 
-It is important to recognize also that it is not correct to simply suppress
-the carbuncle phenomenon in all situations, since it is related to a physical
-instability in the Euler equations; see \cite{moschetta2001carbuncle,elling2009carbuncle}.
-A test problem was contrived in \cite{elling2009carbuncle} to illustrate this.
-In the present work, we provide a test problem that possesses a genuine instability
-that leads to carbuncles in many numerical approximations.  This is an ideal
-test for assessing numerical methods, since neither the presence of carbuncles
-nor the complete absence of instability represents the correct behavior.
+Since Roe's solver is not entropy stable, the blended solver is also
+not guaranteed to be entropy stable.  We therefore include an additional term
+that is chosen to give entropy stability, at least if the second-order correction
+fluxes are neglected.  The complete proposed scheme can be written in
+flux-differencing form 
+We propose
+a new scheme consisting of the second-order LWL method \eqref{flux-differencing-form} using
+the first-order fluxes
+\begin{align} \label{blended-flux-1}
+  \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
+  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-\theta_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfq_{i+1}-\bfq_{i})
+\end{align}
+and correction fluxes \eqref{correction-fluxes} based on the Roe waves and
+the modified speeds
+\begin{align} \label{s_p}
+    s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p
+\end{align}
+with
+\begin{align}\label{lambda_p}
+  \lambda_{i+1/2}^p := \theta_{i+1/2}\lambda_{i+1/2}^{\max} + (1-\theta_{i+1/2})|\hat \lambda_{i+1/2}^p| + \lambda_\iph^{\min}.
+\end{align}
+Here $\mathbb{I}\in\mathbb{R}^{3\times 3}$ is the identity matrix,
+$\lambda_{i+1/2}^{\max}$ is the
+upper bound on the wave speed used in Rusanov's Riemann solver (see \S \ref{sec:rusanov}),
+and $\hat A_{i+1/2}$ is Roe's averaged flux Jacobian (see \S \ref{sec:roe}).
+
+In the following sections we describe the properties of this solver and
+explain how $\theta_\iph$ and $\lambda^{\min}_\iph$ are chosen.
+
+%It is important to recognize also that it is not correct to simply suppress
+%the carbuncle phenomenon in all situations, since it is related to a physical
+%instability in the Euler equations; see \cite{moschetta2001carbuncle,elling2009carbuncle}.
+%A test problem was contrived in \cite{elling2009carbuncle} to illustrate this.
+%In the present work, we provide a test problem that possesses a genuine instability
+%that leads to carbuncles in many numerical approximations.  This is an ideal
+%test for assessing numerical methods, since neither the presence of carbuncles
+%nor the complete absence of instability represents the correct behavior.
 
@@
 where $(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$ and $d$ is the number of physical
-dimensions. In the next section, we use $\theta_i$ to blend Rusanov's and Roe's Riemann solvers.
-
-\subsection{Blended Riemann solver}\label{sec:blended_rs}
-In what follows, we use $\W^{p,\Rus}$ to denote the waves in the Rusanov
-solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.  We propose
-a new scheme consisting of the second-order LWL method \eqref{second-order_via_fluct} using
-the first-order fluxes
-\begin{align}
-  \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
-  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-\theta_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfq_{i+1}-\bfq_{i})
-\end{align}
-and correction fluxes \eqref{correction-fluxes} where
-\begin{align} \label{s_p}
-    s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p
-\end{align}
-with
-\begin{align}\label{lambda_p}
-  \lambda_{i+1/2}^p := \theta_{i+1/2}\lambda_{i+1/2}^{\max} + (1-\theta_{i+1/2})|\hat \lambda_{i+1/2}^p| + \lambda_\iph^{\min}.
-\end{align}
-The term $\lambda_\iph^{\min}$ is used to ensure that the scheme would be entropy stable
-if the second-order corrections were neglected.  The normalized entropy
-residual $\theta$ controls whether the scheme behaves more like that of Rusanov
-or Roe.  Specifically, if $\lambda_\iph^{\min}=0$ and $\theta_\iph=0$, this scheme
-is equivalent to that of Rusanov, while if $\lambda_\iph^{\min}=0$ and $\theta_\iph=1$, 
-this scheme is equivalent to that of Roe.
-
-In this section, we use \eqref{Ri}, which is a normalized measurement of the entropy residual \eqref{ent_residual},
-to blend Rusanov's and Roe's Riemann solvers. To do this, we define the blended numerical flux:
-\begin{align}\label{num_flux_ev}
-  F_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
-  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max} I + (1-\theta_{i+1/2})|\hat A_{i+1/2}| \right)(\bfq_{i+1}-\bfq_{i}).
-\end{align}
-Here $I\in\mathbb{R}^{3\times 3}$ is the identity matrix,
-$\theta_{i+1/2}=\max\{\theta_{i+1},\theta_i\}$, $\lambda_{i+1/2}^{\max}$ is the
-upper bound on the wave speed used in Rusanov's Riemann solver (see \S \ref{sec:rusanov}),
-and $\hat A_{i+1/2}$ is Roe's averaged flux Jacobian (see \S \ref{sec:roe}).
-This flux is equal to the Rusanov flux if $\theta_\iph=0$ and the Roe flux if $\theta_\iph=1$.
-We expect that $\theta_\iph\approx 0 $ in smooth regions, while $\theta_\iph\approx 1$ near
-shocks.
-
-%The first-order method is given simply by \eqref{first-order_FV}
-%with the numerical flux $F_{i+1/2}$ given by \eqref{num_flux_ev}.
-
-The first-order method can also be written in terms of fluctuations via \eqref{first-order_via_fluct}.
-To do this we define the blended wave speed
-\begin{align}\label{lambda_p}
-  \lambda_{i+1/2}^p := \theta_{i+1/2}\lambda_{i+1/2}^{\max} + (1-\theta_{i+1/2})|\hat \lambda_{i+1/2}^p|,
-\end{align}
-where $\hat\lambda_{i+1/2}^p$ is the $p$-th eigenvalue of Roe's solver corresponding to the
-interface $i+1/2$. Note that $\lambda_{i+1/2}^p\geq 0$.
-The fluctuations are then given by
+dimensions.  In \eqref{blended-flux-1} we need values of $\theta$ at the cell interfaces, for
+which we use
+$$
+    \theta_\iph = \max(\theta_i, \theta_{i+1}).
+$$
+We expect that $\theta_\iph\approx 0 $ in smooth regions,
+while $\theta_\iph\approx 1$ near shocks.
+The value of $\theta_\iph$ controls whether the first-order flux \eqref{blended-flux-1}
+behaves more like that of Rusanov or Roe.  Specifically, if
+$\lambda_\iph^{\min}=0$ and $\theta_\iph=1$, the flux is equivalent to that of
+Rusanov, while if $\lambda_\iph^{\min}=0$ and $\theta_\iph=0$, it is equivalent
+to that of Roe.
+
+In fact, with the choice \eqref{s_p}, the correction fluxes also match
+those of Rusanov or Roe in the appropriate limit, as shown in the next section.
+
+\subsection{The correction fluxes}\label{sec:blended_rs}
+In this section we explain the choice of wave speeds \eqref{s_p}.
+For the moment, we consider \eqref{blended-flux-1}-\eqref{lambda_p} without
+entropy stabilization; i.e. we set $\lambda^{\min}_\iph=0$
+for now.
+We use $\W^{p,\Rus}$ to denote the waves in the Rusanov
+solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.
+The first-order method given by \eqref{blended-flux-1} can also be written in
+terms of fluctuations via \eqref{first-order_via_fluct}:
 \begin{subequations}\label{ev_fluctuations}
@@
 \end{subequations}
-These can be used in \eqref{first-order_via_fluct} to obtain a first-order scheme.
 To implement the correction fluxes required for the second-order scheme \eqref{second-order_via_fluct},
@@
 Using only the waves from the Roe solver, it is in general not possible
-to choose speeds that yield the fluctuations \eqref{ev_fluctuations}.
-%In the framework of LeVeque, for a given interface $i+1/2$,
-%we would need to consider a wave speed $s_{i+1/2}^p$ so that
-%\begin{align*}
-%  \hat \lambda_{i+1/2}^p+\lambda_{i+1/2}^p = s_{i+1/2}^p+|s_{i+1/2}^p|, \qquad
-%  \hat \lambda_{i+1/2}^p-\lambda_{i+1/2}^p = s_{i+1/2}^p-|s_{i+1/2}^p|,
-%\end{align*}
-%which is not possible to obtain in general (since the problem is overdetermined).
-Instead, we set
-\begin{align} \label{s_p}
-    s_{i+1/2}^p=\sgn(\hat\lambda_{i+1/2}^p)\lambda_{i+1/2}^p
-\end{align}
-where $\sgn(z)$ is the sign function of $z$.
+to choose speeds that yield the fluctuations \eqref{ev_fluctuations},
+and we instead use \eqref{s_p}.
 Using this in \eqref{correction-fluxes} and \eqref{ev_fluctuations} in 
@@
 \end{align*}
-Note that if $\theta_{i\pm 1/2}=0$ we recover the standard second-order Lax-Wendroff method based on
-Roe's Riemann solver.
-Similarly, if $\theta_{i\pm 1/2}=1$, we recover the Lax-Wendroff method based on Rusanov's Riemann solver,
-which can be seen by writing the fluctuations in Rusanov's method in terms of Roe's waves and speeds.
+It is clear again that if $\theta_{i\pm 1/2}=0$ we recover the standard
+second-order Lax-Wendroff method based on Roe's Riemann solver.
+We now show that if $\theta_{i\pm 1/2}=1$, we recover the Lax-Wendroff method
+based on Rusanov's Riemann solver.
 
@@
 
-\subsection{Entropy-stable Riemann solver}\label{sec:entropy_stable}
-The Riemann solver presented in the previous section blends the solvers of Rusanov and Roe based on a
-local measurement of the entropy residual. The objective is to induce dissipation of entropy
-near discontinuities. However, we have no guarantee that methods \eqref{first-order_via_fluct}
-or \eqref{second-order_via_fluct} based on this Riemann solver are (locally) entropy stable.
-In this section, we further modify the first-order method
-\eqref{first-order_via_fluct} in order to guarantee local entropy stability.
+\subsection{Entropy stabilization}\label{sec:entropy_stable}
+In the previous section we neglected the term $\lambda^{\min}_\iph$ in
+\eqref{blended-flux-1}.  In this section, we show how this term is computed,
+in order to guarantee local entropy stability of the scheme given by \eqref{blended-flux-1}
+when the correction fluxes are neglected.
 
@@
 
-To enforce \eqref{es_cond} we further modify the flux \eqref{num_flux_ev}:
-\begin{align*}
-  \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
-  - \frac{1}{2} \left( \theta_{i+1/2}\lambda_{i+1/2}^{\max}\mathbb{I} + (1-\theta_{i+1/2})|\hat A_{i+1/2}| +\lambda_{i+1/2}^{\min}\mathbb{I}\right)(\bfq_{i+1}-\bfq_{i}).
-\end{align*}
-Here $\lambda_{i+1/2}^{\min}$ is yet to be defined.
-We can write the numerical fluxes in terms of $\lambda^p_{i\pm 1/2}$ \eqref{lambda_p}
-and Roe's waves $\W_{i\pm 1/2}^p$.
-Doing so, we get
-\begin{align*}
-  \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
-  - \frac{1}{2}\sum_p(\lambda_{i+1/2}^p+\lambda_{i+1/2}^{\min})\W_{i+1/2}^p.
-\end{align*}
-
-We set
+After comparing \eqref{blended-flux-1} and the condition above, we arrive at the value
 \begin{align}\label{lambda_min}
-  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\left[\bff(\bfq_{i+1})+\bff(\bfq_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\efp_{i+1}-\efp_i)}{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}
-\end{align}
-which guarantees \eqref{es_cond}, and therefore \eqref{ent_ineq}.
+  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\left[\bff(\bfq_{i+1})+\bff(\bfq_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\efp_{i+1}-\efp_i)}{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}.
+\end{align}
+With this choice, \eqref{blended-flux-1} guarantees \eqref{es_cond}, and therefore \eqref{ent_ineq}.
 
@@
 We follow the setup in \cite[\S 4.1.2]{delestre2013swashes}.
-The domain is given by $\Omega=(0,L=10)$, and the initial condition is $hu(x,0)=0$ and
+The domain is given by $x\in(0,10)$, and the initial condition is $hu(x,0)=0$ and
 \begin{align*}
@@
   \begin{cases}
-    h_l, & \mbox{ if }, x\leq x_0=5 \\
-    h_r, &\mbox{ otherwise},
+    h_l & x\leq x_0=5 \\
+    0 & x >5,
   \end{cases}
 \end{align*}
-with $h_l=5\times 10^{-3}$ and $h_r=0$. In our experiments we use $h_r=1\times 10^{-15}$ to avoid division by zero.
+with $x_0=5$ and $h_l=5\times 10^{-3}$.  In our experiments we use $h_r=1\times
+10^{-15}$ to avoid division by zero.
 The exact solution, which can also be found in \cite{delestre2013swashes} and references therein, is
@@
   \begin{cases}
-    0, &\mbox{ if } x\leq x_A(t), \\
-    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
-    0, &\mbox{ if } x_B(t)\leq x,
+    0, &\mbox{ if } x\leq x_-(t), \\
+    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_-(t) < x\leq x_+(t), \\
+    0, &\mbox{ if } x_+(t)\leq x,
   \end{cases}
 \end{align*}
-with $x_A(t)=x_0-t\sqrt{gh_l}$ and $x_B(t)=x_0+2t\sqrt{gh_l}$.
+with $x_\pm(t)=x_0 \pm t\sqrt{gh_l}$.
 We take $g=1$ and solve the problem up to $t=10$.
@@
 
-For this particular problem, using method \eqref{first-order_via_fluct} with the blended Riemann solver
-leads to $\lambda_{i\pm 1/2}^{\min}=0$ for all the experiments, but this is not true in general.
-Finally, let us demonstrate the effect of $\lambda_{i\pm 1/2}^{\min}$ by imposing $\theta_{i}=0$ in \eqref{lambda_p},
+For this particular problem, using method \eqref{first-order_via_fluct} with
+the blended Riemann solver leads to $\lambda_{i\pm 1/2}^{\min}=0$ for all the experiments.
+We can artificially activate the entropy stabilization by imposing $\theta_{i}=0$ in \eqref{lambda_p},
 which is equivalent to using Roe's solver with extra dissipation given by $\lambda_{i\pm 1/2}^{\min}$.
-In Figure \ref{fig:rp_dry_bed_roe_with_lmin}, we show the solution using different refinements,
-and in the last column of Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test.
-Note that the entropy glitch is greatly reduced.
+In Figure \ref{fig:rp_dry_bed_roe_with_lmin} we show the solution,
+and in the last column of Table \ref{table:rp_dry_bed} we summarize the results of a convergence test.
+The entropy glitch is still noticeable but greatly reduced compared to the
+solution from Roe's method without entropy stabilization.
 To remove completely the entropy glitch we could add high-order entropy dissipation following \cite{tadmor2003entropy}
@@
 We follow the setup in \cite[\S 4.1.1]{delestre2013swashes}.
-The domain is $\Omega=(0,L)$ with $L=10$, the initial condition is given by
+The domain is $x\in(0,10)$ and the initial condition is given by
 $hu(x,0)=0$ and
@@
   \begin{cases}
-    h_l & \mbox{ if } x\in(0,x_0], \\
-    h_r & \mbox{ if } x\in(x_0,L),
+    h_l & x < x_0 \\
+    h_r & x > x_0
   \end{cases}
 \end{align*}
-with $x_0=5$, $h_l=0.005$ and $h_r=0.001$.
-The exact solution, which can be found in \cite{delestre2013swashes} and references therein,
-is given by
+with $x_0=5$, $h_l=0.005$, and $h_r=0.001$.  The exact solution, which can be found in
+\cite{delestre2013swashes} and references therein, is given by
 \begin{align*}
@@
 $-8gh_rc_m^2\left(\sqrt{gh_l}-c_m\right)^2+\left(c_m^2-gh_r\right)^2\left(c_m^2+gh_r\right)=0$.
-We use $g=1$ and solve the problem up to the final time $t=5$ considering method \eqref{second-order_via_fluct}
+We use $g=1$ and solve the problem up to the final time $t=5$ using the
+second-order method \eqref{second-order_via_fluct}
 with Roe's, Rusanov's and the blended solvers. The solution with different refinement levels
@@
 results with the entropy dissipative blended solver are comparable to the results using Roe's solver.
-That is, inducing entropy dissipation via the blended Riemann solver does not degrade the high-order
+That is, imposing entropy dissipation via the blended Riemann solver does not degrade the high-order
 accuracy properties of Roe's solver.
@@
 \subsection{Steady outflow}\label{sec:steady_outflow}
-Let us consider a one-dimensional problem with a constant inflow and a steady outflow.
-The domain is given by $\Omega=(0.1,1)$, the initial condition is
-$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$,
-and the right boundary condition is set to outflow.
+Under rotational symmetry, the two-dimensional shallow water equations
+can be reduced to the system
+\begin{subequations} \label{eq:rsw}
+\begin{align}
+    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
+    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0, \label{mom1}
+\end{align}
+\end{subequations}
+where $r=\sqrt{x^2+y^2}$ is the radial coordinate.  Here we consider a problem
+in the annulus $r\in(0.1,1)$ with constant inflow at $r=0.1$ and outflow at $r=1$.
+The initial condition is
+$h(r,t=0)=0.1$, $hu(r,t=0)=0$, the inner boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$,
+and the outer boundary condition is set to outflow.
 We refer to \cite[\S 21.8.5]{leveque2002finite} for details about the outflow boundary condition.
@@
 Although the chosen initial
-condition leads to shock formation, the steady state is smooth and close to
+condition leads initially to shock formation, the steady state is smooth and close to
 second order convergence is observed for the Roe and blended solvers.
@@
     In the second row we plot $\theta_i$, which is given by \eqref{Ri}.
-    In all simulations we consider $\Delta x=1/400$.
+    In all simulations we take $\Delta x=1/400$.
     \label{fig:steady_outflow}}
@@
 By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
-\begin{subequations} \label{eq:rsw}
-\begin{align}
-    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
-    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
-\end{align}
-\end{subequations}
+\eqref{eq:rsw},
 where the depth $h$ and radial velocity $u$ are functions of the single spatial coordinate $r$.

### 2021-02-10 numerics_paper/version2/numerical_chj.tex
Corrections to last commit.

@@
 carbuncles.
+
+In the present work, we provide a test problem that possesses a genuine instability
+that leads to carbuncles in many numerical approximations.  This is an ideal
+test for assessing numerical methods, since neither the presence of carbuncles
+nor the complete absence of instability represents the correct behavior.
 
@@
 %A test problem was contrived in \cite{elling2009carbuncle} to illustrate this.
-%In the present work, we provide a test problem that possesses a genuine instability
-%that leads to carbuncles in many numerical approximations.  This is an ideal
-%test for assessing numerical methods, since neither the presence of carbuncles
-%nor the complete absence of instability represents the correct behavior.
+%
 
@@
 solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.
-The first-order method given by \eqref{blended-flux-1} can also be written in
+The first-order method given by the numerical flux \eqref{blended-flux-1} can also be written in
 terms of fluctuations via \eqref{first-order_via_fluct}:
@@
     h_l & x\leq x_0=5 \\
-    0 & x >5,
+    h_r & x >5,
   \end{cases}
@@
 \end{align*}
-with $x_\pm(t)=x_0 \pm t\sqrt{gh_l}$.
+with $x_-(t)=x_0 - t\sqrt{gh_l}$ and $x_+=x_0+2t\sqrt{gh_l}$.
 We take $g=1$ and solve the problem up to $t=10$.

### 2021-02-15 numerics_paper/version2/numerical_chj.tex

@@
 \newcommand{\entvar}{\eta'}
-\newcommand{\bfq}{{q}}
+\newcommand{\bfq}{{Q}}
 \newcommand{\bfx}{{x}}
@@
 Here $\tilde{\W}_{i\pm 1/2}^p$ is a TVD-limited version of $\W_{i\pm 1/2}^p$.
-We will sometimes work with
-the first-order method obtained by setting the correction fluxes to zero:
-\begin{align}\label{first-order_via_fluct}
-  Q_i^{n+1} = Q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
-\end{align}
+
 We note that, for the conservative Riemann solvers we use herein, the LWL scheme
@@
 fluxes $\tilde{F}_{i-1/2}$.
+We will sometimes work with
+the first-order method obtained by setting the correction fluxes to zero:
+\begin{align}\label{first-order_via_fluct}
+  Q_i^{n+1} = Q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
+ = Q_i^n -\frac{\Delta t}{\Delta x}\left[\bfF_{i+1/2}-\bfF_{i-1/2}\right].
+\end{align}
 
@@
 solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.
-The first-order method given by the numerical flux \eqref{blended-flux-1} can also be written in
-terms of fluctuations via \eqref{first-order_via_fluct}:
+The first-order method given by \eqref{first-order_via_fluct} with the numerical
+flux \eqref{blended-flux-1} can also be written in
+terms of fluctuations:
 \begin{subequations}\label{ev_fluctuations}
@@
   \sum_p
-  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}(\lambda_{i+1/2}^p)^2 \right)\W_{i+1/2}^p
+  \left[\left(\hat\lambda_{i+1/2}^p -\frac{\Delta t}{\Delta x}(\lambda_{i+1/2}^p)^2 \right)\W_{i+1/2}^{p,\Roe}
   +
-  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}(\lambda_{i-1/2}^p)^2 \right)\W_{i-1/2}^p\right],
+  \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}(\lambda_{i-1/2}^p)^2 \right)\W_{i-1/2}^{p,\Roe}\right],
 \end{align*}
@@
 From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get
-$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfq_i-\bfq_{i-1}=\sum_p\W_{i-1/2}^p$.
-Using \eqref{rs_conservation}, \eqref{rus_rs_aux} becomes
+$\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfq_i-\bfq_{i-1}=\sum_p\W_{i-1/2}^{p,\Roe}$.
+Using \eqref{rs_conservation} and \eqref{ev_fluctuations}, \eqref{rus_rs_aux} becomes
 \begin{align*}
   \lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus}
-  = \sum_p \left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p
+  = \sum_p \left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^{p,\Roe}
   -\lambda_{i-1/2}^{\max}\W_{i-1/2}^{2,\Rus},
@@
 \begin{align}
-  \A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^p,
+  \A^{+,\Rus}\Delta Q_{i-1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i-1/2}^p + \lambda_{i-1/2}^{\max}\right)\W_{i-1/2}^{p,\Roe},
 \end{align}
@@
 \begin{align}
-  \A^{-,\Rus}\Delta Q_{i+1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i+1/2}^p - \lambda_{i+1/2}^{\max}\right)\W_{i+1/2}^p.
+  \A^{-,\Rus}\Delta Q_{i+1/2}=\frac{1}{2}\sum_p\left(\hat \lambda_{i+1/2}^p - \lambda_{i+1/2}^{\max}\right)\W_{i+1/2}^{p,\Roe}.
 \end{align}
@@
 \eqref{blended-flux-1}.  In this section, we show how this term is computed,
-in order to guarantee local entropy stability of the scheme given by \eqref{blended-flux-1}
-when the correction fluxes are neglected.
+in order to guarantee local entropy stability of the scheme given by using
+\eqref{blended-flux-1} in \eqref{first-order_via_fluct}.
 
@@
 With this choice, \eqref{blended-flux-1} guarantees \eqref{es_cond}, and therefore \eqref{ent_ineq}.
+Note that here we have used the identities
+\begin{align*}
+|\hat{A}_\iph|(Q_{i+1}-Q_i) & = \sum_p |\hat{\lambda}^p_\iph| \W^{p,\Roe}_\iph, & 
+Q_{i+1}-Q_i & = \sum_p \W^{p,\Roe}_\iph.
+\end{align*}
 
@@
 \begin{remark}[About the effect of $\lambda_{i\pm 1/2}^{\min}$ for Roe's solver.]
-  If we set $\theta_{i\pm 1/2}=0, ~\forall i$ in \eqref{lambda_p}, then the Riemann solver becomes Roe's solver.
+  If we set $\theta_{i\pm 1/2}=0$ and $\lambda^{\min}_\iph=0, ~\forall i$ in \eqref{lambda_p}, then the Riemann solver becomes Roe's solver.
   In this case, it is well known that transonic rarefactions develop non-physical shocks,

### 2021-02-16 numerics_paper/version2/numerical_chj.tex
update paper (not finished yet)

@@
   \apdq_\imh := \sum_p\left(s_{i-1/2}^p\right)^+\W_{i-1/2}^p, \qquad
-  \amdq_\iph := \sum_p\left(s_{i+1/2}^p\right)^-\W_{i+1/2}^p, \\
+  \amdq_\iph := \sum_p\left(s_{i+1/2}^p\right)^-\W_{i+1/2}^p, 
 \end{align}
@@
 \subsection{Rusanov's Riemann solver}\label{sec:rusanov}
-The Rusanov solver approximates the Riemann solution by with a single wave traveling in each
+The Rusanov's solver approximates the Riemann solution by with a single wave traveling in each
 direction.  Both waves are assumed to travel with speed $\lambda^{\max}_\imh$, which is an
@@
 \begin{align}\label{system_for_alphas}
-  \left[\hat \bfr^1_{i-1/2} ~\hat \bfr^2_{i-1/2} ~\hat \bfr^3_{i-1/2}\right]
+  \sum_p\W_{i-1/2}^{p,\Roe} = \left[\hat \bfr^1_{i-1/2} ~\hat \bfr^2_{i-1/2} ~\hat \bfr^3_{i-1/2}\right]
   \begin{bmatrix}
@@
   \end{bmatrix}
-  =\Delta Q_{i-1/2}:=q_i-q_{i-1}.
+  =\Delta Q_{i-1/2}:=\bfq_i-\bfq_{i-1}.
 \end{align}
@@
 
-\section{An entropy-based blending of Rusanov and Roe}
+\section{An entropy-based blending of Rusanov and Roe}\label{sec:blended_rs}
 
@@
 Roe Riemann solver are prone to exhibit the carbuncle instability.
-In contrast, methods that use a Rusanov Riemann solver exhibit no carbuncles. 
+In contrast, methods that use Rusanov's Riemann solver exhibit no carbuncles. 
 However, such methods tend to dissipate other important physical features of
@@
 
-\subsection{The correction fluxes}\label{sec:blended_rs}
+\subsection{The correction fluxes}
 In this section we explain the choice of wave speeds \eqref{s_p}.
@@
 for now.
-We use $\W^{p,\Rus}$ to denote the waves in the Rusanov
+We use $\W^{p,\Rus}$ to denote the waves in the Rusanov's
 solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.
@@
 the second-order scheme
-\begin{align*}
-  q_i^{n+1}=q_i^n
+\begin{align}\label{LWL}
+  \bfq_i^{n+1}=\bfq_i^n
   &-\frac{\Delta t}{2 \Delta x}
@@
   \left(\hat\lambda_{i-1/2}^p +\frac{\Delta t}{\Delta x}(\lambda_{i-1/2}^p)^2 \right)\W_{i-1/2}^{p,\Roe}\right],
-\end{align*}
+\end{align}
 It is clear again that if $\theta_{i\pm 1/2}=0$ we recover the standard
@@
 \end{align}
-From \eqref{system_for_alphas} and \eqref{rusanov_waves} we get
+From \eqref{rusanov_waves} and \eqref{system_for_alphas} we get
 $\W^{1,\Rus}_{i-1/2}+\W^{2,\Rus}_{i-1/2}=\bfq_i-\bfq_{i-1}=\sum_p\W_{i-1/2}^{p,\Roe}$.
@@
 \end{subequations}
+To get the second-order Lax-Wendroff method based on Rusanov's Riemann solver,
+plug the fluctuations \eqref{fluct_rus} into \eqref{second-order_via_fluct} where 
+$\tilde F_{i\pm 1/2}$ is given by \eqref{correction-fluxes} with $|s^p_{i\pm 1/2}|=\lambda_{i\pm 1/2}^{\max}$. 
+By doing this, we get \eqref{LWL} (since $\theta_{i\pm 1/2}=1 \implies \lambda_{i\pm 1/2}^p=\lambda_{i\pm 1/2}^{\max}$).
 
@@
 In the previous section we neglected the term $\lambda^{\min}_\iph$ in
-\eqref{blended-flux-1}.  In this section, we show how this term is computed,
-in order to guarantee local entropy stability of the scheme given by using
-\eqref{blended-flux-1} in \eqref{first-order_via_fluct}.
+\eqref{blended-flux-1}.  In this section, we follow \cite{kuzmin2020algebraic}
+and show how this term is computed, in order to guarantee local entropy stability
+of the scheme given by using \eqref{blended-flux-1} in \eqref{first-order_via_fluct}.
 
@@
 
-After comparing \eqref{blended-flux-1} and the condition above, we arrive at the value
+After plugging \eqref{blended-flux-1} into the condition \eqref{es_cond}, we arrive at the value
 \begin{align}\label{lambda_min}
@@
 of cell averages with centered point values is a sufficient way to test their accuracy.
-
-We consider a total of six problems. We start in \S \ref{sec:rp_dry_bed} with a
+In all experiments we use $g=1$.
+
+We consider a total of five problems. We start in \S \ref{sec:rp_dry_bed} with a
 one-dimensional Riemann problem over a dry bed.
@@
 In \S \ref{sec:rp_wet_bed} we apply the second-order methods to a dam-break problem
-with a wet bed.
+with a wet bed. This problem contains strong shocks. We observe similar accuracy with the
+blended solver or the Roe solver; this suggests that the extra numerical entropy
+dissipation that the blended solver introduces near the shocks does not degrade the accuracy
+of the underlying Roe solver. This extra dissipation, however, prevents the formation of carbuncles in other experiments,
+which we demonstrate in \S \ref{sec:bow_shock} and \S\ref{sec:2D_chj}.
 In \S \ref{sec:steady_outflow} we consider a one-dimensional problem with a smooth steady state solution. We observe the expected
@@
 is not degraded since the extra dissipation is not applied in smooth regions.
-In \S\ref{sec:rp_wet_bed} and \S\ref{sec:1D_chj} we consider one-dimensional problems with strong shocks. We observe similar accuracy with the
-blended solver or the Roe solver; this suggests that the numerical entropy
-dissipation near the shocks does not degrade the accuracy
-of the underlying Roe solver. This extra dissipation, however, prevents the formation of carbuncles in other experiments,
-which we demonstrate in \S \ref{sec:bow_shock} and \S\ref{sec:2D_chj}.
 The main focus of this work is in the formation of carbuncle instabilities in the two-dimensional CHJ. We present
 an extensive set of experiments for this problem in \S \ref{sec:2D_chj}.
-We consider not only Roe's, Rusanov's and the blended solvers, but also the HLLE solver and a solver proposed in \cite{kemm2014note},
+We consider not only Roe's, Rusanov's and the blended solvers,
+%but also the HLLE solver and a solver proposed in \cite{kemm2014note},
+but also the solver proposed in \cite{kemm2014note},
 which is designed to remove the carbuncle instabilities in the shallow water equations.
@@
 \end{align*}
-with $x_0=5$ and $h_l=5\times 10^{-3}$.  In our experiments we use $h_r=1\times
+with $h_l=5\times 10^{-3}$.  In our experiments we use $h_r=1\times
 10^{-15}$ to avoid division by zero.
@@
   \begin{cases}
-    0, &\mbox{ if } x\leq x_-(t), \\
-    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_-(t) < x\leq x_+(t), \\
-    0, &\mbox{ if } x_+(t)\leq x,
+    0, &\mbox{ if } x\leq x_A(t), \\
+    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
+    0, &\mbox{ if } x_B(t) < x,
   \end{cases}
 \end{align*}
-with $x_-(t)=x_0 - t\sqrt{gh_l}$ and $x_+=x_0+2t\sqrt{gh_l}$.
-We take $g=1$ and solve the problem up to $t=10$.
+with $x_A(t)=x_0 - t\sqrt{gh_l}$ and $x_B=x_0+2t\sqrt{gh_l}$.
+We solve the problem up to $t=10$.
 In Figures \ref{fig:rp_dry_bed_roe}-\ref{fig:rp_dry_bed_blended},
@@
   \begin{cases}
-    h_l & x < x_0 \\
+    h_l & x \leq x_0 \\
     h_r & x > x_0
@@
     2(\sqrt{gh_l}-c_m), & \mbox{ if } x_B(t)<x\leq x_C(t), \\
-    0, &\mbox{ if } x\leq x_C(t),
+    0, &\mbox{ if } x < x_C(t),
   \end{cases}
@@
 $-8gh_rc_m^2\left(\sqrt{gh_l}-c_m\right)^2+\left(c_m^2-gh_r\right)^2\left(c_m^2+gh_r\right)=0$.
-We use $g=1$ and solve the problem up to the final time $t=5$ using the
+We solve the problem up to the final time $t=5$ using the
 second-order method \eqref{second-order_via_fluct}
@@
 \begin{align}
-    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
-    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0, \label{mom1}
+    (rh)_t + (rhu)_r &= 0, \label{mass1} \\
+    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r &= 0, \label{mom1}
 \end{align}
@@
 uniform Cartesian grid.
-
-We show results at $t=80$, after the flow has reached
-a steady state.  With a Roe solver, negative values are generated at an early
+%
+In Figure \ref{fig:steady_outflow}, we show results at $t=80$, after the flow has reached
+a steady state.  With Roe's solver, negative values (of the water height) are generated at an early
 time, leading to failure of the solver.  Therefore, we show results only for
-HLLE, HLLEMCC, and the blended solver.  These three methods give very similar
+Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.  These three methods give very similar
 results.
@@
   {\scriptsize
-    \subfloat[HLLE Solver.]{\includegraphics[scale=0.4]{figures/cylinder_HLLE.png}}
+    %\subfloat[Rusanov's solver.]{\includegraphics[scale=0.4]{figures/cylinder_hlle.png}}
+    \subfloat[Rusanov's solver.]{\includegraphics[scale=0.4]{figures/cylinder_rusanov.png}}
     \quad \quad \quad \quad \quad \quad
-    \subfloat[HLLEMCC Solver.]{\includegraphics[scale=0.4]{figures/cylinder_HLLEMCC.png}}
+    \subfloat[HLLEMCC solver.]{\includegraphics[scale=0.4]{figures/cylinder_hllemcc.png}}
     \quad \quad \quad \quad \quad \quad
@@
   \caption{
-    Flow past a cylinder.  Results for HLLE, HLLEMCC, and the blended solver.
+    Flow past a cylinder.  Results for Rusanov's, HLLEMCC, and the blended solvers.
     \label{fig:cylinder_flow}}
@@
 We have also conducted a more challenging version of this test, in which
-the velocity within the domain is initially zero.  In that case, the HLLE
-and blended solvers give carbuncle-free results similar to Figure \ref{fig:steady_outflow},
+the velocity within the domain is initially zero.  In that case, Rusanov's 
+and the blended solvers give carbuncle-free results similar to Figure \ref{fig:steady_outflow},
 but HLLEMCC exhibits a carbuncle.
@@
 In Figure \ref{fig:chj_r1_diff}, we show solutions at $t=3$ for
-Rusanov's, Kemm's, and the blended Riemann solver.
+Rusanov's, HLLEMCC, and the blended Riemann solvers.
 For this test case, these three solvers give qualitatively similar solutions,
@@
     \begin{tabular}{ccc} 
-      \includegraphics[scale=0.3]{figures/chj_r1_roe_p1.png} &
-      \includegraphics[scale=0.3]{figures/chj_r1_roe_p2.png} &
-      \includegraphics[scale=0.3]{figures/chj_r1_roe_p3.png}
+      \includegraphics[scale=0.38]{figures/chj_r1_roe_p1.png} &
+      \includegraphics[scale=0.38]{figures/chj_r1_roe_p2.png} &
+      \includegraphics[scale=0.38]{figures/chj_r1_roe_p3.png}
     \end{tabular}
@@
 
-  \subfloat[Solution at $t=3$ using (left) Rusanov's, (middle) Kemm's and (right) the blended solvers.\label{fig:chj_r1_diff}]{
+  \subfloat[Solution at $t=3$ using (left) Rusanov's, (middle) HLLEMCC and (right) the blended solvers.\label{fig:chj_r1_diff}]{
     \begin{tabular}{ccc}
-      \includegraphics[scale=0.3]{figures/chj_r1_llf_p3.png} &
-      \includegraphics[scale=0.5]{figures/cross.png} &
-      \includegraphics[scale=0.3]{figures/chj_r1_blended_p3.png}
+      \includegraphics[scale=0.38]{figures/chj_r1_llf_p3.png} &
+      \includegraphics[scale=0.38]{figures/chj_r1_hllemcc_p3.png} &
+      \includegraphics[scale=0.38]{figures/chj_r1_blended_p3.png}
     \end{tabular}
@@
       In all cases we show the Schlieren plot for the water height $h$.
-    %The color scale in the plots go from (blue) $0.0746$ to (red) $0.3738$,
-    %which corresponds to the minimum and maximum values of the momentum at $t=0$.
+      In (a), for the plot that corresponds to $t=3$, we zoom close to the jump
+      and plot the y-momentum. 
     \label{fig:chj_r1}}
@@
 \eqref{second-order_via_fluct} with Roe's Riemann solver. Again the solution develops
-carbuncle instabilities.  In Figure \ref{fig:chj_r2_diff}, we 
-see that with Rusanov's, Kemm's and the blended solver, the carbuncles are avoided.
-{\color{red} TODO: finish this part depending on the results with Kemm's solver.}
+carbuncle instabilities. For this regime, we obtained negative values for the water height with
+the HLLEMCC solver, which lead to failure of the solver. Therefore, we only show results with
+Rusanov's and the blended solvers, see Figure \ref{fig:chj_r2_diff}. With both solvers, the carbuncles are avoided.
 %
@@
   \vspace{10pt}
-  \subfloat[Solution at $t=0.1$ using (left) Rusanov's, (middle) Kemm's and (right) the blended solvers.\label{fig:chj_r2_diff}]{
+  \subfloat[Solution at $t=0.1$ using (left) Rusanov's and (right) the blended solvers.\label{fig:chj_r2_diff}]{
     \begin{tabular}{ccc}
-      \includegraphics[scale=0.3]{figures/chj_r2_llf_t0p1.png} &
-      \quad \includegraphics[scale=0.5]{figures/cross.png} \quad &
-      \includegraphics[scale=0.3]{figures/chj_r2_blended_t0p1.png}
+      \includegraphics[scale=0.3]{figures/chj_r2_llf_t0p1.png} \qquad &
+      %\quad \includegraphics[scale=0.5]{figures/cross.png} \quad &
+      \qquad \includegraphics[scale=0.3]{figures/chj_r2_blended_t0p1.png}
     \end{tabular}
@@
     We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
-    (middle) Kemm's and (right) the blended solvers. 
+    (middle) HLLEMCC and (right) the blended solvers. 
     In all cases we superimpose the Schlieren plot for the water height $h$

### 2021-02-16 numerics_paper/version2/numerical_chj.tex
more changes

@@
 method \eqref{second-order_via_fluct} with Roe's Riemann solver.
-The solution clearly develops carbuncle instabilities.
+The solution clearly develops carbuncle instabilities, which are evident in the inset figure at $t=3$.
 In Figure \ref{fig:chj_r1_diff}, we show solutions at $t=3$ for
@@
     \begin{tabular}{ccc} 
-      \includegraphics[scale=0.38]{figures/chj_r1_roe_p1.png} &
-      \includegraphics[scale=0.38]{figures/chj_r1_roe_p2.png} &
-      \includegraphics[scale=0.38]{figures/chj_r1_roe_p3.png}
+      \includegraphics[scale=0.3]{figures/chj_r1_roe_p1.png} &
+      \includegraphics[scale=0.3]{figures/chj_r1_roe_p2.png} &
+      \includegraphics[scale=0.3]{figures/chj_r1_roe_p3.png}
     \end{tabular}
@@
     \begin{tabular}{ccc}
-      \includegraphics[scale=0.38]{figures/chj_r1_llf_p3.png} &
-      \includegraphics[scale=0.38]{figures/chj_r1_hllemcc_p3.png} &
-      \includegraphics[scale=0.38]{figures/chj_r1_blended_p3.png}
+      \includegraphics[scale=0.3]{figures/chj_r1_llf_p3.png} &
+      \includegraphics[scale=0.3]{figures/chj_r1_hllemcc_p3.png} &
+      \includegraphics[scale=0.3]{figures/chj_r1_blended_p3.png}
     \end{tabular}
@@
 \eqref{second-order_via_fluct} with Roe's Riemann solver. Again the solution develops
-carbuncle instabilities. For this regime, we obtained negative values for the water height with
+carbuncle instabilities, which are clearly seen in the inset figure at $t=3$.
+For this regime, we obtained negative values for the water height with
 the HLLEMCC solver, which lead to failure of the solver. Therefore, we only show results with
@@
 high enough that numerical perturbations start to trigger a response around the CHJ
-(see the middle and right panels of Figure \ref{fig:chj_r2_diff}).
+(see the right panel of Figure \ref{fig:chj_r2_diff}).
 For the resolution that we consider, this behavior is dissipated by the extra numerical viscosity introduced by
@@
     In all cases we show the Schlieren plot for the water height $h$.
-    %The color scale in the plots go from (blue) $0.45$ to (red) $4.5$,
-    %which corresponds to the minimum and maximum values of the momentum at $t=0$.
-  }
+    In (a), for the plot that corresponds to $t=0.11$, we zoom close to the jump and plot the y-momentum.}
 \end{figure}
@@
   \centering
-  \subfloat[$t=5$]{
     \includegraphics[scale=0.3]{figures/chj_r2_rusanov_t5p0.png} \quad
-    \includegraphics[scale=0.5]{figures/cross.png} \quad
-    \includegraphics[scale=0.3]{figures/chj_r2_blended_t5p0.png}
-  }
+    %\includegraphics[scale=0.5]{figures/cross.png} \quad
+    \quad \includegraphics[scale=0.3]{figures/chj_r2_blended_t5p0.png}
   \caption{CHJ at $t=5$ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
-    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's,
-    (middle) HLLEMCC and (right) the blended solvers. 
+    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's
+    and (right) the blended solvers. 
     In all cases we superimpose the Schlieren plot for the water height $h$
@@
 the blended Riemann solvers, respectively.
+In the right panel of Figure \ref{fig:chj_r2wp_blended}, we show an inset figure 
+with the y-momentum. Note that even though the solution is highly unstable 
+(with the blended Riemann solver), no visible carbuncle instabilities are developed.
 
@@
   \caption{CHJ at different times with a random perturbation at the inflow boundary.
-    We consider method \eqref{second-order_via_fluct} with (a) Roe's and (b) the blended
-    Riemann solvers. In all cases we show the Schlieren plot for the water height $h$.}
+    We consider method \eqref{second-order_via_fluct} with different Riemann solvers. 
+    In all cases we show the Schlieren plot for the water height $h$.
+    In (b), for the plot that corresponds to $t=5$, we zoom close to the jump and plot the y-momentum.}
 \end{figure}

### 2021-02-17 numerics_paper/version2/numerical_chj.tex

@@
 \subsection{Rusanov's Riemann solver}\label{sec:rusanov}
-The Rusanov's solver approximates the Riemann solution by with a single wave traveling in each
+Rusanov's solver approximates the Riemann solution by with a single wave traveling in each
 direction.  Both waves are assumed to travel with speed $\lambda^{\max}_\imh$, which is an
@@
 for now.
-We use $\W^{p,\Rus}$ to denote the waves in the Rusanov's
+We use $\W^{p,\Rus}$ to denote the waves in the Rusanov
 solver and $\W^{p,\Roe}$ to denote the waves in the Roe solver.

### 2021-02-17 numerics_paper/version2/numerical_chj.tex
Make section 5 flow better.
Use consistent notation for the domain.

@@
 
+\newcommand{\out}{\infty}
+\newcommand{\jet}{{\textup{jet}}}
 \newcommand{\W}{{\mathcal W}}
@@
 
-\subsection{HLLEMCC Riemann solver} \label{sec:kemm}
+\subsection{Kemm's Riemann solver} \label{sec:kemm}
 Several specialized Riemann solvers have been proposed in order to deal
@@
 
-\subsection{Steady outflow}\label{sec:steady_outflow}
-Under rotational symmetry, the two-dimensional shallow water equations
-can be reduced to the system
+\subsection{Flow past a cylinder}\label{sec:bow_shock}
+In this section we consider the formation of a bow shock when a
+uniform flow encounters a cylindrical obstacle.  This problem has
+been studied previously in the context of carbuncle formation in
+several works for the Euler equations and in
+\cite{kemm2014note,bader2014carbuncle} for the shallow water equations.
+We present results for existing solvers as a form of verification,
+along with results for the new blended solver.  The main question of
+interest is the formation of carbuncles.  It is known, for instance,
+that the Roe solver incorrectly generates a carbuncle at the center
+of the bow shock.
+
+We model only the flow on the upstream side of the cylinder, since
+our interest is in the resolution of the bow shock.  We take the domain
+$[0,40]\times[0,100]$ with a cylinder of radius 20 centered at $(40,50)$.
+Reflecting boundary conditions are imposed at the surface of the cylinder,
+along with outflow conditions along the rest of the right edge of the domain.
+The depth and velocity are set initially and (for all time) at the other
+boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.  We use a $160 \times 400$
+uniform Cartesian grid.
+%
+In Figure \ref{fig:steady_outflow}, we show results at $t=80$, after the flow has reached
+a steady state.  With Roe's solver, negative values (of the water height) are generated at an early
+time, leading to failure of the solver.  Therefore, we show results only for
+Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.  These three methods give very similar
+results.
+
+\begin{figure}[!h]
+  {\scriptsize
+    %\subfloat[Rusanov's solver.]{\includegraphics[scale=0.4]{figures/cylinder_hlle.png}}
+    \subfloat[Rusanov's solver.]{\includegraphics[scale=0.4]{figures/cylinder_rusanov.png}}
+    \quad \quad \quad \quad \quad \quad
+    \subfloat[HLLEMCC solver.]{\includegraphics[scale=0.4]{figures/cylinder_hllemcc.png}}
+    \quad \quad \quad \quad \quad \quad
+    \subfloat[Blended solver.]{\includegraphics[scale=0.4]{figures/cylinder_ES.png}}
+  }
+  \caption{
+    Flow past a cylinder.  Results for Rusanov's, HLLEMCC, and the blended solvers.
+    \label{fig:cylinder_flow}}
+\end{figure}
+
+We have also conducted a more challenging version of this test, in which
+the velocity within the domain is initially zero.  In that case, Rusanov's 
+and the blended solvers give carbuncle-free results similar to Figure \ref{fig:steady_outflow},
+but HLLEMCC exhibits a carbuncle.
+
+\section{Numerical study of the circular hydraulic jump}
+
+\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
+In this section, we consider the initial boundary value problem consisting of the
+shallow water model \eqref{eq:sw} in an annular domain
+\begin{align} \label{eq:annulus}
+r_\textup{jet} \le r \le r_\infty
+\end{align}
+where $r = \sqrt{x^2+y^2}$, with prescribed inflow at $r=r_\textup{jet}$
+and prescribed outflow at $r=r_\infty$.
+The domain and boundary conditions are rotationally symmetric.
+By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
 \begin{subequations} \label{eq:rsw}
@@
 \end{subequations}
-where $r=\sqrt{x^2+y^2}$ is the radial coordinate.  Here we consider a problem
+where the depth $h$ and radial velocity $u$ are functions of $r$ and $t$.
+By direct integration one finds that steady-state solutions of
+\eqref{eq:rsw} satisfy
+\begin{subequations}\label{steady}
+\begin{align}
+    rhu & = C \\
+    h'(r) & = \frac{h(r)}{\frac{g}{\beta^2} r^3 (h(r))^3 -r} = \frac{h(r)}{r} \cdot \frac{(F(r))^2}{1-(F(r))^2} \label{eq:dh}
+\end{align}
+\end{subequations}
+for some $C$ independent of $r$.  Here $F(r)=|u(r)|/\sqrt{gh(r)}$ is the Froude number.
+We see that two types of steady profiles exist, depending on whether the flow
+is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
+include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.
+
+\subsection{Numerical test: steady outflow}\label{sec:steady_outflow}
+We now test the numerical methods by using a time-dependent simulation
+to compute the steady flow solution just described, 
 in the annulus $r\in(0.1,1)$ with constant inflow at $r=0.1$ and outflow at $r=1$.
@@
 $h(r,t=0)=0.1$, $hu(r,t=0)=0$, the inner boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$,
-and the outer boundary condition is set to outflow.
-We refer to \cite[\S 21.8.5]{leveque2002finite} for details about the outflow boundary condition.
+and the outer boundary condition is set to outflow
+(see \cite[\S 21.8.5]{leveque2002finite} for details).
+The computational mesh is logically quadrilateral, of the type shown in
+Figure \ref{fig:mesh_chj}.
+\begin{figure}[!h]
+  \centering 
+  \includegraphics[scale=0.15]{figures/mesh.png}
+  \caption{
+    Example of a grid for the problem with the two-dimensional CHJ.
+    \label{fig:mesh_chj}}
+\end{figure}
 Regardless of the initial condition, the exact solution converges to a
@@
 
-\subsection{Flow past a cylinder}\label{sec:bow_shock}
-In this section we consider the formation of a bow shock when a
-uniform flow encounters a cylindrical obstacle.  This problem has
-been studied previously in the context of carbuncle formation in
-several works for the Euler equations and in
-\cite{kemm2014note,bader2014carbuncle} for the shallow water equations.
-We present results for existing solvers as a form of verification,
-along with results for the new blended solver.  The main question of
-interest is the formation of carbuncles.  It is known, for instance,
-that the Roe solver incorrectly generates a carbuncle at the center
-of the bow shock.
-
-We model only the flow on the upstream side of the cylinder, since
-our interest is in the resolution of the bow shock.  We take the domain
-$[0,40]\times[0,100]$ with a cylinder of radius 20 centered at $(40,50)$.
-Reflecting boundary conditions are imposed at the surface of the cylinder,
-along with outflow conditions along the rest of the right edge of the domain.
-The depth and velocity are set initially and (for all time) at the other
-boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.  We use a $160 \times 400$
-uniform Cartesian grid.
-%
-In Figure \ref{fig:steady_outflow}, we show results at $t=80$, after the flow has reached
-a steady state.  With Roe's solver, negative values (of the water height) are generated at an early
-time, leading to failure of the solver.  Therefore, we show results only for
-Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.  These three methods give very similar
-results.
-
-\begin{figure}[!h]
-  {\scriptsize
-    %\subfloat[Rusanov's solver.]{\includegraphics[scale=0.4]{figures/cylinder_hlle.png}}
-    \subfloat[Rusanov's solver.]{\includegraphics[scale=0.4]{figures/cylinder_rusanov.png}}
-    \quad \quad \quad \quad \quad \quad
-    \subfloat[HLLEMCC solver.]{\includegraphics[scale=0.4]{figures/cylinder_hllemcc.png}}
-    \quad \quad \quad \quad \quad \quad
-    \subfloat[Blended solver.]{\includegraphics[scale=0.4]{figures/cylinder_ES.png}}
-  }
-  \caption{
-    Flow past a cylinder.  Results for Rusanov's, HLLEMCC, and the blended solvers.
-    \label{fig:cylinder_flow}}
-\end{figure}
-
-We have also conducted a more challenging version of this test, in which
-the velocity within the domain is initially zero.  In that case, Rusanov's 
-and the blended solvers give carbuncle-free results similar to Figure \ref{fig:steady_outflow},
-but HLLEMCC exhibits a carbuncle.
-
-\section{Numerical study of the circular hydraulic jump}
-
-\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}
-In this section, we consider the initial boundary value problem consisting of the
-shallow water model \eqref{eq:sw} in an annular domain
-\begin{align}
-r_\textup{jet} \le r \le r_\infty
-\end{align}
-where $r = \sqrt{x^2+y^2}$, with prescribed inflow at $r=r_\textup{jet}$
-and prescribed outflow at $r=r_\infty$.
-The domain and boundary conditions are rotationally symmetric.
-By assuming rotational symmetry in \eqref{eq:sw}, one obtains the system
-\eqref{eq:rsw},
-where the depth $h$ and radial velocity $u$ are functions of the single spatial coordinate $r$.
-Steady-state solutions of \eqref{eq:rsw} satisfy
-\begin{subequations}\label{steady}
-\begin{align}
-    rhu & = C \\
-    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
-\end{align}
-\end{subequations}
-for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
-We see that two types of steady profiles exist, depending on whether the flow
-is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
-include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.
-
+\subsection{Location of the jump} \label{sec:jumploc}
 The steady, rotationally-symmetric circular hydraulic jump involves supercritical
@@
 
-Numerical tests suggest that the solution of \eqref{eq:rsw} tends rapidly approaches that
-just described under general initial conditions as long as the flow at the jet is
-supercritical and the outer outflow is subcritical.  This solution remains steady
+Numerical tests suggest that the solution of \eqref{eq:rsw} rapidly approaches that
+just described under general initial conditions as long as the inflow at the jet is
+supercritical and the outflow at $r_\infty$ is subcritical.  Subcritical outflow
+can be enforced by an appropriate outer boundary condition.  This solution remains steady
 if the rate of inflow and outflow are matched.  The stability of this solution in
@@
 only in the shallow water context but for more realistic fluid models and physically.
-It will play an important role in the results we present later.
+It will play an important role in the results we present below.
 The rotationally-symmetric steady state is a useful initial condition for studies
@@
 Let us finally consider the numerical experiments for the CHJ in two dimensions.
-The domain is given by the annulus $\Omega=\{(x,y)~|~ 0.1\leq\sqrt{x^2+y^2}\leq 1\}$.
-The computational mesh is logically quadrilateral, of the type shown in Figure \ref{fig:mesh_chj}.
-For all the experiments in this section, we use a mesh with 1000 cells in each (radial
-and angular) direction.
-Let
-\begin{align*}
-  \partial\Omega_I:=\{(x,y)\in\Omega ~|~ \sqrt{x^2+y^2}=0.1\}, \qquad \mbox{ and } \qquad
-  \partial\Omega_o:=\{(x,y)\in\Omega ~|~ \sqrt{x^2+y^2}=1\},
-\end{align*}
-denote the inflow and outflow boundaries, respectively. 
-The boundary conditions are 
+The domain is again given by the annulus \eqref{eq:annulus}, but now we solve
+the fully 2D shallow water equations \eqref{eq:sw}.  For all of the following
+experiments, we use a mesh with 1000 cells in each (radial and angular) direction.
+The boundary conditions at the jet and the outer boundary are given by
 \begin{subequations}\label{bcs_chj}
   \begin{align}
-  h(x,y,t)=h_I, \quad
-  u(x,y,t)=|\bfu_I| \left(\frac{x}{r_I}\right), \quad
-  v(x,y,t)=|\bfu_I| \left(\frac{y}{r_I}\right), \qquad \forall (x,y)\in\partial\Omega_I, \\
+  h(x,y,t)=h_\jet, \quad
+  u(x,y,t)=|\bfu_\jet| \left(\frac{x}{r_\jet}\right), \quad
+  v(x,y,t)=|\bfu_\jet| \left(\frac{y}{r_\jet}\right), \qquad \sqrt{x^2+y^2}=r_\jet \\
   %
-  h(x,y,t)=h_o, \quad 
-  u(x,y,t)= \frac{\beta }{r_o h}\left(\frac{x}{r_o}\right), \quad
-  v(x,y,t)= \frac{\beta }{r_o h}\left(\frac{y}{r_o}\right), \qquad \forall (x,y)\in\partial\Omega_o, \label{bcs_chj_outflow}
+  h(x,y,t)=h_\out, \quad 
+  u(x,y,t)= \frac{\beta }{r_\out h}\left(\frac{x}{r_\out}\right), \quad
+  v(x,y,t)= \frac{\beta }{r_\out h}\left(\frac{y}{r_\out}\right), \qquad \sqrt{x^2+y^2}=r_\out,
   \end{align}
 \end{subequations}
-where $h_I=0.3$, $r_I=0.1$, $r_o=1$ and $\beta=r_Ih_I|\bfu_I|$.
-We choose $|\bfu_I|$ and $h_o$ to generate a radially symmetric shock at $r_s:=\sqrt{x^2+y^2}=0.3$, as described in \S \ref{sec:steady_chj}.
-By adjusting the boundary conditions $|\bfu_I|$ and $h_o$ we can study different flow regimes.
+where $h_\jet=0.3$, $r_\jet=0.1$, $r_\out=1$ and $\beta=r_\jet h_\jet|\bfu_\jet|$.
+We choose $|\bfu_\jet|$ and $h_\out$ such that the steady roationally symmetric
+solution involves a symmetric shock at $r_s=0.3$, as described
+in \S \ref{sec:steady_chj}.
+By adjusting the boundary conditions $|\bfu_\jet|$ and $h_\out$ we can study different flow regimes.
 We focus on two cases.
@@
 
-\begin{figure}[!h]
-  \centering 
-  \includegraphics[scale=0.15]{figures/mesh.png}
-  \caption{
-    Example of a grid for the problem with the two-dimensional CHJ.
-    \label{fig:mesh_chj}}
-\end{figure}
-
-\subsubsection{Regime I}
+\subsubsection{Regime I ($F_\jet\approx1.37$)}
 In this case, the boundary conditions are given by \eqref{bcs_chj} with
 \begin{align}\label{bcs_chj_r1}
-  |\bfu_I|=0.75, \qquad h_o=0.37387387318873766. 
-\end{align}
-The initial condition is a symmetric CHJ located at $\sqrt{x^2+y^2}=0.3$, generated as described in \S \ref{sec:steady_chj}.
+  |\bfu_\jet|=0.75, \qquad h_\out=0.37387387318873766. 
+\end{align}
+The initial condition is again a circular hydraulic jump located at $r=0.3$,
+generated as described in \S \ref{sec:jumploc}.
 In Figure \ref{fig:chj_r1_roe}, we show the solution at different times using
@@
   \centering
-  \subfloat[Solution at (left) $t=1$, (middle) $t=1.5$ and (right) $t=3$ using Roe's Riemann solver.\label{fig:chj_r1_roe}]{
+  \subfloat[Solution at (left) $t=1$, (middle) $t=1.5$ and (right) $t=3$ using Roe's Riemann solver.
+    The main plots are Schlieren plots of the depth $h$, while the inset in the right
+    figure shows the magnitude of the momentum.\label{fig:chj_r1_roe}]{
     \begin{tabular}{ccc} 
@@
     }
-    \caption{CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
+    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
       We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
@@
 
-\subsubsection{Regime II}
-We now consider a more aggresive regime. The boundary conditions are given by \eqref{bcs_chj} with 
+\subsubsection{Regime II ($F_\jet\approx27.39$)}
+We now consider a higher-Froude number regime. The boundary conditions are
+given by \eqref{bcs_chj} with 
 \begin{align}\label{bcs_chj_r2}
-  |\bfu_I|=15, \qquad h_o=6.6845019298155357.
-\end{align}
-Again we follow \S \ref{sec:steady_chj} to generate an initial condition with a symmetric CHJ located at
-$\sqrt{x^2+y^2}=0.3$. In Figure \ref{fig:chj_r2_roe}, we show the solution at different times using method
+  |\bfu_\jet|=15, \qquad h_\out=6.6845019298155357.
+\end{align}
+Again we follow \S \ref{sec:jumploc} to generate an initial condition with a circular jump located at
+$r=0.3$. In Figure \ref{fig:chj_r2_roe}, we show the solution at different times using method
 \eqref{second-order_via_fluct} with Roe's Riemann solver. Again the solution develops
@@
 For this regime, we obtained negative values for the water height with
-the HLLEMCC solver, which lead to failure of the solver. Therefore, we only show results with
-Rusanov's and the blended solvers, see Figure \ref{fig:chj_r2_diff}. With both solvers, the carbuncles are avoided.
-%
-In contrast to what we observed in the first test case, in this situation the Froude number is
-high enough that numerical perturbations start to trigger a response around the CHJ
-(see the right panel of Figure \ref{fig:chj_r2_diff}).
-For the resolution that we consider, this behavior is dissipated by the extra numerical viscosity introduced by
-Rusanov's solver (see the left panel of Figure \ref{fig:chj_r2_diff}).
-In Figure \ref{fig:chj_r2_later}, we show the solution at a later time considering the same solvers as before.
-This time, we plot not only the Schlieren plot for the water height, but also the magnitude of the momentum. 
-Clearly, using Rusanov's solver dissipates the physical response to these numerical perturbations.
+the HLLEMCC solver, which lead to failure of the solver. Therefore, in Figure
+\ref{fig:chj_r2_diff} we only show results with Rusanov's and the blended
+solvers.
+
+The solutions obtained with these solvers show no carbuncles.  The Rusanov
+solution remains very close to the initial symmetric equilibrium state,
+whereas the blended solver solution includes perturbations that appear
+just downstream from the jump.
+In Figure \ref{fig:chj_r2_later}, we show the Rusanov and blended solutions at
+a much later time of $t=5$.  In addition to the Schlieren plot of the water
+height, we superimpose a color plot of the magnitude of the momentum. 
+It is evident that the visible perturbations in the blended solution are
+completely dissipated in the Rusanov solution, at least when using the grid employed here.
+
+It is natural to ask whether these perturbations are meaningful; i.e., whether
+the symmetric equilibrium is unstable.  To investigate this, we conduct one
+more test.
 
@@
 \subsubsection{Regime II with random perturbations}
-Finally, we introduce a physical random perturbation to the inflow boundary condition.
+Given the symmetry of the grid and initial data in the section above,
+the non-rotationally-symmetric perturbations seen when using the blended
+solver above must arise due to the influence of numerical errors.
+In order to understand whether these are evidence of a true instability,
+we now introduce a non-symmetric perturbation at the inflow boundary.
 We use the same mean values \eqref{bcs_chj_r2}, but now we set
@@
   \begin{align}
-    h(x,y,t)=\tilde h_I, \quad
-    u(x,y,t)=|\tilde \bfu_I| \left(\frac{x}{r_I}\right), \quad
-    v(x,y,t)=|\tilde \bfu_I| \left(\frac{y}{r_I}\right), \qquad \forall (x,y)\in\partial\Omega_I,
+    h(x,y,t)=\tilde h_\jet, \quad
+    u(x,y,t)=|\tilde \bfu_\jet| \left(\frac{x}{r_\jet}\right), \quad
+    v(x,y,t)=|\tilde \bfu_\jet| \left(\frac{y}{r_\jet}\right), \qquad r=r_\jet,
   \end{align}
 \end{subequations}
-with $\tilde h_I = \frac{h_I}{1+\epsilon(x,y,t)}$ and $|\tilde\bfu_I|=|\bfu_I|(1+\epsilon(x,y,t))$.
+with $\tilde h_\jet = \frac{h_\jet}{1+\epsilon(x,y,t)}$ and $|\tilde\bfu_\jet|=|\bfu_\jet|(1+\epsilon(x,y,t))$.
 Here $\epsilon(x,y,t)$ is chosen at each inflow boundary ghost cell and at each

### 2021-02-18 numerics_paper/version2/numerical_chj.tex
More small changes.
Different changes discussed on Feb 17

@@
 
-\newcommand{\out}{\infty}
+\newcommand{\out}{\textup{out}}
 \newcommand{\jet}{{\textup{jet}}}
@@
 flow at the inner and outer boundaries as follows:
-
 \begin{enumerate}
@@
 and subcritical at the outer boundary.
+
+In this section we described how the location of the jump for a steady, rotationally-symmetric CHJ
+is determined by the boundary conditions. Following similar steps, one can choose inner boundary conditions
+and find outflow boundary conditions that lead to a CHJ at a prescribed location.
+This a convenient approach to construct initial conditions for numerical experiments at different flow regimes. 
+Let us consider two flow regimes and construct the corresponding CHJs, which we use in the following sections.
+Consider the following boundary conditions:
+\begin{subequations}\label{bcs_chj}
+  \begin{align}
+  h(x,y,t)=h_\jet, \quad
+  u(x,y,t)=|\bfu_\jet| \left(\frac{x}{r_\jet}\right), \quad
+  v(x,y,t)=|\bfu_\jet| \left(\frac{y}{r_\jet}\right), \qquad \sqrt{x^2+y^2}=r_\jet \\
+  %
+  h(x,y,t)=h_\out, \quad 
+  u(x,y,t)= \frac{\beta }{r_\out h}\left(\frac{x}{r_\out}\right), \quad
+  v(x,y,t)= \frac{\beta }{r_\out h}\left(\frac{y}{r_\out}\right), \qquad \sqrt{x^2+y^2}=r_\out,
+  \end{align}
+\end{subequations}
+where $h_\jet=0.3$, $r_\jet=0.1$, $r_\out=1$ and $\beta=r_\jet h_\jet|\bfu_\jet|$.
+We choose $|\bfu_\jet|$ and $h_\out$ such that the steady roationally-symmetric
+solution involves a symmetric shock at $r_s=0.3$; see Table \ref{table:regimes}.
+In Figure \ref{fig:regimes}, we show the water depth $h$ along $y=0$ for these regimes.
+
+\begin{table}[!ht]
+  \begin{center}
+    \begin{tabular}{||c||c|c|c||} \hline
+      & $F(r_\jet)$ &  $|\bfu_\jet|$ & $h_\out$ \\ \hline
+      Regime I & 1.37 & 0.75 & 0.37387387318873766 \\ \hline
+      Regime II & 27.39 & 15 & 6.6845019298155357 \\ \hline
+    \end{tabular}
+    \caption{Boundary data $|u_\jet|$ and $h_\out$ which along with \eqref{bcs_chj}
+      produce CHJs located at $r_s=0.3$ for two different flow regimes. \label{table:regimes}}
+  \end{center}
+\end{table}
+
+\begin{figure}[!h]
+  \centering
+  \subfloat[$F(r_\jet)\approx 1.37$\label{fig:ic_regime_I}]{\includegraphics[scale=0.3]{figures/chj_r1_init.png}} \qquad
+  \subfloat[$F(r_\jet)\approx 27.39$\label{fig:ic_regime_II}]{\includegraphics[scale=0.3]{figures/chj_r2_init.png}}
+  \caption{Radially symmetric CHJs created by solving \eqref{steady} with a jump given by \eqref{eq:RH}.
+    We show a slice along $y=0$.
+    In Section \ref{sec:regime_i}, we consider the CHJ in (a) as initial condition;
+    and in Sections \ref{sec:regime_ii} and \ref{sec:regime_iiwp},
+    we consider the CHJ in (b) as initial condition. 
+    \label{fig:regimes}}
+\end{figure}
 
@@
 experiments, we use a mesh with 1000 cells in each (radial and angular) direction.
-The boundary conditions at the jet and the outer boundary are given by
-\begin{subequations}\label{bcs_chj}
-  \begin{align}
-  h(x,y,t)=h_\jet, \quad
-  u(x,y,t)=|\bfu_\jet| \left(\frac{x}{r_\jet}\right), \quad
-  v(x,y,t)=|\bfu_\jet| \left(\frac{y}{r_\jet}\right), \qquad \sqrt{x^2+y^2}=r_\jet \\
+The boundary conditions at the jet and the outer boundary are given by \eqref{bcs_chj}.
+By adjusting the boundary conditions $|\bfu_\jet|$ and $h_\out$ we can study different flow regimes.
+We focus on the two cases in Table \ref{table:regimes}.
 %
-  h(x,y,t)=h_\out, \quad 
-  u(x,y,t)= \frac{\beta }{r_\out h}\left(\frac{x}{r_\out}\right), \quad
-  v(x,y,t)= \frac{\beta }{r_\out h}\left(\frac{y}{r_\out}\right), \qquad \sqrt{x^2+y^2}=r_\out,
-  \end{align}
-\end{subequations}
-where $h_\jet=0.3$, $r_\jet=0.1$, $r_\out=1$ and $\beta=r_\jet h_\jet|\bfu_\jet|$.
-We choose $|\bfu_\jet|$ and $h_\out$ such that the steady roationally symmetric
-solution involves a symmetric shock at $r_s=0.3$, as described
-in \S \ref{sec:steady_chj}.
-By adjusting the boundary conditions $|\bfu_\jet|$ and $h_\out$ we can study different flow regimes.
-We focus on two cases.
 %For the first case, the Froude number at the inflow boundary is $\Fr\approx 1.3693$.
@@
 %For this reason, we believe this problem is more challenging than the standard bow shock problem.
-
+%
 For most of the experiments we show a Schlieren plot for the water height.
 That is, we plot $||\nabla h||_{\ell^2}$ with a greyscale logarithmic colormap.
-For some experiments we also plot the magnitude of the momentum $|h\bfu|=h\sqrt{u^2+v^2}$.
-
-\subsubsection{Regime I ($F_\jet\approx1.37$)}
+
+\subsubsection{Regime I ($F_\jet\approx1.37$)}\label{sec:regime_i}
 In this case, the boundary conditions are given by \eqref{bcs_chj} with
@@
 \end{align}
-The initial condition is again a circular hydraulic jump located at $r=0.3$,
-generated as described in \S \ref{sec:jumploc}.
+The initial condition is a circular hydraulic jump located at $r_s=0.3$; see Figure \ref{fig:ic_regime_I}. 
 In Figure \ref{fig:chj_r1_roe}, we show the solution at different times using
 method \eqref{second-order_via_fluct} with Roe's Riemann solver.
-The solution clearly develops carbuncle instabilities, which are evident in the inset figure at $t=3$.
+The solution clearly develops carbuncle instabilities, which are evident in the inset figure.
 In Figure \ref{fig:chj_r1_diff}, we show solutions at $t=3$ for
@@
 
-  \subfloat[Solution at $t=3$ using (left) Rusanov's, (middle) HLLEMCC and (right) the blended solvers.\label{fig:chj_r1_diff}]{
-    \begin{tabular}{ccc}
-      \includegraphics[scale=0.3]{figures/chj_r1_llf_p3.png} &
-      \includegraphics[scale=0.3]{figures/chj_r1_hllemcc_p3.png} &
-      \includegraphics[scale=0.3]{figures/chj_r1_blended_p3.png}
-    \end{tabular}
+  \subfloat[Solution at $t=3$ using (left) Rusanov's,
+    (upper-right) HLLEMCC and (bottom-right) the blended solvers.\label{fig:chj_r1_diff}]{
+    \includegraphics[scale=0.3]{figures/chj_r1_quadrants.png}
+    %\begin{tabular}{ccc}
+    %  \includegraphics[scale=0.3]{figures/chj_r1_llf_p3.png} &
+    %  \includegraphics[scale=0.3]{figures/chj_r1_hllemcc_p3.png} &
+    %  \includegraphics[scale=0.3]{figures/chj_r1_blended_p3.png}
+    %\end{tabular}
   }
+  \qquad\qquad
+  \subfloat[Solution at $t=3$ along $y=0$ using the same solvers as in (b).]{
+    \includegraphics[scale=0.3]{figures/chj_r1_slices.png}}
+
     \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
@@
 
-\subsubsection{Regime II ($F_\jet\approx27.39$)}
+\subsubsection{Regime II ($F_\jet\approx27.39$)}\label{sec:regime_ii}
 We now consider a higher-Froude number regime. The boundary conditions are
@@
 \end{align}
-Again we follow \S \ref{sec:jumploc} to generate an initial condition with a circular jump located at
-$r=0.3$. In Figure \ref{fig:chj_r2_roe}, we show the solution at different times using method
+The initial condition is a circular hydraulic jump located at $r_s=0.3$; see Figure \ref{fig:ic_regime_II}.
+In Figure \ref{fig:chj_r2_roe}, we show the solution at different times using method
 \eqref{second-order_via_fluct} with Roe's Riemann solver. Again the solution develops
@@
 
-\subsubsection{Regime II with random perturbations}
+\subsubsection{Regime II with random perturbations}\label{sec:regime_iiwp}
 Given the symmetry of the grid and initial data in the section above,

### 2021-02-21 numerics_paper/version2/numerical_chj.tex

@@
     The main plots are Schlieren plots of the depth $h$, while the inset in the right
-    figure shows the magnitude of the momentum.\label{fig:chj_r1_roe}]{
+    figure shows the momentum in the radial direction.\label{fig:chj_r1_roe}]{
     \begin{tabular}{ccc} 
@@
       We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
-      In all cases we show the Schlieren plot for the water height $h$.
-      In (a), for the plot that corresponds to $t=3$, we zoom close to the jump
-      and plot the y-momentum. 
+      In (a) and (b) we show the Schlieren plot for the water height $h$.
     \label{fig:chj_r1}}
@@
   \vspace{10pt}
-  \subfloat[Solution at $t=0.1$ using (left) Rusanov's and (right) the blended solvers.\label{fig:chj_r2_diff}]{
+  \subfloat[Solution at $t=0.11$ using (left) Rusanov's and (right) the blended solvers.\label{fig:chj_r2_diff}]{
     \begin{tabular}{ccc}
@@
     In all cases we show the Schlieren plot for the water height $h$.
-    In (a), for the plot that corresponds to $t=0.11$, we zoom close to the jump and plot the y-momentum.}
+    The inset figures show the momentum in the radial direction.}
 \end{figure}
@@
     which corresponds to the minimum and maximum values at $t=0$.
+    The inset figure shows the momentum in the radial direction.
     \label{fig:chj_r2_later}}
@@
 In the right panel of Figure \ref{fig:chj_r2wp_blended}, we show an inset figure 
-with the y-momentum. Note that even though the solution is highly unstable 
+with the momentum in the radial direction. Note that even though the solution is highly unstable 
 (with the blended Riemann solver), no visible carbuncle instabilities are developed.
@@
     In all cases we show the Schlieren plot for the water height $h$.
-    In (b), for the plot that corresponds to $t=5$, we zoom close to the jump and plot the y-momentum.}
+    The inset figure shows the momentum in the radial direction.}
 \end{figure}

### 2021-02-22 numerics_paper/version2/numerical_chj.tex
Add draft of conclusion.

@@
 \section{Conclusions}
+In this work we have introduced a new Riemann solver for the shallow water
+equations, described in Section \ref{sec:blended_rs}.  Through numerical tests we have
+shown that the solver gives accuracy similar to that of Roe's method and
+robustness similar to that of Rusanov's method.  Although the full
+second-order method we have proposed is not provably entropy stable or
+positivity preserving, it gives improved results for test problems
+where both of these properties are important.
+
+We have introduced two new test problems for numerical shallow water solvers,
+both consisting of flow in an annulus, with inflow from a jet
+at the center and outflow at the outer boundary.  The first test
+problem, describe in Section \ref{sec:steady_outflow}, has a smooth solution that can
+be computed by solving an ODE and thus serves as a useful test of
+accuracy.  The second problem, the circular hydraulic jump, involves
+a standing shock wave that can be physically unstable but is
+also susceptible numerically to the non-physical carbuncle instability.
+Additionally, it involves high-velocity low-depth flow regions where
+it is challenging to maintain positivity.
+This makes it an excellent problem for testing that schemes are both
+robust and not overly dissipative.
 

### 2021-03-03 numerics_paper/version2/numerical_chj.tex
Editorial work on the paper.

@@
 \begin{abstract}
-The numerical artifact known as a carbuncle in the context of the Euler
-equations has been shown to appear also in shallow water simulations.
-We study how to deal with carbuncles in the shallow water context,
-with an emphasis on a new test problem involving a steady circular hydraulic
+We investigate the numerical artifact known as a carbuncle, in the solution
+of the shallow water equations.  We propose a new Riemann solver that is based
+on a local measure of the entropy residual and aims to avoid carbuncles while
+maintaining high accuracy.
+We propose a new challenging test problem for shallow water codes, 
+consisting of a steady circular hydraulic
 jump that can be physically unstable.  We show that numerical methods are prone to
 either suppress the instability completely or form carbuncles.
-We test existing cures for the carbuncle as well as a new method,
-and show results that exhibit the shock instability without unphysical carbuncles.
+We test existing cures for the carbuncle and show that only the proposed
+method is able to avoid unphysical carbuncles without suppressing the
+physical instability.
 \end{abstract}
@@
 \cite{quirk1997contribution,pandolfi2001numerical,dumbser2004matrix,chauvat2005shock,ismail2009proposed,shen2014stability}.
-Most notably, the HLLE Riemann solver suppresses the carbuncle instability \cite{quirk1997contribution}.
+Most notably, dissipative Riemann solvers like HLLE and LLF suppress the carbuncle instability \cite{quirk1997contribution}.
 For a recent review of numerical shock instability and work to alleviate it,
@@
 The shallow water carbuncle behaves similarly to the
-Euler carbuncle: it appears when the Roe solver is used, but not when
-the HLL solver is used, and can be suppressed by similar numerical methods
-\cite{kemm2014note,bader2014carbuncle}.
+Euler carbuncle; for instance, it appears when the Roe solver is used, but not when
+the HLLE solver is used, and can also be suppressed by particular modifications
+of the Roe solver \cite{kemm2014note,bader2014carbuncle}.
 
@@
 In both of these problems, the correct behavior is the formation of a stable
-shock profile, which is achieved easily by simple methods with excessive
-diffusion.  This means these problems are not adequate on their own to
+shock profile without carbuncles.  This is achieved by certain methods
+designed specifically to avoid carbuncles, but also by by typical first-order
+accurate methods.  Thus these problems are not adequate on their own to
 evaluate methods for practical calculations.
-
 Elling \cite{elling2009carbuncle} proposed instead a problem specifically
@@
 
-Herein we introduce a new and more exacting test problem.  Like some
+Herein we introduce a new and more exacting test problem that arises in a
+common physical setting.  Like some
 of the problems above, it includes an equilibrium solution consisting of a
@@
 nor the complete absence of instability represents the correct behavior.
+This test problem is the circular hydraulic jump.
 
@@
 ``may usually be seen whenever a stream of water from a tap strikes a horizontal
-surface".  This phenomenon that is familiar in the everyday kitchen sink, is in
-fact highly nonlinear and complex.  Near the jet, the flow is shallow and
+surface''.  This phenomenon that is familiar in the everyday kitchen sink, is in
+fact highly nonlinear and unintuitive.  Near the jet, the flow is shallow and
 supercritical, while further away it is deeper and subcritical.  The transition from supercritical
@@
 
+In this work, we describe and study the circular hydraulic jump as a new
+test problem for shallow water discretizations.
+
 \subsection{Outline}
-In this work, we describe and study the circular hydraulic jump as a new
-test problem for shallow water discretizations.  We find that although some
+In Section \ref{sec2}, we review some existing numerical methods for
+the shallow water equations, focusing on certain Riemann solvers.
+In Section \ref{sec:blended_rs}, we propose a new Riemann solver that blends
+those of Roe and Rusanov in order to avoid carbuncles with being
+excessively dissipative.  In Section \ref{sec:num}, we use Clawpack to
+compare the performance of the new Riemann solver to existing
+solvers on several standard shallow water test problems.  In Section \ref{sec:chj}, we
+study the circular hydraulic jump using the newly proposed solver.
+We find that although some
 existing methods behave acceptably on previous test problems, they
 are not capable of providing accurate solutions for the circular hydraulic jump
-across the range of flow regimes we study.  We develop a new entropy-stable
-Riemann solver for the shallow water equations that generates solutions with
-the correct behavior.
-
-\section{Numerical methods for the shallow water equations}
+across the range of flow regimes we study.
+Some conclusions and future directions are discussed in Section \ref{sec:conclusion}.
+
+\section{Numerical methods for the shallow water equations} \label{sec2}
 We consider the shallow water model in two horizontal dimensions:
@@
 \begin{align}
-    q_t + f(q)_x + g(q)_y & = 0.
-\end{align}
+    q_t + \nabla \cdot f(q)_x & = 0,
+\end{align}
+where $q=[h, hu, hv]^T$ and the flux functions $f$ is defined in accordance
+with \eqref{eq:sw}.
 
@@
 \end{align}
-Here $\tilde{\W}_{i\pm 1/2}^p$ is a TVD-limited version of $\W_{i\pm 1/2}^p$.
+Here $\tilde{\W}_{i\pm 1/2}^p$ is a limited version of $\W_{i\pm 1/2}^p$, usually
+based on a total variation diminishing limiter function.
 
 We note that, for the conservative Riemann solvers we use herein, the LWL scheme
-\eqref{second-order_via_fluct} can also be written purely in flux-differencing
+\eqref{second-order_via_fluct} can be written equivalently in flux-differencing
 form:
@@
 Roe.  We refer to \cite{ketcheson2020riemann} and references therein for
-a full and detailed description of these two Riemann solvers in the context of
+a detailed description of these two Riemann solvers in the context of
 the shallow water equations.
@@
 fluxes are neglected.  The complete proposed scheme can be written in
-flux-differencing form 
-We propose
-a new scheme consisting of the second-order LWL method \eqref{flux-differencing-form} using
+flux-differencing form \eqref{flux-differencing-form} using
 the first-order fluxes
@@
 \end{align}
-and correction fluxes \eqref{correction-fluxes} based on the Roe waves and
-the modified speeds
+and correction fluxes \eqref{correction-fluxes} based on the Roe waves with
+modified speeds
 \begin{align} \label{s_p}
@@
 \end{align}
-with
+where
 \begin{align}\label{lambda_p}
@@
 \end{align*}
-and $D_i$ being an upper bound of $R_i$ that acts as normalization so that $0\leq \theta_i\leq 1$.
+and $D_i$ being an upper bound on $R_i$ so that $0\leq \theta_i\leq 1$.
 Note that $R_i\approx 0$ if $\bfq$ is smooth in $S_i$.
@@
 In most of the experiments we use these Riemann solvers with the LWL method reviewed in \S\ref{sec:waveprop}.
-When the exact solution is available, we report a convergence test based on the $L^1$-error for the water height
+When the exact solution is available, we report convergence results based on the $L^1$-error for the water height
 \begin{align*}
@@
 
-\section{Numerical study of the circular hydraulic jump}
+\section{Numerical study of the circular hydraulic jump}\label{sec:chj}
 
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusion}
 In this work we have introduced a new Riemann solver for the shallow water
@@
 at the center and outflow at the outer boundary.  The first test
-problem, describe in Section \ref{sec:steady_outflow}, has a smooth solution that can
+problem, described in Section \ref{sec:steady_outflow}, has a smooth solution that can
 be computed by solving an ODE and thus serves as a useful test of
@@
 a standing shock wave that can be physically unstable but is
-also susceptible numerically to the non-physical carbuncle instability.
+also susceptible to the numerical carbuncle instability.
 Additionally, it involves high-velocity low-depth flow regions where

### 2021-03-06 numerics_paper/version2/numerical_chj.tex
last revision

@@
 \newcommand{\Roe}{{\rm Roe}}
+\newcommand{\EV}{{\rm EV}}
 \newcommand{\efp}{\psi}
@@
 either suppress the instability completely or form carbuncles.
-We test existing cures for the carbuncle and show that only the proposed
+We test existing cures for the carbuncle. In our experiments, only the proposed
 method is able to avoid unphysical carbuncles without suppressing the
@@
 \cite{quirk1997contribution,pandolfi2001numerical,dumbser2004matrix,chauvat2005shock,ismail2009proposed,shen2014stability}.
-Most notably, dissipative Riemann solvers like HLLE and LLF suppress the carbuncle instability \cite{quirk1997contribution}.
+Most notably, dissipative Riemann solvers like HLLE and Rusanov suppress the carbuncle instability \cite{quirk1997contribution}.
 For a recent review of numerical shock instability and work to alleviate it,
@@
 Euler carbuncle; for instance, it appears when the Roe solver is used, but not when
-the HLLE solver is used, and can also be suppressed by particular modifications
+the HLLE or Rusanov solvers are used, and can also be suppressed by particular modifications
 of the Roe solver \cite{kemm2014note,bader2014carbuncle}.
+In this work, we propose a new Riemann solver that blends those of Rusanov and Roe
+based on a local measure of the entropy residual, 
+as in \cite{guermond2011entropy,guermond2018second,guermond2018well}.
+Using this Riemann solver (within the second-order Lax-Wendroff-LeVeque 
+finite volume scheme \cite{leveque1997wave, leveque2002finite}) supresses the formation of carbuncles 
+while maintaining an accuracy similar to that of Roe solver.
 
@@
 shock profile without carbuncles.  This is achieved by certain methods
-designed specifically to avoid carbuncles, but also by by typical first-order
-accurate methods.  Thus these problems are not adequate on their own to
+designed specifically to avoid carbuncles, but also by typical first-order
+accurate methods.  Thus these test problems are not adequate on their own to
 evaluate methods for practical calculations.
@@
 test problem for shallow water discretizations.
+We solve this test problem using a novel Riemann solver that supresses the 
+formation of carbuncles without dissipating important features of the solution.
 
@@
 In Section \ref{sec:blended_rs}, we propose a new Riemann solver that blends
-those of Roe and Rusanov in order to avoid carbuncles with being
+those of Roe and Rusanov in order to avoid carbuncles without being
 excessively dissipative.  In Section \ref{sec:num}, we use Clawpack to
@@
 \begin{align}
-    q_t + \nabla \cdot f(q)_x & = 0,
+    q_t + \nabla \cdot f(q) & = 0,
 \end{align}
@@
 Neither of these solvers deals with the carbuncle instability in a fully
-satisfactory way.  Rusanov's solver suppresses the carbuncle but is known to
+satisfactory way.  Rusanov's solver suppresses the carbuncle but 
+(due to the use of excessive numerical dissipation)
+is known to
 also suppress related real instabilities, while Roe's solver exhibits carbuncles.
@@
 \subsection{Rusanov's Riemann solver}\label{sec:rusanov}
-Rusanov's solver approximates the Riemann solution by with a single wave traveling in each
+Rusanov's solver approximates the Riemann solution with a single wave traveling in each
 direction.  Both waves are assumed to travel with speed $\lambda^{\max}_\imh$, which is an
@@
 \begin{align*}
-  \hat\lambda_{i-1/2}^1=\hat u_{i-1/2}-g\sqrt{g\bar h_{i-1/2}}, \qquad
+  \hat\lambda_{i-1/2}^1=\hat u_{i-1/2}-\sqrt{g\bar h_{i-1/2}}, \qquad
   \hat\lambda_{i-1/2}^2=\hat u_{i-1/2}, \qquad
-  \hat\lambda_{i-1/2}^3=\hat u_{i-1/2}+g\sqrt{g\bar h_{i-1/2}}. \\
+  \hat\lambda_{i-1/2}^3=\hat u_{i-1/2}+\sqrt{g\bar h_{i-1/2}}. \\
 \end{align*}
@@
 the HLLE and Roe solvers in such a way that the resulting method behaves
-like Roe's when away from potential carbuncle instabilities, and like HLLE
+like Roe's away from potential carbuncle instabilities, and like HLLE
 where the potential for such instabilities arises.   This is done using
@@
 Let $\entvar(\bfq)$ and $\entflux(\bfq)$ denote the entropy variable and entropy flux.
+Recall that $Q=[h,hu,hv]^t$. We use the total energy as entropy; that is, we consider 
+\begin{align*}
+  \eta(\bfq)=\frac{1}{2} gh^2 + \frac{(hu)^2+(hv)^2}{2h}, \qquad
+  \entflux(\bfq)=
+    \frac{\eta(\bfq)}{h}
+    \begin{bmatrix} hu \\ hv \end{bmatrix}.
+\end{align*}
 Based on \cite{guermond2011entropy}, we consider
@@
 \end{align}
-as a measurement of the entropy production in cell $i$ (here $S_i=[x_\imh,x_\iph]$.
+as a measurement of the entropy production in cell $i$ (here $S_i=[x_\imh,x_\iph]$).
 To avoid the need to compute the time derivative of the entropy, we follow
@@
 For the moment, we consider \eqref{blended-flux-1}-\eqref{lambda_p} without
-entropy stabilization; i.e. we set $\lambda^{\min}_\iph=0$
+entropy stabilization; i.e., we set $\lambda^{\min}_\iph=0$
 for now.
@@
 plug the fluctuations \eqref{fluct_rus} into \eqref{second-order_via_fluct} where 
-$\tilde F_{i\pm 1/2}$ is given by \eqref{correction-fluxes} with $|s^p_{i\pm 1/2}|=\lambda_{i\pm 1/2}^{\max}$. 
+$\tilde F_{i\pm 1/2}$ is given by \eqref{correction-fluxes} with $|s^p_{i\pm 1/2}|=\lambda_{i\pm 1/2}^{\max}$ and $\tilde\W^p_{i\pm 1/2}=\W_{i\pm 1/2}^{p,\Roe}$.
 By doing this, we get \eqref{LWL} (since $\theta_{i\pm 1/2}=1 \implies \lambda_{i\pm 1/2}^p=\lambda_{i\pm 1/2}^{\max}$).
@@
 \end{align*}
-is the discrete entropy flux.
+is a discretization of the entropy flux.
 
@@
 \begin{align}\label{lambda_min}
-  \lambda_{i+1/2}^{\min} = \max\left\{0,\frac{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\left[\bff(\bfq_{i+1})+\bff(\bfq_{i})-\sum_p\lambda_{i+1/2}^p\W_{i+1/2}^p\right]-(\efp_{i+1}-\efp_i)}{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\sum_p\W_{i+1/2}^p}\right\}.
-\end{align}
+  \lambda_{i+1/2}^{\min} 
+  = \max\left\{0,\frac{\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot
+    \left[\bff(\bfq_{i+1})+\bff(\bfq_{i})-\sum_p\lambda_{i+1/2}^{p,\EV}\W_{i+1/2}^{p,\Roe}\right]
+    -(\efp_{i+1}-\efp_i)}
+  {\frac{1}{2}(\entvar_{i+1}-\entvar_{i})\cdot\sum_p\W_{i+1/2}^{p,\Roe}}\right\},
+\end{align}
+where 
+\begin{align*}
+  \lambda_{i+1/2}^{p,\EV}=\theta_{i+1/2}\lambda_{i+1/2}^{\max}+(1-\theta_{i+1/2})|\hat{\lambda}_{i+1/2}^p|.
+\end{align*}
 With this choice, \eqref{blended-flux-1} guarantees \eqref{es_cond}, and therefore \eqref{ent_ineq}.
@@
 \begin{align*}
-|\hat{A}_\iph|(Q_{i+1}-Q_i) & = \sum_p |\hat{\lambda}^p_\iph| \W^{p,\Roe}_\iph, & 
+|\hat{A}_\iph|(Q_{i+1}-Q_i) = \sum_p |\hat{\lambda}^p_\iph| \W^{p,\Roe}_\iph, \qquad
 Q_{i+1}-Q_i & = \sum_p \W^{p,\Roe}_\iph.
@@
 
-\clearpage
+\begin{remark}[About the design philosophy.]
+  The work by Tadmor \cite{tadmor2003entropy} is based on imposing condition \eqref{ent_ineq} as equality. 
+  By doing this, the method is said to be entropy conservative. To improve the nonlinear 
+  stability properties of the numerical scheme, a consistent (and potentially high-order) 
+  artifical entropy dissipative operator is included. 
+  In this work, we add dissipation on the conserved variables based on 
+  a measurement of the entropy residual 
+  (as in \cite{guermond2011entropy,guermond2018second,guermond2018well}).
+  Doing this, however, does not guarantee \eqref{ent_ineq}. 
+  Indeed, some linear stabilization techniques that add artificial dissipation on the conserved 
+  variables are known to produce entropy; see for example \cite{ern2013weighting,kuzmin2020entropy}. 
+  To guarantee \eqref{ent_ineq}, we add extra dissipation on the conserved variables via 
+  \eqref{lambda_min}. 
+  Doing this counteracts entropy production created by any component of the Riemann solver;
+  see for example \cite{kuzmin2020entropy} (in the context of $C^0$ finite elements). 
+\end{remark}
+
 \section{Numerical results}\label{sec:num}
@@
     h_l & x\leq x_0=5 \\
-    h_r & x >5,
+    h_r & x >x_0,
   \end{cases}

### 2021-03-07 numerics_paper/version2/numerical_chj.tex
changes requested by Daviid

@@
 \newtheorem{remark}{Remark}
+\newtheorem*{remark*}{Remark}{}
 
@@
 Euler carbuncle; for instance, it appears when the Roe solver is used, but not when
-the HLLE or Rusanov solvers are used, and can also be suppressed by particular modifications
+the HLLE or Rusanov solver is used, and can also be suppressed by particular modifications
 of the Roe solver \cite{kemm2014note,bader2014carbuncle}.
@@
 satisfactory way.  Rusanov's solver suppresses the carbuncle but 
-(due to the use of excessive numerical dissipation)
+(unless the mesh is highly refined)
 is known to
@@
 \subsection{The entropy residual}
-Let $\entvar(\bfq)$ and $\entflux(\bfq)$ denote the entropy variable and entropy flux.
-Recall that $Q=[h,hu,hv]^t$. We use the total energy as entropy; that is, we consider 
+Let $\entvar(q)$ and $\entflux(q)$ denote the entropy variable and entropy flux.
+Recall that $q=[h,hu,hv]^T$. We use the total energy as entropy; that is, 
 \begin{align*}
-  \eta(\bfq)=\frac{1}{2} gh^2 + \frac{(hu)^2+(hv)^2}{2h}, \qquad
-  \entflux(\bfq)=
-    \frac{\eta(\bfq)}{h}
+  \eta(q)=\frac{1}{2} gh^2 + \frac{(hu)^2+(hv)^2}{2h}, \qquad
+  \entflux(q)=
+    \frac{\eta(q)}{h}
     \begin{bmatrix} hu \\ hv \end{bmatrix}.
@@
 \begin{align}\label{ent_residual}
-  \int_{S_i} \entvar(\bfq) \cdot \left[ \frac{\partial \bfq}{\partial t} + \nabla\cdot \bff(\bfq)\right]d\bfx
-  =\int_{S_i} \left[\frac{\partial \eta(\bfq)}{\partial t} + \entvar(\bfq) \cdot \nabla\cdot \bff(\bfq)\right] d\bfx
+  \int_{S_i} \entvar(q) \cdot \left[ \frac{\partial q}{\partial t} + \nabla\cdot \bff(q)\right]d\bfx
+  =\int_{S_i} \left[\frac{\partial \eta(q)}{\partial t} + \entvar(q) \cdot \nabla\cdot \bff(q)\right] d\bfx
 \end{align}
@@
 \cite{guermond2018second, guermond2018well} and use
-$\int \frac{\partial\eta(\bfq)}{\partial t}d\bfx=-\int \nabla\cdot\entflux(\bfq) d\bfx$,
+$\int \frac{\partial\eta(q)}{\partial t}d\bfx=-\int \nabla\cdot\entflux(q) d\bfx$,
 which holds in smooth regions. Then we define
@@
   R_i=
-  \left|\int_{S_i} \left[\entvar(\bfq) \cdot \nabla\cdot \bff(\bfq) - \nabla\cdot\entflux(\bfq) \right] d\bfx\right|,
+  \left|\int_{S_i} \left[\entvar(q) \cdot \nabla\cdot \bff(q) - \nabla\cdot\entflux(q) \right] d\bfx\right|,
 \end{align*}
 and $D_i$ being an upper bound on $R_i$ so that $0\leq \theta_i\leq 1$.
-Note that $R_i\approx 0$ if $\bfq$ is smooth in $S_i$.
+Note that $R_i\approx 0$ if $q$ is smooth in $S_i$.
 In our implementation, we use
 \begin{align*}
-  R_i\approx \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(\bfq)d\bfx
-  -\int_{S_i}\nabla\cdot\entflux(\bfq) d\bfx\right|,
+  R_i\approx \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(q)d\bfx
+  -\int_{S_i}\nabla\cdot\entflux(q) d\bfx\right|,
   \quad
@@
   %+\left|\int_{S_i}\nabla\cdot\bfq(\bfq)d\bfx\right|.
-  D_i = \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\int_{S_i}\left(\nabla\cdot\bff(\bfq)\right)_kd\bfx\right|
-  +\left|\int_{S_i}\nabla\cdot\entflux(\bfq)d\bfx\right|,
+  D_i = \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\int_{S_i}\left(\nabla\cdot\bff(q)\right)_kd\bfx\right|
+  +\left|\int_{S_i}\nabla\cdot\entflux(q)d\bfx\right|,
 \end{align*}
@@
 
-\begin{remark}[About the design philosophy.]
+\begin{remark*}[About the design philosophy.]
   The work by Tadmor \cite{tadmor2003entropy} is based on imposing condition \eqref{ent_ineq} as equality. 
@@
   see for example \cite{kuzmin2020entropy} (in the context of $C^0$ finite elements). 
-\end{remark}
+\end{remark*}
 

### 2021-03-07 numerics_paper/version2/numerical_chj.tex
more changes

@@
 Neither of these solvers deals with the carbuncle instability in a fully
-satisfactory way.  Rusanov's solver suppresses the carbuncle but 
-(unless the mesh is highly refined)
-is known to
-also suppress related real instabilities, while Roe's solver exhibits carbuncles.
+satisfactory way.  
+Rusanov's solver suppresses the carbuncle instability but may 
+introduce excessive numerical dissipation, while Roe's solver exhibits carbuncles.
 Later we will combine these two solvers in order to better deal with shock instability.

### 2021-03-07 numerics_paper/version2/numerical_chj.tex
Work the remarks more naturally into the text.

--- 2021-03-07 numerics_paper/version2/numerical_chj.tex
@@
 \end{align}
-where $q=[h, hu, hv]^T$ and the flux functions $f$ is defined in accordance
+where $q=[h, hu, hv]^T$ and the flux function $f$ is defined in accordance
 with \eqref{eq:sw}.
@@
 \begin{align*}
-  R_i\approx \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(q)d\bfx
-  -\int_{S_i}\nabla\cdot\entflux(q) d\bfx\right|,
+  R_i\approx \Delta x \left|\entvar(Q_i)\cdot \nabla\cdot f(Q_i)
+  -\nabla\cdot\entflux(Q_i) \right|,
   \quad
@@
   %+\left|\int_{S_i}\nabla\cdot\bfq(\bfq)d\bfx\right|.
-  D_i = \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\int_{S_i}\left(\nabla\cdot\bff(q)\right)_kd\bfx\right|
-  +\left|\int_{S_i}\nabla\cdot\entflux(q)d\bfx\right|,
+  D_i = \Delta x \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\left(\nabla\cdot\bff(Q_i)\right)_k\right|
+  +\left|\nabla\cdot\entflux(Q_i)\right|,
 \end{align*}
@@
 \end{align}
-then the following entropy inequality holds
+then we have the entropy inequality
 \begin{align}\label{ent_ineq}
@@
 is a discretization of the entropy flux.
-
-After plugging \eqref{blended-flux-1} into the condition \eqref{es_cond}, we arrive at the value
+If \eqref{ent_ineq} holds with equality, the scheme is entropy-conservative
+\cite{tadmor2003entropy}.
+
+The added dissipation described in the previous sections does not guarantee \eqref{ent_ineq}. 
+Indeed, some linear stabilization techniques that add artificial dissipation of the conserved 
+variables are known to produce entropy; see for example \cite{ern2013weighting,kuzmin2020entropy}. 
+To guarantee \eqref{ent_ineq}, we add extra dissipation of the conserved variables via 
+\eqref{lambda_min}. 
+Doing this counteracts entropy production created by any component of the Riemann solver;
+see for example \cite{kuzmin2020entropy} (in the context of $C^0$ finite elements).
+
+Plugging \eqref{blended-flux-1} into the condition \eqref{es_cond} yields the required value
 \begin{align}\label{lambda_min}
@@
 
-\begin{remark}
-In terms of fluctuations, the entropy stability condition \eqref{es_cond} can be written as
+We close this section with two remarks.  First, note that 
+the entropy stability condition \eqref{es_cond} can be written equivalently in terms of fluctuations:
 \begin{align*}
@@
 \end{align*}
-\end{remark}
-
-\begin{remark}[About the effect of $\lambda_{i\pm 1/2}^{\min}$ for Roe's solver.]
-  If we set $\theta_{i\pm 1/2}=0$ and $\lambda^{\min}_\iph=0, ~\forall i$ in \eqref{lambda_p}, then the Riemann solver becomes Roe's solver.
-  In this case, it is well known that transonic rarefactions develop non-physical shocks,
-  a phenomenon known as `entropy glitch'. To fix this problem, many entropy fixes to Roe's
-  solver have been developed. Enhancing the dissipation via \eqref{lambda_min}
-  also serves as a transonic entropy fix for Roe's solver, which we demonstrate
+
+Second, the additional dissipation introduced by $\lambda_{i\pm 1/2}^{\min}$ can also
+serve independently as an entropy fix for Roe's solver, as we demonstrate
 via a numerical experiment in \S \ref{sec:rp_dry_bed}.
-\end{remark}
-
-\begin{remark*}[About the design philosophy.]
-  The work by Tadmor \cite{tadmor2003entropy} is based on imposing condition \eqref{ent_ineq} as equality. 
-  By doing this, the method is said to be entropy conservative. To improve the nonlinear 
-  stability properties of the numerical scheme, a consistent (and potentially high-order) 
-  artifical entropy dissipative operator is included. 
-  In this work, we add dissipation on the conserved variables based on 
-  a measurement of the entropy residual 
-  (as in \cite{guermond2011entropy,guermond2018second,guermond2018well}).
-  Doing this, however, does not guarantee \eqref{ent_ineq}. 
-  Indeed, some linear stabilization techniques that add artificial dissipation on the conserved 
-  variables are known to produce entropy; see for example \cite{ern2013weighting,kuzmin2020entropy}. 
-  To guarantee \eqref{ent_ineq}, we add extra dissipation on the conserved variables via 
-  \eqref{lambda_min}. 
-  Doing this counteracts entropy production created by any component of the Riemann solver;
-  see for example \cite{kuzmin2020entropy} (in the context of $C^0$ finite elements). 
-\end{remark*}
 

### 2021-03-07 numerics_paper/version2/numerical_chj.tex
changes we discussed

@@
 \newcommand{\bfx}{{x}}
+\newcommand{\bfs}{{s}}
+\newcommand{\bfn}{{n}}
 \newcommand{\sgn}{{\rm sgn}}
@@
 
-\subsection{The entropy residual}
+\subsection{The entropy residual}\label{sec:entres}
 Let $\entvar(q)$ and $\entflux(q)$ denote the entropy variable and entropy flux.
@@
 In our implementation, we use
-\begin{align*}
-  R_i\approx \Delta x \left|\entvar(Q_i)\cdot \nabla\cdot f(Q_i)
-  -\nabla\cdot\entflux(Q_i) \right|,
-  \quad
-  %D_i = ||\entvar(\bfq_i)||_{\ell^2}\left|\left|\int_{S_i}\nabla\cdot\bff(\bfq)d\bfx\right|\right|_{\ell^2}
-  %+\left|\int_{S_i}\nabla\cdot\bfq(\bfq)d\bfx\right|.
-  D_i = \Delta x \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\left(\nabla\cdot\bff(Q_i)\right)_k\right|
-  +\left|\nabla\cdot\entflux(Q_i)\right|,
-\end{align*}
-where $(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$ and $d$ is the number of physical
-dimensions.  In \eqref{blended-flux-1} we need values of $\theta$ at the cell interfaces, for
+\begin{subequations}\label{Ri_and_Di}
+\begin{align}
+  R_i 
+  \approx 
+  \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(q)d\bfx
+  -\int_{S_i}\nabla\cdot\entflux(q) d\bfx\right| 
+  = 
+  \left|\entvar(\bfq_i)\cdot \int_{\partial S_i}\bfn_i(s) \cdot \bff(q) d\bfs
+  -\int_{\partial S_i}\bfn_i(s)\cdot\entflux(q) d\bfs\right|, 
+\end{align}
+and 
+\begin{align}
+  D_i 
+  = 
+  \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\int_{\partial S_i}\left(\bfn_i(s)\cdot\bff(q)\right)_kd\bfs\right|
+  +\left|\int_{\partial S_i}\bfn_i(s)\cdot\entflux(q) d\bfs\right|,
+\end{align}
+\end{subequations}
+where $\bfn_i$ denotes the outward-pointing normal vector at the boundary of cell $i$, 
+$(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$, 
+and $d$ is the number of physical dimensions.  
+In the integrands of \eqref{Ri_and_Di}, $q$ is the average of $Q$ 
+from each side of the boundary $\partial S_i$. 
+In \eqref{blended-flux-1} we need values of $\theta$ at the cell interfaces, for
 which we use
@@
 
-The added dissipation described in the previous sections does not guarantee \eqref{ent_ineq}. 
+The approach in \cite{tadmor2003entropy}, is based on first developing an
+entropy-conservative scheme and then adding entropy dissipation to enforce \eqref{ent_ineq}.
+On the other hand, herein we have added dissipation (as described in Section \ref{sec:entres})
+that does not guarantee \eqref{ent_ineq}.
 Indeed, some linear stabilization techniques that add artificial dissipation of the conserved 
@@
 positivity preserving, it gives improved results for test problems
-where both of these properties are important.
+where both of these properties are important.  The approach used in Section
+\ref{sec:blended_rs} could be applied to enforce entropy stability for any Riemann
+solver of the type used in Clawpack.  The same techniques could also
+be used more generally to avoid carbuncles in the solution of the
+Euler equations.
 

### 2021-03-07 numerics_paper/version2/numerical_chj.tex
two versions of Ri and Di

@@
 \newcommand{\bfn}{{n}}
+\newcommand{\f}{{\mathcal{F}}}
 \newcommand{\sgn}{{\rm sgn}}
@@
 Note that $R_i\approx 0$ if $q$ is smooth in $S_i$.
-In our implementation, we use
+In our implementation, we consider
+{\color{blue}
+\begin{align}
+  R_i 
+  &\approx 
+  \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(q)d\bfx
+  -\int_{S_i}\nabla\cdot\entflux(q) d\bfx\right| 
+  = 
+  \left|\entvar(\bfq_i)\cdot \int_{\partial S_i}\bfn_i(s) \cdot \bff(q) d\bfs
+  -\int_{\partial S_i}\bfn_i(s)\cdot\entflux(q) d\bfs\right|,
+\end{align}
+where $\bfn_i$ denotes the unit vector normal to $\partial S_i$ pointing outward from cell $i$. 
+To compute these integrals, we first split $\partial S_i$ into its faces $\{\f_i\in\partial S_i\}$.
+Let $|\f_i|$ be the length of $\f_i$, $\bar Q_{\f_i}$ be the average of $Q$ 
+from both sides of $\f_i$ and $\bfn_{\f_i}$ be the unit vector normal to $\f_i$ 
+pointing outward from cell $i$.
+Then, we use
 \begin{subequations}\label{Ri_and_Di}
+\begin{align}
+  R_i = 
+  \left|\sum_{\f_i\in\partial S_i} |\f_i|
+  \left[\entvar(\bfq_i)\cdot \left(\bfn_{\f_i} \cdot \bff\left(\bar Q_{\f_i}\right)\right)-\bfn_{\f_i} \cdot\entflux \left(\bar Q_{\f_i}\right)\right] \right|,
+\end{align}
+and 
+\begin{align}
+  D_i 
+  = 
+  \sum_{k=1}^{d+1}\left|\entvar_k\left(\bfq_i\right)\right|
+  \left|\sum_{\f_i\in\partial S_i}|\f_i|\left(\bfn_{\f_i}\cdot\bff\left(\bar Q_{\f_i}\right)\right)_k\right|
+  +\left|\sum_{\f_i\in \partial S_i} |\f_i| \bfn_{\f_i}\cdot\entflux\left(\bar Q_{\f_i}\right) \right|,
+\end{align}
+\end{subequations}
+where $(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$, 
+and $d$ is the number of physical dimensions.}
+{\color{red}
+\begin{subequations}\label{Ri_and_Di_tmp}
 \begin{align}
@@
 \end{subequations}
-where $\bfn_i$ denotes the outward-pointing normal vector at the boundary of cell $i$, 
+where $\bfn_i$ denotes the unit vector normal to $\partial S_i$ pointing outward from cell $i$,
 $(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$, 
 and $d$ is the number of physical dimensions.  
-In the integrands of \eqref{Ri_and_Di}, $q$ is the average of $Q$ 
-from each side of the boundary $\partial S_i$. 
+In the integrands of \eqref{Ri_and_Di_tmp}, $q$ is the average of $Q$ 
+from each side of the boundary $\partial S_i$.}
 In \eqref{blended-flux-1} we need values of $\theta$ at the cell interfaces, for

### 2021-03-07 numerics_paper/version2/numerical_chj.tex
remove older definition
Simplify notation in section 3.1.

@@
 Note that $R_i\approx 0$ if $q$ is smooth in $S_i$.
-In our implementation, we consider
-{\color{blue}
+In our implementation, we use the approximation
 \begin{align}
@@
 where $\bfn_i$ denotes the unit vector normal to $\partial S_i$ pointing outward from cell $i$. 
-To compute these integrals, we first split $\partial S_i$ into its faces $\{\f_i\in\partial S_i\}$.
-Let $|\f_i|$ be the length of $\f_i$, $\bar Q_{\f_i}$ be the average of $Q$ 
-from both sides of $\f_i$ and $\bfn_{\f_i}$ be the unit vector normal to $\f_i$ 
-pointing outward from cell $i$.
-Then, we use
+To compute the boundary integrals, we approximate $q$ on each cell edge $\f_{ij}$ by the average of the two
+neighboring cell averages $\bar Q_{ij} = (Q_i+Q_j)/2$, resulting in
 \begin{subequations}\label{Ri_and_Di}
@@
   R_i = 
-  \left|\sum_{\f_i\in\partial S_i} |\f_i|
-  \left[\entvar(\bfq_i)\cdot \left(\bfn_{\f_i} \cdot \bff\left(\bar Q_{\f_i}\right)\right)-\bfn_{\f_i} \cdot\entflux \left(\bar Q_{\f_i}\right)\right] \right|,
-\end{align}
-and 
+  \left|\sum_{\f_{ij}\in\partial S_i} |\f_{ij}|
+  \left[\entvar(\bfq_i)\cdot \left(\bfn_{ij} \cdot \bff\left(\bar Q_{ij}\right)\right)-\bfn_{ij} \cdot\entflux \left(\bar Q_{ij}\right)\right] \right|,
+\end{align}
+where $\bfn_{ij}$ and $|\f_{ij}|$ are
+the unit outward-facing normal and edge length of edge $\f_{ij}$, respectively.
+The normalization factor is similarly computed as
 \begin{align}
@@
   \sum_{k=1}^{d+1}\left|\entvar_k\left(\bfq_i\right)\right|
-  \left|\sum_{\f_i\in\partial S_i}|\f_i|\left(\bfn_{\f_i}\cdot\bff\left(\bar Q_{\f_i}\right)\right)_k\right|
-  +\left|\sum_{\f_i\in \partial S_i} |\f_i| \bfn_{\f_i}\cdot\entflux\left(\bar Q_{\f_i}\right) \right|,
+  \left|\sum_{\f_{ij}\in\partial S_i}|\f_{ij}|\left(\bfn_{ij}\cdot\bff\left(\bar Q_{ij}\right)\right)_k\right|
+  +\left|\sum_{\f_i\in \partial S_i} |\f_{ij}| \bfn_{ij}\cdot\entflux\left(\bar Q_{ij}\right) \right|,
 \end{align}
@@
 where $(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$, 
-and $d$ is the number of physical dimensions.}
-{\color{red}
-\begin{subequations}\label{Ri_and_Di_tmp}
-\begin{align}
-  R_i 
-  \approx 
-  \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(q)d\bfx
-  -\int_{S_i}\nabla\cdot\entflux(q) d\bfx\right| 
-  = 
-  \left|\entvar(\bfq_i)\cdot \int_{\partial S_i}\bfn_i(s) \cdot \bff(q) d\bfs
-  -\int_{\partial S_i}\bfn_i(s)\cdot\entflux(q) d\bfs\right|, 
-\end{align}
-and 
-\begin{align}
-  D_i 
-  = 
-  \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\int_{\partial S_i}\left(\bfn_i(s)\cdot\bff(q)\right)_kd\bfs\right|
-  +\left|\int_{\partial S_i}\bfn_i(s)\cdot\entflux(q) d\bfs\right|,
-\end{align}
-\end{subequations}
-where $\bfn_i$ denotes the unit vector normal to $\partial S_i$ pointing outward from cell $i$,
-$(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$, 
 and $d$ is the number of physical dimensions.
-In the integrands of \eqref{Ri_and_Di_tmp}, $q$ is the average of $Q$ 
-from each side of the boundary $\partial S_i$.}
+%\begin{subequations}\label{Ri_and_Di_tmp}
+%\begin{align}
+%  R_i 
+%  \approx 
+%  \left|\entvar(\bfq_i)\cdot \int_{S_i}\nabla\cdot \bff(q)d\bfx
+%  -\int_{S_i}\nabla\cdot\entflux(q) d\bfx\right| 
+%  = 
+%  \left|\entvar(\bfq_i)\cdot \int_{\partial S_i}\bfn_i(s) \cdot \bff(q) d\bfs
+%  -\int_{\partial S_i}\bfn_i(s)\cdot\entflux(q) d\bfs\right|, 
+%\end{align}
+%and 
+%\begin{align}
+%  D_i 
+%  = 
+%  \sum_{k=1}^{d+1}|\entvar_k(\bfq_i)|\left|\int_{\partial S_i}\left(\bfn_i(s)\cdot\bff(q)\right)_kd\bfs\right|
+%  +\left|\int_{\partial S_i}\bfn_i(s)\cdot\entflux(q) d\bfs\right|,
+%\end{align}
+%\end{subequations}
+%where $\bfn_i$ denotes the unit vector normal to $\partial S_i$ pointing outward from cell $i$,
+%$(z)_k$ denotes the $k$-th component of $z\in\mathbb{R}^{d+1}$, 
+%and $d$ is the number of physical dimensions.  
+%In the integrands of \eqref{Ri_and_Di_tmp}, $q$ is the average of $Q$ 
+%from each side of the boundary $\partial S_i$.
 In \eqref{blended-flux-1} we need values of $\theta$ at the cell interfaces, for

### 2021-03-07 numerics_paper/version2/numerical_chj.tex
add minor change

@@
 where $\bfn_i$ denotes the unit vector normal to $\partial S_i$ pointing outward from cell $i$. 
-To compute the boundary integrals, we approximate $q$ on each cell edge $\f_{ij}$ by the average of the two
+To compute the boundary integrals, we approximate $q$ on each cell edge $\f_{ij}\in\partial S_i$ by the average of the two
 neighboring cell averages $\bar Q_{ij} = (Q_i+Q_j)/2$, resulting in
@@
 \end{align}
-where $\bfn_{ij}$ and $|\f_{ij}|$ are
-the unit outward-facing normal and edge length of edge $\f_{ij}$, respectively.
+where $|\f_{ij}|$ is the length of $\f_{ij}$ and 
+$\bfn_{ij}$ is the unit vector normal to $\f_{ij}$ pointing outward from cell $i$.
+%where $n_{ij}$ and $|\f_{ij}|$ are the unit outward-facing normal and edge length of edge $\f_{ij}$, respectively.
 The normalization factor is similarly computed as

### 2021-03-09 numerics_paper/version2/numerical_chj.tex
Add affiliations.

@@
 \author{
-    David I. Ketcheson \and
+    David I. Ketcheson
+  \thanks{
+    Computer, Electrical, and Mathematical Sciences \& Engineering Division,
+    King Abdullah University of Science and Technology, 4700 KAUST, Thuwal
+    23955, Saudi Arabia. (\{manuel.quezada, david.ketcheson\}@kaust.edu.sa).
+  }
+    \and
     Manuel Quezada de Luna
+  \footnotemark[1]
 }
@@
 where $\bfn_i$ denotes the unit vector normal to $\partial S_i$ pointing outward from cell $i$. 
-To compute the boundary integrals, we approximate $q$ on each cell edge $\f_{ij}\in\partial S_i$ by the average of the two
+To compute the boundary integrals, we approximate $q$ on each cell edge $\f_{ij}$ by the average of the two
 neighboring cell averages $\bar Q_{ij} = (Q_i+Q_j)/2$, resulting in

### 2021-03-09 mql/FV/chj/cartesian/hydraulic_jump_1D.py
setup for steady outflow problem with Rusanov

@@ -1,0 +1,445 @@
+"""
+Code to simulate 2D circular hydraulic jump in PyClaw.
+
+Currently uses a Cartesian grid.
+
+In order to get a jump, you must either introduce a bathymetric barrier (circular ring),
+include bottom friction, or impose a subcritical flow at the outer boundary.
+To work with bathymetry, you have to use a Riemann solver that handles variable
+bathymetry; for now this means the GeoClaw solver.
+
+We could also get a jump by introducing bottom friction; we haven't tried this yet.
+
+Options:
+    - Riemann solver: GeoClaw, Roe, HLLE, HLLEMCC
+"""
+
+from matplotlib import animation
+import matplotlib.pyplot as plt
+from clawpack import pyclaw
+import numpy as np
+from scipy import integrate
+from scipy.interpolate import interp1d
+from ent_residual import *
+
+def step_radial_src(solver,state,dt):
+    """
+    Geometric source terms for SW equations with cylindrical symmetry.
+    Integrated using a 2-stage, 2nd-order Runge-Kutta method.
+    This is a Clawpack-style source term routine, which approximates
+    the integral of the source terms over a step.
+    """
+    dt2 = dt/2.
+    q = state.q
+    rad = state.grid.p_centers[0]
+
+    h = q[0,:,:]
+    u = q[1,:,:]/h
+
+    qstar = np.empty(q.shape)
+    qstar[0,:,:] = q[0,:,:] - dt2/rad * h*u
+    qstar[1,:,:] = q[1,:,:] - dt2/rad * h*u*u
+
+    h = qstar[0,:]
+    u = qstar[1,:]/h
+    
+    q[0,:,:] = q[0,:,:] - dt/rad * h*u
+    q[1,:,:] = q[1,:,:] - dt/rad * h*u*u
+#
+
+def get_FrOutflow(hInflow,rInflow,velInflow,rOutflow,hOutflow,grav):
+    beta = hInflow * rInflow * velInflow
+    velOutflow = beta/(rOutflow*hOutflow)
+    return velOutflow/np.sqrt(grav*hOutflow)
+#
+
+def b4step(self, state, old_state=None):
+    """
+    Set solution value inside the centeral jet region.
+    This should really be a boundary condition, but it's
+    simpler to impose this way.
+    """
+    state.get_qbc_from_q(2,self.qbc)
[…]

### 2021-03-09 mql/FV/chj/cartesian/plot_chj_1D.py

@@ -1,0 +1,265 @@
+from clawpack.petclaw.solution import Solution
+#from petclaw.io.petsc import read_petsc
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pylab as pl
+from matplotlib import rc
+import numpy as np
+import os
+from ent_residual import *
+
+def plot_q(frame,
+           with_jump=True,
+           file_prefix='claw',
+           path='./_output/'):
+    import sys
+    sys.path.append('.')
+
+    sol=Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
+    x=sol.state.grid.x.centers; y=sol.state.grid.y.centers
+    yy,xx = np.meshgrid(y,x)
+    mx=len(x); my=len(y)
+
+    h=sol.state.q[0,:,:]
+    q1=sol.state.q[1,:,:]
+    Ri = get_ent_residual(sol.state)
+
+    if frame < 10:
+        str_frame = "000"+str(frame)
+    elif frame < 100:
+        str_frame = "00"+str(frame)
+    elif frame < 1000:
+        str_frame = "0"+str(frame)
+    else:
+        str_frame = str(frame)
+    #
+
+    fig, axs = pl.subplots(2,1)
+    axs[0].plot(x,h[:,my/2],lw=2)
+    axs[1].plot(x,Ri[:,my/2],'-k',lw=2)
+    
+    pl.suptitle("t= "+str(sol.state.t),fontsize=20)
+    pl.savefig('./_plots_chj_1D/h_'+str_frame+'_slices.png')
+    pl.close()
+#
+
+def plot_frame(frame,
+               file_prefix='claw',
+               path='./_output/'):
+    import sys
+    sys.path.append('.')
+
+    sol=Solution(frame,file_format='petsc',read_aux=False,path=path,file_prefix=file_prefix)
+    x=sol.state.grid.x.centers; y=sol.state.grid.y.centers
+    yy,xx = np.meshgrid(y,x)
+    mx=len(x); my=len(y)
+
+    h=sol.state.q[0,:,:]
+
+    if frame < 10:
+        str_frame = "000"+str(frame)
+    elif frame < 100:
+        str_frame = "00"+str(frame)
+    elif frame < 1000:
+        str_frame = "0"+str(frame)
+    else:
+        str_frame = str(frame)
+    #
+    pl.plot(x,h[:,my/2],lw=2)
+#
+
+def steady_rhs(h,r,beta,g=1.):
[…]

### 2021-03-09 mql/FV/chj/cartesian/rpn2_shallow_ev.f90
push riemann solvers

@@ -1,0 +1,254 @@
+subroutine get_fZ(grav,dh,dhZ,fZ)
+  double precision grav,dh,dhZ,fZ
+  if (dh<= dhZ) then
+     fZ = 2.d0*(dsqrt(grav*dh)-dsqrt(grav*dhZ))
+  else
+     fZ = (dh-dhZ)*dsqrt(0.5d0*grav*(dh+dhZ)/dh/dhZ)
+  endif
+  return
+end subroutine get_fZ
+
+subroutine get_f(grav,dh,dhL,dhR,velL,velR,f)
+  double precision grav,dh,dhL,dhR,velL,velR,dfL,dfR,f
+  call get_fZ(grav,dh,dhL,dfL)
+  call get_fZ(grav,dh,dhR,dfR) 
+  f = dfL + dfR + velR - velL
+  return
+end subroutine get_f
+
+subroutine get_lambda_llf(grav,dhL,dhR,velL,velR,dlambda_llf)
+  double precision grav,dhL,dhR,velL,velR,dlambda_llf
+  double precision dhMin,dhMax,x0,dfMin,dfMax,dhStar,dlambda1,dlambda3
+  dhMin = min(dhL,dhR)
+  dhMax = max(dhL,dhR)            
+  x0 = (2.d0*dsqrt(2.d0)-1.d0)**2.d0
+  call get_f(grav,x0*dhMin,dhL,dhR,velL,velR,dfMin)
+  call get_f(grav,x0*dhMax,dhL,dhR,velL,velR,dfMax)
+  if (0 <= dfMin) then
+     dhStar = (max(0.d0,velL-velR+2.d0*dsqrt(grav)*(dsqrt(dhL)+dsqrt(dhR))))**2.d0/(16.d0*grav)
+  else
+     if (0 <= dfMax) then
+        dhStar = (-dsqrt(2.d0*dhMin)+dsqrt(3.d0*dhMin+2.d0*dsqrt(2.d0*dhMin*dhMax)+dsqrt(2.d0/grav)*(velL-velR)*dsqrt(dhMin)))**2.d0
+     else ! fMax < 0
+        dhStar = dsqrt(dhMin*dhMax)*(1.d0+(dsqrt(2.d0)*(velL-velR))/(dsqrt(grav*dhMin)+dsqrt(grav*dhMax)))
+     endif
+  endif
+  ! Compute max wave speed based on hStar0
[…]

### 2021-03-09 mql/FV/chj/cartesian/rpn2_shallow_hllc.f90

@@ -1,0 +1,165 @@
+! =====================================================
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =====================================================
+
+! HLLC solver for the 2D shallow water equations.
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x-momentum
+!       3 y-momentum
+
+    implicit none
+
+    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
+    
+    double precision, dimension(meqn, mwaves, 1-mbc:maxm+mbc), intent(out) :: wave
+    double precision, dimension(mwaves, 1-mbc:maxm+mbc), intent(out) :: s
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(out) :: amdq, apdq
+    
+    double precision, dimension(2,1) :: q_hat
+    double precision :: u_l, u_r, c_l, c_r, u_hat, c_hat, v_l, v_r, v_star, v_hat, h_hat
+    double precision :: h_l, h_r, hsqrt_l, hsqrt_r, hsq2
+    double precision :: grav, a, b, c, aprime, bprime, root0, root
+    double precision :: h_ml, hu_ml, hv_ml, h_mr, hu_mr, hv_mr, s1, s2, s_star, u_ml, u_mr
+    integer :: depth, mu, mv
+    integer :: i, m, mw, j, maxiter
+    
+
+    common /cparam/  grav
+
+!     # set mu to point to  the component of the system that corresponds
+!     # to momentum in the direction of this slice, mv to the orthogonal
+!     # momentum:
+
+    ! max num of iteration for the root-finding algorithm
+    maxiter = 20   
+    depth = 1
+    if (ixy.eq.1) then
+        mu = 2
+        mv = 3
+    else
+        mu = 3
+        mv = 2
+    endif
+
+    do i=2-mbc,mx+mbc
+        h_l = qr(depth,i-1)
+        h_r = ql(depth,i)
[…]

### 2021-03-09 mql/FV/chj/cartesian/rpn2_shallow_llf.f90

@@ -1,0 +1,155 @@
+! =====================================================
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =====================================================
+
+! Roe-solver for the 2D shallow water equations
+! solve Riemann problems along one slice of data.
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x_momentum
+!       3 y_momentum
+
+! On input, ql contains the state vector at the left edge of each cell
+!           qr contains the state vector at the right edge of each cell
+
+! This data is along a slice in the x-direction if ixy=1
+!                            or the y-direction if ixy=2.
+
+! On output, wave contains the waves, s the speeds,
+! and amdq, apdq the decomposition of the flux difference
+!   f(qr(i-1)) - f(ql(i))
+! into leftgoing and rightgoing parts respectively.
+! With the Roe solver we have
+!    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
+! where A is the Roe matrix.  An entropy fix can also be incorporated
+! into the flux differences.
+
+! Note that the i'th Riemann problem has left state qr(:,i-1)
+!                                    and right state ql(:,i)
+! From the basic clawpack routines, this routine is called with ql = qr
+
+
+! This Riemann solver differs from the original clawpack Riemann solver
+! for the interleaved indices
+
+    implicit double precision (a-h,o-z)
+
+    dimension   ql(meqn,           1-mbc:maxm+mbc)
+    dimension   qr(meqn,           1-mbc:maxm+mbc)
+    dimension   auxl(maux,           1-mbc:maxm+mbc)
+    dimension   auxr(maux,           1-mbc:maxm+mbc)    
+    dimension   s(mwaves,         1-mbc:maxm+mbc)
+    dimension   wave(meqn,   mwaves, 1-mbc:maxm+mbc)
+    dimension   apdq(meqn,          1-mbc:maxm+mbc)
+    dimension   amdq(meqn,          1-mbc:maxm+mbc)
+
[…]

### 2021-03-09 mql/FV/chj/cartesian/rpn2_shallow_roe_exp.f90

@@ -1,0 +1,363 @@
+subroutine get_f(grav,dh,dhZ,f)
+  double precision grav,dh,dhZ,f
+  if (dh<= dhZ) then
+     f = 2.0d0*(dsqrt(grav*dh)-dsqrt(grav*dhZ))
+  else
+     f = (dh-dhZ)*sqrt(0.5d0*grav*(dh+dhZ)/dh/dhZ)
+  endif
+  return
+end subroutine get_f
+
+subroutine get_phi(grav,dh,dhL,dhR,velL,velR,phi)
+  double precision grav,dh,dhL,dhR,velL,velR,dfL,dfR,phi
+  call get_f(g,h,hL,dfL)
+  call get_f(g,h,hR,dfR) 
+  phi = dfL + dfR + velR - velL
+  return
+end subroutine get_phi
+
+subroutine get_lambda_llf(dhL,dhR,velL,velR,dlambda_llf)            
+  dhMin = min(dhL,dhR)
+  dhMax = max(dhL,dhR)            
+  x0 = (2.0d0*dsqrt(2.0d0)-1.0d0)**2.0d0
+  call get_phi(grav,x0*dhMin,dhL,dhR,velL,velR,dfMin)
+  call get_phi(grav,x0*dhMax,dhL,dhR,velL,velR,dfMax)
+  if (0 <= dfMin) then
+     dhStar = (max(0.0d0,velL-velR+2.0d0*sqrt(grav)*(sqrt(hL)+sqrt(hR))))**2.0d0/16.0d0/grav
+  else
+     if (0 <= dfMax) then
+        dhStar = (-sqrt(2.*dhMin)+sqrt(3.*dhMin+2.*sqrt(2.*dhMin*dhMax)+sqrt(2./grav)*(velL-velR)*sqrt(dhMin)))**2.d0
+     else ! fMax < 0
+        dhStar = sqrt(dhMin*dhMax)*(1.+(sqrt(2.d0)*(velL-velR))/(sqrt(grav*dhMin)+sqrt(grav*dhMax)))
+     endif
+  endif
+  ! Compute max wave speed based on hStar0
+  dlambda1 = ( velL - sqrt(grav*dhL)*sqrt( (1.+max((dhStar-dhL)/2.d0/dhL,0.)) * (1.+max((dhStar-dhL)/dhL,0.))))
[…]

### 2021-03-09 mql/FV/chj/cartesian/rpn2_shallow_roe_with_efix.f90

@@ -1,0 +1,152 @@
+! =====================================================
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =====================================================
+
+! Roe-solver for the 2D shallow water equations
+! solve Riemann problems along one slice of data.
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x_momentum
+!       3 y_momentum
+
+! On input, ql contains the state vector at the left edge of each cell
+!           qr contains the state vector at the right edge of each cell
+
+! This data is along a slice in the x-direction if ixy=1
+!                            or the y-direction if ixy=2.
+
+! On output, wave contains the waves, s the speeds,
+! and amdq, apdq the decomposition of the flux difference
+!   f(qr(i-1)) - f(ql(i))
+! into leftgoing and rightgoing parts respectively.
+! With the Roe solver we have
+!    amdq  =  A^- \Delta q    and    apdq  =  A^+ \Delta q
+! where A is the Roe matrix.  An entropy fix can also be incorporated
+! into the flux differences.
+
+! Note that the i'th Riemann problem has left state qr(:,i-1)
+!                                    and right state ql(:,i)
+! From the basic clawpack routines, this routine is called with ql = qr
+
+
+! This Riemann solver differs from the original clawpack Riemann solver
+! for the interleaved indices
+
+    implicit double precision (a-h,o-z)
+
+    dimension   ql(meqn,           1-mbc:maxm+mbc)
+    dimension   qr(meqn,           1-mbc:maxm+mbc)
+    dimension    s(mwaves,         1-mbc:maxm+mbc)
+    dimension wave(meqn,   mwaves, 1-mbc:maxm+mbc)
+    dimension  apdq(meqn,          1-mbc:maxm+mbc)
+    dimension  amdq(meqn,          1-mbc:maxm+mbc)
+
+!   # Gravity constant set in the shallow1D.py or setprob.f90 file
+    common /cparam/ grav
+
+!   # Roe averages quantities of each interface
+    parameter (maxm2 = 1800)
[…]

### 2021-03-09 mql/FV/chj/cartesian/rpn2_swq_hllemcc.f90

@@ -1,0 +1,235 @@
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+
+! Roe-solver for the 2D shallow water equations on a mapped 
+! (logically quadrilateral) grid
+
+! waves: 3
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x-momentum
+!       3 y-momentum
+
+    implicit none
+
+    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
+    double precision, dimension(meqn,mwaves,1-mbc:maxm+mbc), intent(out) :: wave
+    double precision, dimension(mwaves,1-mbc:maxm+mbc), intent(out) :: s
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(out) :: amdq, apdq
+
+    ! local arrays -- common block comroe is passed to rpt2sh
+    ! ------------
+
+    double precision, dimension(mwaves) :: splus, sminus
+    integer :: depth, mu, mv
+    integer :: i, m, mw
+    integer :: inx, iny, ilenrat
+    double precision :: ffroude, a1, a2, a3
+
+    integer, parameter :: maxm2 = 1602  ! assumes at most 1000x1000 grid with mbc=2
+    double precision, dimension(3) :: delta
+    logical :: efix, cfix
+    double precision, dimension(-1:maxm2) :: u, v, a, h, unorl, unorr, utanl, utanr, alf, beta
+    double precision :: h_l, h_r, u_l, u_r, v_l, v_r, hsqrt_l, hsqrt_r, hsq2, c_l, c_r
+    double precision :: grav
+    double precision :: s1l, s3r, saux, sm1fix, sm1roe, sp3fix, sp3roe, zfroude
+    double precision :: renore, rere1, reremu, reremv, rhind
+    double precision :: kalpha, kbeta, kepsilon  ! Kemm's parameters for carbuncle fix
+
+    common /cparam/ grav, kalpha, kbeta, kepsilon
+    common /comroe/ u, v, a, h
+
[…]

### 2021-03-09 mql/FV/chj/cartesian/rpn2_shallow_llf_dk.f90
Add DK version of LLF solver.

@@ -1,0 +1,103 @@
+! =====================================================
+subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apdq)
+! =====================================================
+
+! HLLE solver for the 2D shallow water equations.
+
+! waves: 2
+! equations: 3
+
+! Conserved quantities:
+!       1 depth
+!       2 x-momentum
+!       3 y-momentum
+
+    implicit none
+
+    integer, intent(in) :: ixy, maxm, meqn, mwaves, maux, mbc, mx
+    double precision, dimension(meqn,1-mbc:maxm+mbc), intent(in) :: ql, qr
+    double precision, dimension(maux,1-mbc:maxm+mbc), intent(in) :: auxl, auxr
+    double precision, dimension(meqn, mwaves, 1-mbc:maxm+mbc), intent(out) :: wave
+    double precision, dimension(mwaves, 1-mbc:maxm+mbc), intent(out) :: s
+    double precision, dimension(meqn, 1-mbc:maxm+mbc), intent(out) :: amdq, apdq
+
+    double precision :: u_l, u_r, c_l, c_r, u_hat, c_hat, v_l, v_r, v_hat, h_hat
+    double precision :: h_l, h_r, hsqrt_l, hsqrt_r, hsq2
+    double precision :: grav
+    double precision :: h_m, hu_m, hv_m, s1, s2, a
+    integer :: depth, mu, mv
+    integer :: i, m, mw
+
+    common /cparam/  grav
+
+!     # set mu to point to  the component of the system that corresponds
+!     # to momentum in the direction of this slice, mv to the orthogonal
+!     # momentum:
+!
+
+    depth = 1
+    if (ixy.eq.1) then
+        mu = 2
+        mv = 3
+    else
+        mu = 3
+        mv = 2
+    endif
+
+    do i=2-mbc,mx+mbc
+        h_l = qr(depth,i-1)
+        h_r = ql(depth,i)
+        ! Velocity
+        u_l = qr(mu,i-1) / qr(depth,i-1)
+        u_r = ql(mu,i  ) / ql(depth,i  )
[…]

### 2021-03-10 mql/FV/chj/cartesian/rpn2_shallow_llf_dk.f90
fix small bug in Davids LLF RS

@@ -67,7 +67,8 @@ subroutine rpn2(ixy,maxm,meqn,mwaves,maux,mbc,mx,ql,qr,auxl,auxr,wave,s,amdq,apd
         ! Speeds of non-shear waves
         s1 = min(u_l - c_l, u_hat - c_hat)
         s2 = max(u_r + c_r, u_hat + c_hat)
-        a = max(dabs(s1,s2))
+        !a = max(dabs(s1,s2))
+        a = max(dabs(s1),dabs(s2))
 
         ! middle" state
         h_m = -(ql(mu,i)-qr(mu,i-1))/(2.d0*a) + (h_r + h_l)/2.d0

### 2021-03-15 subsonic_outflow/plot_subsonic_outflow.py
Make code run in Python 3.

@@ -34,8 +34,8 @@ def plot_q(frame,
     #
     
     # Plot numerical solution
-    pl.plot(x,h[:,my/2],'-k',lw=2)
-    pl.plot(r0+dx/2.0,h[0,my/2],'ko',lw=10)
+    pl.plot(x,h[:,my//2],'-k',lw=2)
+    pl.plot(r0+dx/2.0,h[0,my//2],'ko',lw=10)
 
     # plot exact solution in the inner ghost cells
     pl.plot(r_backward,hh_backward,'--r')
@@ -51,7 +51,7 @@ def plot_q(frame,
     pl.close()
     
     # compute L1 error
-    return dx*np.sum(np.abs(hInterp_forward(x)-h[:,my/2]))
+    return dx*np.sum(np.abs(hInterp_forward(x)-h[:,my//2]))
 #
 
 if __name__== "__main__":
@@ -70,7 +70,7 @@ if __name__== "__main__":
     if not os.path.exists('./_plots'): os.mkdir('_plots')
     from_frame = 0
     to_frame   = 100
-    frames=xrange(from_frame,to_frame+1)
+    frames=range(from_frame,to_frame+1)
     Ncells_interp = 10000
     
     # ********************************************************* #

### 2021-03-15 subsonic_outflow/subsonic_outflow.py
r0=0.8

@@ -67,7 +67,7 @@ def b4step(self, state, old_state=None):
 
 def boundary_jet(state, dim, _, qbc, __, num_ghost):
     x = state.grid.x.centers
-    dx = x[1]-x[0]
+    dx = state.grid.x.delta
         
     r0 =  state.problem_data['r0']
     h0 =  state.problem_data['h0']
@@ -250,10 +250,10 @@ def setup(
     state.problem_data['rOutflow'] = rOutflow
 
     # ***** COMPUTE EXACT SOLUTION TO USE AT BOUNDARIES ***** #    
-    dx=state.grid.x.centers[1]-state.grid.x.centers[0]
+    dx=state.grid.x.delta
     beta = r0*h0*u0
     # inner boundary
-    dh=ode(steady_rhs_backward,
+    dh=ode(steady_rhs,
            steady_rhs_backward_jacobian).set_integrator('vode',
                                                         method='bdf',
                                                         nsteps=1E5)
@@ -313,7 +313,7 @@ if __name__ == "__main__":
     # inflow 
     hInflow = 0.3
     velInflow = 0.05
-    rInflow = 0.1
+    rInflow = 0.8
     # outflow
     rOutflow = 1.0
     

### 2021-07-07 numerics_paper/version2/numerical_chj.tex
Add acknowledgment.
some changes for the first revision

@@
 \usepackage{bm, amsfonts}     % writing bold mathematical symbols
-\usepackage{xcolor}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{fixmath}
@@
 10^{-15}$ to avoid division by zero.
+{\color{OliveGreen}
+If we use Rusanov's solver with the first-order method \eqref{first-order_via_fluct}, 
+the water height is guaranteed to remain positive. Using Roe or the blended solvers 
+can introduce violations of positivity. Alternatively, if we consider the second-order scheme 
+\eqref{second-order_via_fluct}, the water height could become negative regardless of the Riemann solver.
+To reduce the occurance of loss of positivity, we only use the first-order method 
+\eqref{first-order_via_fluct} for this test problem. 
+Although using Roe or the blended solvers might still lead to loss of positivity, we 
+did not observe that in our experiments.}
 The exact solution, which can also be found in \cite{delestre2013swashes} and references therein, is
@@
 In Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test.
-Note that using the blended solver leads to more accurate results.
+{\color{OliveGreen}
+    Due to the entropy glitch, we do not converge to the right solution with Roe solver,
+    hence the drop in rate of convergence. Using any of the other solvers, we converge to the 
+    correct solution. Note that using the blended solver leads to more accurate results. 
+    Since the method we use is first-order accurate, we get no better than first-order convergence.}
 
@@
 Euler equations.
+{\color{OliveGreen}
+  For future work, we plan to extend these results to the shallow water equations 
+  with source terms that model non-flat bathymetry and friction. In addition, we 
+  are interested in guaranteeing positivity preservation for the water height
+  via algebraic flux limiters, like those proposed by Boris and Book \cite{boris1973flux} 
+  and later extended by Zalesak \cite{zalesak1979fully}.
+}
 
@@
 
+{\bf Acknowledgment.}  We thank Prof. Friedemann Kemm for sharing helpful code
+with us and for reviewing an early version of this manuscript.
 %\bibliographystyle{apalike}

### 2021-07-07 numerics_paper/version2/Response_to_Reviewers.tex
add response to reviewers

\documentclass[english,11pt]{article}

\usepackage{enumitem}
\usepackage{cite}
\usepackage{bbm}
\usepackage{ulem}
\usepackage{amsthm}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage[hmargin=0.9in, vmargin=0.75in]{geometry}

\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage[dvipsnames]{xcolor}

\usepackage{url}
\usepackage{bm}

\usepackage{amsmath}
\usepackage[numbers]{natbib}
\let\cite=\citet
\newcommand{\bfx}{{\bf x}}

\makeatletter
\providecommand{\tabularnewline}{\\}
\makeatother

%%---------------------------------------------------%%
%%---------------------- TITLE ----------------------%%
%%---------------------------------------------------%%
\begin{document}
\title{Response to Reviewers}
\maketitle

\vspace{1.5cm}
\noindent
Michael Dumbser \\
Editor \\
International Journal for Numerical Methods in Fluids \\

\vspace{1.5cm}

\noindent
Manuscript \# FLD-21-0079
\vspace{1.5cm}

\noindent
Dear Prof. Michael Dumbser,

\vspace{0.25cm}
\bigskip
The manuscript entitled
\begin{center}
``Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump''
\end{center}
by David I. Ketcheson and Manuel Quezada de Luna
has been revised following the comments made by the referees.

\bigskip
Letters detailing how the comments of each reviewer have been
addressed are enclosed. To facilitate your perusal of the revised
manuscript, changes concerning the comments of the first and second
reviewers are highlighted in red and green, respectively.

\vspace{0.25cm}
\bigskip\noindent 
With our best regards,

\vspace{0.5cm}
\bigskip\noindent
David I. Ketcheson and Manuel Quezada de Luna

\newpage
%%---------------------------------------------------------%%
%%-------------------- FIRST REVIEWER -------------------- %%
%%---------------------------------------------------------%%
\section*{Revision of Manuscript \# FLD-21-0079 -- Reviewer $\#1$}
We thank the reviewer for the comments and suggestions.
We have incorporated all of them.  For your convenience,
the modifications related to the comments made by reviewer $\#1$ and $\#2$
appear in red and green respectively in the new version of the manuscript. 
Statements in red below are comments from the reviewer.
In this letter, the comments made by the reviewer refer to the original
manuscript and our responses refer to the new manuscript.

\medskip
\noindent
The major change in the new manuscript is XX.

\bigskip
{\color{red}
  \begin{itemize}
    \item[(1)]
      Firstly, it is not clear why blending the Rusanov and Roe methods (Section 3) would fix the carbuncle
      using entropy-residual as the guiding principle. From literature, it is obvious that Roe-type methods suffer
      from carbuncle whereas Rusanov-type schemes do not. But why using the entropy-residual somehow
      makes the hybrid of the two methods remove the carbuncle? Is entropy-stability a factor (Eq. 13)? If yes,
      why?
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{red}
  \begin{itemize}
  \item[(2)]
    Assuming there is a solid reason for the proposed fix in the paper. It is also unclear how the indicator
    ($\theta$, after Eq. 19b) works to actively switch between Rusanov to Roe and vice-versa. The authors
    mention that $\theta=0$ means Roe-flux whereas $\theta=1$ recovers the Rusanov flux. What is the guiding
    principle to select the 'appropriate' $\theta$ values? This is somewhat discussed in 3.1 in between Eq. 
    16-19b but it is not intuitively clear to me how the entropy-residual is used as a guide to select $\theta$.
    Moreover, what does the value in between $0<\theta<1$ means?
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{red}
  \begin{itemize}
  \item[(3)]
    If we continue to assume that (2) above can also be justified, I strongly suggest that the one or two of
    the test cases to be further evaluated. Dumbser et al. (04) and Kitamura et al. (AIAA 09) showed that the
    instability may manifest itself for a certain 'disturbance' in the test cases either via initial point values
    disturbance or grid disturbances, even if the fix works nicely on the non-disturbed conditions. I strongly
    suggest to vary the mesh as done in Kitamura et al. for the cylinder case (at least 3 different meshes for
    cylinder) and to include the residual plot. This is because the residual plot for convergence can easily show
    potential instability. The contour plot solution may look nice but the residual plot may show that it is
    moving toward instability.
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{red}
  \begin{itemize}
  \item[(4)]
    The hydraulic jumps with random perturbations (Section 5.4.3). This is good start for introducing a
    'disturbance' but may not be sufficient to justify carbuncle-free fix. Please try out the following
    disturbance. Let $(u_L, h_L)$ to be the initial left and $(u_R, h_R)$ the right states of the jump. 
    Introduce the following disturbance at the last cell prior to the jump such that 
    $u=\delta { u_L} + (1-\delta)u_R, h=\delta {h_L} + (1-\delta)$ where $0\leq{\delta}\leq{1}$. 
    Please try a discrete increment such that $\delta = 0.0, 0.1, 0.2,...,0.9, 1.0$. 
    Details can be found in Kitamura et al. (AIAA 2007, 2009). Run the proposed range of $\delta$
    for at least one of the hydraulic jump cases.
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{red}
  \begin{itemize}
  \item[(5)]
    I am a bit concerned when the authors mention in the conclusion that the proposed fix may not be fully
    entropy-stable. This is somewhat counter-intuitive of using entropy control to remove the instability as
    proposed throughout the paper. This comes back to my first question. Why is your proposed fix herein
    removes the carbuncle problem?
  \end{itemize}
}
\noindent
XX

\clearpage
%\bibliographystyle{plain}
\bibliographystyle{abbrvnat}
\bibliography{ref.bib}

%%----------------------------------------------------------%%
%%-------------------- SECOND REVIEWER -------------------- %%
%%----------------------------------------------------------%%
\newpage
\section*{Revision of Manuscript \# FLD-21-0079 -- Reviewer $\#2$}
We thank the reviewer for the comments and suggestions.
We have incorporated all of them.  For your convenience,
the modifications related to the comments made by reviewer $\#1$ and $\#2$
appear in red and green respectively in the new version of the manuscript.
Statements in green below are comments from the reviewer.
In this letter, the comments made by the reviewer refer to the original
manuscript and our responses refer to the new manuscript.

\medskip
\noindent
The major change in the new manuscript is XX.

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(1)]
    When dealing with strong transcritical (or transonic, in the case of Euler equations) shocks, 
    the slowly-moving shock anomaly may appear 
    (see the thesis from Zaide \url{https://www.danielzaide.com/PDF/zaide_thesis.pdf})
    This anomaly is related to the discrete shock structure and appears when having a non-linear Hugoniot locus. 
    A spike in the momentum is observed downstream of the shock wave. Some authors have related the carbuncle to 
    the discrete shock structure and the unphysical values obtained in the cell containing the shock 
    (See Xie et al. \url{https://www.sciencedirect.com/science/article/pii/S0021999117306447?via\%3Dihub},
    Zaide and Roe \url{https://arc.aiaa.org/doi/10.2514/6.2011-3686}, etc). I recommend the authors to mention 
    this in the introduction and examine if the novel blended solver suffers from this anomaly (see comments below). 
    The performance of carbuncle-free solvers based on entropy arguments for the reduction of the slowly-moving 
    shock anomaly has not been yet assessed (to my knowledge).
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(2)]
    A numerical fix for the slowly-moving shock anomaly which avoids adding additional dissipation is proposed 
    by Zaide in \url{https://www.danielzaide.com/PDF/zaide_thesis.pdf}.
    The authors are recommended to include this reference in the introduction, 
    since it includes a very detailed review on numerical shockwave anomalies.
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(3)]
    The proposed method is designed for the homogeneous SWE. 
    Please explain if this idea can be easily extended to the SWE with bed elevation, friction, etc. 
    Is there any limitation when dealing with source terms?
  \end{itemize}
}
\noindent
We are interested in considering source terms and guaranteeing positivity preservation but 
we believe that constitutes a whole new project and, as a result, we would prefer to leave that 
for future work. We have included a remark about this in the conclusions.

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(4)]
    There is a paper by Navas-Montilla and Murillo 
    \url{https://www.sciencedirect.com/science/article/pii/S0021999118307496?via\%3Dihub}
    where the carbuncle and the slowly-moving shock anomaly for the SWE with bed elevation source term 
    have been examined together and some numerical fixes are proposed. This seems to be the only work 
    where source terms are considered in the study of shock anomalies for the SWE. 
    Please mention it in the introduction.
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(5)]
    In test case 4.1, flow over dry bed is considered but the details of the wet-dry interface 
    capturing method are omitted. Can you please mention or cite the algorithm for the wet-dry treatment?
  \end{itemize}
}
\noindent
In the experiments, we set $h_r=10^{-15}$. Therefore, we do not deal with wet-dry interfaces. 
Related to this, it is reasonable to wonder if the algorithm could yield negative water height values, 
which would be unphysical and would make the code to crash. 
The first-order method with Rusanov's solver is positivity preserving (under a CFL-like time step restriction). 
Using the other Riemann solvers and/or the high-order corrections might introduce violations of positivity.
To minimize the loss of positivity, we used the first-order scheme for this test problem. 
Importantly, Roe's and the blended solvers might still introduce violations of positivity, but 
we did not observe that in our experiments.

In the original manuscript, we mentioned that we use method (7) for this problem, 
which is the first-order scheme. However, we believe we were not clear enough and, 
consequently, have revised the text to improve clarity.

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(6)]
    The convergence rates in Table 1 are suboptimal (around 0.4 for Roe solver), 
    whereas those in Table 2 achieve 1-st order of accuracy. Is this suboptimality caused by 
    the wet-dry interface? Please add a comment in the text.
  \end{itemize}
}
\noindent
Due to the entropy glitch, we do not converge to the right solution with Roe solver,
hence the drop in rate of convergence.
Using any of the other solvers, we converge to the correct solution. The convergence rate
for those cases is still smaller than in Table 2 because Table 1 is obtained using the first-order scheme. 
We revised the text in Section 4.1 to improve clarity.

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(7)]
    In test case 4.3, can you add a plot of the solution for hu along x (at y=50)? 
    This will allow to visualize the possible presence of unphysical intermediate shock values 
    (slowly-moving shock anomaly) in the traditional solvers and determine if the blended solver 
    is able to overcome this issue. 
    See Case 4.1.2 from \url{https://www.sciencedirect.com/science/article/pii/S0021999118307496?via\%3Dihub}
    and include a plot like figure 4, where the evolution in time of hu is depicted and the shedding of 
    oscillations observed.
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(8)]
    I think the presence of the aforementioned unphysical values (spike in hu) are observed in Figure 8b (right) 
    and Figure 9 (right), as there is a ring of amplitude above (red color) for hu separating the supercritical 
    and subcritical regions. Can you include some comments about this?
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(9)]
    In figure 7c, can you also plot $hu$?
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(10)]
    In figure 7c, the label of the $Y$ axis is missing.
  \end{itemize}
}
\noindent
XX

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(11)]
    In case 5.4.3, the authors mention that the observed instabilities have a physical base and are not 
    produced by the numerical method. When saying ``physical instabilities'', what do you mean? 
    The SWE is a 2D model which consider depth-averaged quantities. Note that the real physics may involve 
    a more developed hydraulic jump (which involves vertical velocities) which cannot be represented by a 
    pure discontinuity as done with the SWE. Please clarify in the text the meaning of ``physical instabilities''.
  \end{itemize}
}
\noindent
XX

\end{document}

### 2021-07-14 numerics_paper/version2/Response_to_Reviewers.tex
update revision

@@
 \usepackage[numbers]{natbib}
-\let\cite=\citet
+%\let\cite=\citet
 \newcommand{\bfx}{{\bf x}}
@@
 manuscript and our responses refer to the new manuscript.
-
-\medskip
-\noindent
-The major change in the new manuscript is XX.
 
@@
 
-\clearpage
 %\bibliographystyle{plain}
-\bibliographystyle{abbrvnat}
-\bibliography{ref.bib}
+%\bibliography{refs}
+%\clearpage
+%\bibliographystyle{plain}
+%#\bibliographystyle{abbrvnat}
+%\bibliography{ref.bib}
 
@@
 We are interested in considering source terms and guaranteeing positivity preservation but 
-we believe that constitutes a whole new project and, as a result, we would prefer to leave that 
+we believe that constitutes a new project and, as a result, we would prefer to leave that 
 for future work. We have included a remark about this in the conclusions.
@@
 In the original manuscript, we mentioned that we use method (7) for this problem, 
-which is the first-order scheme. However, we believe we were not clear enough and, 
+which is the first-order scheme. However, we now believe we were not clear enough and, 
 consequently, have revised the text to improve clarity.
@@
 \noindent
-XX
+We appreciate the reviewer's suggestion. 
+We have included the requested plots using the Rusanov and the blended solvers.
+%In both cases, we obtain the non-physical spike in the momentum, which is expected due to the non-linear Hugoniot locus
+%and the presence of intermediate states through the numerical shock.
+In \cite[Section 4.1.2]{navas2019improved}, it is remarked that spurious oscillations in the transverse direction
+are introduced with Roe's solver. If HLL(S) is employed, these oscillations are damped but are still present.
+The authors also conjecture that increasing dissipation of the shear waves would reduce this effect even further.
+Rusanov's solver adds more dissipation to the shear waves and, as conjectured by the authors,
+the spurious oscillations are eliminated; see the top-right panel of our newly added Figure 4.
+With the blended solver there are still small oscillations.
+
+Importantly, with the Rusanov or the blended solvers, the position of the shock appears to remain constant
+(once the steady state is achieved). This is suggested by the parabolic profile of the space-time evolution of $hu(x=18,y)$.
+To further confirm this, we run the simulation up to $t=160$ and plot the location of the shock along $y=50$.
+Note that after $t\approx 55$, the location of the shock remains constant.
 
@@
 \noindent
-XX
+We agree with the reviewer. We have included a remark about this in Section 5.4.2 and in the conclusions.
+Our blended solver is a convex combination of Rusanov's and Roe's solvers,
+both of which produce the spike. Consequently, the blended solver also produces the spike. 
+We believe that, within our framework, the key ingredient to solve this problem is to modify
+Rusanov's fluxes to eliminate the spike following for example \cite[Section 5.1]{navas2019improved}.
+Afterwards, we can blend the modified Rusanov's fluxes with Roe's fluxes.
+We leave this for future research but make a corresponding remark in the conclusions.
 
@@
 \noindent
-XX
+We included the requested plot. As expected, the non-physical spike is clear.
+We also added a remark about this in Section 5.4.1.
 
@@
 \noindent
-XX
+Fixed as requested.
 
@@
 \noindent
-XX
+In Sections 5.1-5.3, we assume rotational symmetry in the shallow water equations and derive a semi-analytical
+profile for the hydraulic jump.
+When we solve the shallow water equations without the assumption on rotational symmetry,
+we might obtain a shock that remains steady or an unsteady shock that deviates from the initial condition.
+When the latter happens, we say the hydraulic jump is physically unstable since we believe this behavior is driven by
+the nonlinear effects of the shallow water equations.
+We revised the text in Section 5.4.3 to improve clarity.
+
+\bibliographystyle{plain}
+\bibliography{refs}
 

### 2021-07-14 numerics_paper/version2/numerical_chj.tex
update revision

@@
 \usepackage{subfig}
+\usepackage[export]{adjustbox}
 \usepackage{graphicx}
@@
 
+{\color{OliveGreen}
+  We inspect further the behavior of the blended solver across the shock. 
+  We are interested in the possible presence of an unphysical spike in the momentum.
+  This numerical artifact might lead to a slow displacement of the position of the shock,
+  which is known as slowly-moving shock anomaly; see for instance \cite{navas2019improved, zaide2012numerical}.
+  In the left panels of Figure \ref{fig:evol_of_hu}, we show the time evolution of $hu(x,y=50)$ using the Rusanov and the blended solvers.
+  Clearly, with both solvers we obtain the spike in $hu$. This behavior is expected due to the non-linear Hugoniot locus
+  and the presence of intermediate states through the numerical shock \cite[Section 4.1]{navas2019improved}.
+
+  In \cite{navas2019improved}, it is remarked that spurious oscillations in the transverse direction are introduced with Roe's solver.
+  The authors observed these oscillations by plotting the time evolution of a slice along a fixed value of $x$.
+  The authors conjectured that increasing dissipation of the shear waves reduces these oscillations.
+  Rusanov's solver adds more dissipation (than the solvers they use) to the shear waves; therefore,
+  it eliminates the spurious oscillations,
+  as seen in the right-top panel of Figure \ref{fig:evol_of_hu}. Small oscillations are present with the blended solver.
+  %
+  Importantly, with the Rusanov and the blended solvers, the position of the shock appears to remain constant
+  (once the steady state is achieved). This is suggested by the parabolic profile of the space-time evolution of $hu(x=18,y)$.
+  To further confirm this, we run the simulation up to $t=160$ and plot in Figure \ref{fig:pos_of_shock} the location of the shock along $y=50$.
+  Note that after $t\approx 55$, the location of the shock remains constant.
+
+  \begin{figure}[!h]
+    \centering
+    \subfloat[Evolution of $hu$ along (left) $y=50$ and (right) $x=18$.
+      We use the Rusanov and the blended solvers in the first and second rows, respectively.\label{fig:evol_of_hu}]{
+      \begin{tabular}{ccc}
+      \includegraphics[scale=0.25,valign=t]{figures/slices_of_hu_llf_at_y50.png}
+      & &
+      \includegraphics[scale=0.25,valign=t]{figures/slices_of_hu_llf_at_x18.png}
+      \\
+      \includegraphics[scale=0.25,valign=t]{figures/slices_of_hu_blended_at_y50.png}
+      & &
+      \includegraphics[scale=0.25,valign=t]{figures/slices_of_hu_blended_at_x18.png}
+      \end{tabular}
+    }
+    \quad
+    \subfloat[Location of the shock along $y=50$. \label{fig:pos_of_shock}]{
+      \includegraphics[scale=0.475,valign=c]{figures/pos_shock.png}}
+    \caption{Study of the slowly-moving shock anomaly. \label{fig:slowly_moving_anomaly}}
+\end{figure}
+}
+
 We have also conducted a more challenging version of this test, in which
@@
 all of which are free from carbuncles.
-%however, we expect Rusanov's solver to be less accurate.
+{\color{OliveGreen} In Figure \ref{fig:chj_r1_slices}, we plot the water height and the x-momentum along $y=0$.
+  Clearly, the solution based on Rusanov's solver is more dissipated. From the plot, it is also clear the presence
+  of the non-physical spike in the momentum. Just as in Section \ref{sec:bow_shock}, in our numerical experiments,
+  this spike does not lead to a slowly-moving shock.}
 
@@
     (upper-right) HLLEMCC and (bottom-right) the blended solvers.\label{fig:chj_r1_diff}]{
-    \includegraphics[scale=0.3]{figures/chj_r1_quadrants.png}
+    \includegraphics[scale=0.3,valign=c]{figures/chj_r1_quadrants.png}
     %\begin{tabular}{ccc}
@@
   \qquad\qquad
-  \subfloat[Solution at $t=3$ along $y=0$ using the same solvers as in (b).]{
-    \includegraphics[scale=0.3]{figures/chj_r1_slices.png}}
-
+  \subfloat[(Top) water height and (bottom) x-momentum at $t=3$ along $y=0$ using the same solvers as in (b).\label{fig:chj_r1_slices}]{
+    \begin{tabular}{c}
+      \includegraphics[scale=0.27,valign=c]{figures/chj_r1_slices.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_hu_r1_slices.png}
+    \end{tabular}
+  }
     \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
@@
 
-The solutions obtained with these solvers show no carbuncles.  The Rusanov
-solution remains very close to the initial symmetric equilibrium state,
+The solutions obtained with these solvers show no carbuncles.
+{\color{OliveGreen} Note that, similar to the experiments in Section \ref{sec:bow_shock},
+  the non-physical spike in the momentum is present in the numerical solution.}
+The Rusanov solution remains very close to the initial symmetric equilibrium state,
 whereas the blended solver solution includes perturbations that appear
@@
 time step as an i.i.d. random variable from the uniform distribution $[-0.01,0.01]$.
+
 In this case, due to the large Froude number, the random perturbations in the inflow
-may trigger physical instabilities, which we want to preserve. At the same time, we
+may trigger physical instabilities;
+{\color{OliveGreen} i.e., the shape of the hydraulic jump might deviate from a rotationally symmetric profile and become unsteady}.
+We want to preserve this behavior. At the same time, we
 do not want the solution to develop carbuncle instabilities.
@@
   and later extended by Zalesak \cite{zalesak1979fully}.
+  Finally, we are interested in reducing (or eliminating) the non-physical spike in
+  the momentum. To do this we plan to modify Rusanov's solver based on \cite[Section 5.1]{navas2019improved}. 
 }

### 2021-07-28 numerics_paper/version2/Response_to_Reviewers.tex
update revision and push references

@@
 \noindent
-XX
+As the reviewer remarks, Rusanov-type solvers do not suffer from carbuncle instabilities. Therefore,
+our aim is to use Rusanov's solver near shocks without degrading the accuracy of Roe's solver.
+The task is to find a blending criteria that can robustly and sharply locate shocks.
+The entropy-residual has proven to be a successful tool for this; see for instance \cite{guermond2011,guermond2018}. 
+%
+By using $\theta_i\approx 1$, our blended solver induces dissipation of entropy near shocks,
+which is a property of the exact solution. Moreover, if the solver is used with the first order method (7),
+the solution is provably entropy stable, as discussed in Section 3.3.
+Using entropy stability to avoid carbuncle formation is not a new idea. For example, in \cite{ismail2009,ismail2009v2},
+the authors obtain carbuncle free solvers for Euler by either inducing entropy dissipation or guaranteeing entropy stability. 
+To our knowledge, no work has been done exploring similar ideas for the shallow water equations. However, based on
+the success in the aforementioned references, it is reasonable to expect that entropy stability is a good tool to
+eliminate carbuncle formation in the shallow water equations. 
+%
+We have revised the introduction to emphasize our motivation for using the entropy-residual as blending
+criteria between Roe's and Rusanov's solvers.
 
@@
 \noindent
-XX
+We provide a precise definition of $\theta_i$ in (17) and (19). The guiding principle to select $\theta_i$
+is based on the entropy-residual. As discussed before, using the entropy-residual to locate shocks has proven
+its worth in many works. 
+By selecting $\theta_i=\{0,1\}$, the numerical flux (13) is either Roe's or Rusanov's.
+Having $0<\theta<1$ produces a solver that combines the dissipation of both solvers; i.e.,
+for those cases, the blended solver is more dissipative than Roe's but not as much as Rusanov's. 
+We have added a remark in Section 3.1 to discuss the meaning of $0<\theta_i<1$.
 
@@
 \noindent
-XX
+In Section 3.3, we enhance the artificial dissipation via $\lambda_{i+1/2}^{\min}$ to guarantee the first-order method (7)
+is entropy stable; that is, the entropy inequality (25) is satisfied. When we add the second-order flux corrections to get method (6),
+we can't guarantee the entropy inequality (25) still holds. When we mentioned in the conclusions that we do not have
+entropy stability, we meant that we do not have a proof that the entropy inequality (25) holds when using the blended solver
+with the second-order method (6). It is important to clarify that the second-order method based on the blended solver
+still induces dissipation of entropy. We have revised the conclusions to improve clarity.
+
+\begin{thebibliography}{99}
+  \bibitem{guermond2011}
+    JL. Guermond, R. Pasquetti, B Popov,
+    Entropy viscosity method for nonlinear conservation laws.
+    {\it J. Comput. Phys.}, {\bf 230} (2011) 4248--4267.
+
+    \bibitem{guermond2018}
+      JL. Guermond, M. Nazarov, B. Popov, I. Tomas,
+      Second-order invariant domain preserving approximation of the Euler equations using convex limiting.
+      {\it SIAM J. Sci. Comput.}, {\bf 40(5)} (2018) A3211--A3239.
+
+    \bibitem{ismail2009}
+      F. Ismail and P.L. Roe,
+      Affordable, entropy-consistent Euler flux functions II: Entropy production at shocks.
+      {\it J. Comput. Phys.}, {\bf 228(15)} (2009) 5410--5436.
+
+    \bibitem{ismail2009v2}
+      F. Ismail, P.L. Roe, and H. Nishikawa,
+      A proposed cure to the carbuncle phenomenon.
+      {\it In Computational Fluid Dynamics}, (2006), 149--154.
+\end{thebibliography}
 

### 2021-08-08 numerics_paper/version2/numerical_chj.tex
update

@@
 of the Roe solver \cite{kemm2014note,bader2014carbuncle}.
-In this work, we propose a new Riemann solver that blends those of Rusanov and Roe
-based on a local measure of the entropy residual, 
-as in \cite{guermond2011entropy,guermond2018second,guermond2018well}.
+In this work, we propose a new Riemann solver that blends those of Rusanov and Roe.
+%based on a local measure of the entropy residual, 
+%as in \cite{guermond2011entropy,guermond2018second,guermond2018well}.
+{\color{red} Since Rusanov's solver does not suffer from carbuncle instabilities,
+  our aim is to use it near shocks without degrading the good accuracy of Roe's solver. 
+  To do this, we need to find a blending criteria that can robustly and sharply locate shocks.
+  The entropy-residual has proven to be a successful tool for this task; see for instance
+  \cite{guermond2011entropy,guermond2018second,guermond2018well}.
+  By doing this, our blended solver induces dissipation of entropy.
+  Entropy dissipative Riemann solvers have been used to produce carbuncle free solvers for the Euler equations;
+  see for instance \cite{ismail2009affordable,ismail2009proposed}.
+  However, to the best of our knowledge, this idea has not been used before to supress carbuncle instabilities for the
+  shallow water equations. 
+}
 Using this Riemann solver (within the second-order Lax-Wendroff-LeVeque 
@@
   Q_i^{n+1} = Q_i^n-\frac{\Delta t}{\Delta x}\left[\apdq_\imh+\amdq_\iph\right]
-  -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right],
+  -\frac{\Delta t}{\Delta x}\left[\tilde{F}_{i+1/2}-\tilde{F}_{i-1/2}\right].
 \end{align}
@@
 on Rusanov's solver are given by \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.
+{\color{red}
+  The second-order method can also be written in flux-differencing form \eqref{flux-differencing-form}, where 
+  \begin{align}\label{first-order_fluxes_LLF}
+    \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2}
+    - \frac{\lambda_{i+1/2}^{\max}}{2}(\bfq_{i+1}-\bfq_{i}).
+  \end{align}
+}
 
@@
 \end{align}
+{\color{red}
+  The first-order Roe's fluxes used in \eqref{flux-differencing-form} and \eqref{first-order_via_fluct} are given as follows:
+  \begin{align}\label{first-order_fluxes_Roe}
+    \bfF_{i+1/2} = \frac{\bff(\bfq_{i+1})+\bff(\bfq_i)}{2} - \frac{1}{2} |\hat A_{i+1/2}|(\bfq_{i+1}-\bfq_{i}).
+  \end{align}
+}
 
@@
 The value of $\theta_\iph$ controls whether the first-order flux \eqref{blended-flux-1}
-behaves more like that of Rusanov or Roe.  Specifically, if
-$\lambda_\iph^{\min}=0$ and $\theta_\iph=1$, the flux is equivalent to that of
-Rusanov, while if $\lambda_\iph^{\min}=0$ and $\theta_\iph=0$, it is equivalent
-to that of Roe.
-
-In fact, with the choice \eqref{s_p}, the correction fluxes also match
+behaves more like that of Rusanov or Roe.
+{\color{red}
+  Specifically, if $\lambda_\iph^{\min}=0$ and $\theta_\iph=1$, the flux \eqref{blended-flux-1}
+  is equivalent to that of Rusanov \eqref{first-order_fluxes_LLF},
+  while if $\lambda_\iph^{\min}=0$ and $\theta_\iph=0$, it is equivalent to that of Roe \eqref{first-order_fluxes_Roe}.
+  With $0<\theta<1$ the blended solver is more dissipative than Roe's but less than Rusanov's.
+}
+With the choice \eqref{s_p}, the correction fluxes also match
 those of Rusanov or Roe in the appropriate limit, as shown in the next section.
@@
 shown that the solver gives accuracy similar to that of Roe's method and
-robustness similar to that of Rusanov's method.  Although the full
-second-order method we have proposed is not provably entropy stable or
+robustness similar to that of Rusanov's method.
+{\color{red}
+  Using this Riemann solver with the first-order method \eqref{first-order_via_fluct} we
+  prove that the discrete entropy inequality \eqref{ent_ineq} holds.
+  In principle, we could apply the same idea to further modify the artificial dissipation
+  via $\lambda^{\min}_{i+1/2}$ to guarantee the inequality holds when using the second-order method \eqref{second-order_via_fluct}. 
+  However, in our numerical experiments, we obtained the desired behavior 
+  so, to avoid further complexity, we did not pursue this idea. 
+}
+Although the full second-order method we have proposed is not 
 positivity preserving, it gives improved results for test problems
-where both of these properties are important.  The approach used in Section
+where this property is important.  The approach used in Section
 \ref{sec:blended_rs} could be applied to enforce entropy stability for any Riemann

### 2021-08-09 numerics_paper/version2/Response_to_Reviewers.tex
update
Add some citations and revise discussion of physical instability.

@@
 As the reviewer remarks, Rusanov-type solvers do not suffer from carbuncle instabilities. Therefore,
-our aim is to use Rusanov's solver near shocks without degrading the accuracy of Roe's solver.
+our aim is to use Rusanov's solver near shocks (where the carbuncle instability might appear) 
+while using the less-diffusive Roe's solver elsewhere.
 The task is to find a blending criteria that can robustly and sharply locate shocks.
-The entropy-residual has proven to be a successful tool for this; see for instance \cite{guermond2011,guermond2018}. 
+The entropy residual is a natural choice since it is expected to vanish away
+from shocks and has proven to be a successful tool for this; see for instance
+\cite{guermond2011,guermond2018}. 
 %
-By using $\theta_i\approx 1$, our blended solver induces dissipation of entropy near shocks,
-which is a property of the exact solution. Moreover, if the solver is used with the first order method (7),
+Near shocks the entropy residual is non-zero, so our blended solver induces dissipation of entropy near shocks,
+similarly to the exact solution. Moreover, if the solver is used with the first order method (7),
 the solution is provably entropy stable, as discussed in Section 3.3.
@@
 %
-We have revised the introduction to emphasize our motivation for using the entropy-residual as blending
+We have revised the introduction to emphasize our motivation for using the entropy residual as blending
 criteria between Roe's and Rusanov's solvers.
@@
 for those cases, the blended solver is more dissipative than Roe's but not as much as Rusanov's. 
-We have added a remark in Section 3.1 to discuss the meaning of $0<\theta_i<1$.
+We have revised the text in Section 3.1 accordingly.
 
@@
 \noindent
-In Section 3.3, we enhance the artificial dissipation via $\lambda_{i+1/2}^{\min}$ to guarantee the first-order method (7)
-is entropy stable; that is, the entropy inequality (25) is satisfied. When we add the second-order flux corrections to get method (6),
-we can't guarantee the entropy inequality (25) still holds. When we mentioned in the conclusions that we do not have
-entropy stability, we meant that we do not have a proof that the entropy inequality (25) holds when using the blended solver
-with the second-order method (6). It is important to clarify that the second-order method based on the blended solver
-still induces dissipation of entropy. We have revised the conclusions to improve clarity.
+We have revised the conclusions to make it clear that we do propose an entropy-stable carbuncel-free scheme;
+namely, the first-order method (7).  It is only the second-order method (6) that is not entropy-stable.
+Furthermore, the second-order scheme retains the modification made to the first-order scheme, which is
+presumably the reason that it does not exhibit carbuncles.
+In the design of numerical methods, it is often the case that some compromise must be made between conflicting
+objectives, and this is such a case.  In the construction of the second order
+method, we have chosen to give up the strict entropy-stable property in order
+to achieve greater accuracy without compromising the carbuncle-free property.
 
@@
 \noindent
-XX
+Thank you for bringing this reference to our attention.  It is indeed highly relevant; we have
+added a citation in the introduction as well as in appropriate places later in the manuscript.
 
@@
 \noindent
-XX
+Thank you for bringing to our attention this extremely relevant work.  We have added a reference in
+the introduction and have also expanded and modified our discussion of later results taking
+this work into account.
 
@@
 \noindent
-In the experiments, we set $h_r=10^{-15}$. Therefore, we do not deal with wet-dry interfaces. 
-Related to this, it is reasonable to wonder if the algorithm could yield negative water height values, 
-which would be unphysical and would make the code to crash. 
-The first-order method with Rusanov's solver is positivity preserving (under a CFL-like time step restriction). 
-Using the other Riemann solvers and/or the high-order corrections might introduce violations of positivity.
-To minimize the loss of positivity, we used the first-order scheme for this test problem. 
-Importantly, Roe's and the blended solvers might still introduce violations of positivity, but 
-we did not observe that in our experiments.
+Although we use the term "dry bed" in the section title, we note clearly in the text that we
+use a non-zero depth $h_r=10^{-15}$.  Although we do not have a proof of positivity preservation
+for our method, we do not observe violations of positivity.  Thus we avoid the
+need to deal with wet-dry interfaces.  Note that we use the first-order method
+(7) in order to reduce the potential for the appearance of negative depth
+values.  This was already stated in the original manuscript, but we have made a revision
+that emphasizes this.
 
-In the original manuscript, we mentioned that we use method (7) for this problem, 
-which is the first-order scheme. However, we now believe we were not clear enough and, 
-consequently, have revised the text to improve clarity.
+%This does raise the question of how well the solver is able to avoid negative values of $h$
+%in general.
+%The first-order method with Rusanov's solver is positivity preserving (under a CFL-like time step restriction). 
+%Using the other Riemann solvers and/or the high-order corrections might introduce violations of positivity.
+%To minimize the loss of positivity, we used the first-order scheme for this test problem. 
+%Importantly, Roe's and the blended solvers might still introduce violations of positivity, but 
+%we did not observe that in our experiments.
 
@@
 \noindent
-In Sections 5.1-5.3, we assume rotational symmetry in the shallow water equations and derive a semi-analytical
-profile for the hydraulic jump.
-When we solve the shallow water equations without the assumption on rotational symmetry,
-we might obtain a shock that remains steady or an unsteady shock that deviates from the initial condition.
-When the latter happens, we say the hydraulic jump is physically unstable since we believe this behavior is driven by
-the nonlinear effects of the shallow water equations.
-We revised the text in Section 5.4.3 to improve clarity.
+We have revised the discussion of this result to be more clear and to avoid
+referring to the instability in the simulations as "physical".  We do refer by
+way of comparison to a physical instability of the CHJ that has been known for
+some time
+(see e.g. [8, Figure 2]).  While we agree that the shallow water model omits
+details that may be important to such an instability, we are not aware of any result that
+would preclude a close connection between what is observed here and in experiments.
+We believe this is an interesting result that merits further research.
 

### 2021-08-09 numerics_paper/version2/numerical_chj.tex

@@
 For a recent review of numerical shock instability and work to alleviate it,
-we refer to \cite[Section 2.5]{simonnumerical}.
+we refer to \cite[Section 2.5]{simonnumerical} and \cite{zaide2012numerical}.
 
@@
 the HLLE or Rusanov solver is used, and can also be suppressed by particular modifications
-of the Roe solver \cite{kemm2014note,bader2014carbuncle}.
+of the Roe solver \cite{kemm2014note,bader2014carbuncle}.  Another approach to avoid
+shallow water carbuncles has been proposed in \cite{navas2019improved}, where it is
+studied together with the slowly-moving shock anomaly and in the presence of varying bathymetry.
 In this work, we propose a new Riemann solver that blends those of Rusanov and Roe.
@@
   To do this, we need to find a blending criteria that can robustly and sharply locate shocks.
-  The entropy-residual has proven to be a successful tool for this task; see for instance
+  The entropy residual has proven to be a successful tool for this task; see for instance
   \cite{guermond2011entropy,guermond2018second,guermond2018well}.
@@
 
-In this case, due to the large Froude number, the random perturbations in the inflow
-may trigger physical instabilities;
-{\color{OliveGreen} i.e., the shape of the hydraulic jump might deviate from a rotationally symmetric profile and become unsteady}.
-We want to preserve this behavior. At the same time, we
-do not want the solution to develop carbuncle instabilities.
-In figures \ref{fig:chj_r2wp_rusanov} and \ref{fig:chj_r2wp_blended}, we show the solution
-at different times considering method \eqref{second-order_via_fluct} with Rusanov's and
-the blended Riemann solvers, respectively.
+{\color{OliveGreen} 
+Solutions are shown in Figures \ref{fig:chj_r2wp_rusanov} and \ref{fig:chj_r2wp_blended},
+obtained with Rusanov's method and the blended solver, respectively.
+It is known from experiments that the circular hydraulic jump in a low-viscosity fluid
+exhibits instability at large Froude number; see e.g. \cite[Figure~2]{craik1981circular}.
+Although the physics of this instability is still a matter of study, the blended solver
+produces a solution that bears a qualitative similarity to what is seen in experiments,
+without generating carbuncles.  The Rusanov solver, on the other hand, appears to suppress
+the manifestation of this instability.
+}
+
 In the right panel of Figure \ref{fig:chj_r2wp_blended}, we show an inset figure 

### 2021-08-09 numerics_paper/version2/numerical_chj.tex
add one of the missing experiments

@@
 
+{\color{red}
+  Finally, following \cite[Section III B]{kitamura2009evaluation}, we repeat the same experiment using
+  the blended solver with an initial condition where the shock,
+  located at $r_s=0.3$, is aligned with a grid interface. Afterward, we perturb the initial condition near the shock to misalign it
+  from the grid interface. We use a mesh with $90$ elements in the radial direction and $90$ elements in the angular direction;
+  see Figure \ref{fig:mesh_chj}.
+  For this mesh, $\Delta r=(r_\out-r_\jet)/90=0.01$. 
+  Let $(x_i,y_i)$ denote the coordinates of the center of the $i$-th element and $r_i:=\sqrt{x_i^2+y_i^2}$.
+  For all elements where $r_i\not\in[r_s,r_s+\Delta r]$, we set
+  \begin{align*}
+    h_i = \frac{1}{\Delta r}\int_{r_i-\Delta r/2}^{r_i+\Delta r/2}h(r) dr,
+    \qquad
+    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
+    \qquad
+    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
+  \end{align*}
+  where $\beta=r_\jet h_\jet |u_\jet|$ and $h(r)$ is the solution of the ODE \eqref{steady} from either
+  $r=r_\jet$ to $r=r_s^-$ or $r_s^+$ to $r_\out$.
+  If $r_i = r_s+\Delta r/2$, we set
+  \begin{align*}
+    h_i = \delta h_L + (1+\delta) h_R,
+    \qquad
+    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
+    \qquad
+    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
+  \end{align*}
+  where $\delta=\{0,0.1,0.2,\dots,0.9,1\}$ and $h_L$ and $h_R$ are the values of the water height
+  at the left and right cells from the cell with $r_i=r_s+\Delta r/2$, respectively.
+  In Figure \ref{stab_shock_init_cond}, we plot $h(r,\theta_j)$ with $j=\{1,2,\dots,90\}$; that is,
+  we plot the water height as a function of $r$ for every value of $\theta$.
+  In Figures \ref{stab_shock_roe} and \ref{stab_shock_blended}, we plot the corresponding solutions at
+  $t=3$ using method \eqref{second-order_via_fluct} with the Roe solver and with the blended solver,
+  respectively. With the Roe solver, the carbuncle instability appears; as a consequence, the
+  location of the shock changes for the different values of $\theta$.
+  With the blended solver, no carbuncle instability appears and, therefore, all slices are aligned.
+  In the figures we only show the results using $\delta=\{0,0.5,1\}$ but we
+  obtained similar results with the other values of $\delta$.
+
+  \begin{figure}
+    \subfloat[Initial condition.\label{stab_shock_init_cond}]{
+    \begin{tabular}{c}
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0.png} 
+    \end{tabular}
+    }
+    ~
+    \subfloat[Roe's solver.\label{stab_shock_roe}]{
+    \begin{tabular}{c}
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_roe.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_roe.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_roe.png} 
+    \end{tabular}
+    }
+    ~
+    \subfloat[Blended solver.\label{stab_shock_blended}]{
+    \begin{tabular}{c}
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_blended.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_blended.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_blended.png} 
+    \end{tabular}
+  }    
+    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
+      We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
+      In (a) we show the different initial conditions that we consider.
+      In (b) and (c) we show the corresponding solutions at $t=3$ using method
+      \eqref{second-order_via_fluct} with the Roe solver and with the blended solver, respectively. 
+      \label{fig:stab_shock}}
+  \end{figure}
+}
+
 \subsubsection{Regime II ($F_\jet\approx27.39$)}\label{sec:regime_ii}

### 2021-08-10 numerics_paper/version2/numerical_chj.tex
Revise discussion of shock perturbation experiments.

--- 2021-08-09 numerics_paper/version2/numerical_chj.tex
@@
 
+{\color{red}
+  Next we perform a set of experiments similar to those of \cite[Section III B]{kitamura2009evaluation}.
+  In these experiments, we perturb the initial condition in just one layer of cells, immediately behind the shock.
+  We use an annular mesh (see Figure \ref{fig:mesh_chj}) with $90$ elements in
+  the radial direction and $90$ elements in the angular direction.  Let $h(r)$ denote the exact solution
+  of the ODE \eqref{steady}, and let $r_i$ denote the radius of a given cell center.  Then for all cells
+  away from the shock (i.e. those with $r_i \notin [0.3+\Delta r]$) we use the exact solution averaged over the grid:
+  \begin{align*}
+    h_i = \frac{1}{\Delta r}\int_{r_i-\Delta r/2}^{r_i+\Delta r/2}h(r) dr,
+    \qquad
+    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
+    \qquad
+    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
+  \end{align*}
+  where $\beta=r_\jet h_\jet |u_\jet|$ and $(x_i,y_i)$ are the cartesian
+  coordinates of the cell center.  For the layer of cells just behind the shock (with $0.3 \le r \le 0.3+\Delta r$),
+  we set
+  \begin{align*}
+    h_i = \delta h_{i-1} + (1+\delta) h_{i+1},
+    \qquad
+    u_i = \frac{\beta}{h_ir_i}\left(\frac{x_i}{r_i}\right),
+    \qquad
+    v_i = \frac{\beta}{h_ir_i}\left(\frac{y_i}{r_i}\right),
+  \end{align*}
+  where $\delta=\{0,0.1,0.2,\dots,0.9,1\}$ and $h_L$ and $h_R$ are the values of the water height
+  at the left and right cells from the cell with $r_i=r_s+\Delta r/2$, respectively.
+  In Figure \ref{stab_shock_init_cond}, we plot representative results of these experiments, showing
+  the initial data in the first column and solutions at $t=3$ in the second and third columns.
+  we show only the results using $\delta=\{0,0.5,1\}$ since other results were similar.
+  The solution obtained with the Roe solver, in which the carbuncle instability appears,
+  is sensitive to the perturbation just described, while the solution obtained with the
+  blended solver is radially symmetric and essentially the same for all values of $\delta$.
+  \begin{figure}
+    \subfloat[Initial condition ($t=0$)\label{stab_shock_init_cond}]{
+    \begin{tabular}{c}
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0.png} 
+    \end{tabular}
+    }
+    ~
+    \subfloat[Roe's solver ($t=3$)\label{stab_shock_roe}]{
+    \begin{tabular}{c}
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_roe.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_roe.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_roe.png} 
+    \end{tabular}
+    }
+    ~
+    \subfloat[Blended solver ($t=3$)\label{stab_shock_blended}]{
+    \begin{tabular}{c}
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p0_blended.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta0p5_blended.png} \\ \\
+      \includegraphics[scale=0.27,valign=c]{figures/chj_slices_delta1p0_blended.png} 
+    \end{tabular}
+  }    
+    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1},
+      using the 2nd-order method \eqref{second-order_via_fluct} with two Riemann solvers. The depth $h$
+      is shown as a function of $r$.  Note that the plots in the third column contain include lines for
+      each angle along the grid, but only a single line is visible due to the solution's symmetry.
+      \label{fig:stab_shock}}
+  \end{figure}
+}
+
 \subsubsection{Regime II ($F_\jet\approx27.39$)}\label{sec:regime_ii}

### 2021-08-10 numerics_paper/version2/numerical_chj.tex
add experiments with residuals

--- 2021-08-09 numerics_paper/version2/numerical_chj.tex
@@
 The depth and velocity are set initially and (for all time) at the other
-boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.  We use a $160 \times 400$
-uniform Cartesian grid.
+boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.
+We use a $160 \times 400$ uniform Cartesian grid.
 %
@@
 time, leading to failure of the solver.  Therefore, we show results only for
-Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.  These three methods give very similar
-results.
+Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.
+{\color{red}
+  These three methods give carbuncle free results.
+  It is worth noting that the solution with Rusanov's solver is the most dissipated solution. 
+  Let 
+  \begin{align}\label{res}
+    \xi(t^{n+1}):=\frac{||h^{n+1}-h^n||_{\ell^2}}{||h^0||_{\ell^2}}
+  \end{align}
+  denote the residual based on the water height. In Figure \ref{fig:residuals_cylinder_2nd}, we plot $\xi(t)$ considering
+  the three Riemann solvers. With HLLEMCC or the blended solver, the residual stagnates. 
+  In Figure \ref{fig:residuals_cylinder_1st} we repeat the same experiment but using the first-order method \eqref{first-order_via_fluct} instead.
+  In this case, the residual stagnates only when we consider the HLLEMCC solver.
+  Clearly, the ability of the blended solver to converge to the steady state solution is affected only by the second-order corrections.
+  In Figure \ref{fig:cylinder_blended}, we zoom close to the front of the shock and show the grid we use.
+  Note that the grid is cartesian and that the shock is resolved within few cells. Therefore, the ability of the method to deliver
+  carbuncle free solutions does not depend on the shock being aligned with a grid interface. 
+}
 
@@
     \quad \quad \quad \quad \quad \quad
-    \subfloat[Blended solver.]{\includegraphics[scale=0.4]{figures/cylinder_ES.png}}
+    \subfloat[Blended solver. \label{fig:cylinder_blended}]{\includegraphics[scale=0.4]{figures/cylinder_ES.png}}
   }
@@
     \label{fig:cylinder_flow}}
+\end{figure}
+
+\begin{figure}[!h]
+  {\scriptsize
+    \subfloat[First-order method \eqref{first-order_via_fluct} \label{fig:residuals_cylinder_1st}]
+             {\includegraphics[scale=0.4]{figures/residuals_cylinder_order1.png}}
+    \subfloat[Second-order method \eqref{second-order_via_fluct} \label{fig:residuals_cylinder_2nd}]
+             {\includegraphics[scale=0.4]{figures/residuals_cylinder_order2.png}}
+  }
+  \caption{
+    Residual $\xi(t)$ \eqref{res} with different Riemann solvers for the problem of the flow past a cylinder.
+    \label{fig:residuals_cylinder}}
 \end{figure}

### 2021-08-14 numerics_paper/version2/numerical_chj.tex

@@
   \end{align}
-  denote the residual based on the water height. In Figure \ref{fig:residuals_cylinder_2nd}, we plot $\xi(t)$ considering
-  the three Riemann solvers. With HLLEMCC or the blended solver, the residual stagnates. 
+  denote the residual based on the water height. In Figure \ref{fig:residuals_cylinder_2nd}, we plot $\xi(t)$ for the
+  solution obtained with each Riemann solver. With HLLEMCC or the blended solver, the residual stagnates. 
   In Figure \ref{fig:residuals_cylinder_1st} we repeat the same experiment but using the first-order method \eqref{first-order_via_fluct} instead.
-  In this case, the residual stagnates only when we consider the HLLEMCC solver.
-  Clearly, the ability of the blended solver to converge to the steady state solution is affected only by the second-order corrections.
-  In Figure \ref{fig:cylinder_blended}, we zoom close to the front of the shock and show the grid we use.
-  Note that the grid is cartesian and that the shock is resolved within few cells. Therefore, the ability of the method to deliver
+  We obtain similar results with Rusanov and HLLEMCC, but now the solution with the blended solver continues to converge
+  at a rate similar to that of Rusanov.  This suggests that the lack of convergence for the 2nd-order method with the
+  blended solver is due to the correction terms and not to the Riemann solver itself.
+  In Figure \ref{fig:cylinder_blended}, we show a closeup of the front of the shock and show the grid we use.
+  Note that the grid is cartesian and that the shock is resolved across a few
+  cells. Therefore, the ability of the method to deliver
   carbuncle free solutions does not depend on the shock being aligned with a grid interface. 
@@
 {\color{OliveGreen}
-  We inspect further the behavior of the blended solver across the shock. 
-  We are interested in the possible presence of an unphysical spike in the momentum.
-  This numerical artifact might lead to a slow displacement of the position of the shock,
+  In Figure \ref{fig:evol_of_hu} we show the time evolution of the momentum in the region around the shock.
+  We see that the momentum is not monotone across the shock; this is a common feature of numerical shock
+  solutions and in some cases is associated with a slow displacement of the position of the shock,
   which is known as slowly-moving shock anomaly; see for instance \cite{navas2019improved, zaide2012numerical}.
-  In the left panels of Figure \ref{fig:evol_of_hu}, we show the time evolution of $hu(x,y=50)$ using the Rusanov and the blended solvers.
-  Clearly, with both solvers we obtain the spike in $hu$. This behavior is expected due to the non-linear Hugoniot locus
-  and the presence of intermediate states through the numerical shock \cite[Section 4.1]{navas2019improved}.
-
-  In \cite{navas2019improved}, it is remarked that spurious oscillations in the transverse direction are introduced with Roe's solver.
-  The authors observed these oscillations by plotting the time evolution of a slice along a fixed value of $x$.
-  The authors conjectured that increasing dissipation of the shear waves reduces these oscillations.
+  This anomaly is, however, not present in the blended solver solution;
+  the position of the shock appears to remain constant (once the steady state
+  is achieved).
+  To further confirm this, we run the simulation up to $t=160$ and plot in
+  Figure \ref{fig:pos_of_shock} the location of the shock along $y=50$.
+  Note that after $t\approx 55$, the location of the shock remains constant.
+
+  Another numerical artifact that has been observed in this context is that of
+  spurious oscillations in the transverse direction, behind the shock; these have been
+  observed for instance when using Roe's solver \cite{navas2019improved}.
   Rusanov's solver adds more dissipation (than the solvers they use) to the shear waves; therefore,
@@
   %
-  Importantly, with the Rusanov and the blended solvers, the position of the shock appears to remain constant
-  (once the steady state is achieved). This is suggested by the parabolic profile of the space-time evolution of $hu(x=18,y)$.
-  To further confirm this, we run the simulation up to $t=160$ and plot in Figure \ref{fig:pos_of_shock} the location of the shock along $y=50$.
-  Note that after $t\approx 55$, the location of the shock remains constant.
 

### 2021-08-14 numerics_paper/version2/numerical_chj.tex
more changes

@@
 we refer to \cite[Section 2.5]{simonnumerical} and \cite{zaide2012numerical}.
+{\color{OliveGreen}
+  Another important shock anomaly is the presence of a spike in the momentum downstream 
+  from the shock. This unphysical behavior might lead to the so called 
+  slowly-moving shock anomaly. We refer to \cite{zaide2012numerical} for a comprehensive 
+  discussion about this phenomenon.
+  In \cite{navas2019improved}, the authors focus on this problem when solving the shallow water equations. 
+}
 
@@
 while maintaining an accuracy similar to that of Roe solver.  
+{\color{OliveGreen}
+  Using Rusanov's solver leads to a non-monotone (and unphysical) profile for the momentum. 
+  Nevertheless, based on our numerical experiments, this behavior does not lead to the slowly-moving 
+  anomaly when using Rusanov's solver. Our blended solver inherets the same properties.
+}
 
@@
 along with outflow conditions along the rest of the right edge of the domain.
-The depth and velocity are set initially and (for all time) at the other
-boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.
+{\color{red} The depth and velocity are set initially and (for all time) at the other
+  boundaries to $h_0=1$, $u_0=5$, $v_0=0$ for a Froude number of 5. Therefore, we do not
+  start with a shock; instead, at the beginning of the simulation, the shock forms
+  near the cylinder and travels to its position at steady state.}
   We use a $160 \times 400$ uniform Cartesian grid.

### 2021-08-15 numerics_paper/version2/numerical_chj.tex

--- 2021-08-14 numerics_paper/version2/numerical_chj.tex
@@
 we refer to \cite[Section 2.5]{simonnumerical} and \cite{zaide2012numerical}.
+{\color{OliveGreen}
+  The slowly-moving shock anomaly is another important unphysical behavior that might be present 
+  with some numerical schemes. We refer to \cite{zaide2012numerical} for a comprehensive 
+  discussion about this phenomenon.
+  In \cite{navas2019improved}, the authors focus on this problem when solving the shallow water equations. 
+}
 
@@
 while maintaining an accuracy similar to that of Roe solver.  
+{\color{OliveGreen}
+  Based on our numerical experiments, using Rusanov's solver does not lead to the 
+  slowly-moving shock anomaly. Our blended solver inherits the same property.
+}
 
@@
 along with outflow conditions along the rest of the right edge of the domain.
-The depth and velocity are set initially and (for all time) at the other
-boundaries to $h_0=1$, $u_0=5$, for a Froude number of 5.
+{\color{red} The depth and velocity are set initially and (for all time) at the other
+  boundaries to $h_0=1$, $u_0=5$, $v_0=0$ for a Froude number of 5. Therefore, 
+  the initial condition does not contain a shock, but due to the presence of the 
+  cylinder a bow shock rapidly forms and travels to its eventual steady-state position.}
   We use a $160 \times 400$ uniform Cartesian grid.

### 2021-08-15 numerics_paper/version2/numerical_chj.tex

@@
 {\color{red}
-  Finally, following \cite[Section III B]{kitamura2009evaluation}, we repeat the same experiment using
-  the blended solver with an initial condition where the shock,
-  located at $r_s=0.3$, is aligned with a grid interface. Afterward, we perturb the initial condition near the shock to misalign it
-  from the grid interface. We use a mesh with $90$ elements in the radial direction and $90$ elements in the angular direction;
-  see Figure \ref{fig:mesh_chj}.
-  For this mesh, $\Delta r=(r_\out-r_\jet)/90=0.01$. 
-  Let $(x_i,y_i)$ denote the coordinates of the center of the $i$-th element and $r_i:=\sqrt{x_i^2+y_i^2}$.
-  For all elements where $r_i\not\in[r_s,r_s+\Delta r]$, we set
+  Next we perform a set of experiments similar to those of \cite[Section III B]{kitamura2009evaluation}.
+  In these experiments, we perturb the initial condition in just one layer of cells, immediately behind the shock.
+  We use an annular mesh (see Figure \ref{fig:mesh_chj}) with $90$ elements in
+  the radial direction and $90$ elements in the angular direction.  Let $h(r)$ denote the exact solution
+  of the ODE \eqref{steady}, and let $r_i$ denote the radius of a given cell center.  Then for all cells
+  away from the shock (i.e. those with $r_i \notin [0.3+\Delta r]$) we use the exact solution averaged over the grid:
   \begin{align*}
@@
   \end{align*}
-  where $\beta=r_\jet h_\jet |u_\jet|$ and $h(r)$ is the solution of the ODE \eqref{steady} from either
-  $r=r_\jet$ to $r=r_s^-$ or $r_s^+$ to $r_\out$.
-  If $r_i = r_s+\Delta r/2$, we set
+  where $\beta=r_\jet h_\jet |u_\jet|$ and $(x_i,y_i)$ are the cartesian
+  coordinates of the cell center.  For the layer of cells just behind the shock (with $0.3 \le r \le 0.3+\Delta r$),
+  we set
   \begin{align*}
-    h_i = \delta h_L + (1+\delta) h_R,
+    h_i = \delta h_{i-1} + (1+\delta) h_{i+1},
     \qquad
@@
   at the left and right cells from the cell with $r_i=r_s+\Delta r/2$, respectively.
-  In Figure \ref{stab_shock_init_cond}, we plot $h(r,\theta_j)$ with $j=\{1,2,\dots,90\}$; that is,
-  we plot the water height as a function of $r$ for every value of $\theta$.
-  In Figures \ref{stab_shock_roe} and \ref{stab_shock_blended}, we plot the corresponding solutions at
-  $t=3$ using method \eqref{second-order_via_fluct} with the Roe solver and with the blended solver,
-  respectively. With the Roe solver, the carbuncle instability appears; as a consequence, the
-  location of the shock changes for the different values of $\theta$.
-  With the blended solver, no carbuncle instability appears and, therefore, all slices are aligned.
-  In the figures we only show the results using $\delta=\{0,0.5,1\}$ but we
-  obtained similar results with the other values of $\delta$.
-
+  In Figure \ref{stab_shock_init_cond}, we plot representative results of these experiments, showing
+  the initial data in the first column and solutions at $t=3$ in the second and third columns.
+  we show only the results using $\delta=\{0,0.5,1\}$ since other results were similar.
+  The solution obtained with the Roe solver, in which the carbuncle instability appears,
+  is sensitive to the perturbation just described, while the solution obtained with the
+  blended solver is radially symmetric and essentially the same for all values of $\delta$.
   \begin{figure}
-    \subfloat[Initial condition.\label{stab_shock_init_cond}]{
+    \subfloat[Initial condition ($t=0$)\label{stab_shock_init_cond}]{
     \begin{tabular}{c}
@@
     ~
-    \subfloat[Roe's solver.\label{stab_shock_roe}]{
+    \subfloat[Roe's solver ($t=3$)\label{stab_shock_roe}]{
     \begin{tabular}{c}
@@
     ~
-    \subfloat[Blended solver.\label{stab_shock_blended}]{
+    \subfloat[Blended solver ($t=3$)\label{stab_shock_blended}]{
     \begin{tabular}{c}
@@
   }    
-    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
-      We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
-      In (a) we show the different initial conditions that we consider.
-      In (b) and (c) we show the corresponding solutions at $t=3$ using method
-      \eqref{second-order_via_fluct} with the Roe solver and with the blended solver, respectively. 
+    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1},
+      using the 2nd-order method \eqref{second-order_via_fluct} with two Riemann solvers. The depth $h$
+      is shown as a function of $r$.  Note that the plots in the third column contain include lines for
+      each angle along the grid, but only a single line is visible due to the solution's symmetry.
       \label{fig:stab_shock}}

### 2021-08-15 numerics_paper/version2/Response_to_Reviewers.tex

@@
 \noindent
-We provide a precise definition of $\theta_i$ in (17) and (19). The guiding principle to select $\theta_i$
+We provide a precise definition of $\theta_i$ in (19) and (21). The guiding principle to select $\theta_i$
 is based on the entropy-residual. As discussed before, using the entropy-residual to locate shocks has proven
 its worth in many works. 
-By selecting $\theta_i=\{0,1\}$, the numerical flux (13) is either Roe's or Rusanov's.
+By selecting $\theta_i=\{0,1\}$, the numerical flux (15) is either Roe's or Rusanov's.
 Having $0<\theta<1$ produces a solver that combines the dissipation of both solvers; i.e.,
@@
 \noindent
-XX
+Our initial condition is given by $h_0=1$, $u_0=5$ and $v_0=0$; i.e., we do not start with a shock. 
+At the beginning of the simulation, the shock forms near the cylinder and travels to its position at 
+steady state. Once at steady state, the shock is not aligned with a grid interface. Instead, the 
+shock is resolved within a few cells (see Figure 3c). 
+Since we start from a condition far from the steady state, 
+perturbing the initital condition or the grid has no effect in the solution. We modified the text in
+Section 4.3 to explain this. In addition, we added a plot of the residual; see equation (29) and 
+Figure 4. 
+The residual stagnates when we use the blended solver with the second-order method (6); however, it 
+continues to decrease if the first-order scheme (7) is considered. This suggests that the residual stagnates 
+not due to the Riemann solver but due to the second-order corrections.
 
@@
 \noindent
-XX
+We added the requested simulation (and cited the corresponding reference) 
+for the first hydraulic jump case; see Figure 10. We obtain a similar result as before; that is, 
+when the Roe solver is used with any value of $\delta$, we obtain the carbuncle instability. 
+With the blended solver, we do not obtain carbuncles for any value of $\delta$.
 
@@
 \noindent
-XX
+We appreciate you brought these references as well as the slowly-moving anomaly to our attention.
+We modified the introduction as suggested. In addition, in the new manuscript we investigate 
+if our blended solver suffers from this problem. 
+Based on our numercal experiments, the profile of the momentum contains the non-physical spike.
+However, the location of steady shocks remains fixed; i.e., our solver does not suffer from 
+the slowly-moving anomaly.
 
@@
 \noindent
-Although we use the term "dry bed" in the section title, we note clearly in the text that we
+Although we use the term ``dry bed'' in the section title, we note clearly in the text that we
 use a non-zero depth $h_r=10^{-15}$.  Although we do not have a proof of positivity preservation
@@
 We have revised the discussion of this result to be more clear and to avoid
-referring to the instability in the simulations as "physical".  We do refer by
+referring to the instability in the simulations as ``physical''.  We do refer by
 way of comparison to a physical instability of the CHJ that has been known for

### 2021-08-15 numerics_paper/version2/numerical_chj.tex

@@
   \begin{align*}
-    h_i = \delta h_{i-1} + (1+\delta) h_{i+1},
+    h_i = \delta h_{L} + (1+\delta) h_{R},
     \qquad
@@
   the initial data in the first column and solutions at $t=3$ in the second and third columns.
-  we show only the results using $\delta=\{0,0.5,1\}$ since other results were similar.
+  We show only the results using $\delta=\{0,0.5,1\}$ since other results were similar.
   The solution obtained with the Roe solver, in which the carbuncle instability appears,
@@
       using the 2nd-order method \eqref{second-order_via_fluct} with two Riemann solvers. The depth $h$
-      is shown as a function of $r$.  Note that the plots in the third column contain include lines for
+      is shown as a function of $r$.  Note that the plots in the third column include lines for
       each angle along the grid, but only a single line is visible due to the solution's symmetry.

### 2021-08-19 numerics_paper/version2/Response_to_Reviewers.tex

@@
       Firstly, it is not clear why blending the Rusanov and Roe methods (Section 3) would fix the carbuncle
-      using entropy-residual as the guiding principle. From literature, it is obvious that Roe-type methods suffer
-      from carbuncle whereas Rusanov-type schemes do not. But why using the entropy-residual somehow
+      using entropy residual as the guiding principle. From literature, it is obvious that Roe-type methods suffer
+      from carbuncle whereas Rusanov-type schemes do not. But why using the entropy residual somehow
       makes the hybrid of the two methods remove the carbuncle? Is entropy-stability a factor (Eq. 13)? If yes,
@@
 \noindent
+Thank you for raising this import point. 
+We have revised the introduction to emphasize our motivation for using the entropy residual as blending
+criteria between Roe's and Rusanov's solvers.
+
 As the reviewer remarks, Rusanov-type solvers do not suffer from carbuncle instabilities. Therefore,
@@
 Near shocks the entropy residual is non-zero, so our blended solver induces dissipation of entropy near shocks,
-similarly to the exact solution. Moreover, if the solver is used with the first order method (7),
+similarly to the exact solution. Moreover, if the blended Riemann solver is used with the first order method (7),
 the solution is provably entropy stable, as discussed in Section 3.3.
 Using entropy stability to avoid carbuncle formation is not a new idea. For example, in \cite{ismail2009,ismail2009v2},
-the authors obtain carbuncle free solvers for Euler by either inducing entropy dissipation or guaranteeing entropy stability. 
+the authors obtain carbuncle-free solvers for the Euler equations by either inducing entropy dissipation or guaranteeing entropy stability. 
 To our knowledge, no work has been done exploring similar ideas for the shallow water equations. However, based on
@@
 %
-We have revised the introduction to emphasize our motivation for using the entropy residual as blending
-criteria between Roe's and Rusanov's solvers.
 
@@
     principle to select the 'appropriate' $\theta$ values? This is somewhat discussed in 3.1 in between Eq. 
-    16-19b but it is not intuitively clear to me how the entropy-residual is used as a guide to select $\theta$.
+    16-19b but it is not intuitively clear to me how the entropy residual is used as a guide to select $\theta$.
     Moreover, what does the value in between $0<\theta<1$ means?
@@
 We provide a precise definition of $\theta_i$ in (19) and (21). The guiding principle to select $\theta_i$
-is based on the entropy-residual. As discussed before, using the entropy-residual to locate shocks has proven
+is based on the entropy residual. As discussed before, using the entropy residual to locate shocks has proven
 its worth in many works. 
-By selecting $\theta_i=\{0,1\}$, the numerical flux (15) is either Roe's or Rusanov's.
 Having $0<\theta<1$ produces a solver that combines the dissipation of both solvers; i.e.,
@@
 \noindent
-Our initial condition is given by $h_0=1$, $u_0=5$ and $v_0=0$; i.e., we do not start with a shock. 
-At the beginning of the simulation, the shock forms near the cylinder and travels to its position at 
-steady state. Once at steady state, the shock is not aligned with a grid interface. Instead, the 
+For this problem, we do not initialize the flow with the steady state solution; we start
+with a uniform flow given by $h_0=1$, $u_0=5$ and $v_0=0$ (i.e., we do not start with a shock).
+After a very short time, a shock forms near the cylinder and travels to its steady state position.
+Once at steady state, the shock is not aligned with a grid interface. Instead, the 
 shock is resolved within a few cells (see Figure 3c). 
 Since we start from a condition far from the steady state, 
-perturbing the initital condition or the grid has no effect in the solution. We modified the text in
+perturbing the initital condition or the grid has no effect on the eventual shock that forms. We modified the text in
 Section 4.3 to explain this.  In addition, we added a plot of the residual; see equation (29) and 
@@
 \noindent
-We have revised the conclusions to make it clear that we do propose an entropy-stable carbuncel-free scheme;
+We have revised the conclusions to make it clear that we do propose an entropy-stable carbuncle-free scheme;
 namely, the first-order method (7).  It is only the second-order method (6) that is not entropy-stable.
-Furthermore, the second-order scheme retains the modification made to the first-order scheme, which is
+The second-order scheme does incorporate the modification made to the first-order scheme, which is
 presumably the reason that it does not exhibit carbuncles.
@@
 \noindent
-We appreciate you brought these references as well as the slowly-moving anomaly to our attention.
+Thank you for bringing these references to our attention.
 We modified the introduction as suggested. In addition, in the new manuscript we investigate 
-if our blended solver suffers from this problem. 
-Based on our numercal experiments, the profile of the momentum contains the non-physical spike.
+whether our blended solver suffers from this problem. 
+Based on our numercal experiments, the profile of the momentum across the shock is non-monotone.
 However, the location of steady shocks remains fixed; i.e., our solver does not suffer from 
@@
 \noindent
-Due to the entropy glitch, we do not converge to the right solution with Roe solver,
+Due to the entropy glitch, Roe's solver does not converge to the correct solution with Roe solver,
 hence the drop in rate of convergence.
-Using any of the other solvers, we converge to the correct solution. The convergence rate
-for those cases is still smaller than in Table 2 because Table 1 is obtained using the first-order scheme. 
+The other solvers do converge to the correct solution, but at a slower rate than
+in Table 2 because Table 1 is obtained using the first-order scheme. 
 We revised the text in Section 4.1 to improve clarity.
@@
 Importantly, with the Rusanov or the blended solvers, the position of the shock appears to remain constant
-(once the steady state is achieved). This is suggested by the parabolic profile of the space-time evolution of $hu(x=18,y)$.
+(once the steady state is achieved).
 To further confirm this, we run the simulation up to $t=160$ and plot the location of the shock along $y=50$.
@@
 both of which produce the spike. Consequently, the blended solver also produces the spike. 
-We believe that, within our framework, the key ingredient to solve this problem is to modify
+We believe that, within our framework, the key ingredient to solve this problem would be to modify
 Rusanov's fluxes to eliminate the spike following for example \cite[Section 5.1]{navas2019improved}.

### 2021-09-21 numerics_paper/version2/numerical_chj.tex
correct small typo regarding setting hu instead of u

@@
 We follow the setup in \cite[\S 4.1.2]{delestre2013swashes}.
-The domain is given by $x\in(0,10)$, and the initial condition is $hu(x,0)=0$ and
+The domain is given by $x\in(0,10)$, and the initial condition is ${\color{cyan}u(x,0)=0}$ and
 \begin{align*}
@@
 The domain is $x\in(0,10)$ and the initial condition is given by
-$hu(x,0)=0$ and
+${\color{cyan} u(x,0)=0}$ and
 \begin{align*}
@@
 The initial condition is
-$h(r,t=0)=0.1$, $hu(r,t=0)=0$, the inner boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$,
+$h(r,t=0)=0.1$, ${\color{cyan}u(r,t=0)=0}$, the inner boundary condition is $h(0.1,t)=0.3$, ${\color{cyan}u(0.1,t)=0.75}$,
 and the outer boundary condition is set to outflow

### 2021-10-18 numerics_paper/version2/Response_to_Reviewers_R2.tex
make first changes related to the 2nd revision

\documentclass[english,11pt]{article}

\usepackage{enumitem}
\usepackage{cite}
\usepackage{bbm}
\usepackage{ulem}
\usepackage{amsthm}
\usepackage[latin9]{inputenc}
\usepackage{babel}
\usepackage[hmargin=0.9in, vmargin=0.75in]{geometry}

\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage[dvipsnames]{xcolor}

\usepackage{url}
\usepackage{bm}

\usepackage{amsmath}
\usepackage[numbers]{natbib}
%\let\cite=\citet
\newcommand{\bfx}{{\bf x}}

\makeatletter
\providecommand{\tabularnewline}{\\}
\makeatother

%%---------------------------------------------------%%
%%---------------------- TITLE ----------------------%%
%%---------------------------------------------------%%
\begin{document}
\title{Response to Reviewers}
\maketitle

\vspace{1.5cm}
\noindent
Michael Dumbser \\
Editor \\
International Journal for Numerical Methods in Fluids \\

\vspace{1.5cm}

\noindent
Manuscript \# FLD-21-0079.R1
\vspace{1.5cm}

\noindent
Dear Prof. Michael Dumbser,

\vspace{0.25cm}
\bigskip
The manuscript entitled
\begin{center}
``Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump''
\end{center}
by David I. Ketcheson and Manuel Quezada de Luna
has been revised following the comments made by the referees.

\bigskip
Letters detailing how the comments of each reviewer have been
addressed are enclosed. To facilitate your perusal of the revised
manuscript, changes concerning the comments of the first and second
reviewers are highlighted in red and green, respectively.

\vspace{0.25cm}
\bigskip\noindent 
With our best regards,

\vspace{0.5cm}
\bigskip\noindent
David I. Ketcheson and Manuel Quezada de Luna

\newpage
%%---------------------------------------------------------%%
%%-------------------- FIRST REVIEWER -------------------- %%
%%---------------------------------------------------------%%
\section*{Revision of Manuscript \# FLD-21-0079.R1 -- Reviewer $\#1$}
We thank the reviewer for the comments and suggestions.
We have incorporated all of them.  For your convenience,
the modifications related to the comments made by reviewer $\#1$ and $\#2$
appear in red and green respectively in the new version of the manuscript. 
Statements in red below are comments from the reviewer.
In this letter, the comments made by the reviewer refer to the original
manuscript and our responses refer to the new manuscript.

{\color{red}
  The authors have closely applied the reviewer's suggestions 
  and the quality of the paper has been improved. This manuscript contains 
  a highly valuable work, very relevant to the journal, which may spur 
  some new ideas in the field of numerical shockwave anomalies. 
  Therefore, I do recommend this paper for publication after a minor correction (see below).

  Before publication, I would ask the authors to remove the sentence in lines 14-15. 
  As far as I understand, the slowly moving shock anomaly does not necessarily 
  involve the oscillation of a steady shock. It also involves the presence of the spike. 
  When the shock is not steady and moves slowly, the spike swings and sheds oscillations 
  downstream. However, when the shock is steady, the spike may remain fixed.
}

\noindent
%
%Following the reviewer's suggestion, we have removed the sentence in lines 14-15.

%%----------------------------------------------------------%%
%%-------------------- SECOND REVIEWER -------------------- %%
%%----------------------------------------------------------%%
\newpage
\section*{Revision of Manuscript \# FLD-21-0079.R1 -- Reviewer $\#2$}
We thank the reviewer for the comments and suggestions.
We have incorporated all of them.  For your convenience,
the modifications related to the comments made by reviewer $\#1$ and $\#2$
appear in red and green respectively in the new version of the manuscript.
Statements in green below are comments from the reviewer.
In this letter, the comments made by the reviewer refer to the original
manuscript and our responses refer to the new manuscript.

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(1)]
    To strengthen the notion of using entropy-control as a means to remove the carbuncle, 
    I strongly suggest to cite the most recent work in removing shock instability using 
    entropy control. Chizari et al.
    (CAMWA 2021) developed a more general (and more multidimensional) entropy-stable finite volume
    scheme which has many benefits including minimal sensitivity to grid skewness and does not 
    suffer from shock-instability as shown for the Quirk's problem (Section 5.3). 
    Perhaps, citation of this paper can be done in page 2 in the authors' paper where 
    references [19,20] are cited.

    Cell-vertex entropy-stable finite volume methods for the system of Euler equations 
    on unstructured grids, CAMWA 2021, Vol 98, pp 261-279.
  \end{itemize}
}

\noindent
Thank you for bringing this reference to our attention. We have included the reference as suggested.

\bigskip
{\color{OliveGreen}
  \begin{itemize}
  \item[(2)]
    For the flow past a cylinder problem, I still believe that the choice grids will be a 
    crucial factor in fully understanding the instability unless one can develop a scheme 
    which is minimally sensitive to grid changes as done by Chizari et al (CAMWA 2021). 
    As shown in that paper, the classic entropy-stable FV scheme will eventually succumb to 
    shock instability if the grids are 'bad' enough but the newly developed entropystable 
    FV method (residual-distribution based) therein does not, since the latter is proven to be 
    almost non-affected by grid degradations.

    In the authors' case. Even on a nice grid like the one used in 4.3, there are some 
    issues with convergence (Figure 4b) using the blended 2nd order scheme where there 
    seem to be limit cycle in the residual convergence. The limit cycle may be an onset 
    for instability if the simulation is ran long enough. This is the classic instability 
    issue that has troubled the carbuncle community for decades. The authors' claim
    that this is due to the second order influence, which is fair enough.

    \begin{itemize}
    \item[(a)]
      However, how do you guarantee that the residual of the second order method will completely decay?
      Fig 4b does not indicate that this will happen. In other words, how do you know if the 
      second-order residual limit cycle will not manifest into shock instability if you ran the 
      simulations longer?      
    \end{itemize}
  \end{itemize}
}
\noindent
For a given mesh, we believe the residual will not decay anymore. 
This is a common problem with many 2nd order schemes that make use of limiters 
based on a posteriori fixes; see for instance \cite{kuzmin2020monolithic}. As stated in the 
paper, we believe the reason for the lack of convergence to a steady state is related the second-order 
correction (including the limiters). This is suggested by the fact that the blended solver with 
the first-order scheme does not prevent convergence to a steady state and since it is well-known that 
some a posteriori limiters inhibit convergence to steady state solutions.

We have run the simulation for a much longer time ($t=16000$ instead of $t=160$) and we still do not 
observe a shock instability. Of course this is not a guarantee that a problem will never occur. 
We have modified the text to discuss this problem.

{\color{OliveGreen}
  \begin{itemize}
    \item[]
    \begin{itemize}
      \item[(b)]
        Looking at the magnitude of residual of $\mathcal{O}(10^{-3})$ for the blended 
        2nd order scheme. This is of the same magnitude with the residuals of the HLLEMCC 
        (Fig 4b). The authors state in page 14 that the HLLEMCC residual stagnates. 
        Is this is a bad thing? If yes, how would it be any different from the similar 
        residual behavior for the second order blended method?
    \end{itemize}
  \end{itemize}
}
\noindent
In the paper we state ``With HLLEMCC or the blended solver, the residual stagnates.'' That is, 
we acknowledge that with both solvers the residual stagnates. In our opinion, the fact that the 
residual stagnates is not necessarily a bad thing, assuming this does not lead to a 
serious shock instability, as remarked by the reviewer. We only have numerical evidence that 
this does not occur. In any case, having a method for which the residual of a steady state solution 
decays to zero (up to machine precision) is desirable. We have started a new line of research to 
fix this problem; however, we believe this new research should be included in a different publication. 
We have modified the text (in Section 4.3 and in the conclusions) to discuss this problem and 
to comment on the new line of work we are pursuing.

{\color{OliveGreen}
  \begin{itemize}
    \item[]
    \begin{itemize}
      Is the blended solution in Fig 3c based on the first or second order method? 
      I suggest to included the first order (Fig 3c?) and second order (Fig 3d?) 
      blended shock profiles to demonstrate the difference between the two on a contour plot.
    \end{itemize}
  \end{itemize}
}
\noindent
The solution is based on the second-order method. As suggested by the reviewer, we have included 
a new figure using the first-order method.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}

### 2021-10-18 numerics_paper/version2/numerical_chj.tex
change beta for C
make first changes related to the 2nd revision

@@
 we refer to \cite[Section 2.5]{simonnumerical} and \cite{zaide2012numerical}.
-{\color{OliveGreen}
 The slowly-moving shock anomaly is another important unphysical behavior that might be present 
@@
 In \cite{navas2019improved}, the authors focus on this problem when solving the shallow water equations.
-}
 
@@
 %as in \cite{guermond2011entropy,guermond2018second,guermond2018well}.
-{\color{red} Since Rusanov's solver does not suffer from carbuncle instabilities,
+Since Rusanov's solver does not suffer from carbuncle instabilities,
 our aim is to use it near shocks without degrading the good accuracy of Roe's solver. 
@@
 Entropy dissipative Riemann solvers have been used to produce carbuncle free solvers for the Euler equations;
-  see for instance \cite{ismail2009affordable,ismail2009proposed}.
-  However, to the best of our knowledge, this idea has not been used before to supress carbuncle instabilities for the
-  shallow water equations. 
-}
+{\color{OliveGreen} see for instance \cite{ismail2009affordable,ismail2009proposed, chizari2021cell}.}
+However, to the best of our knowledge, this idea has not been used before to supress carbuncle instabilities for the shallow water equations.
+
 Using this Riemann solver (within the second-order Lax-Wendroff-LeVeque 
@@
 while maintaining an accuracy similar to that of Roe solver.  
-{\color{OliveGreen}
 Based on our numerical experiments, using Rusanov's solver does not lead to the 
 slowly-moving shock anomaly. Our blended solver inherits the same property.
-}
 
@@
 on Rusanov's solver are given by \eqref{first-order_via_fluct} and \eqref{second-order_via_fluct}, respectively.
-{\color{red}
 The second-order method can also be written in flux-differencing form \eqref{flux-differencing-form}, where 
@@
 \end{align}
-}
 
@@
 \end{align}
-{\color{red}
 The first-order Roe's fluxes used in \eqref{flux-differencing-form} and \eqref{first-order_via_fluct} are given as follows:
@@
 behaves more like that of Rusanov or Roe.
-{\color{red}
 Specifically, if $\lambda_\iph^{\min}=0$ and $\theta_\iph=1$, the flux \eqref{blended-flux-1}
@@
 With $0<\theta<1$ the blended solver is more dissipative than Roe's but less than Rusanov's.
-}
 With the choice \eqref{s_p}, the correction fluxes also match
@@
 10^{-15}$ to avoid division by zero.
-{\color{OliveGreen}
 If we use Rusanov's solver with the first-order method \eqref{first-order_via_fluct}, 
@@
 Although using Roe or the blended solvers might still lead to loss of positivity, we 
-did not observe that in our experiments.}
+did not observe that in our experiments.
 The exact solution, which can also be found in \cite{delestre2013swashes} and references therein, is
@@
 In Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test.
-{\color{OliveGreen}
 Due to the entropy glitch, we do not converge to the right solution with Roe solver,
@@
 correct solution. Note that using the blended solver leads to more accurate results. 
-    Since the method we use is first-order accurate, we get no better than first-order convergence.}
+Since the method we use is first-order accurate, we get no better than first-order convergence.
 
@@
 along with outflow conditions along the rest of the right edge of the domain.
-{\color{red} The depth and velocity are set initially and (for all time) at the other
+The depth and velocity are set initially and (for all time) at the other
 boundaries to $h_0=1$, $u_0=5$, $v_0=0$ for a Froude number of 5. Therefore, 
 the initial condition does not contain a shock, but due to the presence of the 
-  cylinder a bow shock rapidly forms and travels to its eventual steady-state position.}
+cylinder a bow shock rapidly forms and travels to its eventual steady-state position.
 We use a $160 \times 400$ uniform Cartesian grid.
 %
-In Figure \ref{fig:steady_outflow}, we show results at $t=80$, after the flow has reached
+In Figure \ref{fig:cylinder_flow}, we show results at $t=80$, after the flow has reached
 a steady state.  With Roe's solver, negative values (of the water height) are generated at an early
-time, leading to failure of the solver.  Therefore, we show results only for
-Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.
-{\color{red}
+time, leading to failure of the solver.  Therefore, we show results only with the 
+{\color{OliveGreen} second-order method}
+\eqref{second-order_via_fluct} using Rusanov's, HLLEMCC by \cite{kemm2014note}, and the blended solvers.
 These three methods give carbuncle free results.
 It is worth noting that the solution with Rusanov's solver is the most dissipated solution. 
+{\color{OliveGreen}
+For comparison, we also show in Figure \ref{fig:cylinder_blended_1st_order} 
+the solution using the first-order method \eqref{first-order_via_fluct}
+with the blended solver.}
 Let 
@@
 blended solver is due to the correction terms and not to the Riemann solver itself.
+{\color{OliveGreen} We have performed a similar experiment (not shown here) where we ran the simulation
+up to a final time of $t=16000$ using the second-order method with the blended solver. Although, the 
+residual $\xi(t)$ did not converge to zero, the simulation did not become unstable. Indeed, the 
+final solution is indistinguishable from the solution in Figure \ref{fig:cylinder_blended}.}
+
 In Figure \ref{fig:cylinder_blended}, we show a closeup of the front of the shock and show the grid we use.
@@
 carbuncle free solutions does not depend on the shock being aligned with a grid interface.
-}
 
@@
     \subfloat[Rusanov's solver.]{\includegraphics[scale=0.4]{figures/cylinder_rusanov.png}}
-    \quad \quad \quad \quad \quad \quad
+    \quad ~~
     \subfloat[HLLEMCC solver.]{\includegraphics[scale=0.4]{figures/cylinder_hllemcc.png}}
-    \quad \quad \quad \quad \quad \quad
+    \quad ~~
     \subfloat[Blended solver. \label{fig:cylinder_blended}]{\includegraphics[scale=0.4]{figures/cylinder_ES.png}}
+    \quad ~~
+    \subfloat[Blended solver. \label{fig:cylinder_blended_1st_order}]{\includegraphics[scale=0.4]{figures/cylinder_ES_1st_order.png}}
   }
   \caption{
-    Flow past a cylinder.  Results for Rusanov's, HLLEMCC, and the blended solvers.
+    Flow past a cylinder. In (a)-(c), results using \eqref{second-order_via_fluct} 
+    with Rusanov's, HLLEMCC, and the blended solvers.
+    In (d), results using \eqref{first-order_via_fluct} with the blended solver.
     \label{fig:cylinder_flow}}
@@
 
-{\color{OliveGreen}
 In Figure \ref{fig:evol_of_hu} we show the time evolution of the momentum in the region around the shock.
@@
 \end{figure}
-}
 
@@
     rhu & = C \\
-    h'(r) & = \frac{h(r)}{\frac{g}{\beta^2} r^3 (h(r))^3 -r} = \frac{h(r)}{r} \cdot \frac{(F(r))^2}{1-(F(r))^2} \label{eq:dh}
+    h'(r) & = \frac{h(r)}{\frac{g}{{\color{cyan} C}^2} r^3 (h(r))^3 -r} = \frac{h(r)}{r} \cdot \frac{(F(r))^2}{1-(F(r))^2} \label{eq:dh}
 \end{align}
@@
 all of which are free from carbuncles.
-{\color{OliveGreen} In Figure \ref{fig:chj_r1_slices}, we plot the water height and the x-momentum along $y=0$.
+In Figure \ref{fig:chj_r1_slices}, we plot the water height and the x-momentum along $y=0$.
 Clearly, the solution based on Rusanov's solver is more dissipated. From the plot, it is also clear the presence
 of the non-physical spike in the momentum. Just as in Section \ref{sec:bow_shock}, in our numerical experiments,
-  this spike does not lead to a slowly-moving shock.}
+this spike does not lead to a slowly-moving shock.
 
@@
 
-{\color{red}
 Next we perform a set of experiments similar to those of \cite[Section III B]{kitamura2009evaluation}.
@@
 The solutions obtained with these solvers show no carbuncles.
-{\color{OliveGreen} Note that, similar to the experiments in Section \ref{sec:bow_shock},
-  the non-physical spike in the momentum is present in the numerical solution.}
+Note that, similar to the experiments in Section \ref{sec:bow_shock},
+the non-physical spike in the momentum is present in the numerical solution.
 The Rusanov solution remains very close to the initial symmetric equilibrium state,
@@
 
-{\color{OliveGreen} 
 Solutions are shown in Figures \ref{fig:chj_r2wp_rusanov} and \ref{fig:chj_r2wp_blended},
@@
 the manifestation of this instability.
-}
 
@@
 robustness similar to that of Rusanov's method.
-{\color{red}
 Using this Riemann solver with the first-order method \eqref{first-order_via_fluct} we
@@
 so, to avoid further complexity, we did not pursue this idea. 
-}
 Although the full second-order method we have proposed is not 
@@
 Euler equations.
-{\color{OliveGreen}
 For future work, we plan to extend these results to the shallow water equations 
@@
 and later extended by Zalesak \cite{zalesak1979fully}.
+{\color{OliveGreen}
+We are also interested in obtaining a method that converges (up to machine precision) 
+to steady states. To do this we plan to use the monolithic limiter from \cite{kuzmin2020monolithic}.}
 Finally, we are interested in reducing (or eliminating) the non-physical spike in
 the momentum. To do this we plan to modify Rusanov's solver based on \cite[Section 5.1]{navas2019improved}.
-}
 

### 2021-10-24 numerics_paper/version2/numerical_chj.tex
Add modification based on the comments from the 1st reviwer

@@
 while maintaining an accuracy similar to that of Roe solver.  
-Based on our numerical experiments, using Rusanov's solver does not lead to the 
-slowly-moving shock anomaly. Our blended solver inherits the same property.
+Based on our numerical experiments, 
+{\color{red} the Rusanov's solver does not perturb the location of steady shocks.} 
+Our blended solver inherits the same property.
 

### 2021-10-25 numerics_paper/version2/numerical_chj.tex
Remove cyan coloring.

@@
 We follow the setup in \cite[\S 4.1.2]{delestre2013swashes}.
-The domain is given by $x\in(0,10)$, and the initial condition is ${\color{cyan}u(x,0)=0}$ and
+The domain is given by $x\in(0,10)$, and the initial condition is $u(x,0)=0$ and
 \begin{align*}
@@
 The domain is $x\in(0,10)$ and the initial condition is given by
-${\color{cyan} u(x,0)=0}$ and
+$u(x,0)=0$ and
 \begin{align*}
@@
     rhu & = C \\
-    h'(r) & = \frac{h(r)}{\frac{g}{{\color{cyan} C}^2} r^3 (h(r))^3 -r} = \frac{h(r)}{r} \cdot \frac{(F(r))^2}{1-(F(r))^2} \label{eq:dh}
+    h'(r) & = \frac{h(r)}{\frac{g}{C^2} r^3 (h(r))^3 -r} = \frac{h(r)}{r} \cdot \frac{(F(r))^2}{1-(F(r))^2} \label{eq:dh}
 \end{align}
@@
 The initial condition is
-$h(r,t=0)=0.1$, ${\color{cyan}u(r,t=0)=0}$, the inner boundary condition is $h(0.1,t)=0.3$, ${\color{cyan}u(0.1,t)=0.75}$,
+$h(r,t=0)=0.1$, $u(r,t=0)=0$, the inner boundary condition is $h(0.1,t)=0.3$, $u(0.1,t)=0.75$,
 and the outer boundary condition is set to outflow

### 2022-01-26 numerics_paper/version2/numerical_chj.tex
Add references to simulations of reverse CHJ.

@@
 can also be completely steady for sufficiently small velocities and depths.
+The reverse flow (with a fast, inward flow leading to a similar
+circular jump, but with the region of fast on the outside of the jump) has also been
+studied using the shallow water model \cite{foglizzo2012shallow} and
+an extendion for flows with shear \cite{ivanova2019structure,busto2021thermodynamically},
+with instability observed in both cases.
 

### 2022-01-27 numerics_paper/version2/numerical_chj.tex
Fix typo.

--- 2021-10-25 numerics_paper/version2/numerical_chj.tex
@@
 can also be completely steady for sufficiently small velocities and depths.
+The reverse flow (with a fast, inward flow leading to a similar
+circular jump, but with the region of fast on the outside of the jump) has also been
+studied using the shallow water model \cite{foglizzo2012shallow} and
+an extension for flows with shear \cite{ivanova2019structure,busto2021thermodynamically},
+with instability observed in both cases.
 
